### 2021-04-11 dev_vignettes/Bushplus_strains.Rmd
testing

---
title: "Biodiversity, eco-evo feedback, and the stability of responses to environmental change"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(here)
library(deSolve)
library(rootSolve)

set.seed(13)

run_sims <- FALSE
```

# Introduction

This is relevant: https://www.biorxiv.org/content/10.1101/2021.03.22.436420v1

Aims:

* Investigate the effects of biodiversity within each of the three functional groups on responses to environmental change.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

We have added:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.
* Option to include biodiversity within functional groups.

# Methods

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Dynamic model

```{r}
source(here("functions/bushplus_dynamic_model.r"))
## contains various functions:
## bushplus_dynamic_model
## bushplus_dynamic_model_strains
```

## Defaults

### Default parameter values

To do:

* Make a function that can create parameter value sets, including for example controlling the number and types of strains within each functional group.

```{r}
default_parameter_values <- c(

  ## Maximum growth rate
  g_max_CB_0 = 0.05,
  g_max_CB_1 = 0.05,
  g_max_PB_0 = 0.07,
  g_max_PB_1 = 0.07,
  g_max_SB_0 = 0.10,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB_0 = 300,
  h_SR_CB_1 = 300,
  h_O_PB_0 = 100,
  h_O_PB_1 = 100,
  h_O_SB_0 = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  P_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB = 0,
  i_PB = 0,
  i_SB = 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5

  )
```

### Default initial state

To do:

* Make a function that can create initial conditions corresponding to the number of strains in the parameter values, and assign initial state values according to some rule (e.g. total abundance and even distribution of among strains).

Initial conditions favouring anoxic

```{r}
default_initial_state <- c(
  CB_0 = 5e1,
  CB_1 = 0,
  PB_0 = 1e7,
  PB_1 = 0,
  SB_0 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1)
```

### Default event

#### Default event definition

Used to add noise and to ensure minimum abundance.

Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

```{r}
source(here("functions/default_event_definition.r"))
## contains two functions:
## default_event_definition
## default_event_definition_strains -- this should work for any number of strains
```

#### Default event schedule

This is when an event happens

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimums

```{r}
## we only have one minimum abundance for each functional group
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB",
                                       "PB",
                                       "SB") 
```

### Default simulation duration

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 10
```

### Default oxygen diffusivity (a) forcing

Here we specify no forcing of a_0 (used unless set elsewhere).

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
default_log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(default_sim_duration),
                                            length=length(default_log10a_series))),
                                default_log10a_series))

```

## Simulation function

```{r}
source(here("functions/run_simulation.r"))
```

## Helper functions

```{r}
source(here("functions/plot_dynamics.r"))
source(here("functions/ss_expt_a_N.r"))
source(here("functions/get_nonlinearity.r"))
source(here("functions/get_hysteresis_range.r"))
```

## Code test

```{r}
dynamic_model = bushplus_dynamic_model_strains
event_definition = default_event_definition_strains
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_forcing = default_log10a_forcing
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing_func <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing_func, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
                           event_definition = default_event_definition_strains,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation()

plot_dynamics_strains(sim_res)

```

# Results: Reproducing previous results

Here we use the default parameter values, which includes two strains, but with same parameters. Results should be the same as with only one strain.

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(

  CB_0 = 5e1,
  CB_1 = 0,
  PB_0 = 1e7,
  PB_1 = 0,
  SB_0 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1

  )

sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
plot_dynamics_strains(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

## Initial conditions favouring anoxic, with noise, no forcing

Figure 2a, b in Bush et al, but with noise

```{r}
## add some noise
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          noise_sigma = noise_sigma_0.01)
plot_dynamics_strains(sim_res)
#ggsave("high_SB_PB.png")

noise_sigma_0.01 <- 0.00

```

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r}
noise_sigma_0.01 <- 0.00

init_state_favouring_oxic <- c(
  CB_0 = 1e7,
  CB_1 = 0,
  PB_0 = 1e5,
  PB_1 = 0,
  SB_0 = 1e5,
  SO = 500,
  SR = 50,
  O = 30,
  P = 10
  )

sim_res <- run_simulation(initial_state = init_state_favouring_oxic)

plot_dynamics_strains(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Initial conditions favouring oxic, with noise, no forcing

Figure 2c, d in Bush et al. but with noise.

```{r} 
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          noise_sigma = noise_sigma_0.01)

plot_dynamics_strains(sim_res)
##ggsave("high_CB.png")
noise_sigma_0.01 <- 0.00
```

# Results: two different CB strains

## Starting oxic steady state

```{r}
names(default_parameter_values)
two_CB_strains_parvals <- default_parameter_values
two_CB_strains_parvals["g_max_CB_1"] <- 0.0495
two_CB_strains_parvals["h_SR_CB_1"] <- 315
two_CB_strains_parvals

two_CB_init_state_favouring_oxic <- c(
  CB_0 = 0.75e7,
  CB_1 = 0.25e7,
  PB_0 = 1e5,
  PB_1 = 0,
  SB_0 = 1e5,
  SO = 500,
  SR = 50,
  O = 30,
  P = 10
  )

sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = two_CB_strains_parvals)

plot_dynamics_strains(sim_res)

```

## Forcing
```{r}

## simulation timing
simulation_duration <- 70000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -10, -10, -10, -10, -10, -10)

## run simulation with both CB strains having default growth rate
## and tolerance
sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = default_parameter_values,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)

plot_dynamics_strains(sim_res)
```

```{r}
## run simulation with CB strains having different growth rate
## and tolerance
sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
                          parameter_values = two_CB_strains_parvals,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)

plot_dynamics_strains(sim_res)

```

# TO DO:

* Check what happens when there are more strains.
* Be able to find the stable states.

# Results: Composition/diversity and nature of response

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 100
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

### Each composition

#### No organisms

```{r eval = run_sims}

initial_CBs <- 0
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_no_biology <- ss_by_a_N(expt)
saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
```

```{r}
ss_no_biology <- readRDS("data/ss_no_biology.RDS")
```

```{r}
ss_no_biology %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_no_biology %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_CB <- ss_by_a_N(expt)
saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
```

```{r}
ss_only_CB <- readRDS("data/ss_only_CB.RDS")
```

```{r}
tempx1 <- ss_only_CB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  geom_path(data = filter(tempx1, Species == "PB"),
            aes(y=0.2), size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only CB")

ggsave("figures/CB.png", width = 3, height = 2)
```

```{r}
ss_only_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 0
initial_SBs <- 10^seq(0, 8, length=grid_num_N)
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_SB <- ss_by_a_N(expt)

saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
```

```{r}
ss_only_SB <- readRDS("data/ss_only_SB.RDS")
```

```{r}
ss_only_SB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_PB <- ss_by_a_N(expt)

saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
```

```{r}
ss_only_PB <- readRDS("data/ss_only_PB.RDS")
```

```{r}
tempx1 <- ss_only_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size = 0.8) +
  theme_light() +
  ylim(0,10) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  geom_path(data = filter(tempx1, Species == "CB"),
            aes(y=0.2), size = 0.8)+
  geom_path(data = filter(tempx1, Species == "PB"),
            size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only PB")

ggsave("figures/PB.png", width = 3, height = 2)

```

```{r}
ss_only_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_PB

ss_only_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_CB

ss_only_SB_CB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:5) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_CB <- expt$N_PB

ss_only_CB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB, SB, and PB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB") 
ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### All response measures

```{r eval = run_sims}
all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
                    composition = rep("No organisms", grid_num_a*grid_num_N),
                    direction = c(rep("up", grid_num_a),
                                  rep("down", grid_num_a)))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB.RDS"),
                              composition = rep("CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB.RDS"),
                              composition = rep("SB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_PB.RDS"),
                              composition = rep("PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
                              composition = rep("SB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
                              composition = rep("SB-CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
                              composition = rep("CB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
                              composition = rep("CB-SB-PB",
                                                grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))

x_long <- all_ss %>%
  gather(key = Species, value = Quantity, 2:8)
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(composition, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(composition = fct_relevel(composition,
                                   "No organisms",
                                   "CB",
                                   "SB",
                                   "PB",
                                   "CB-PB",
                                   "SB-CB",
                                   "SB-PB",
                                   "CB-SB-PB"),
         num_func_groups = case_when(composition =="No organisms" ~ 0,
                                     composition == "CB" ~ 1,
                                     composition == "SB" ~ 1,
                                     composition == "PB" ~ 1,
                                     composition == "CB-PB" ~ 2,
                                     composition == "SB-CB" ~ 2,
                                     composition == "SB-PB" ~ 2,
                                     composition == "CB-SB-PB" ~ 3
                                     ),
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
saveRDS(resp_summary, "data/resp_summary.RDS")
```

```{r}
resp_summary <- readRDS("data/resp_summary.RDS")
resp_summary %>%
  ggplot() +
  geom_point(aes(x = composition, y = value)) +
  facet_grid(Species ~ resp_measure, scales = "free") +
  coord_flip()

```

### Nonlinearity

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(composition, Species, num_func_groups, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(composition, ave_nl)) +
  facet_wrap( ~ Species)

```

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = sum(value) / unique(num_func_groups))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, composition) %>%
  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(composition) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_mean_hyst)) +
  xlab("Functional groups present") +
  ylab("Amount of hysteresis") +
  ylim(0,2) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)

```

# Results: Varying strength of inhibition

## Introduction

Three parameters set the strength of inhibition (half-inhibition constants):

* h_SR_CB = 300,
* h_O_PB = 100,
* h_O_SB = 100,

These work by the inhibition function:

$\frac{1}{1 + \frac{X}{h_X}}$

```{r}
inhibition
```

Plot some examples:

```{r}
default_parameter_values["h_SR_CB"]
SR <- 10^seq(1.4, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
  ylab("Multiplicative effect\non CB growth rate")
```

```{r}
default_parameter_values["h_O_PB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
  ylab("Multiplicative effect\non PB growth rate")
```

```{r}
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

```{r}
proportion_change <- 2
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

## Experiment

Increasing the inhibition half saturation constants weakens the inhibition.

```{r}
ssfind_minimum_abundances <- default_minimum_abundances
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = NA,
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

grid_num_mult <- 20
grid_num_N <- 2
grid_num_a <- 100

a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5

expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    inhib_const_mult = inhib_const_mults,
                    a_O = a_Os)
```

```{r eval = run_sims}
ss_h1 <- ss_by_a_h_N(expt)

saveRDS(ss_h1, "data/ss_h1.RDS")
```

```{r}
ss_h1 <- readRDS("data/ss_h1.RDS")
```

## Results

### Close to default inhibition

```{r}
inhib_const_mults
focal_inhibition_multiplier <- inhib_const_mults[14]

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Double strength inhibition

```{r}
## YES! 0.5 does mean twice as strong inhibition
focal_inhibition_multiplier <- 0.5

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Weaker inhibition

```{r}
## YES! 2 does mean half as strong inhibition
focal_inhibition_multiplier <- 2^0.4

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Nonlinearity

```{r}
x_long <- ss_h1 %>%
  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
  gather(key = Species, value = Quantity, 3:9) 
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(inhib_const_mult, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
```

#### All species nonlinearity

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(inhib_const_mult, Species, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_nl)) +
  facet_wrap( ~ Species)

```

#### Substrate nonlinearity

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### Organism nonlinearity

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

#### System nonlinearity

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```

### Hysteresis

#### Substrate hysteresis

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### Organism hysteresis

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult,  resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```

#### System hysteresis (separate)

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, inhib_const_mult) %>%
  summarise(mean_hyst = mean(value))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```

#### System hysteresis (combined)

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(inhib_const_mult) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_mean_hyst)) +
  xlab("Organismal tolerance") +
  ylab("Amount of hysteresis") +
  theme_light() +
  ylim(0,2)
ggsave("figures/inhibition_system_hysteresis.png", width = 2.5, height = 2)

```

# Results: Temporal state switching

## Switch anoxic-oxic-anoxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_anoxic

## simulation timing
simulation_duration <- 5000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          log10a_series = lhl_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/lhl_with_defaults.RDS")
```

```{r}
lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
plot_dynamics(lhl_with_defaults)
```

```{r eval = run_sims}
lhl_with_defaults$result %>%
  filter(between(time, 1000, 73000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

## Switch oxic-anoxic-oxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_oxic

## simulation timing
simulation_duration <- 50000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/hlh_with_defaults.RDS")
```

```{r}
hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
plot_dynamics(hlh_with_defaults)
```

```{r eval = run_sims}
hlh_with_defaults$result %>%
  filter(between(time, 1000, 100000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

# Appendices

## runsteady function not finding correct values

```{r eval = FALSE}

init_state_anoxic_2 <- c(N_CB = 1e1,
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 300,
           SR = 300,
           O = 1e-5,
           P = 1e1)

RS <- runsteady(y = init_state_anoxic_2,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y
RS <- runsteady(y = init_state_favouring_oxic,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y

```

## Separatrix

Note that this section not working / updated since the code was made into functions

### Effect of N_CB and oxygen diffusability

```{r eval = FALSE}

noise_sigma <- 0.0

get_final_states_N_CB <- function(x) {

  #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_CB"] <- x["N_CB"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)

  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                    times = times,
                    func = micro_mod_var,
                    parms = parameters,
                    method = "radau",
                    events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 50000)

state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_CB and oxygen diffusability
grid_num <- 50
expt <- expand.grid(N_CB = 10^seq(0, 8, length=grid_num),
                    a_O = 10^seq(-3.5, -2, length=grid_num))

result <- apply(expt, 1, function(x) get_final_states_N_CB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_CB=expt$N_CB,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_CB>2e8, 'high', 'low'))

result <- result %>%
  mutate(N_CB = ifelse(N_CB < minimum_abundance, minimum_abundance, N_CB))

result$highlow <- as.factor(result$highlow)
saveRDS(result, "data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result <- readRDS("data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}

#plot N_CB / P_CB against a_0
p <- ggplot(result) + 
  geom_line(aes(x=log10(a_O), y=log10(N_CB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial cyanobacteria abundance')
p <- p + geom_point(data=result, aes(x=log10(a_O), y=log10(initial_N_CB), color=highlow)) + labs(color = "Final CB concentration")
plot(p)
#ggsave("initial conditions_1.png")
```

```{r eval = FALSE}
#show hysterisis effects for all the parameters
#plot oxygen against diffusability
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(O))))+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxygen concentration]')+
  geom_segment(aes(x=log10(result$a_O[839]), y=log10(result$O[839]), xend=log10(result$a_O[838]), yend=log10(result$O[838])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=log10(result$a_O[1761]), y=log10(result$O[1761]), xend=log10(result$a_O[1762]), yend=log10(result$O[1762])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=-4.25, y=-0.45, xend=-3.75, yend=0.1), arrow=arrow(), colour="blue", size=0.5)+
  geom_curve(aes(x = -3.75, y = 1.9, xend = -4.25, yend = 1.5), arrow=arrow(), colour='blue', size = 0.5, curvature = -0.2)
#ggsave("hysterisis_CB_1.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(P))))+
  ggtitle("Hysterisis effect of phoshate concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[phosphate concentration]')
#ggsave("hysterisis_CB_2.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SO))))+
  ggtitle("Hysterisis effect of oxidized sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxydized sulfur concentration]')
#ggsave("hysterisis_CB_3.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SR))))+
  ggtitle("Hysterisis effect of reduced sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[reduced sulfur] concentration]')
#ggsave("hysterisis_CB_4.png")
```

These graphs show an at least qualitative match to the result in the paper.

### Effect of PB/SB depending on oxygen diffusability

```{r eval = FALSE}
noise_sigma <- 0.00
get_final_states_N_PB <- function(x) { 
  #give in N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_PB"] <- x["var"]
  state["N_SB"] <- x["var"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)
  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                         times = times,
                         func = micro_mod_var,
                         parms = parameters,
                         method = "radau", events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 5000)
state <- c(N_CB = 5, 
           N_PB = NA,#this one is varied
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_PB and oxygen diffusivity
grid_num <- 50
expt <- expand.grid(var=10^seq(0, 7, length=grid_num),
                    a_O = 10^seq(-3, -2, length=grid_num))

result2 <- apply(expt, 1, function(x) get_final_states_N_PB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_PB=expt$var,
                                 initial_N_SB=expt$var,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_PB>1e8, 'high', 'low'))

result2 <- result2 %>%
  mutate(N_PB = ifelse(N_PB < minimum_abundance, minimum_abundance, N_PB))

result2$highlow <- as.factor(result2$highlow)
saveRDS(result2, "data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result2 <- readRDS("data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
#plot N_PB against a_0
p <- ggplot(result2) + 
  geom_line(aes(x=log10(a_O), y=log10(N_PB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial PS and SR bacteria concentration')
p <- p + geom_point(data=result2,
                    aes(x=log10(a_O), y=log10(initial_N_PB), color=highlow))
p <- p + labs(color = "Final PB and SB concentration")
plot(p)
#ggsave("initial conditions_2.png")
```

## EWS analyses

Evaluation of the switch: 
Get the functions: 
```{r eval = F}
source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
```

Get state shifts and calculate parameters: 
```{r eval = F}
x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)

#first switch
name_1='anoxic_oxic 1'
trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
out_df_1_N_CB <- data.frame(trans1$N_CB)
colnames(out_df_1_N_CB) <- c("N_CB")
out_df_1_N_PB <- data.frame(trans1$N_PB)
colnames(out_df_1_N_PB) <- c("N_PB")
out_df_1_N_SB <- data.frame(trans1$N_SB)
colnames(out_df_1_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
  ggtitle("Anoxic_oxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 1.png")

#second switch
name_2='oxic_anoxic 1'
trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
n_23 <- round(length(trans2$time)*x3/100) 
t_23 <- length(trans2$time) - n_23 + 1
n_25 <- round(length(trans2$time)*x5/100) 
t_25 <- length(trans2$time) - n_25 + 1
out_df_2_N_CB <- data.frame(trans2$N_CB)
colnames(out_df_2_N_CB) <- c("N_CB")
out_df_2_N_PB <- data.frame(trans2$N_PB)
colnames(out_df_2_N_PB) <- c("N_PB")
out_df_2_N_SB <- data.frame(trans2$N_SB)
colnames(out_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
  ggtitle("Oxic_anoxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 1.png")

#third switch
name_3='oxic_anoxic 2'
trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
n_33 <- round(length(trans3$time)*x3/100) 
t_33 <- length(trans3$time) - n_33 + 1
n_35 <- round(length(trans3$time)*x5/100) 
t_35 <- length(trans3$time) - n_35 + 1
out2_df_1_N_CB <- data.frame(trans3$N_CB)
colnames(out2_df_1_N_CB) <- c("N_CB")
out2_df_1_N_SB <- data.frame(trans3$N_SB)
colnames(out2_df_1_N_SB) <- c("N_SB")
out2_df_1_N_PB <- data.frame(trans3$N_PB)
colnames(out2_df_1_N_PB) <- c("N_PB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
  ggtitle("Oxic_anoxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 2.png")

#fourth switch
name_4='anoxic_oxic 2'
trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
n_43 <- round(length(trans4$time)*x3/100) 
t_43 <- length(trans4$time) - n_43 + 1
n_45 <- round(length(trans4$time)*x5/100) 
t_45 <- length(trans4$time) - n_45 + 1
out2_df_2_N_CB <- data.frame(trans4$N_CB)
colnames(out2_df_2_N_CB) <- c("N_CB")
out2_df_2_N_PB <- data.frame(trans4$N_PB)
colnames(out2_df_2_N_PB) <- c("N_PB")
out2_df_2_N_SB <- data.frame(trans4$N_SB)
colnames(out2_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
  ggtitle("Anoxic_oxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 2.png")
```

Rolling window analysis
```{r eval = F}
#analysis for 1st
rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)

```

Metric based indicators from the earlywarnings package
```{r eval = F}
#analysis for 1st
earlywarnings_plot(out_df_1_N_CB, name_1, x3)
earlywarnings_plot(out_df_1_N_PB, name_1, x3)
earlywarnings_plot(out_df_1_N_SB, name_1, x3)

#analysis for 2nd
earlywarnings_plot(out_df_2_N_CB, name_2, x3)
earlywarnings_plot(out_df_2_N_PB, name_2, x3)
earlywarnings_plot(out_df_2_N_SB, name_2, x3)

#analysis for 3rd
earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
earlywarnings_plot(out2_df_1_N_SB, name_3, x3)

#analysis for 4th
earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
```

Conditional Heteroskedasticity 
```{r eval = F}
#analysis for 1st
ch <- condhet_analysis(out_df_1_N_CB, name_1)
ch <- condhet_analysis(out_df_1_N_PB, name_1)
ch <- condhet_analysis(out_df_1_N_SB, name_1)

#analysis for 2nd
ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
ch2 <- condhet_analysis(out_df_2_N_SB, name_2)

#analysis for 3rd
ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)

#analysis for 4th
ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
```

Detrending fluctuation analysis 
```{r eval = F}
#analysis for 1st
dfa_analysis(trans1, 2, n_15, t_15, name_1)
dfa_analysis(trans1, 3, n_15, t_15, name_1)
dfa_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
dfa_analysis(trans2, 2, n_25, t_25, name_2)
dfa_analysis(trans2, 3, n_25, t_25, name_2)
dfa_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
dfa_analysis(trans3, 2, n_35, t_35, name_3)
dfa_analysis(trans3, 3, n_35, t_35, name_3)
dfa_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
dfa_analysis(trans4, 2, n_45, t_45, name_4)
dfa_analysis(trans4, 3, n_45, t_45, name_4)
dfa_analysis(trans4, 4, n_45, t_45, name_4)
```

Drift Diffusion Jump Nonparametrics Early Warning Signals:
```{r eval = F}
#shift 1
ddj11<- ddj_model(out_df_1_N_CB)
plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
plot_ddj_2(out_df_1_N_CB, ddj11, name_1)

ddj12<- ddj_model(out_df_1_N_PB)
plot_ddj_1(out_df_1_N_PB, ddj12, name_1)
plot_ddj_2(out_df_1_N_PB, ddj12, name_1)

ddj13<- ddj_model(out_df_1_N_SB)
plot_ddj_1(out_df_1_N_SB, ddj13, name_1)
plot_ddj_2(out_df_1_N_SB, ddj13, name_1)

#shift 2
ddj21<- ddj_model(out_df_2_N_CB)
plot_ddj_1(out_df_2_N_CB, ddj21, name_2)
plot_ddj_2(out_df_2_N_CB, ddj21, name_2)

ddj22<- ddj_model(out_df_2_N_PB)
plot_ddj_1(out_df_2_N_PB, ddj22, name_2)
plot_ddj_2(out_df_2_N_PB, ddj22, name_2)

ddj23<- ddj_model(out_df_2_N_SB)
plot_ddj_1(out_df_2_N_SB, ddj23, name_2)
plot_ddj_2(out_df_2_N_SB, ddj23, name_2)

#shift 3
ddj31<- ddj_model(out2_df_1_N_CB)
plot_ddj_1(out2_df_1_N_CB, ddj31, name_3)
plot_ddj_2(out2_df_1_N_CB, ddj31, name_3)

ddj32<- ddj_model(out2_df_1_N_PB)
plot_ddj_1(out2_df_1_N_PB, ddj32, name_3)
plot_ddj_2(out2_df_1_N_PB, ddj32, name_3)

ddj33<- ddj_model(out2_df_1_N_SB)
plot_ddj_1(out2_df_1_N_SB, ddj33, name_3)
plot_ddj_2(out2_df_1_N_SB, ddj33, name_3)

#shift 4
ddj41<- ddj_model(out2_df_2_N_CB)
plot_ddj_1(out2_df_2_N_CB, ddj41, name_4)
plot_ddj_2(out2_df_2_N_CB, ddj41, name_4)

ddj42<- ddj_model(out2_df_2_N_PB)
plot_ddj_1(out2_df_2_N_PB, ddj42, name_4)
plot_ddj_2(out2_df_2_N_PB, ddj42, name_4)

ddj43<- ddj_model(out2_df_2_N_SB)
plot_ddj_1(out2_df_2_N_SB, ddj43, name_4)
plot_ddj_2(out2_df_2_N_SB, ddj43, name_4)
```

### 2021-04-11 various_stuff/dev_vignettes/Bushplus.Rmd
reorganise

--- 2021-04-11 dev_vignettes/Bushplus_strains.Rmd
@@
 ---
-title: "Biodiversity, eco-evo feedback, and the stability of responses to environmental change"
-author: "Owen Petchey"
+title: "Diversity and the nature of responses to environmental change."
+author: "Owen Petchey & Jennifer Probst"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 library(tidyverse)
-library(here)
 library(deSolve)
@@
 
-This is relevant: https://www.biorxiv.org/content/10.1101/2021.03.22.436420v1
-
-Aims:
-
-* Investigate the effects of biodiversity within each of the three functional groups on responses to environmental change.
-
-Here is the original publication:
-https://www.nature.com/articles/s41467-017-00912-x
-
-And here is the supplementary information, including the table of parameter values:
-https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
-
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
-(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
-
-We have added:
+The aims here are:
+
+* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
+* Add an investigation of how changes in tolerances alter the response to environmental change.
+
+We also add:
 
@@
 * Option to include a minimum abundances.
-* Option to include biodiversity within functional groups.
+
+Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups.
+That is, there is only one parameter set per functional group.
+
+Here is the original publication:
+https://www.nature.com/articles/s41467-017-00912-x
+
+And here is the supplementary information, including the table of parameter values:
+https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
+
+Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
+(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
 
@@
 ```{r}
-source(here("functions/bushplus_dynamic_model.r"))
-## contains various functions:
-## bushplus_dynamic_model
-## bushplus_dynamic_model_strains
+source("functions/bushplus_dynamic_model.r")
 ```
@@
 
+The framework used for running a simulation, and for varying aspects of a simulation run is:
+
+* Define the default simulation conditions.
+* Use these defaults unless other conditions are specified.
+* The function call `run_simulation()` would therefore run a simulation with default conditions.
+* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.
+
 ### Default parameter values
 
-To do:
-
-* Make a function that can create parameter value sets, including for example controlling the number and types of strains within each functional group.
-
 ```{r}
@@
 
-  ## Maximum growth rate
-  g_max_CB_0 = 0.05,
-  g_max_CB_1 = 0.05,
-  g_max_PB_0 = 0.07,
-  g_max_PB_1 = 0.07,
-  g_max_SB_0 = 0.10,
+  ## maximum specific growth rates
+  g_max_CB = 0.05,
+  g_max_PB = 0.07,
+  g_max_SB = 0.1,
 
@@
   ## half-inhibition constant
-  h_SR_CB_0 = 300,
-  h_SR_CB_1 = 300,
-  h_O_PB_0 = 100,
-  h_O_PB_1 = 100,
-  h_O_SB_0 = 100,
+  h_SR_CB = 300,
+  h_O_PB = 100,
+  h_O_SB = 100,
 
@@
   ## immigration rates (cells per time)
-  i_CB = 0,
-  i_PB = 0,
-  i_SB = 0,
+  i_CB <- 0,
+  i_PB <- 0,
+  i_SB <- 0,
 
@@
   c = 4e-5
-
   )
@@
 
+### Default event
+
+#### Default event definition
+
+Used to add noise and to ensure minimum abundance.
+Noise is added to the resource concentrations, and not to the organism abundances.
+Minimum organism abundance is set.
+I.e.
+organismal abundance cannot go below some very low level.
+
+```{r}
+source("functions/default_event_definition.r")
+```
+
+#### Default event schedule
+
+This is when an event happens
+
+```{r}
+default_event_interval <- 10
+```
+
+#### Default noise event options
+
+Default to no noise:
+
+```{r}
+default_noise_sigma <- 0
+```
+
+#### Default minimum abundance event
+
+Default to following minimums
+
+```{r}
+default_minimum_abundances <- rep(1e2, 3)
+names(default_minimum_abundances) <- c("CB", "PB", "SB") 
+```
+
+### Default simulation duration
+
+```{r}
+default_sim_duration <- 2000
+default_sim_sample_interval <- 1
+```
+
+### Default oxygen diffusivity (a) forcing
+
+Here we specify no forcing of a_0 (used unless set elsewhere).
+
+```{r}
+default_log10a_series <- c(log10(default_parameter_values["a_O"]),
+                           log10(default_parameter_values["a_O"]))
+```
+
 ### Default initial state
 
-To do:
-
-* Make a function that can create initial conditions corresponding to the number of strains in the parameter values, and assign initial state values according to some rule (e.g.
-total abundance and even distribution of among strains).
-
-Initial conditions favouring anoxic
-
-```{r}
-default_initial_state <- c(
-  CB_0 = 5e1,
-  CB_1 = 0,
-  PB_0 = 1e7,
-  PB_1 = 0,
-  SB_0 = 1e7,
+Initial conditions favouring anoxic.
+
+```{r}
+default_initial_state <- c(N_CB = 5e1,
+           N_PB = 1e7,
+           N_SB = 1e7,
            SO = 300,
@@
 
-### Default event
-
-#### Default event definition
-
-Used to add noise and to ensure minimum abundance.
-
-Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.
-
-```{r}
-source(here("functions/default_event_definition.r"))
-## contains two functions:
-## default_event_definition
-## default_event_definition_strains -- this should work for any number of strains
-```
-
-#### Default event schedule
-
-This is when an event happens
-
-```{r}
-default_event_interval <- 10
-```
-
-#### Default noise event options
-
-Default to no noise:
-
-```{r}
-default_noise_sigma <- 0
-```
-
-#### Default minimum abundance event
-
-Default to following minimums
-
-```{r}
-## we only have one minimum abundance for each functional group
-default_minimum_abundances <- rep(1e2, 3)
-names(default_minimum_abundances) <- c("CB",
-                                       "PB",
-                                       "SB") 
-```
-
-### Default simulation duration
-
-```{r}
-default_sim_duration <- 2000
-default_sim_sample_interval <- 10
-```
-
-### Default oxygen diffusivity (a) forcing
-
-Here we specify no forcing of a_0 (used unless set elsewhere).
-
-```{r}
-default_log10a_series <- c(log10(default_parameter_values["a_O"]),
-                           log10(default_parameter_values["a_O"]))
-default_log10a_forcing <- matrix(ncol=2,
-                         byrow=F,
-                         data=c(ceiling(seq(0,
-                                            max(default_sim_duration),
-                                            length=length(default_log10a_series))),
-                                default_log10a_series))
-
-```
-
 ## Simulation function
@@
 ```{r}
-source(here("functions/run_simulation.r"))
+source("functions/run_simulation.r")
 ```
@@
 ```{r}
-source(here("functions/plot_dynamics.r"))
-source(here("functions/ss_expt_a_N.r"))
-source(here("functions/get_nonlinearity.r"))
-source(here("functions/get_hysteresis_range.r"))
+source("functions/plot_dynamics.r")
+source("functions/ss_expt_a_N.r")
+source("functions/get_nonlinearity.r")
+source("functions/get_hysteresis_range.r")
 ```
@@
 ```{r}
-dynamic_model = bushplus_dynamic_model_strains
-event_definition = default_event_definition_strains
-parameter_values = default_parameter_values
-event_interval = default_event_interval
-noise_sigma = default_noise_sigma
-minimum_abundances = default_minimum_abundances
-sim_duration = default_sim_duration
-sim_sample_interval = default_sim_sample_interval
-log10a_forcing = default_log10a_forcing
-initial_state = default_initial_state
-solver_method = "radau"
-
-times <- seq(0,
-             sim_duration,
-             by = sim_sample_interval)
-
-event_times <- seq(min(times),
-                   max(times),
-                   by = event_interval)
-
-log10a_forcing_func <- approxfun(x = log10a_forcing[,1],
-                            y = log10a_forcing[,2],
-                            method = "linear",
-                            rule = 2)
-assign("log10a_forcing_func", log10a_forcing_func, envir = .GlobalEnv)
-
-out <- as.data.frame(ode(y = initial_state,
-                         times = times,
-                         func = dynamic_model,
-                         parms = parameter_values,
-                         method = solver_method,
-                         events = list(func=event_definition,
-                                       time=event_times)))
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
-                           event_definition = default_event_definition_strains,
-                           parameter_values = default_parameter_values,
-                           event_interval = default_event_interval,
-                           noise_sigma = default_noise_sigma,
-                           minimum_abundances = default_minimum_abundances,
-                           sim_duration = default_sim_duration,
-                           sim_sample_interval = default_sim_sample_interval,
-                           log10a_series = default_log10a_series,
-                           initial_state = default_initial_state,
-                           solver_method = "radau")
-
-sim_res <- run_simulation()
-
-plot_dynamics_strains(sim_res)
+# dynamic_model = bushplus_dynamic_model
+# event_definition = default_event_definition
+# parameter_values = default_parameter_values
+# event_interval = default_event_interval
+# noise_sigma = default_noise_sigma
+# minimum_abundances = default_minimum_abundances
+# sim_duration = default_sim_duration
+# sim_sample_interval = default_sim_sample_interval
+# log10a_forcing = default_log10a_forcing
+# initial_state = default_initial_state
+# solver_method = "radau"
+# 
+# 
+# times <- seq(0,
+#              sim_duration,
+#              by = sim_sample_interval)
+# 
+# event_times <- seq(min(times),
+#                    max(times),
+#                    by = event_interval)  
+# 
+# log10a_forcing <- approxfun(x = log10a_forcing[,1],
+#                             y = log10a_forcing[,2],
+#                             method = "linear",
+#                             rule = 2)
+# assign("log10a_forcing", log10a_forcing, envir = .GlobalEnv)
+# 
+# out <- as.data.frame(ode(y = initial_state,
+#                          times = times,
+#                          func = dynamic_model,
+#                          parms = parameter_values,
+#                          method = solver_method,
+#                          events = list(func=event_definition,
+#                                        time=event_times)))
+# 
+# sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+#                            event_definition = default_event_definition,
+#                            parameter_values = default_parameter_values,
+#                            event_interval = default_event_interval,
+#                            noise_sigma = default_noise_sigma,
+#                            minimum_abundances = default_minimum_abundances,
+#                            sim_duration = default_sim_duration,
+#                            sim_sample_interval = default_sim_sample_interval,
+#                            log10a_forcing = default_log10a_forcing,
+#                            initial_state = default_initial_state,
+#                            solver_method = "radau")
+
+#sim_res <- run_simulation()
+
+#plot_dynamics(sim_res)
 
@@
 
-Here we use the default parameter values, which includes two strains, but with same parameters. Results should be the same as with only one strain.
-
 ## Initial conditions favouring anoxic, no noise, no forcing
@@
 
-init_state_favouring_anoxic <- c(
-
-  CB_0 = 5e1,
-  CB_1 = 0,
-  PB_0 = 1e7,
-  PB_1 = 0,
-  SB_0 = 1e7,
+init_state_favouring_anoxic <- c(N_CB = 5e1,
+           N_PB = 1e7,
+           N_SB = 1e7,
            SO = 300,
@@
            O = 1e1,
-  P = 1e1
-
-  )
+           P = 1e1)
 
 sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
-plot_dynamics_strains(simulation_result = sim_res)
+plot_dynamics(simulation_result = sim_res)
 
@@
 ```
+
+Population dynamics seem as in the paper.
+Resource dynamics fluctuate more than in the paper.
 
@@
                           noise_sigma = noise_sigma_0.01)
-plot_dynamics_strains(sim_res)
+plot_dynamics(sim_res)
 #ggsave("high_SB_PB.png")
-
-noise_sigma_0.01 <- 0.00
-
 ```
@@
 ```{r} 
-noise_sigma_0.01 <- 0.00
-
-init_state_favouring_oxic <- c(
-  CB_0 = 1e7,
-  CB_1 = 0,
-  PB_0 = 1e5,
-  PB_1 = 0,
-  SB_0 = 1e5,
+init_state_favouring_oxic <- c(N_CB = 1e7,
+           N_PB = 1e5,
+           N_SB = 1e5,
            SO = 500,
@@
            O = 30,
-  P = 10
-  )
+           P = 10)
 
@@
 
-plot_dynamics_strains(sim_res)
+plot_dynamics(sim_res)
 ##ggsave("high_CB.png")
@@
 ```{r}
-noise_sigma_0.01 <- 0.01
+
 sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
@@
 
-plot_dynamics_strains(sim_res)
+plot_dynamics(sim_res)
 ##ggsave("high_CB.png")
-noise_sigma_0.01 <- 0.00
-```
-
-# Results: two different CB strains
-
-## Starting oxic steady state
-
-```{r}
-names(default_parameter_values)
-two_CB_strains_parvals <- default_parameter_values
-two_CB_strains_parvals["g_max_CB_1"] <- 0.0495
-two_CB_strains_parvals["h_SR_CB_1"] <- 315
-two_CB_strains_parvals
-
-two_CB_init_state_favouring_oxic <- c(
-  CB_0 = 0.75e7,
-  CB_1 = 0.25e7,
-  PB_0 = 1e5,
-  PB_1 = 0,
-  SB_0 = 1e5,
-  SO = 500,
-  SR = 50,
-  O = 30,
-  P = 10
-  )
-
-sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
-                          parameter_values = two_CB_strains_parvals)
-
-plot_dynamics_strains(sim_res)
-
-```
-
-## Forcing
-```{r}
-
-## simulation timing
-simulation_duration <- 70000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-hlh_log10a_series <- c(-2, -2, -10, -10, -10, -10, -10, -10, -10)
-
-## run simulation with both CB strains having default growth rate
-## and tolerance
-sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
-                          parameter_values = default_parameter_values,
-                          log10a_series = hlh_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-
-plot_dynamics_strains(sim_res)
-```
-
-```{r}
-## run simulation with CB strains having different growth rate
-## and tolerance
-sim_res <- run_simulation(initial_state = two_CB_init_state_favouring_oxic,
-                          parameter_values = two_CB_strains_parvals,
-                          log10a_series = hlh_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-
-plot_dynamics_strains(sim_res)
-
-```
-
-# TO DO:
-
-* Check what happens when there are more strains.
-* Be able to find the stable states.
+```
 
@@
 
+Rainer: don't change what comes below
+
 ## runsteady function not finding correct values

### 2021-04-13 various_stuff/dev_vignettes/Bush_basic_reproduction_new.Rmd
lots

---
title: "Reproducing some results of Bush et al 2017"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(deSolve)
library(rootSolve)
library(here)
library(microxanox)

## Set random number generator seed
## Is redundant if there are no stochastic components in the model
set.seed(13)

run_steadystate_sims <- FALSE
```

# Introduction

The aims here are:

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
* Add an investigation of how changes in tolerances alter the response to environmental change.

We also add:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.

Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups. That is, there is only one parameter set per functional group.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

# Methods

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Modelling framework

The framework used for running a simulation, and for varying aspects of a simulation run is:

* Define the default simulation conditions.
* Use these defaults unless other conditions are specified.
* The function call `run_simulation()` would therefore run a simulation with default conditions.
* Whereas, for example, the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values.

### Dynamic model

The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model_strains()`. This is an augemented version that can handle multiple strains within each of the three functional groups. The basic version that has only one strain per functional group is `bushplus_dynamic_model()`.

### Default parameter values

```{r}
default_parameter_values <- c(

  ## maximum specific growth rates
  g_max_CB = 0.05,
  g_max_PB = 0.07,
  g_max_SB = 0.1,

  ## half-saturation constants
  k_PB_SR = 10,
  k_SB_SO = 5,
  k_CB_P = 0.2,
  k_PB_P = 0.5,
  k_SB_P = 0.5,

  ## half-inhibition constant
  h_SR_CB = 300,
  h_O_PB = 100,
  h_O_SB = 100,

  ## yield (cells per micromole)
  Y_SO_SB = 3.33e7,  
  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
  y_P_CB = 1.67e8,
  y_P_PB = 1.67e8,
  y_P_SB = 1.67e8,

  ## oxygen production per microbial cell
  Pr_CB = 6e-9,

  ## mortality rate
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## immigration rates (cells per time)
  i_CB <- 0,
  i_PB <- 0,
  i_SB <- 0,

  ## oxidisation rate of reduced sulphur
  c = 4e-5
  )
```

### Default event

#### Default event definition

Used to add noise and to ensure minimum abundance. Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

The event definition is made in the form of a function. The default definition is in the function `default_event_definition()`

#### Default event schedule

At the moment this is specified by giving the time interval between events:

```{r}
default_event_interval <- 10
```

#### Default noise event options

Default to no noise added to substrate concentrations:

```{r}
default_noise_sigma <- 0
```

#### Default minimum abundance event

Default to following minimum abundances:

```{r}
default_minimum_abundances <- rep(1e2, 3)
names(default_minimum_abundances) <- c("CB", "PB", "SB") 
```

### Default simulation duration

Default to the following duration and interval between samples:

```{r}
default_sim_duration <- 2000
default_sim_sample_interval <- 1
```

### Default oxygen diffusivity (a) forcing

Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.

```{r}
default_log10a_series <- c(log10(default_parameter_values["a_O"]),
                           log10(default_parameter_values["a_O"]))
```

### Default initial state

Specify the default set of initial conditions (here favouring anoxic organisms).

```{r}
default_initial_state <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)
```

## Code test

```{r}
dynamic_model = bushplus_dynamic_model
event_definition = default_event_definition
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_series = default_log10a_series
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(times),
                                            length=length(log10a_series))),
                                log10a_series))

log10a_forcing <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

bushplus_dynamic_model(1, initial_state, default_parameter_values)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
                           event_definition = default_event_definition,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)

plot_dynamics(sim_res)

```

# Results

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)

parameters <- default_parameter_values
state <- init_state_favouring_anoxic
bushplus_dynamic_model(100, state, parameters)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

Population dynamics seem as in the paper.
Resource dynamics fluctuate more than in the paper.

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r} 
init_state_favouring_oxic <- c(N_CB = 1e7,
           N_PB = 1e5,
           N_SB = 1e5,
           SO = 500,
           SR = 50,
           O = 30,
           P = 10)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_oxic)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Stable states

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 2
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

```{r eval = run_steadystate_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
```

```{r}
#ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB")  %>%
  print()
##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```

```{r}
p1 <- ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
print(p1)
```

# Repeat the results above using the functions that allow simulation of multiple strains per functional group

## Test with one strain per functional group

Constructing parameter set.

```{r}
CB_strain_params <- data.frame(strain_name = c("CB_1"),
                       g_max_CB = c(0.05),
                       k_CB_P = c(0.2),
                       h_SR_CB = c(300),
                       y_P_CB = c(1.67e8),
                       Pr_CB = c(6e-9),
                       m_CB = c(0.02),
                       i_CB = c(0))

PB_strain_params <- data.frame(strain_name = c("PB_1"),
                       g_max_PB = c(0.07),
                       k_PB_SR = 10,
                       k_PB_P = 0.5,
                       h_O_PB = 100,
                       y_SR_PB = 1.25e7,
                       y_P_PB = 1.67e8,
                       m_PB = 0.028,
                       i_PB = 0)

SB_strain_params <- data.frame(strain_name = c("SB_1"),
                       g_max_SB = c(0.1),
                       k_SB_SO = 5,
                       k_SB_P = 0.5,
                       h_O_SB = 100,
                       y_SO_SB = 3.33e7,  
                       y_P_SB = 1.67e8,
                       m_SB = 0.04,
                       i_SB = 0)

all_other_params <- c(

  ## substrate diffusivity
  a_S = 0.001,
  a_O = 8e-4,
  a_P = 0.01,

  ## background substrate concentration
  back_SR = 300,
  back_SO = 300,
  back_O = 300,
  back_P = 9.5,

  ## oxidisation rate of reduced sulphur
  c = 4e-5

)

one_strain_params <- construct_parameter_set(CB_strain_params,
                                    PB_strain_params,
                                    SB_strain_params,
                                    all_other_params)

one_strain_initial_state <- c(
  CB_1 = 5e1,
  PB_1 = 1e7,
  SB_1 = 1e7,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1
  )

simulation_result <- run_simulation(dynamic_model =
                                      bushplus_dynamic_model_strains,
                                    parameter_values = one_strain_params,
                                    initial_state = one_strain_initial_state)

plot_dynamics_strains(simulation_result)

```

### Test with N strains per functional group

```{r}
num_CB_strains <- 3
num_PB_strains <- 3
num_SB_strains <- 3

CB_strain_params <- data.frame(strain_name = paste0("CB_", 1:num_CB_strains),
                       g_max_CB = rep(0.05, num_CB_strains),
                       k_CB_P = rep(0.2, num_CB_strains),
                       h_SR_CB = rep(300, num_CB_strains),
                       y_P_CB = rep(1.67e8, num_CB_strains),
                       Pr_CB = rep(6e-9, num_CB_strains),
                       m_CB = rep(0.02, num_CB_strains),
                       i_CB = rep(0, num_CB_strains))

PB_strain_params <- data.frame(strain_name = paste0("PB_", 1:num_PB_strains),
                       g_max_PB = rep(0.07, num_PB_strains),
                       k_PB_SR = rep(10, num_PB_strains),
                       k_PB_P = rep(0.5, num_PB_strains),
                       h_O_PB = rep(100, num_PB_strains),
                       y_SR_PB = rep(1.25e7, num_PB_strains),
                       y_P_PB = rep(1.67e8, num_PB_strains),
                       m_PB = rep(0.028, num_PB_strains),
                       i_PB = rep(0, num_PB_strains))

SB_strain_params <- data.frame(strain_name = paste0("SB_", 1:num_SB_strains),
                       g_max_SB = rep(0.1, num_SB_strains),
                       k_SB_SO = rep(5, num_SB_strains),
                       k_SB_P = rep(0.5, num_SB_strains),
                       h_O_SB = rep(100, num_SB_strains),
                       y_SO_SB = rep(3.33e7, num_SB_strains),  
                       y_P_SB = rep(1.67e8, num_SB_strains),
                       m_SB = rep(0.04, num_SB_strains),
                       i_SB = rep(0, num_SB_strains))

multi_strain_params <- construct_parameter_set(CB_strain_params,
                                    PB_strain_params,
                                    SB_strain_params,
                                    all_other_parms)

multi_strain_initial_state <- c(
  CB_1 = 5e1,
  CB_2 = 0,
  CB_3 = 0,
  PB_1 = 1e7,
  PB_2 = 1e6,
  PB_3 = 1e5,
  SB_1 = 1e7,
  SB_2 = 1e6,
  SB_3 = 1e5,
  SO = 300,
  SR = 300,
  O = 1e1,
  P = 1e1)

simulation_result <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
               parameter_values = multi_strain_params,
               initial_state = multi_strain_initial_state)
plot_dynamics_strains(simulation_result)

```

### 2021-04-13 various_stuff/dev_vignettes/Bush_basic_reproduction_new.Rmd
various

@@
 
-```{r}
+```{r eval = run_steadystate_sims}
 ssfind_minimum_abundances <- rep(0, 3)
@@
 grid_num_N <- 2
-```
-
-```{r eval = run_steadystate_sims}
 
@@
 save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
-```
-
-```{r}
-#ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
 ```

### 2021-04-14 various_stuff/EWS functions/EWS analyses.Rmd
reorganisation of vignettes
now need to get them working!

--- 2021-04-11 various_stuff/dev_vignettes/Bushplus.Rmd
@@
 ---
-title: "Diversity and the nature of responses to environmental change."
+title: "EWS analysis"
 author: "Owen Petchey & Jennifer Probst"
@@
 ---
-
-# Setup R
-
-```{r setup}
-rm(list=ls())
-knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
-
-library(tidyverse)
-library(deSolve)
-library(rootSolve)
-
-set.seed(13)
-
-run_sims <- FALSE
-```
-
-# Introduction
-
-The aims here are:
-
-* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-* Add an investigation of how changes in tolerances alter the response to environmental change.
-
-We also add:
-
-* Option to include random noise in substrate concentrations.
-* Option to temporally vary oxygen diffusivity.
-* Option to include immigration.
-* Option to include a minimum abundances.
-
-Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups.
-That is, there is only one parameter set per functional group.
-
-Here is the original publication:
-https://www.nature.com/articles/s41467-017-00912-x
-
-And here is the supplementary information, including the table of parameter values:
-https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
-
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
-(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
-
-# Methods
-
-## Model dimensions
-
-* Time: hours
-* Volume: litres
-* Substrate quantity: micromoles
-* Organism quantity: cells
-
-## Dynamic model
-
-```{r}
-source("functions/bushplus_dynamic_model.r")
-```
-
-## Defaults
-
-The framework used for running a simulation, and for varying aspects of a simulation run is:
-
-* Define the default simulation conditions.
-* Use these defaults unless other conditions are specified.
-* The function call `run_simulation()` would therefore run a simulation with default conditions.
-* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.
-
-### Default parameter values
-
-```{r}
-default_parameter_values <- c(
-
-  ## maximum specific growth rates
-  g_max_CB = 0.05,
-  g_max_PB = 0.07,
-  g_max_SB = 0.1,
-
-  ## half-saturation constants
-  k_PB_SR = 10,
-  k_SB_SO = 5,
-  k_CB_P = 0.2,
-  k_PB_P = 0.5,
-  k_SB_P = 0.5,
-
-  ## half-inhibition constant
-  h_SR_CB = 300,
-  h_O_PB = 100,
-  h_O_SB = 100,
-
-  ## yield (cells per micromole)
-  Y_SO_SB = 3.33e7,  
-  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
-  y_P_CB = 1.67e8,
-  y_P_PB = 1.67e8,
-  y_P_SB = 1.67e8,
-
-  ## oxygen production per microbial cell
-  P_CB = 6e-9,
-
-  ## mortality rate
-  m_CB = 0.02,
-  m_PB = 0.028,
-  m_SB = 0.04,
-
-  ## substrate diffusivity
-  a_S = 0.001,
-  a_O = 8e-4,
-  a_P = 0.01,
-
-  ## background substrate concentration
-  back_SR = 300,
-  back_SO = 300,
-  back_O = 300,
-  back_P = 9.5,
-
-  ## immigration rates (cells per time)
-  i_CB <- 0,
-  i_PB <- 0,
-  i_SB <- 0,
-
-  ## oxidisation rate of reduced sulphur
-  c = 4e-5
-  )
-```
-
-### Default event
-
-#### Default event definition
-
-Used to add noise and to ensure minimum abundance.
-Noise is added to the resource concentrations, and not to the organism abundances.
-Minimum organism abundance is set.
-I.e.
-organismal abundance cannot go below some very low level.
-
-```{r}
-source("functions/default_event_definition.r")
-```
-
-#### Default event schedule
-
-This is when an event happens
-
-```{r}
-default_event_interval <- 10
-```
-
-#### Default noise event options
-
-Default to no noise:
-
-```{r}
-default_noise_sigma <- 0
-```
-
-#### Default minimum abundance event
-
-Default to following minimums
-
-```{r}
-default_minimum_abundances <- rep(1e2, 3)
-names(default_minimum_abundances) <- c("CB", "PB", "SB") 
-```
-
-### Default simulation duration
-
-```{r}
-default_sim_duration <- 2000
-default_sim_sample_interval <- 1
-```
-
-### Default oxygen diffusivity (a) forcing
-
-Here we specify no forcing of a_0 (used unless set elsewhere).
-
-```{r}
-default_log10a_series <- c(log10(default_parameter_values["a_O"]),
-                           log10(default_parameter_values["a_O"]))
-```
-
-### Default initial state
-
-Initial conditions favouring anoxic.
-
-```{r}
-default_initial_state <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-```
-
-## Simulation function
-
-```{r}
-source("functions/run_simulation.r")
-```
-
-## Helper functions
-
-```{r}
-source("functions/plot_dynamics.r")
-source("functions/ss_expt_a_N.r")
-source("functions/get_nonlinearity.r")
-source("functions/get_hysteresis_range.r")
-```
-
-## Code test
-
-```{r}
-# dynamic_model = bushplus_dynamic_model
-# event_definition = default_event_definition
-# parameter_values = default_parameter_values
-# event_interval = default_event_interval
-# noise_sigma = default_noise_sigma
-# minimum_abundances = default_minimum_abundances
-# sim_duration = default_sim_duration
-# sim_sample_interval = default_sim_sample_interval
-# log10a_forcing = default_log10a_forcing
-# initial_state = default_initial_state
-# solver_method = "radau"
-# 
-# 
-# times <- seq(0,
-#              sim_duration,
-#              by = sim_sample_interval)
-# 
-# event_times <- seq(min(times),
-#                    max(times),
-#                    by = event_interval)  
-# 
-# log10a_forcing <- approxfun(x = log10a_forcing[,1],
-#                             y = log10a_forcing[,2],
-#                             method = "linear",
-#                             rule = 2)
-# assign("log10a_forcing", log10a_forcing, envir = .GlobalEnv)
-# 
-# out <- as.data.frame(ode(y = initial_state,
-#                          times = times,
-#                          func = dynamic_model,
-#                          parms = parameter_values,
-#                          method = solver_method,
-#                          events = list(func=event_definition,
-#                                        time=event_times)))
-# 
-# sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-#                            event_definition = default_event_definition,
-#                            parameter_values = default_parameter_values,
-#                            event_interval = default_event_interval,
-#                            noise_sigma = default_noise_sigma,
-#                            minimum_abundances = default_minimum_abundances,
-#                            sim_duration = default_sim_duration,
-#                            sim_sample_interval = default_sim_sample_interval,
-#                            log10a_forcing = default_log10a_forcing,
-#                            initial_state = default_initial_state,
-#                            solver_method = "radau")
-
-#sim_res <- run_simulation()
-
-#plot_dynamics(sim_res)
-
-```
-
-# Results: Reproducing previous results
-
-## Initial conditions favouring anoxic, no noise, no forcing
-
-Figure 2a, b in Bush et al.
-
-```{r}
-
-init_state_favouring_anoxic <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
-plot_dynamics(simulation_result = sim_res)
-
-#ggsave('bacteria_no_noise.png')
-```
-
-Population dynamics seem as in the paper.
-Resource dynamics fluctuate more than in the paper.
-
-## Initial conditions favouring anoxic, with noise, no forcing
-
-Figure 2a, b in Bush et al, but with noise
-
-```{r}
-## add some noise
-noise_sigma_0.01 <- 0.01
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          noise_sigma = noise_sigma_0.01)
-plot_dynamics(sim_res)
-#ggsave("high_SB_PB.png")
-```
-
-## Initial conditions favouring oxic, no noise, no forcing
-
-Figure 2c, d in Bush et al.
-
-```{r} 
-init_state_favouring_oxic <- c(N_CB = 1e7,
-           N_PB = 1e5,
-           N_SB = 1e5,
-           SO = 500,
-           SR = 50,
-           O = 30,
-           P = 10)
-
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-Population and resource dynamics seen as those in the paper.
-
-## Initial conditions favouring oxic, with noise, no forcing
-
-Figure 2c, d in Bush et al. but with noise.
-
-```{r}
-
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          noise_sigma = noise_sigma_0.01)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-
-# Results: Composition/diversity and nature of response
-
-```{r}
-ssfind_minimum_abundances <- rep(0, 3)
-names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-# ssfind_init_state <- c(N_CB = NA, 
-#                        N_PB = NA,
-#                        N_SB = NA,
-#                        SO = 500,
-#                        SR = 50,
-#                        O = 30,
-#                        P = 10)
-##ssfind_init_state <- default_initial_state
-ssfind_init_state <- init_state_favouring_anoxic
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-grid_num_a <- 100
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-grid_num_N <- 2
-```
-
-### Each composition
-
-#### No organisms
-
-```{r eval = run_sims}
-
-initial_CBs <- 0
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_no_biology <- ss_by_a_N(expt)
-saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
-```
-
-```{r}
-ss_no_biology <- readRDS("data/ss_no_biology.RDS")
-```
-
-```{r}
-ss_no_biology %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_no_biology %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_CB <- ss_by_a_N(expt)
-saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
-```
-
-```{r}
-ss_only_CB <- readRDS("data/ss_only_CB.RDS")
-```
-
-```{r}
-tempx1 <- ss_only_CB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.))
-tempx1 %>%
-  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
-  geom_path(size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  geom_path(data = filter(tempx1, Species == "PB"),
-            aes(y=0.2), size = 0.8) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="Only CB")
-
-ggsave("figures/CB.png", width = 3, height = 2)
-```
-
-```{r}
-ss_only_CB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 0
-initial_SBs <- 10^seq(0, 8, length=grid_num_N)
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_SB <- ss_by_a_N(expt)
-
-saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
-```
-
-```{r}
-ss_only_SB <- readRDS("data/ss_only_SB.RDS")
-```
-
-```{r}
-ss_only_SB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_PB <- ss_by_a_N(expt)
-
-saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
-```
-
-```{r}
-ss_only_PB <- readRDS("data/ss_only_PB.RDS")
-```
-
-```{r}
-tempx1 <- ss_only_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.))
-tempx1 %>%
-  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
-  geom_path(size = 0.8) +
-  theme_light() +
-  ylim(0,10) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  geom_path(data = filter(tempx1, Species == "CB"),
-            aes(y=0.2), size = 0.8)+
-  geom_path(data = filter(tempx1, Species == "PB"),
-            size = 0.8) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="Only PB")
-
-ggsave("figures/PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-ss_only_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_SB <- expt$N_PB
-
-ss_only_SB_PB <- ss_by_a_N(expt)
-saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
-```
-
-```{r}
-ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB and CB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_SB <- expt$N_CB
-
-ss_only_SB_CB <- ss_by_a_N(expt)
-saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
-```
-
-```{r}
-ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
-```
-
-```{r}
-ss_only_SB_CB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_CB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:5) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_CB <- expt$N_PB
-
-ss_only_CB_PB <- ss_by_a_N(expt)
-saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
-```
-
-```{r}
-ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
-```
-
-```{r}
-ss_only_CB_PB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB, SB, and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_CB_SB_PB <- ss_by_a_N(expt)
-saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
-```
-
-```{r}
-ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
-```
-
-```{r}
-ss_CB_SB_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species),
-            size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="CB, SB, & PB") 
-ggsave("figures/CB_SB_PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-ss_CB_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### All response measures
-
-```{r eval = run_sims}
-all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
-                    composition = rep("No organisms", grid_num_a*grid_num_N),
-                    direction = c(rep("up", grid_num_a),
-                                  rep("down", grid_num_a)))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_CB.RDS"),
-                              composition = rep("CB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB.RDS"),
-                              composition = rep("SB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_PB.RDS"),
-                              composition = rep("PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
-                              composition = rep("SB-PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
-                              composition = rep("SB-CB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
-                              composition = rep("CB-PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
-                              composition = rep("CB-SB-PB",
-                                                grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-
-x_long <- all_ss %>%
-  gather(key = Species, value = Quantity, 2:8)
-x_wide <- x_long %>%
-  select(-initial_N_CB) %>%
-  spread(key = direction, value=Quantity, drop=T)
-
-resp_summary <- x_wide %>%
-  group_by(composition, Species) %>%
-  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
-            hyst_2 = get_hysteresis_range(up, down, a),
-            nl_up = get_nonlinearity(a, log10(up)),
-            nl_down = get_nonlinearity(a, log10(down))) %>%
-  ungroup() %>%
-  gather(key = resp_measure, value = value, 3:6) %>%
-  mutate(composition = fct_relevel(composition,
-                                   "No organisms",
-                                   "CB",
-                                   "SB",
-                                   "PB",
-                                   "CB-PB",
-                                   "SB-CB",
-                                   "SB-PB",
-                                   "CB-SB-PB"),
-         num_func_groups = case_when(composition =="No organisms" ~ 0,
-                                     composition == "CB" ~ 1,
-                                     composition == "SB" ~ 1,
-                                     composition == "PB" ~ 1,
-                                     composition == "CB-PB" ~ 2,
-                                     composition == "SB-CB" ~ 2,
-                                     composition == "SB-PB" ~ 2,
-                                     composition == "CB-SB-PB" ~ 3
-                                     ),
-         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
-         )
-saveRDS(resp_summary, "data/resp_summary.RDS")
-```
-
-```{r}
-resp_summary <- readRDS("data/resp_summary.RDS")
-resp_summary %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = value)) +
-  facet_grid(Species ~ resp_measure, scales = "free") +
-  coord_flip()
-
-```
-
-### Nonlinearity
-
-#### All species nonlinearity
-
-```{r}
-resp_summ1 <- resp_summary %>%
-  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
-  group_by(composition, Species, num_func_groups, var_type) %>%
-  summarise(ave_nl = mean(value))
-
-resp_summ1 %>%
-  ggplot() +
-  geom_point(aes(composition, ave_nl)) +
-  facet_wrap( ~ Species)
-
-```
-
-#### Substrate nonlinearity
-
-```{r}
-focus_on <- "Substrate"
-ave_nl_substrate <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_substrate %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### Organism nonlinearity
-
-```{r}
-focus_on <- "Organism"
-ave_nl_organism <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
-ave_nl_organism %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### System nonlinearity
-
-```{r}
-focus_on <- "System"
-ave_nl_system <- resp_summ1 %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
-ave_nl_system %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-ggsave("figures/system_nonlinearity.png", width = 3, height = 2)
-
-```
-
-### Hysteresis
-
-#### Substrate hysteresis
-
-```{r}
-focus_on <- "Substrate"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(composition, num_func_groups, resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### Organism hysteresis
-
-```{r}
-focus_on <- "Organism"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(composition, num_func_groups, resp_measure, var_type) %>%
-  summarise(ave_hyst = sum(value) / unique(num_func_groups))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### System hysteresis (separate)
-
-```{r}
-resp_summ4 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
-  group_by(resp_measure, composition) %>%
-  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))
-
-resp_summ4 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = mean_hyst)) +
-  facet_wrap( ~ resp_measure) +
-  ggtitle("Hysteresis in system response.")
-
-```
-
-#### System hysteresis (combined)
-
-```{r}
-resp_summ5 <- resp_summ4 %>%
-  group_by(composition) %>%
-  summarise(mean_mean_hyst = mean(mean_hyst))
-
-resp_summ5 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = mean_mean_hyst)) +
-  xlab("Functional groups present") +
-  ylab("Amount of hysteresis") +
-  ylim(0,2) +
-  theme_light() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)
-
-```
-
-# Results: Varying strength of inhibition
-
-## Introduction
-
-Three parameters set the strength of inhibition (half-inhibition constants):
-
-* h_SR_CB = 300,
-* h_O_PB = 100,
-* h_O_SB = 100,
-
-These work by the inhibition function:
-
-$\frac{1}{1 + \frac{X}{h_X}}$
-
-```{r}
-inhibition
-```
-
-Plot some examples:
-
-```{r}
-default_parameter_values["h_SR_CB"]
-SR <- 10^seq(1.4, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
-  ylab("Multiplicative effect\non CB growth rate")
-```
-
-```{r}
-default_parameter_values["h_O_PB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
-  ylab("Multiplicative effect\non PB growth rate")
-```
-
-```{r}
-default_parameter_values["h_O_SB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
-  ylab("Multiplicative effect\non SB growth rate")
-```
-
-```{r}
-proportion_change <- 2
-default_parameter_values["h_O_SB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
-  ylab("Multiplicative effect\non SB growth rate")
-```
-
-## Experiment
-
-Increasing the inhibition half saturation constants weakens the inhibition.
-
-```{r}
-ssfind_minimum_abundances <- default_minimum_abundances
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = NA,
-           N_SB = NA,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-grid_num_mult <- 20
-grid_num_N <- 2
-grid_num_a <- 100
-
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    inhib_const_mult = inhib_const_mults,
-                    a_O = a_Os)
-```
-
-```{r eval = run_sims}
-ss_h1 <- ss_by_a_h_N(expt)
-
-saveRDS(ss_h1, "data/ss_h1.RDS")
-```
-
-```{r}
-ss_h1 <- readRDS("data/ss_h1.RDS")
-```
-
-## Results
-
-### Close to default inhibition
-
-```{r}
-inhib_const_mults
-focal_inhibition_multiplier <- inhib_const_mults[14]
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Double strength inhibition
-
-```{r}
-## YES! 0.5 does mean twice as strong inhibition
-focal_inhibition_multiplier <- 0.5
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Weaker inhibition
-
-```{r}
-## YES! 2 does mean half as strong inhibition
-focal_inhibition_multiplier <- 2^0.4
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Nonlinearity
-
-```{r}
-x_long <- ss_h1 %>%
-  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
-  gather(key = Species, value = Quantity, 3:9) 
-x_wide <- x_long %>%
-  select(-initial_N_CB) %>%
-  spread(key = direction, value=Quantity, drop=T)
-
-resp_summary <- x_wide %>%
-  group_by(inhib_const_mult, Species) %>%
-  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
-            hyst_2 = get_hysteresis_range(up, down, a),
-            nl_up = get_nonlinearity(a, log10(up)),
-            nl_down = get_nonlinearity(a, log10(down))) %>%
-  ungroup() %>%
-  gather(key = resp_measure, value = value, 3:6) %>%
-  mutate(
-         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
-         )
-```
-
-#### All species nonlinearity
-
-```{r}
-resp_summ1 <- resp_summary %>%
-  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
-  group_by(inhib_const_mult, Species, var_type) %>%
-  summarise(ave_nl = mean(value))
-
-resp_summ1 %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_nl)) +
-  facet_wrap( ~ Species)
-
-```
-
-#### Substrate nonlinearity
-
-```{r}
-focus_on <- "Substrate"
-ave_nl_substrate <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_substrate %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### Organism nonlinearity
-
-```{r}
-focus_on <- "Organism"
-ave_nl_organism <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_organism %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### System nonlinearity
-
-```{r}
-focus_on <- "System"
-ave_nl_system <- resp_summ1 %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_system %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-ggsave("figures/system_nonlinearity.png", width = 3, height = 2)
-
-```
-
-### Hysteresis
-
-#### Substrate hysteresis
-
-```{r}
-focus_on <- "Substrate"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(inhib_const_mult, resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### Organism hysteresis
-
-```{r}
-focus_on <- "Organism"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(inhib_const_mult,  resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### System hysteresis (separate)
-
-```{r}
-resp_summ4 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
-  group_by(resp_measure, inhib_const_mult) %>%
-  summarise(mean_hyst = mean(value))
-
-resp_summ4 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = mean_hyst)) +
-  facet_wrap( ~ resp_measure) +
-  ggtitle("Hysteresis in system response.")
-
-```
-
-#### System hysteresis (combined)
-
-```{r}
-resp_summ5 <- resp_summ4 %>%
-  group_by(inhib_const_mult) %>%
-  summarise(mean_mean_hyst = mean(mean_hyst))
-
-resp_summ5 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = mean_mean_hyst)) +
-  xlab("Organismal tolerance") +
-  ylab("Amount of hysteresis") +
-  theme_light() +
-  ylim(0,2)
-ggsave("figures/inhibition_system_hysteresis.png", width = 2.5, height = 2)
-
-```
-
-# Results: Temporal state switching
-
-## Switch anoxic-oxic-anoxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_anoxic
-
-## simulation timing
-simulation_duration <- 5000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          log10a_series = lhl_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/lhl_with_defaults.RDS")
-```
-
-```{r}
-lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
-plot_dynamics(lhl_with_defaults)
-```
-
-```{r eval = run_sims}
-lhl_with_defaults$result %>%
-  filter(between(time, 1000, 73000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
-## Switch oxic-anoxic-oxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_oxic
-
-## simulation timing
-simulation_duration <- 50000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          log10a_series = hlh_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/hlh_with_defaults.RDS")
-```
-
-```{r}
-hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
-plot_dynamics(hlh_with_defaults)
-```
-
-```{r eval = run_sims}
-hlh_with_defaults$result %>%
-  filter(between(time, 1000, 100000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
-# Appendices
-
-Rainer: don't change what comes below
-
-## runsteady function not finding correct values
-
-```{r eval = FALSE}
-
-init_state_anoxic_2 <- c(N_CB = 1e1,
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 300,
-           SR = 300,
-           O = 1e-5,
-           P = 1e1)
-
-RS <- runsteady(y = init_state_anoxic_2,
-                fun = bushplus_dynamic_model,
-                parms = default_parameter_values,
-                times = c(0, 1e5),
-                stol=1)
-RS$y
-RS <- runsteady(y = init_state_favouring_oxic,
-                fun = bushplus_dynamic_model,
-                parms = default_parameter_values,
-                times = c(0, 1e5),
-                stol=1)
-RS$y
-
-```
-
-## Separatrix
-
-Note that this section not working / updated since the code was made into functions
-
-### Effect of N_CB and oxygen diffusability
-
-```{r eval = FALSE}
-
-noise_sigma <- 0.0
-
-get_final_states_N_CB <- function(x) {
-
-  #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
-  state["N_CB"] <- x["N_CB"]
-  parameters["a_O"] <- x["a_O"]
-
-  ## update forcing function
-  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
-                                             log10(parameters["a_O"]),
-                                             max(times),
-                                             log10(parameters["a_O"])))
-  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
-                             y = a_forcing1[,2],
-                             method = "linear",
-                             rule = 2)
-
-  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
-  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)
-
-  as.data.frame(ode(y = state,
-                    times = times,
-                    func = micro_mod_var,
-                    parms = parameters,
-                    method = "radau",
-                    events = list(func=noise_events, time=times)))[2,-1] 
-  #gives you values for the last time step
-}
-
-times <- seq(0, 50000, by = 50000)
-
-state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-#different combinations of N_CB and oxygen diffusability
-grid_num <- 50
-expt <- expand.grid(N_CB = 10^seq(0, 8, length=grid_num),
-                    a_O = 10^seq(-3.5, -2, length=grid_num))
-
-result <- apply(expt, 1, function(x) get_final_states_N_CB(x)) %>%
-                          tibble() %>%
-                          unnest() %>%
-                          mutate(initial_N_CB=expt$N_CB,
-                                 a_O=expt$a_O,
-                                 highlow=ifelse(N_CB>2e8, 'high', 'low'))
-
-result <- result %>%
-  mutate(N_CB = ifelse(N_CB < minimum_abundance, minimum_abundance, N_CB))
-
-result$highlow <- as.factor(result$highlow)
-saveRDS(result, "data/CB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-result <- readRDS("data/CB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-
-#plot N_CB / P_CB against a_0
-p <- ggplot(result) + 
-  geom_line(aes(x=log10(a_O), y=log10(N_CB), color=highlow), size=1.25) +
-  ggtitle("Stable states of the system")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('Initial cyanobacteria abundance')
-p <- p + geom_point(data=result, aes(x=log10(a_O), y=log10(initial_N_CB), color=highlow)) + labs(color = "Final CB concentration")
-plot(p)
-#ggsave("initial conditions_1.png")
-```
-
-```{r eval = FALSE}
-#show hysterisis effects for all the parameters
-#plot oxygen against diffusability
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(O))))+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[oxygen concentration]')+
-  geom_segment(aes(x=log10(result$a_O[839]), y=log10(result$O[839]), xend=log10(result$a_O[838]), yend=log10(result$O[838])), arrow=arrow(), colour="blue", size=0.5)+
-  geom_segment(aes(x=log10(result$a_O[1761]), y=log10(result$O[1761]), xend=log10(result$a_O[1762]), yend=log10(result$O[1762])), arrow=arrow(), colour="blue", size=0.5)+
-  geom_segment(aes(x=-4.25, y=-0.45, xend=-3.75, yend=0.1), arrow=arrow(), colour="blue", size=0.5)+
-  geom_curve(aes(x = -3.75, y = 1.9, xend = -4.25, yend = 1.5), arrow=arrow(), colour='blue', size = 0.5, curvature = -0.2)
-#ggsave("hysterisis_CB_1.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(P))))+
-  ggtitle("Hysterisis effect of phoshate concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[phosphate concentration]')
-#ggsave("hysterisis_CB_2.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(SO))))+
-  ggtitle("Hysterisis effect of oxidized sulfur concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[oxydized sulfur concentration]')
-#ggsave("hysterisis_CB_3.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(SR))))+
-  ggtitle("Hysterisis effect of reduced sulfur concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[reduced sulfur] concentration]')
-#ggsave("hysterisis_CB_4.png")
-```
-
-These graphs show an at least qualitative match to the result in the paper.
-
-### Effect of PB/SB depending on oxygen diffusability
-
-```{r eval = FALSE}
-noise_sigma <- 0.00
-get_final_states_N_PB <- function(x) { 
-  #give in N_CB, N_SB, N_PB and a_O values and put these into state and parameters
-  state["N_PB"] <- x["var"]
-  state["N_SB"] <- x["var"]
-  parameters["a_O"] <- x["a_O"]
-
-  ## update forcing function
-  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
-                                             log10(parameters["a_O"]),
-                                             max(times),
-                                             log10(parameters["a_O"])))
-  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
-                             y = a_forcing1[,2],
-                             method = "linear",
-                             rule = 2)
-  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
-  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)
-
-  as.data.frame(ode(y = state,
-                         times = times,
-                         func = micro_mod_var,
-                         parms = parameters,
-                         method = "radau", events = list(func=noise_events, time=times)))[2,-1] 
-  #gives you values for the last time step
-}
-
-times <- seq(0, 50000, by = 5000)
-state <- c(N_CB = 5, 
-           N_PB = NA,#this one is varied
-           N_SB = NA,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-#different combinations of N_PB and oxygen diffusivity
-grid_num <- 50
-expt <- expand.grid(var=10^seq(0, 7, length=grid_num),
-                    a_O = 10^seq(-3, -2, length=grid_num))
-
-result2 <- apply(expt, 1, function(x) get_final_states_N_PB(x)) %>%
-                          tibble() %>%
-                          unnest() %>%
-                          mutate(initial_N_PB=expt$var,
-                                 initial_N_SB=expt$var,
-                                 a_O=expt$a_O,
-                                 highlow=ifelse(N_PB>1e8, 'high', 'low'))
-
-result2 <- result2 %>%
-  mutate(N_PB = ifelse(N_PB < minimum_abundance, minimum_abundance, N_PB))
-
-result2$highlow <- as.factor(result2$highlow)
-saveRDS(result2, "data/PBSB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-result2 <- readRDS("data/PBSB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-#plot N_PB against a_0
-p <- ggplot(result2) + 
-  geom_line(aes(x=log10(a_O), y=log10(N_PB), color=highlow), size=1.25) +
-  ggtitle("Stable states of the system")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('Initial PS and SR bacteria concentration')
-p <- p + geom_point(data=result2,
-                    aes(x=log10(a_O), y=log10(initial_N_PB), color=highlow))
-p <- p + labs(color = "Final PB and SB concentration")
-plot(p)
-#ggsave("initial conditions_2.png")
-```
 

### 2021-04-14 various_stuff/dev_vignettes/User manual.Rmd
reorganisation of vignettes
now need to get them working!

--- 2021-04-13 various_stuff/dev_vignettes/Bush_basic_reproduction_new.Rmd
@@
 ---
-title: "Reproducing some results of Bush et al 2017"
+title: "User manual -- microxanox R Package for microbial oxic and anoxic ecosystem simulations"
 author: "Owen Petchey"
@@
 vignette: >
-  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
+  %\VignetteIndexEntry{User manual -- microxanox R Package for microbial oxic and anoxic ecosystem simulations}
   %\VignetteEngine{knitr::rmarkdown}
@@
 ---
-
-# Setup R
-
-```{r setup}
-rm(list=ls())
-knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
-
-library(tidyverse)
-library(deSolve)
-library(rootSolve)
-library(here)
-library(microxanox)
-
-## Set random number generator seed
-## Is redundant if there are no stochastic components in the model
-set.seed(13)
-
-run_steadystate_sims <- FALSE
-```
 
@@
 
-## Model dimensions
-
-* Time: hours
-* Volume: litres
-* Substrate quantity: micromoles
-* Organism quantity: cells
-
 ## Modelling framework
@@
 
-### Dynamic model
+## Dynamic model
+
+There are two versions...
 
 The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model_strains()`.
-This is an augemented version that can handle multiple strains within each of the three functional groups.
+This is an augmented version that can handle multiple strains within each of the three functional groups.
 The basic version that has only one strain per functional group is `bushplus_dynamic_model()`.
 
-### Default parameter values
+## Model dimensions
 
-```{r}
-default_parameter_values <- c(
-
-  ## maximum specific growth rates
-  g_max_CB = 0.05,
-  g_max_PB = 0.07,
-  g_max_SB = 0.1,
-
-  ## half-saturation constants
-  k_PB_SR = 10,
-  k_SB_SO = 5,
-  k_CB_P = 0.2,
-  k_PB_P = 0.5,
-  k_SB_P = 0.5,
-
-  ## half-inhibition constant
-  h_SR_CB = 300,
-  h_O_PB = 100,
-  h_O_SB = 100,
-
-  ## yield (cells per micromole)
-  Y_SO_SB = 3.33e7,  
-  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
-  y_P_CB = 1.67e8,
-  y_P_PB = 1.67e8,
-  y_P_SB = 1.67e8,
-
-  ## oxygen production per microbial cell
-  Pr_CB = 6e-9,
-
-  ## mortality rate
-  m_CB = 0.02,
-  m_PB = 0.028,
-  m_SB = 0.04,
-
-  ## substrate diffusivity
-  a_S = 0.001,
-  a_O = 8e-4,
-  a_P = 0.01,
-
-  ## background substrate concentration
-  back_SR = 300,
-  back_SO = 300,
-  back_O = 300,
-  back_P = 9.5,
-
-  ## immigration rates (cells per time)
-  i_CB <- 0,
-  i_PB <- 0,
-  i_SB <- 0,
-
-  ## oxidisation rate of reduced sulphur
-  c = 4e-5
-  )
-```
-
-### Default event
-
-#### Default event definition
-
-Used to add noise and to ensure minimum abundance.
-Noise is added to the resource concentrations, and not to the organism abundances.
-Minimum organism abundance is set.
-I.e.
-organismal abundance cannot go below some very low level.
-
-The event definition is made in the form of a function. The default definition is in the function `default_event_definition()`
-
-#### Default event schedule
-
-At the moment this is specified by giving the time interval between events:
-
-```{r}
-default_event_interval <- 10
-```
-
-#### Default noise event options
-
-Default to no noise added to substrate concentrations:
-
-```{r}
-default_noise_sigma <- 0
-```
-
-#### Default minimum abundance event
-
-Default to following minimum abundances:
-
-```{r}
-default_minimum_abundances <- rep(1e2, 3)
-names(default_minimum_abundances) <- c("CB", "PB", "SB") 
-```
-
-### Default simulation duration
-
-Default to the following duration and interval between samples:
-
-```{r}
-default_sim_duration <- 2000
-default_sim_sample_interval <- 1
-```
-
-### Default oxygen diffusivity (a) forcing
-
-Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.
-
-```{r}
-default_log10a_series <- c(log10(default_parameter_values["a_O"]),
-                           log10(default_parameter_values["a_O"]))
-```
-
-### Default initial state
-
-Specify the default set of initial conditions (here favouring anoxic organisms).
-
-```{r}
-default_initial_state <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-```
-
-## Code test
-
-```{r}
-dynamic_model = bushplus_dynamic_model
-event_definition = default_event_definition
-parameter_values = default_parameter_values
-event_interval = default_event_interval
-noise_sigma = default_noise_sigma
-minimum_abundances = default_minimum_abundances
-sim_duration = default_sim_duration
-sim_sample_interval = default_sim_sample_interval
-log10a_series = default_log10a_series
-initial_state = default_initial_state
-solver_method = "radau"
-
-times <- seq(0,
-             sim_duration,
-             by = sim_sample_interval)
-
-event_times <- seq(min(times),
-                   max(times),
-                   by = event_interval)
-
-log10a_forcing <- matrix(ncol=2,
-                         byrow=F,
-                         data=c(ceiling(seq(0,
-                                            max(times),
-                                            length=length(log10a_series))),
-                                log10a_series))
-
-log10a_forcing <- approxfun(x = log10a_forcing[,1],
-                            y = log10a_forcing[,2],
-                            method = "linear",
-                            rule = 2)
-assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
-
-out <- as.data.frame(ode(y = initial_state,
-                         times = times,
-                         func = dynamic_model,
-                         parms = parameter_values,
-                         method = solver_method,
-                         events = list(func=event_definition,
-                                       time=event_times)))
-
-bushplus_dynamic_model(1, initial_state, default_parameter_values)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-                           event_definition = default_event_definition,
-                           parameter_values = default_parameter_values,
-                           event_interval = default_event_interval,
-                           noise_sigma = default_noise_sigma,
-                           minimum_abundances = default_minimum_abundances,
-                           sim_duration = default_sim_duration,
-                           sim_sample_interval = default_sim_sample_interval,
-                           log10a_series = default_log10a_series,
-                           initial_state = default_initial_state,
-                           solver_method = "radau")
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)
-
-plot_dynamics(sim_res)
-
-```
-
-# Results
-
-## Initial conditions favouring anoxic, no noise, no forcing
-
-Figure 2a, b in Bush et al.
-
-```{r}
-
-init_state_favouring_anoxic <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-
-parameters <- default_parameter_values
-state <- init_state_favouring_anoxic
-bushplus_dynamic_model(100, state, parameters)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-  initial_state = init_state_favouring_anoxic)
-plot_dynamics(simulation_result = sim_res)
-
-#ggsave('bacteria_no_noise.png')
-```
-
-Population dynamics seem as in the paper.
-Resource dynamics fluctuate more than in the paper.
-
-## Initial conditions favouring oxic, no noise, no forcing
-
-Figure 2c, d in Bush et al.
-
-```{r} 
-init_state_favouring_oxic <- c(N_CB = 1e7,
-           N_PB = 1e5,
-           N_SB = 1e5,
-           SO = 500,
-           SR = 50,
-           O = 30,
-           P = 10)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-  initial_state = init_state_favouring_oxic)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-Population and resource dynamics seen as those in the paper.
-
-## Stable states
-
-```{r eval = run_steadystate_sims}
-ssfind_minimum_abundances <- rep(0, 3)
-names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-# ssfind_init_state <- c(N_CB = NA, 
-#                        N_PB = NA,
-#                        N_SB = NA,
-#                        SO = 500,
-#                        SR = 50,
-#                        O = 30,
-#                        P = 10)
-##ssfind_init_state <- default_initial_state
-ssfind_init_state <- init_state_favouring_anoxic
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-grid_num_a <- 2
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-grid_num_N <- 2
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_CB_SB_PB <- ss_by_a_N(expt)
-save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
-```
-
-```{r}
-load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
-ss_CB_SB_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species),
-            size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="CB, SB, & PB")  %>%
-  print()
-##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-p1 <- ss_CB_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-print(p1)
-```
-
-# Repeat the results above using the functions that allow simulation of multiple strains per functional group
-
-## Test with one strain per functional group
-
-Constructing parameter set.
-
-```{r}
-CB_strain_params <- data.frame(strain_name = c("CB_1"),
-                       g_max_CB = c(0.05),
-                       k_CB_P = c(0.2),
-                       h_SR_CB = c(300),
-                       y_P_CB = c(1.67e8),
-                       Pr_CB = c(6e-9),
-                       m_CB = c(0.02),
-                       i_CB = c(0))
-
-PB_strain_params <- data.frame(strain_name = c("PB_1"),
-                       g_max_PB = c(0.07),
-                       k_PB_SR = 10,
-                       k_PB_P = 0.5,
-                       h_O_PB = 100,
-                       y_SR_PB = 1.25e7,
-                       y_P_PB = 1.67e8,
-                       m_PB = 0.028,
-                       i_PB = 0)
-
-SB_strain_params <- data.frame(strain_name = c("SB_1"),
-                       g_max_SB = c(0.1),
-                       k_SB_SO = 5,
-                       k_SB_P = 0.5,
-                       h_O_SB = 100,
-                       y_SO_SB = 3.33e7,  
-                       y_P_SB = 1.67e8,
-                       m_SB = 0.04,
-                       i_SB = 0)
-
-all_other_params <- c(
-
-  ## substrate diffusivity
-  a_S = 0.001,
-  a_O = 8e-4,
-  a_P = 0.01,
-
-  ## background substrate concentration
-  back_SR = 300,
-  back_SO = 300,
-  back_O = 300,
-  back_P = 9.5,
-
-  ## oxidisation rate of reduced sulphur
-  c = 4e-5
-
-)
-
-one_strain_params <- construct_parameter_set(CB_strain_params,
-                                    PB_strain_params,
-                                    SB_strain_params,
-                                    all_other_params)
-
-one_strain_initial_state <- c(
-  CB_1 = 5e1,
-  PB_1 = 1e7,
-  SB_1 = 1e7,
-  SO = 300,
-  SR = 300,
-  O = 1e1,
-  P = 1e1
-  )
-
-simulation_result <- run_simulation(dynamic_model =
-                                      bushplus_dynamic_model_strains,
-                                    parameter_values = one_strain_params,
-                                    initial_state = one_strain_initial_state)
-
-plot_dynamics_strains(simulation_result)
-
-```
-
-### Test with N strains per functional group
-
-```{r}
-num_CB_strains <- 3
-num_PB_strains <- 3
-num_SB_strains <- 3
-
-CB_strain_params <- data.frame(strain_name = paste0("CB_", 1:num_CB_strains),
-                       g_max_CB = rep(0.05, num_CB_strains),
-                       k_CB_P = rep(0.2, num_CB_strains),
-                       h_SR_CB = rep(300, num_CB_strains),
-                       y_P_CB = rep(1.67e8, num_CB_strains),
-                       Pr_CB = rep(6e-9, num_CB_strains),
-                       m_CB = rep(0.02, num_CB_strains),
-                       i_CB = rep(0, num_CB_strains))
-
-PB_strain_params <- data.frame(strain_name = paste0("PB_", 1:num_PB_strains),
-                       g_max_PB = rep(0.07, num_PB_strains),
-                       k_PB_SR = rep(10, num_PB_strains),
-                       k_PB_P = rep(0.5, num_PB_strains),
-                       h_O_PB = rep(100, num_PB_strains),
-                       y_SR_PB = rep(1.25e7, num_PB_strains),
-                       y_P_PB = rep(1.67e8, num_PB_strains),
-                       m_PB = rep(0.028, num_PB_strains),
-                       i_PB = rep(0, num_PB_strains))
-
-SB_strain_params <- data.frame(strain_name = paste0("SB_", 1:num_SB_strains),
-                       g_max_SB = rep(0.1, num_SB_strains),
-                       k_SB_SO = rep(5, num_SB_strains),
-                       k_SB_P = rep(0.5, num_SB_strains),
-                       h_O_SB = rep(100, num_SB_strains),
-                       y_SO_SB = rep(3.33e7, num_SB_strains),  
-                       y_P_SB = rep(1.67e8, num_SB_strains),
-                       m_SB = rep(0.04, num_SB_strains),
-                       i_SB = rep(0, num_SB_strains))
-
-multi_strain_params <- construct_parameter_set(CB_strain_params,
-                                    PB_strain_params,
-                                    SB_strain_params,
-                                    all_other_parms)
-
-multi_strain_initial_state <- c(
-  CB_1 = 5e1,
-  CB_2 = 0,
-  CB_3 = 0,
-  PB_1 = 1e7,
-  PB_2 = 1e6,
-  PB_3 = 1e5,
-  SB_1 = 1e7,
-  SB_2 = 1e6,
-  SB_3 = 1e5,
-  SO = 300,
-  SR = 300,
-  O = 1e1,
-  P = 1e1)
-
-simulation_result <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
-               parameter_values = multi_strain_params,
-               initial_state = multi_strain_initial_state)
-plot_dynamics_strains(simulation_result)
-
-```
+* Time: hours
+* Volume: litres
+* Substrate quantity: micromoles
+* Organism quantity: cells

### 2021-04-14 various_stuff/dev_vignettes/inhibition strength study.Rmd
reorganisation of vignettes
now need to get them working!

--- 2021-04-11 various_stuff/dev_vignettes/Bushplus.Rmd
@@
 ---
-title: "Diversity and the nature of responses to environmental change."
-author: "Owen Petchey & Jennifer Probst"
+title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
+author: "Owen Petchey"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  html_document:
-    toc: true
-    toc_depth: 4
-    number_sections: true
-    code_folding: hide
-    toc_float:
-      collapsed: true
-      smooth_scroll: true
+output: rmarkdown::html_vignette
+vignette: >
+  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
+  %\VignetteEngine{knitr::rmarkdown}
+  \usepackage[utf8]{inputenc}
 ---
 
-# Setup R
-
 ```{r setup}
+## Clear R
 rm(list=ls())
+
+## some chunk options
 knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
 
+## load required libraries
 library(tidyverse)
@@
 library(rootSolve)
-
+library(microxanox)
+
+## Set random number generator seed
+## Is redundant if there are no stochastic components in the model
 set.seed(13)
 
-run_sims <- FALSE
+## Should the rather lengthy simulations required to 
+## construct the state-environment relationship
+## be run?
+run_steadystate_sims <- FALSE
 ```
@@
 * Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-* Add an investigation of how changes in tolerances alter the response to environmental change.
-
-We also add:
-
-* Option to include random noise in substrate concentrations.
-* Option to temporally vary oxygen diffusivity.
-* Option to include immigration.
-* Option to include a minimum abundances.
-
-Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups.
-That is, there is only one parameter set per functional group.
-
-Here is the original publication:
-https://www.nature.com/articles/s41467-017-00912-x
-
-And here is the supplementary information, including the table of parameter values:
-https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
-
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
-(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
+
+Please see the User manual vignette for further introduction and general methods.
 
 # Methods
-
-## Model dimensions
-
-* Time: hours
-* Volume: litres
-* Substrate quantity: micromoles
-* Organism quantity: cells
-
-## Dynamic model
-
-```{r}
-source("functions/bushplus_dynamic_model.r")
-```
-
-## Defaults
-
-The framework used for running a simulation, and for varying aspects of a simulation run is:
-
-* Define the default simulation conditions.
-* Use these defaults unless other conditions are specified.
-* The function call `run_simulation()` would therefore run a simulation with default conditions.
-* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.
 
@@
 
-# Results: Reproducing previous results
-
-## Initial conditions favouring anoxic, no noise, no forcing
-
-Figure 2a, b in Bush et al.
-
-```{r}
-
-init_state_favouring_anoxic <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
-plot_dynamics(simulation_result = sim_res)
-
-#ggsave('bacteria_no_noise.png')
-```
-
-Population dynamics seem as in the paper.
-Resource dynamics fluctuate more than in the paper.
-
-## Initial conditions favouring anoxic, with noise, no forcing
-
-Figure 2a, b in Bush et al, but with noise
-
-```{r}
-## add some noise
-noise_sigma_0.01 <- 0.01
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          noise_sigma = noise_sigma_0.01)
-plot_dynamics(sim_res)
-#ggsave("high_SB_PB.png")
-```
-
-## Initial conditions favouring oxic, no noise, no forcing
-
-Figure 2c, d in Bush et al.
-
-```{r} 
-init_state_favouring_oxic <- c(N_CB = 1e7,
-           N_PB = 1e5,
-           N_SB = 1e5,
-           SO = 500,
-           SR = 50,
-           O = 30,
-           P = 10)
-
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-Population and resource dynamics seen as those in the paper.
-
-## Initial conditions favouring oxic, with noise, no forcing
-
-Figure 2c, d in Bush et al. but with noise.
-
-```{r}
-
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          noise_sigma = noise_sigma_0.01)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-
-# Results: Composition/diversity and nature of response
-
-```{r}
-ssfind_minimum_abundances <- rep(0, 3)
-names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-# ssfind_init_state <- c(N_CB = NA, 
-#                        N_PB = NA,
-#                        N_SB = NA,
-#                        SO = 500,
-#                        SR = 50,
-#                        O = 30,
-#                        P = 10)
-##ssfind_init_state <- default_initial_state
-ssfind_init_state <- init_state_favouring_anoxic
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-grid_num_a <- 100
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-grid_num_N <- 2
-```
-
-### Each composition
-
-#### No organisms
-
-```{r eval = run_sims}
-
-initial_CBs <- 0
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_no_biology <- ss_by_a_N(expt)
-saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
-```
-
-```{r}
-ss_no_biology <- readRDS("data/ss_no_biology.RDS")
-```
-
-```{r}
-ss_no_biology %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_no_biology %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_CB <- ss_by_a_N(expt)
-saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
-```
-
-```{r}
-ss_only_CB <- readRDS("data/ss_only_CB.RDS")
-```
-
-```{r}
-tempx1 <- ss_only_CB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.))
-tempx1 %>%
-  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
-  geom_path(size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  geom_path(data = filter(tempx1, Species == "PB"),
-            aes(y=0.2), size = 0.8) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="Only CB")
-
-ggsave("figures/CB.png", width = 3, height = 2)
-```
-
-```{r}
-ss_only_CB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 0
-initial_SBs <- 10^seq(0, 8, length=grid_num_N)
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_SB <- ss_by_a_N(expt)
-
-saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
-```
-
-```{r}
-ss_only_SB <- readRDS("data/ss_only_SB.RDS")
-```
-
-```{r}
-ss_only_SB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_only_PB <- ss_by_a_N(expt)
-
-saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
-```
-
-```{r}
-ss_only_PB <- readRDS("data/ss_only_PB.RDS")
-```
-
-```{r}
-tempx1 <- ss_only_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.))
-tempx1 %>%
-  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
-  geom_path(size = 0.8) +
-  theme_light() +
-  ylim(0,10) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  geom_path(data = filter(tempx1, Species == "CB"),
-            aes(y=0.2), size = 0.8)+
-  geom_path(data = filter(tempx1, Species == "PB"),
-            size = 0.8) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="Only PB")
-
-ggsave("figures/PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-ss_only_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_SB <- expt$N_PB
-
-ss_only_SB_PB <- ss_by_a_N(expt)
-saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
-```
-
-```{r}
-ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### SB and CB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 0
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_SB <- expt$N_CB
-
-ss_only_SB_CB <- ss_by_a_N(expt)
-saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
-```
-
-```{r}
-ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
-```
-
-```{r}
-ss_only_SB_CB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_CB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:5) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 0  # this one is varied
-initial_PBs <- 10^seq(0, 8, length=grid_num_N)
-initial_SBs <- 0
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-expt$N_CB <- expt$N_PB
-
-ss_only_CB_PB <- ss_by_a_N(expt)
-saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
-```
-
-```{r}
-ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
-```
-
-```{r}
-ss_only_CB_PB %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_only_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-#### CB, SB, and PB
-
-```{r eval = run_sims}
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_CB_SB_PB <- ss_by_a_N(expt)
-saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
-```
-
-```{r}
-ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
-```
-
-```{r}
-ss_CB_SB_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species),
-            size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="CB, SB, & PB") 
-ggsave("figures/CB_SB_PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-ss_CB_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### All response measures
-
-```{r eval = run_sims}
-all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
-                    composition = rep("No organisms", grid_num_a*grid_num_N),
-                    direction = c(rep("up", grid_num_a),
-                                  rep("down", grid_num_a)))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_CB.RDS"),
-                              composition = rep("CB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB.RDS"),
-                              composition = rep("SB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_PB.RDS"),
-                              composition = rep("PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
-                              composition = rep("SB-PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
-                              composition = rep("SB-CB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
-                              composition = rep("CB-PB", grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-all_ss <- bind_rows(all_ss,
-                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
-                              composition = rep("CB-SB-PB",
-                                                grid_num_a*grid_num_N),
-                              direction = c(rep("up", grid_num_a),
-                                            rep("down", grid_num_a))))
-
-x_long <- all_ss %>%
-  gather(key = Species, value = Quantity, 2:8)
-x_wide <- x_long %>%
-  select(-initial_N_CB) %>%
-  spread(key = direction, value=Quantity, drop=T)
-
-resp_summary <- x_wide %>%
-  group_by(composition, Species) %>%
-  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
-            hyst_2 = get_hysteresis_range(up, down, a),
-            nl_up = get_nonlinearity(a, log10(up)),
-            nl_down = get_nonlinearity(a, log10(down))) %>%
-  ungroup() %>%
-  gather(key = resp_measure, value = value, 3:6) %>%
-  mutate(composition = fct_relevel(composition,
-                                   "No organisms",
-                                   "CB",
-                                   "SB",
-                                   "PB",
-                                   "CB-PB",
-                                   "SB-CB",
-                                   "SB-PB",
-                                   "CB-SB-PB"),
-         num_func_groups = case_when(composition =="No organisms" ~ 0,
-                                     composition == "CB" ~ 1,
-                                     composition == "SB" ~ 1,
-                                     composition == "PB" ~ 1,
-                                     composition == "CB-PB" ~ 2,
-                                     composition == "SB-CB" ~ 2,
-                                     composition == "SB-PB" ~ 2,
-                                     composition == "CB-SB-PB" ~ 3
-                                     ),
-         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
-         )
-saveRDS(resp_summary, "data/resp_summary.RDS")
-```
-
-```{r}
-resp_summary <- readRDS("data/resp_summary.RDS")
-resp_summary %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = value)) +
-  facet_grid(Species ~ resp_measure, scales = "free") +
-  coord_flip()
-
-```
-
-### Nonlinearity
-
-#### All species nonlinearity
-
-```{r}
-resp_summ1 <- resp_summary %>%
-  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
-  group_by(composition, Species, num_func_groups, var_type) %>%
-  summarise(ave_nl = mean(value))
-
-resp_summ1 %>%
-  ggplot() +
-  geom_point(aes(composition, ave_nl)) +
-  facet_wrap( ~ Species)
-
-```
-
-#### Substrate nonlinearity
-
-```{r}
-focus_on <- "Substrate"
-ave_nl_substrate <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_substrate %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### Organism nonlinearity
-
-```{r}
-focus_on <- "Organism"
-ave_nl_organism <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
-ave_nl_organism %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### System nonlinearity
-
-```{r}
-focus_on <- "System"
-ave_nl_system <- resp_summ1 %>%
-  group_by(composition) %>%
-  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
-ave_nl_system %>%
-  ggplot() +
-  geom_point(aes(composition, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-ggsave("figures/system_nonlinearity.png", width = 3, height = 2)
-
-```
-
-### Hysteresis
-
-#### Substrate hysteresis
-
-```{r}
-focus_on <- "Substrate"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(composition, num_func_groups, resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### Organism hysteresis
-
-```{r}
-focus_on <- "Organism"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(composition, num_func_groups, resp_measure, var_type) %>%
-  summarise(ave_hyst = sum(value) / unique(num_func_groups))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### System hysteresis (separate)
-
-```{r}
-resp_summ4 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
-  group_by(resp_measure, composition) %>%
-  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))
-
-resp_summ4 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = mean_hyst)) +
-  facet_wrap( ~ resp_measure) +
-  ggtitle("Hysteresis in system response.")
-
-```
-
-#### System hysteresis (combined)
-
-```{r}
-resp_summ5 <- resp_summ4 %>%
-  group_by(composition) %>%
-  summarise(mean_mean_hyst = mean(mean_hyst))
-
-resp_summ5 %>%
-  ggplot() +
-  geom_point(aes(x = composition, y = mean_mean_hyst)) +
-  xlab("Functional groups present") +
-  ylab("Amount of hysteresis") +
-  ylim(0,2) +
-  theme_light() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)
-
-```
-
 # Results: Varying strength of inhibition
@@
 
-# Results: Temporal state switching
-
-## Switch anoxic-oxic-anoxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_anoxic
-
-## simulation timing
-simulation_duration <- 5000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          log10a_series = lhl_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/lhl_with_defaults.RDS")
-```
-
-```{r}
-lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
-plot_dynamics(lhl_with_defaults)
-```
-
-```{r eval = run_sims}
-lhl_with_defaults$result %>%
-  filter(between(time, 1000, 73000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
-## Switch oxic-anoxic-oxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_oxic
-
-## simulation timing
-simulation_duration <- 50000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          log10a_series = hlh_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/hlh_with_defaults.RDS")
-```
-
-```{r}
-hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
-plot_dynamics(hlh_with_defaults)
-```
-
-```{r eval = run_sims}
-hlh_with_defaults$result %>%
-  filter(between(time, 1000, 100000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
 # Appendices
@@
 ```
-
-## EWS analyses
-
-Evaluation of the switch: 
-Get the functions: 
-```{r eval = F}
-source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
-```
-
-Get state shifts and calculate parameters: 
-```{r eval = F}
-x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
-x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)
-
-#first switch
-name_1='anoxic_oxic 1'
-trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
-trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
-n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
-t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
-n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
-t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
-out_df_1_N_CB <- data.frame(trans1$N_CB)
-colnames(out_df_1_N_CB) <- c("N_CB")
-out_df_1_N_PB <- data.frame(trans1$N_PB)
-colnames(out_df_1_N_PB) <- c("N_PB")
-out_df_1_N_SB <- data.frame(trans1$N_SB)
-colnames(out_df_1_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
-  ggtitle("Anoxic_oxic 1")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("anoxic_oxic 1.png")
-
-#second switch
-name_2='oxic_anoxic 1'
-trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
-trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
-n_23 <- round(length(trans2$time)*x3/100) 
-t_23 <- length(trans2$time) - n_23 + 1
-n_25 <- round(length(trans2$time)*x5/100) 
-t_25 <- length(trans2$time) - n_25 + 1
-out_df_2_N_CB <- data.frame(trans2$N_CB)
-colnames(out_df_2_N_CB) <- c("N_CB")
-out_df_2_N_PB <- data.frame(trans2$N_PB)
-colnames(out_df_2_N_PB) <- c("N_PB")
-out_df_2_N_SB <- data.frame(trans2$N_SB)
-colnames(out_df_2_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
-  ggtitle("Oxic_anoxic 1")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("oxic_anoxic 1.png")
-
-#third switch
-name_3='oxic_anoxic 2'
-trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
-trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
-n_33 <- round(length(trans3$time)*x3/100) 
-t_33 <- length(trans3$time) - n_33 + 1
-n_35 <- round(length(trans3$time)*x5/100) 
-t_35 <- length(trans3$time) - n_35 + 1
-out2_df_1_N_CB <- data.frame(trans3$N_CB)
-colnames(out2_df_1_N_CB) <- c("N_CB")
-out2_df_1_N_SB <- data.frame(trans3$N_SB)
-colnames(out2_df_1_N_SB) <- c("N_SB")
-out2_df_1_N_PB <- data.frame(trans3$N_PB)
-colnames(out2_df_1_N_PB) <- c("N_PB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
-  ggtitle("Oxic_anoxic 2")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("oxic_anoxic 2.png")
-
-#fourth switch
-name_4='anoxic_oxic 2'
-trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
-trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
-n_43 <- round(length(trans4$time)*x3/100) 
-t_43 <- length(trans4$time) - n_43 + 1
-n_45 <- round(length(trans4$time)*x5/100) 
-t_45 <- length(trans4$time) - n_45 + 1
-out2_df_2_N_CB <- data.frame(trans4$N_CB)
-colnames(out2_df_2_N_CB) <- c("N_CB")
-out2_df_2_N_PB <- data.frame(trans4$N_PB)
-colnames(out2_df_2_N_PB) <- c("N_PB")
-out2_df_2_N_SB <- data.frame(trans4$N_SB)
-colnames(out2_df_2_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
-  ggtitle("Anoxic_oxic 2")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("anoxic_oxic 2.png")
-```
-
-Rolling window analysis
-```{r eval = F}
-#analysis for 1st
-rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
-rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
-rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)
-
-#analysis for 2nd
-rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
-rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
-rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)
-
-#analysis for 3rd
-rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
-rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
-rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)
-
-#analysis for 4th
-rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
-rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
-rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)
-
-```
-
-Metric based indicators from the earlywarnings package
-```{r eval = F}
-#analysis for 1st
-earlywarnings_plot(out_df_1_N_CB, name_1, x3)
-earlywarnings_plot(out_df_1_N_PB, name_1, x3)
-earlywarnings_plot(out_df_1_N_SB, name_1, x3)
-
-#analysis for 2nd
-earlywarnings_plot(out_df_2_N_CB, name_2, x3)
-earlywarnings_plot(out_df_2_N_PB, name_2, x3)
-earlywarnings_plot(out_df_2_N_SB, name_2, x3)
-
-#analysis for 3rd
-earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
-earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
-earlywarnings_plot(out2_df_1_N_SB, name_3, x3)
-
-#analysis for 4th
-earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
-earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
-earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
-```
-
-Conditional Heteroskedasticity 
-```{r eval = F}
-#analysis for 1st
-ch <- condhet_analysis(out_df_1_N_CB, name_1)
-ch <- condhet_analysis(out_df_1_N_PB, name_1)
-ch <- condhet_analysis(out_df_1_N_SB, name_1)
-
-#analysis for 2nd
-ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
-ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
-ch2 <- condhet_analysis(out_df_2_N_SB, name_2)
-
-#analysis for 3rd
-ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
-ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
-ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)
-
-#analysis for 4th
-ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
-ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
-ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
-```
-
-Detrending fluctuation analysis 
-```{r eval = F}
-#analysis for 1st
-dfa_analysis(trans1, 2, n_15, t_15, name_1)
-dfa_analysis(trans1, 3, n_15, t_15, name_1)
-dfa_analysis(trans1, 4, n_15, t_15, name_1)
-
-#analysis for 2nd
-dfa_analysis(trans2, 2, n_25, t_25, name_2)
-dfa_analysis(trans2, 3, n_25, t_25, name_2)
-dfa_analysis(trans2, 4, n_25, t_25, name_2)
-
-#analysis for 3rd
-dfa_analysis(trans3, 2, n_35, t_35, name_3)
-dfa_analysis(trans3, 3, n_35, t_35, name_3)
-dfa_analysis(trans3, 4, n_35, t_35, name_3)
-
-#analysis for 4th
-dfa_analysis(trans4, 2, n_45, t_45, name_4)
-dfa_analysis(trans4, 3, n_45, t_45, name_4)
-dfa_analysis(trans4, 4, n_45, t_45, name_4)
-```
-
-Drift Diffusion Jump Nonparametrics Early Warning Signals:
-```{r eval = F}
-#shift 1
-ddj11<- ddj_model(out_df_1_N_CB)
-plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
-plot_ddj_2(out_df_1_N_CB, ddj11, name_1)
-
-ddj12<- ddj_model(out_df_1_N_PB)
-plot_ddj_1(out_df_1_N_PB, ddj12, name_1)
-plot_ddj_2(out_df_1_N_PB, ddj12, name_1)
-
-ddj13<- ddj_model(out_df_1_N_SB)
-plot_ddj_1(out_df_1_N_SB, ddj13, name_1)
-plot_ddj_2(out_df_1_N_SB, ddj13, name_1)
-
-#shift 2
-ddj21<- ddj_model(out_df_2_N_CB)
-plot_ddj_1(out_df_2_N_CB, ddj21, name_2)
-plot_ddj_2(out_df_2_N_CB, ddj21, name_2)
-
-ddj22<- ddj_model(out_df_2_N_PB)
-plot_ddj_1(out_df_2_N_PB, ddj22, name_2)
-plot_ddj_2(out_df_2_N_PB, ddj22, name_2)
-
-ddj23<- ddj_model(out_df_2_N_SB)
-plot_ddj_1(out_df_2_N_SB, ddj23, name_2)
-plot_ddj_2(out_df_2_N_SB, ddj23, name_2)
-
-#shift 3
-ddj31<- ddj_model(out2_df_1_N_CB)
-plot_ddj_1(out2_df_1_N_CB, ddj31, name_3)
-plot_ddj_2(out2_df_1_N_CB, ddj31, name_3)
-
-ddj32<- ddj_model(out2_df_1_N_PB)
-plot_ddj_1(out2_df_1_N_PB, ddj32, name_3)
-plot_ddj_2(out2_df_1_N_PB, ddj32, name_3)
-
-ddj33<- ddj_model(out2_df_1_N_SB)
-plot_ddj_1(out2_df_1_N_SB, ddj33, name_3)
-plot_ddj_2(out2_df_1_N_SB, ddj33, name_3)
-
-#shift 4
-ddj41<- ddj_model(out2_df_2_N_CB)
-plot_ddj_1(out2_df_2_N_CB, ddj41, name_4)
-plot_ddj_2(out2_df_2_N_CB, ddj41, name_4)
-
-ddj42<- ddj_model(out2_df_2_N_PB)
-plot_ddj_1(out2_df_2_N_PB, ddj42, name_4)
-plot_ddj_2(out2_df_2_N_PB, ddj42, name_4)
-
-ddj43<- ddj_model(out2_df_2_N_SB)
-plot_ddj_1(out2_df_2_N_SB, ddj43, name_4)
-plot_ddj_2(out2_df_2_N_SB, ddj43, name_4)
-```

### 2021-04-14 various_stuff/dev_vignettes/partial reproduction Bushetal2017.Rmd
reorganisation of vignettes
now need to get them working!

--- 2021-04-11 various_stuff/dev_vignettes/Bushplus.Rmd
@@
 ---
-title: "Diversity and the nature of responses to environmental change."
-author: "Owen Petchey & Jennifer Probst"
+title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
+author: "Owen Petchey"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output:
-  html_document:
-    toc: true
-    toc_depth: 4
-    number_sections: true
-    code_folding: hide
-    toc_float:
-      collapsed: true
-      smooth_scroll: true
+output: rmarkdown::html_vignette
+vignette: >
+  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
+  %\VignetteEngine{knitr::rmarkdown}
+  \usepackage[utf8]{inputenc}
 ---
-
-# Setup R
 
@@
 library(rootSolve)
-
+library(here)
+library(microxanox)
+
+## Set random number generator seed
+## Is redundant if there are no stochastic components in the model
 set.seed(13)
 
-run_sims <- FALSE
+run_steadystate_sims <- FALSE
 ```
@@
 * Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-* Add an investigation of how changes in tolerances alter the response to environmental change.
-
-We also add:
-
-* Option to include random noise in substrate concentrations.
-* Option to temporally vary oxygen diffusivity.
-* Option to include immigration.
-* Option to include a minimum abundances.
-
-Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups.
-That is, there is only one parameter set per functional group.
-
-Here is the original publication:
-https://www.nature.com/articles/s41467-017-00912-x
-
-And here is the supplementary information, including the table of parameter values:
-https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
-
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
-(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
+
+Please see the User manual vignette for further introduction and general methods.
 
@@
 
-## Model dimensions
-
-* Time: hours
-* Volume: litres
-* Substrate quantity: micromoles
-* Organism quantity: cells
-
-## Dynamic model
-
-```{r}
-source("functions/bushplus_dynamic_model.r")
-```
-
-## Defaults
-
-The framework used for running a simulation, and for varying aspects of a simulation run is:
-
-* Define the default simulation conditions.
-* Use these defaults unless other conditions are specified.
-* The function call `run_simulation()` would therefore run a simulation with default conditions.
-* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.
-
 ### Default parameter values
+
+Only one set of parameter values is used throughout. It gives the possibility for alternate stable states.
 
@@
   ## oxygen production per microbial cell
-  P_CB = 6e-9,
+  Pr_CB = 6e-9,
 
@@
 
-### Default event
-
-#### Default event definition
-
-Used to add noise and to ensure minimum abundance.
-Noise is added to the resource concentrations, and not to the organism abundances.
-Minimum organism abundance is set.
-I.e.
-organismal abundance cannot go below some very low level.
-
-```{r}
-source("functions/default_event_definition.r")
-```
-
-#### Default event schedule
-
-This is when an event happens
-
-```{r}
-default_event_interval <- 10
-```
-
-#### Default noise event options
-
-Default to no noise:
-
-```{r}
-default_noise_sigma <- 0
-```
-
-#### Default minimum abundance event
-
-Default to following minimums
-
-```{r}
-default_minimum_abundances <- rep(1e2, 3)
-names(default_minimum_abundances) <- c("CB", "PB", "SB") 
-```
-
-### Default simulation duration
-
-```{r}
-default_sim_duration <- 2000
-default_sim_sample_interval <- 1
-```
-
-### Default oxygen diffusivity (a) forcing
-
-Here we specify no forcing of a_0 (used unless set elsewhere).
-
-```{r}
-default_log10a_series <- c(log10(default_parameter_values["a_O"]),
-                           log10(default_parameter_values["a_O"]))
-```
-
 ### Default initial state
 
-Initial conditions favouring anoxic.
+Specify the default set of initial conditions (here favouring anoxic organisms).
 
@@
 
-## Simulation function
-
-```{r}
-source("functions/run_simulation.r")
-```
-
-## Helper functions
-
-```{r}
-source("functions/plot_dynamics.r")
-source("functions/ss_expt_a_N.r")
-source("functions/get_nonlinearity.r")
-source("functions/get_hysteresis_range.r")
+### Default simulation duration
+
+Default to the following duration and interval between samples:
+
+```{r}
+default_sim_duration <- 2000
+default_sim_sample_interval <- 1
+```
+
+### Default oxygen diffusivity (a) forcing
+
+Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.
+
+```{r}
+default_log10a_series <- c(log10(default_parameter_values["a_O"]),
+                           log10(default_parameter_values["a_O"]))
+```
+
+### Default event
+
+#### Default event definition
+
+Default is to ensure abundances are not below minimum, and to add the amount of noise specified below (none in this vignette).
+
+These default are in the definition function `default_event_definition()`
+
+#### Default event schedule
+
+Here set to an event every 10 time steps.
+
+```{r}
+default_event_interval <- 10
+```
+
+#### Default noise event options
+
+Default to no noise added to substrate concentrations:
+
+```{r}
+default_noise_sigma <- 0
+```
+
+#### Default minimum abundance event
+
+Default to following minimum abundances:
+
+```{r}
+default_minimum_abundances <- rep(1e2, 3)
+names(default_minimum_abundances) <- c("CB", "PB", "SB") 
 ```
@@
 
-```{r}
-# dynamic_model = bushplus_dynamic_model
-# event_definition = default_event_definition
-# parameter_values = default_parameter_values
-# event_interval = default_event_interval
-# noise_sigma = default_noise_sigma
-# minimum_abundances = default_minimum_abundances
-# sim_duration = default_sim_duration
-# sim_sample_interval = default_sim_sample_interval
-# log10a_forcing = default_log10a_forcing
-# initial_state = default_initial_state
-# solver_method = "radau"
-# 
-# 
-# times <- seq(0,
-#              sim_duration,
-#              by = sim_sample_interval)
-# 
-# event_times <- seq(min(times),
-#                    max(times),
-#                    by = event_interval)  
-# 
-# log10a_forcing <- approxfun(x = log10a_forcing[,1],
-#                             y = log10a_forcing[,2],
-#                             method = "linear",
-#                             rule = 2)
-# assign("log10a_forcing", log10a_forcing, envir = .GlobalEnv)
-# 
-# out <- as.data.frame(ode(y = initial_state,
-#                          times = times,
-#                          func = dynamic_model,
-#                          parms = parameter_values,
-#                          method = solver_method,
-#                          events = list(func=event_definition,
-#                                        time=event_times)))
-# 
-# sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-#                            event_definition = default_event_definition,
-#                            parameter_values = default_parameter_values,
-#                            event_interval = default_event_interval,
-#                            noise_sigma = default_noise_sigma,
-#                            minimum_abundances = default_minimum_abundances,
-#                            sim_duration = default_sim_duration,
-#                            sim_sample_interval = default_sim_sample_interval,
-#                            log10a_forcing = default_log10a_forcing,
-#                            initial_state = default_initial_state,
-#                            solver_method = "radau")
-
-#sim_res <- run_simulation()
-
-#plot_dynamics(sim_res)
-
-```
-
-# Results: Reproducing previous results
+Remove this section when the rest definitely works.
+
+```{r eval = FALSE}
+dynamic_model = bushplus_dynamic_model
+event_definition = default_event_definition
+parameter_values = default_parameter_values
+event_interval = default_event_interval
+noise_sigma = default_noise_sigma
+minimum_abundances = default_minimum_abundances
+sim_duration = default_sim_duration
+sim_sample_interval = default_sim_sample_interval
+log10a_series = default_log10a_series
+initial_state = default_initial_state
+solver_method = "radau"
+
+times <- seq(0,
+             sim_duration,
+             by = sim_sample_interval)
+
+event_times <- seq(min(times),
+                   max(times),
+                   by = event_interval)
+
+log10a_forcing <- matrix(ncol=2,
+                         byrow=F,
+                         data=c(ceiling(seq(0,
+                                            max(times),
+                                            length=length(log10a_series))),
+                                log10a_series))
+
+log10a_forcing <- approxfun(x = log10a_forcing[,1],
+                            y = log10a_forcing[,2],
+                            method = "linear",
+                            rule = 2)
+assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
+
+out <- as.data.frame(ode(y = initial_state,
+                         times = times,
+                         func = dynamic_model,
+                         parms = parameter_values,
+                         method = solver_method,
+                         events = list(func=event_definition,
+                                       time=event_times)))
+
+bushplus_dynamic_model(1, initial_state, default_parameter_values)
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+                           event_definition = default_event_definition,
+                           parameter_values = default_parameter_values,
+                           event_interval = default_event_interval,
+                           noise_sigma = default_noise_sigma,
+                           minimum_abundances = default_minimum_abundances,
+                           sim_duration = default_sim_duration,
+                           sim_sample_interval = default_sim_sample_interval,
+                           log10a_series = default_log10a_series,
+                           initial_state = default_initial_state,
+                           solver_method = "radau")
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)
+
+plot_dynamics(sim_res)
+
+```
+
+# Results
 
@@
 
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
+parameters <- default_parameter_values
+state <- init_state_favouring_anoxic
+bushplus_dynamic_model(100, state, parameters)
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+  initial_state = init_state_favouring_anoxic)
 plot_dynamics(simulation_result = sim_res)
@@
 Resource dynamics fluctuate more than in the paper.
-
-## Initial conditions favouring anoxic, with noise, no forcing
-
-Figure 2a, b in Bush et al, but with noise
-
-```{r}
-## add some noise
-noise_sigma_0.01 <- 0.01
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          noise_sigma = noise_sigma_0.01)
-plot_dynamics(sim_res)
-#ggsave("high_SB_PB.png")
-```
 
@@
 
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic)
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+  initial_state = init_state_favouring_oxic)
 
@@
 
-## Initial conditions favouring oxic, with noise, no forcing
-
-Figure 2c, d in Bush et al. but with noise.
-
-```{r}
-
+## Stable states
+
+```{r eval = run_steadystate_sims}
+ssfind_minimum_abundances <- rep(0, 3)
+names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
+ssfind_simulation_duration <- 10000
+ssfind_simulation_sampling_interval <- ssfind_simulation_duration
+ssfind_event_interval <- ssfind_simulation_duration
+ssfind_parameters <- default_parameter_values
+# ssfind_init_state <- c(N_CB = NA, 
+#                        N_PB = NA,
+#                        N_SB = NA,
+#                        SO = 500,
+#                        SR = 50,
+#                        O = 30,
+#                        P = 10)
+##ssfind_init_state <- default_initial_state
+ssfind_init_state <- init_state_favouring_anoxic
+ssfind_init_state <- c(N_CB = NA, #this one is varied
+           N_PB = 1e8,
+           N_SB = 1e8,
+           SO = 200,
+           SR = 200,
+           O = 10,
+           P = 9.5)
+grid_num_a <- 2
+a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
+grid_num_N <- 2
+
+initial_CBs <- 10^seq(0, 8, length=grid_num_N)
+initial_PBs <- 1e5
+initial_SBs <- 1e5
+expt <- expand.grid(N_CB = initial_CBs,
+                    N_PB = initial_PBs,
+                    N_SB = initial_SBs,
+                    a_O = a_Os)
+
+ss_CB_SB_PB <- ss_by_a_N(expt)
+save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
+```
+
+```{r}
+load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
+ss_CB_SB_PB %>%
+  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
+  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
+  ggplot() +
+  geom_path(aes(x = a, y = log10(Quantity), col = Species),
+            size=0.8) +
+  ylab(expression(Log[10](Abundance))) +
+  xlab(expression(Log[10](Oxygen~diffusivity))) +
+  theme_light() +
+  ylim(0,10) +
+  theme(plot.title=element_text(size=10, hjust=0.5)) +
+  labs(title="CB, SB, & PB")  %>%
+  print()
+##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)
+
+```
+
+```{r}
+p1 <- ss_CB_SB_PB %>%
+  select(a, SO, SR, O, P) %>%
+  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
+  ggplot() +
+  geom_path(aes(x = a, y = log10(Quantity), col = Species))
+print(p1)
+```
+
+## Temporal state switching
+
+### Anoxic-oxic-anoxic
+
+```{r eval = run_sims}
+## anoxic initial state
+init_state <- init_state_favouring_anoxic
+
+## simulation timing
+simulation_duration <- 5000
+simulation_sampling_interval <- 10
+event_interval_t <- 100
+
+## low-high-low oxgygen diffusivity 
+lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)
+
+## run simulation
+sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
+                          log10a_series = lhl_log10a_series,
+                          sim_duration = simulation_duration,
+                          sim_sample_interval = simulation_sampling_interval,
+                          event_interval = event_interval_t)
+saveRDS(sim_res, "data/lhl_with_defaults.RDS")
+```
+
+```{r}
+lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
+plot_dynamics(lhl_with_defaults)
+```
+
+```{r eval = run_sims}
+lhl_with_defaults$result %>%
+  filter(between(time, 1000, 73000)) %>%
+  ggplot() +
+  geom_path(aes(x=a, y=log10(O)),
+            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
+```
+
+### Oxic-anoxic-oxic
+
+```{r eval = run_sims}
+## anoxic initial state
+init_state <- init_state_favouring_oxic
+
+## simulation timing
+simulation_duration <- 50000
+simulation_sampling_interval <- 10
+event_interval_t <- 100
+
+## low-high-low oxgygen diffusivity 
+hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)
+
+## run simulation
 sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          noise_sigma = noise_sigma_0.01)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-
-# Results: Composition/diversity and nature of response
+                          log10a_series = hlh_log10a_series,
+                          sim_duration = simulation_duration,
+                          sim_sample_interval = simulation_sampling_interval,
+                          event_interval = event_interval_t)
+saveRDS(sim_res, "data/hlh_with_defaults.RDS")
+```
+
+```{r}
+hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
+plot_dynamics(hlh_with_defaults)
+```
+
+```{r eval = run_sims}
+hlh_with_defaults$result %>%
+  filter(between(time, 1000, 100000)) %>%
+  ggplot() +
+  geom_path(aes(x=a, y=log10(O)),
+            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
+```
+
+## Composition and nature of response
 
@@
 ```
-
-# Results: Varying strength of inhibition
-
-## Introduction
-
-Three parameters set the strength of inhibition (half-inhibition constants):
-
-* h_SR_CB = 300,
-* h_O_PB = 100,
-* h_O_SB = 100,
-
-These work by the inhibition function:
-
-$\frac{1}{1 + \frac{X}{h_X}}$
-
-```{r}
-inhibition
-```
-
-Plot some examples:
-
-```{r}
-default_parameter_values["h_SR_CB"]
-SR <- 10^seq(1.4, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
-  ylab("Multiplicative effect\non CB growth rate")
-```
-
-```{r}
-default_parameter_values["h_O_PB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
-  ylab("Multiplicative effect\non PB growth rate")
-```
-
-```{r}
-default_parameter_values["h_O_SB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
-  ylab("Multiplicative effect\non SB growth rate")
-```
-
-```{r}
-proportion_change <- 2
-default_parameter_values["h_O_SB"]
-O <- 10^seq(-0.2, 2.5, length = 100)
-ggplot() +
-  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
-  ylab("Multiplicative effect\non SB growth rate")
-```
-
-## Experiment
-
-Increasing the inhibition half saturation constants weakens the inhibition.
-
-```{r}
-ssfind_minimum_abundances <- default_minimum_abundances
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = NA,
-           N_SB = NA,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-grid_num_mult <- 20
-grid_num_N <- 2
-grid_num_a <- 100
-
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    inhib_const_mult = inhib_const_mults,
-                    a_O = a_Os)
-```
-
-```{r eval = run_sims}
-ss_h1 <- ss_by_a_h_N(expt)
-
-saveRDS(ss_h1, "data/ss_h1.RDS")
-```
-
-```{r}
-ss_h1 <- readRDS("data/ss_h1.RDS")
-```
-
-## Results
-
-### Close to default inhibition
-
-```{r}
-inhib_const_mults
-focal_inhibition_multiplier <- inhib_const_mults[14]
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Double strength inhibition
-
-```{r}
-## YES! 0.5 does mean twice as strong inhibition
-focal_inhibition_multiplier <- 0.5
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Weaker inhibition
-
-```{r}
-## YES! 2 does mean half as strong inhibition
-focal_inhibition_multiplier <- 2^0.4
-
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, N_CB, N_SB, N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-```{r}
-ss_h1 %>%
-  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-```
-
-### Nonlinearity
-
-```{r}
-x_long <- ss_h1 %>%
-  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
-  gather(key = Species, value = Quantity, 3:9) 
-x_wide <- x_long %>%
-  select(-initial_N_CB) %>%
-  spread(key = direction, value=Quantity, drop=T)
-
-resp_summary <- x_wide %>%
-  group_by(inhib_const_mult, Species) %>%
-  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
-            hyst_2 = get_hysteresis_range(up, down, a),
-            nl_up = get_nonlinearity(a, log10(up)),
-            nl_down = get_nonlinearity(a, log10(down))) %>%
-  ungroup() %>%
-  gather(key = resp_measure, value = value, 3:6) %>%
-  mutate(
-         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
-         )
-```
-
-#### All species nonlinearity
-
-```{r}
-resp_summ1 <- resp_summary %>%
-  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
-  group_by(inhib_const_mult, Species, var_type) %>%
-  summarise(ave_nl = mean(value))
-
-resp_summ1 %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_nl)) +
-  facet_wrap( ~ Species)
-
-```
-
-#### Substrate nonlinearity
-
-```{r}
-focus_on <- "Substrate"
-ave_nl_substrate <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_substrate %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### Organism nonlinearity
-
-```{r}
-focus_on <- "Organism"
-ave_nl_organism <- resp_summ1 %>%
-  filter(var_type == focus_on) %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_organism %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-```
-
-#### System nonlinearity
-
-```{r}
-focus_on <- "System"
-ave_nl_system <- resp_summ1 %>%
-  group_by(inhib_const_mult) %>%
-  summarise(ave_ave_nl = mean(ave_nl))
-ave_nl_system %>%
-  ggplot() +
-  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
-  ggtitle(paste("Nonlinearity in", focus_on, "response."))
-ggsave("figures/system_nonlinearity.png", width = 3, height = 2)
-
-```
-
-### Hysteresis
-
-#### Substrate hysteresis
-
-```{r}
-focus_on <- "Substrate"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(inhib_const_mult, resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### Organism hysteresis
-
-```{r}
-focus_on <- "Organism"
-resp_summ3 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2"),
-         var_type == focus_on) %>%
-  group_by(inhib_const_mult,  resp_measure, var_type) %>%
-  summarise(ave_hyst = mean(value))
-
-resp_summ3 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
-  facet_wrap( ~ resp_measure, scales="free") +
-  ggtitle(paste("Hysteresis in", focus_on, "response."))
-
-```
-
-#### System hysteresis (separate)
-
-```{r}
-resp_summ4 <- resp_summary %>%
-  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
-  group_by(resp_measure, inhib_const_mult) %>%
-  summarise(mean_hyst = mean(value))
-
-resp_summ4 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = mean_hyst)) +
-  facet_wrap( ~ resp_measure) +
-  ggtitle("Hysteresis in system response.")
-
-```
-
-#### System hysteresis (combined)
-
-```{r}
-resp_summ5 <- resp_summ4 %>%
-  group_by(inhib_const_mult) %>%
-  summarise(mean_mean_hyst = mean(mean_hyst))
-
-resp_summ5 %>%
-  ggplot() +
-  geom_point(aes(x = inhib_const_mult, y = mean_mean_hyst)) +
-  xlab("Organismal tolerance") +
-  ylab("Amount of hysteresis") +
-  theme_light() +
-  ylim(0,2)
-ggsave("figures/inhibition_system_hysteresis.png", width = 2.5, height = 2)
-
-```
-
-# Results: Temporal state switching
-
-## Switch anoxic-oxic-anoxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_anoxic
-
-## simulation timing
-simulation_duration <- 5000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
-                          log10a_series = lhl_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/lhl_with_defaults.RDS")
-```
-
-```{r}
-lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
-plot_dynamics(lhl_with_defaults)
-```
-
-```{r eval = run_sims}
-lhl_with_defaults$result %>%
-  filter(between(time, 1000, 73000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
-## Switch oxic-anoxic-oxic
-
-```{r eval = run_sims}
-## anoxic initial state
-init_state <- init_state_favouring_oxic
-
-## simulation timing
-simulation_duration <- 50000
-simulation_sampling_interval <- 10
-event_interval_t <- 100
-
-## low-high-low oxgygen diffusivity 
-hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)
-
-## run simulation
-sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
-                          log10a_series = hlh_log10a_series,
-                          sim_duration = simulation_duration,
-                          sim_sample_interval = simulation_sampling_interval,
-                          event_interval = event_interval_t)
-saveRDS(sim_res, "data/hlh_with_defaults.RDS")
-```
-
-```{r}
-hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
-plot_dynamics(hlh_with_defaults)
-```
-
-```{r eval = run_sims}
-hlh_with_defaults$result %>%
-  filter(between(time, 1000, 100000)) %>%
-  ggplot() +
-  geom_path(aes(x=a, y=log10(O)),
-            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
-```
-
-# Appendices
-
-Rainer: don't change what comes below
-
-## runsteady function not finding correct values
-
-```{r eval = FALSE}
-
-init_state_anoxic_2 <- c(N_CB = 1e1,
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 300,
-           SR = 300,
-           O = 1e-5,
-           P = 1e1)
-
-RS <- runsteady(y = init_state_anoxic_2,
-                fun = bushplus_dynamic_model,
-                parms = default_parameter_values,
-                times = c(0, 1e5),
-                stol=1)
-RS$y
-RS <- runsteady(y = init_state_favouring_oxic,
-                fun = bushplus_dynamic_model,
-                parms = default_parameter_values,
-                times = c(0, 1e5),
-                stol=1)
-RS$y
-
-```
-
-## Separatrix
-
-Note that this section not working / updated since the code was made into functions
-
-### Effect of N_CB and oxygen diffusability
-
-```{r eval = FALSE}
-
-noise_sigma <- 0.0
-
-get_final_states_N_CB <- function(x) {
-
-  #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
-  state["N_CB"] <- x["N_CB"]
-  parameters["a_O"] <- x["a_O"]
-
-  ## update forcing function
-  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
-                                             log10(parameters["a_O"]),
-                                             max(times),
-                                             log10(parameters["a_O"])))
-  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
-                             y = a_forcing1[,2],
-                             method = "linear",
-                             rule = 2)
-
-  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
-  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)
-
-  as.data.frame(ode(y = state,
-                    times = times,
-                    func = micro_mod_var,
-                    parms = parameters,
-                    method = "radau",
-                    events = list(func=noise_events, time=times)))[2,-1] 
-  #gives you values for the last time step
-}
-
-times <- seq(0, 50000, by = 50000)
-
-state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-#different combinations of N_CB and oxygen diffusability
-grid_num <- 50
-expt <- expand.grid(N_CB = 10^seq(0, 8, length=grid_num),
-                    a_O = 10^seq(-3.5, -2, length=grid_num))
-
-result <- apply(expt, 1, function(x) get_final_states_N_CB(x)) %>%
-                          tibble() %>%
-                          unnest() %>%
-                          mutate(initial_N_CB=expt$N_CB,
-                                 a_O=expt$a_O,
-                                 highlow=ifelse(N_CB>2e8, 'high', 'low'))
-
-result <- result %>%
-  mutate(N_CB = ifelse(N_CB < minimum_abundance, minimum_abundance, N_CB))
-
-result$highlow <- as.factor(result$highlow)
-saveRDS(result, "data/CB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-result <- readRDS("data/CB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-
-#plot N_CB / P_CB against a_0
-p <- ggplot(result) + 
-  geom_line(aes(x=log10(a_O), y=log10(N_CB), color=highlow), size=1.25) +
-  ggtitle("Stable states of the system")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('Initial cyanobacteria abundance')
-p <- p + geom_point(data=result, aes(x=log10(a_O), y=log10(initial_N_CB), color=highlow)) + labs(color = "Final CB concentration")
-plot(p)
-#ggsave("initial conditions_1.png")
-```
-
-```{r eval = FALSE}
-#show hysterisis effects for all the parameters
-#plot oxygen against diffusability
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(O))))+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[oxygen concentration]')+
-  geom_segment(aes(x=log10(result$a_O[839]), y=log10(result$O[839]), xend=log10(result$a_O[838]), yend=log10(result$O[838])), arrow=arrow(), colour="blue", size=0.5)+
-  geom_segment(aes(x=log10(result$a_O[1761]), y=log10(result$O[1761]), xend=log10(result$a_O[1762]), yend=log10(result$O[1762])), arrow=arrow(), colour="blue", size=0.5)+
-  geom_segment(aes(x=-4.25, y=-0.45, xend=-3.75, yend=0.1), arrow=arrow(), colour="blue", size=0.5)+
-  geom_curve(aes(x = -3.75, y = 1.9, xend = -4.25, yend = 1.5), arrow=arrow(), colour='blue', size = 0.5, curvature = -0.2)
-#ggsave("hysterisis_CB_1.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(P))))+
-  ggtitle("Hysterisis effect of phoshate concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[phosphate concentration]')
-#ggsave("hysterisis_CB_2.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(SO))))+
-  ggtitle("Hysterisis effect of oxidized sulfur concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[oxydized sulfur concentration]')
-#ggsave("hysterisis_CB_3.png")
-```
-
-```{r eval = FALSE}
-ggplot(result) + 
-  geom_point(aes(x=log10(a_O), y=(log10(SR))))+
-  ggtitle("Hysterisis effect of reduced sulfur concentration")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('log[reduced sulfur] concentration]')
-#ggsave("hysterisis_CB_4.png")
-```
-
-These graphs show an at least qualitative match to the result in the paper.
-
-### Effect of PB/SB depending on oxygen diffusability
-
-```{r eval = FALSE}
-noise_sigma <- 0.00
-get_final_states_N_PB <- function(x) { 
-  #give in N_CB, N_SB, N_PB and a_O values and put these into state and parameters
-  state["N_PB"] <- x["var"]
-  state["N_SB"] <- x["var"]
-  parameters["a_O"] <- x["a_O"]
-
-  ## update forcing function
-  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
-                                             log10(parameters["a_O"]),
-                                             max(times),
-                                             log10(parameters["a_O"])))
-  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
-                             y = a_forcing1[,2],
-                             method = "linear",
-                             rule = 2)
-  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
-  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)
-
-  as.data.frame(ode(y = state,
-                         times = times,
-                         func = micro_mod_var,
-                         parms = parameters,
-                         method = "radau", events = list(func=noise_events, time=times)))[2,-1] 
-  #gives you values for the last time step
-}
-
-times <- seq(0, 50000, by = 5000)
-state <- c(N_CB = 5, 
-           N_PB = NA,#this one is varied
-           N_SB = NA,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-
-#different combinations of N_PB and oxygen diffusivity
-grid_num <- 50
-expt <- expand.grid(var=10^seq(0, 7, length=grid_num),
-                    a_O = 10^seq(-3, -2, length=grid_num))
-
-result2 <- apply(expt, 1, function(x) get_final_states_N_PB(x)) %>%
-                          tibble() %>%
-                          unnest() %>%
-                          mutate(initial_N_PB=expt$var,
-                                 initial_N_SB=expt$var,
-                                 a_O=expt$a_O,
-                                 highlow=ifelse(N_PB>1e8, 'high', 'low'))
-
-result2 <- result2 %>%
-  mutate(N_PB = ifelse(N_PB < minimum_abundance, minimum_abundance, N_PB))
-
-result2$highlow <- as.factor(result2$highlow)
-saveRDS(result2, "data/PBSB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-result2 <- readRDS("data/PBSB_oxdiff_hyst.RDS")
-```
-
-```{r eval = FALSE}
-#plot N_PB against a_0
-p <- ggplot(result2) + 
-  geom_line(aes(x=log10(a_O), y=log10(N_PB), color=highlow), size=1.25) +
-  ggtitle("Stable states of the system")+
-  xlab('log[oxygen diffusivity]') +
-  ylab('Initial PS and SR bacteria concentration')
-p <- p + geom_point(data=result2,
-                    aes(x=log10(a_O), y=log10(initial_N_PB), color=highlow))
-p <- p + labs(color = "Final PB and SB concentration")
-plot(p)
-#ggsave("initial conditions_2.png")
-```
-
-## EWS analyses
-
-Evaluation of the switch: 
-Get the functions: 
-```{r eval = F}
-source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
-```
-
-Get state shifts and calculate parameters: 
-```{r eval = F}
-x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
-x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)
-
-#first switch
-name_1='anoxic_oxic 1'
-trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
-trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
-n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
-t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
-n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
-t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
-out_df_1_N_CB <- data.frame(trans1$N_CB)
-colnames(out_df_1_N_CB) <- c("N_CB")
-out_df_1_N_PB <- data.frame(trans1$N_PB)
-colnames(out_df_1_N_PB) <- c("N_PB")
-out_df_1_N_SB <- data.frame(trans1$N_SB)
-colnames(out_df_1_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
-  ggtitle("Anoxic_oxic 1")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("anoxic_oxic 1.png")
-
-#second switch
-name_2='oxic_anoxic 1'
-trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
-trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
-n_23 <- round(length(trans2$time)*x3/100) 
-t_23 <- length(trans2$time) - n_23 + 1
-n_25 <- round(length(trans2$time)*x5/100) 
-t_25 <- length(trans2$time) - n_25 + 1
-out_df_2_N_CB <- data.frame(trans2$N_CB)
-colnames(out_df_2_N_CB) <- c("N_CB")
-out_df_2_N_PB <- data.frame(trans2$N_PB)
-colnames(out_df_2_N_PB) <- c("N_PB")
-out_df_2_N_SB <- data.frame(trans2$N_SB)
-colnames(out_df_2_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
-  ggtitle("Oxic_anoxic 1")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("oxic_anoxic 1.png")
-
-#third switch
-name_3='oxic_anoxic 2'
-trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
-trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
-n_33 <- round(length(trans3$time)*x3/100) 
-t_33 <- length(trans3$time) - n_33 + 1
-n_35 <- round(length(trans3$time)*x5/100) 
-t_35 <- length(trans3$time) - n_35 + 1
-out2_df_1_N_CB <- data.frame(trans3$N_CB)
-colnames(out2_df_1_N_CB) <- c("N_CB")
-out2_df_1_N_SB <- data.frame(trans3$N_SB)
-colnames(out2_df_1_N_SB) <- c("N_SB")
-out2_df_1_N_PB <- data.frame(trans3$N_PB)
-colnames(out2_df_1_N_PB) <- c("N_PB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
-  ggtitle("Oxic_anoxic 2")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("oxic_anoxic 2.png")
-
-#fourth switch
-name_4='anoxic_oxic 2'
-trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
-trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
-n_43 <- round(length(trans4$time)*x3/100) 
-t_43 <- length(trans4$time) - n_43 + 1
-n_45 <- round(length(trans4$time)*x5/100) 
-t_45 <- length(trans4$time) - n_45 + 1
-out2_df_2_N_CB <- data.frame(trans4$N_CB)
-colnames(out2_df_2_N_CB) <- c("N_CB")
-out2_df_2_N_PB <- data.frame(trans4$N_PB)
-colnames(out2_df_2_N_PB) <- c("N_PB")
-out2_df_2_N_SB <- data.frame(trans4$N_SB)
-colnames(out2_df_2_N_SB) <- c("N_SB")
-
-ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
-  ggtitle("Anoxic_oxic 2")+
-  xlab('Time') +
-  ylab('Trans_quantity') +
-  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
-  geom_line()+
-  labs(color = "Organism/Substrate")
-ggsave("anoxic_oxic 2.png")
-```
-
-Rolling window analysis
-```{r eval = F}
-#analysis for 1st
-rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
-rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
-rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)
-
-#analysis for 2nd
-rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
-rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
-rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)
-
-#analysis for 3rd
-rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
-rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
-rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)
-
-#analysis for 4th
-rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
-rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
-rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)
-
-```
-
-Metric based indicators from the earlywarnings package
-```{r eval = F}
-#analysis for 1st
-earlywarnings_plot(out_df_1_N_CB, name_1, x3)
-earlywarnings_plot(out_df_1_N_PB, name_1, x3)
-earlywarnings_plot(out_df_1_N_SB, name_1, x3)
-
-#analysis for 2nd
-earlywarnings_plot(out_df_2_N_CB, name_2, x3)
-earlywarnings_plot(out_df_2_N_PB, name_2, x3)
-earlywarnings_plot(out_df_2_N_SB, name_2, x3)
-
-#analysis for 3rd
-earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
-earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
-earlywarnings_plot(out2_df_1_N_SB, name_3, x3)
-
-#analysis for 4th
-earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
-earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
-earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
-```
-
-Conditional Heteroskedasticity 
-```{r eval = F}
-#analysis for 1st
-ch <- condhet_analysis(out_df_1_N_CB, name_1)
-ch <- condhet_analysis(out_df_1_N_PB, name_1)
-ch <- condhet_analysis(out_df_1_N_SB, name_1)
-
-#analysis for 2nd
-ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
-ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
-ch2 <- condhet_analysis(out_df_2_N_SB, name_2)
-
-#analysis for 3rd
-ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
-ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
-ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)
-
-#analysis for 4th
-ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
-ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
-ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
-```
-
-Detrending fluctuation analysis 
-```{r eval = F}
-#analysis for 1st
-dfa_analysis(trans1, 2, n_15, t_15, name_1)
-dfa_analysis(trans1, 3, n_15, t_15, name_1)
-dfa_analysis(trans1, 4, n_15, t_15, name_1)
-
-#analysis for 2nd
-dfa_analysis(trans2, 2, n_25, t_25, name_2)
-dfa_analysis(trans2, 3, n_25, t_25, name_2)
-dfa_analysis(trans2, 4, n_25, t_25, name_2)
-
-#analysis for 3rd
-dfa_analysis(trans3, 2, n_35, t_35, name_3)
-dfa_analysis(trans3, 3, n_35, t_35, name_3)
-dfa_analysis(trans3, 4, n_35, t_35, name_3)
-
-#analysis for 4th
-dfa_analysis(trans4, 2, n_45, t_45, name_4)
-dfa_analysis(trans4, 3, n_45, t_45, name_4)
-dfa_analysis(trans4, 4, n_45, t_45, name_4)
-```
-
-Drift Diffusion Jump Nonparametrics Early Warning Signals:
-```{r eval = F}
-#shift 1
-ddj11<- ddj_model(out_df_1_N_CB)
-plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
-plot_ddj_2(out_df_1_N_CB, ddj11, name_1)
-
-ddj12<- ddj_model(out_df_1_N_PB)
-plot_ddj_1(out_df_1_N_PB, ddj12, name_1)
-plot_ddj_2(out_df_1_N_PB, ddj12, name_1)
-
-ddj13<- ddj_model(out_df_1_N_SB)
-plot_ddj_1(out_df_1_N_SB, ddj13, name_1)
-plot_ddj_2(out_df_1_N_SB, ddj13, name_1)
-
-#shift 2
-ddj21<- ddj_model(out_df_2_N_CB)
-plot_ddj_1(out_df_2_N_CB, ddj21, name_2)
-plot_ddj_2(out_df_2_N_CB, ddj21, name_2)
-
-ddj22<- ddj_model(out_df_2_N_PB)
-plot_ddj_1(out_df_2_N_PB, ddj22, name_2)
-plot_ddj_2(out_df_2_N_PB, ddj22, name_2)
-
-ddj23<- ddj_model(out_df_2_N_SB)
-plot_ddj_1(out_df_2_N_SB, ddj23, name_2)
-plot_ddj_2(out_df_2_N_SB, ddj23, name_2)
-
-#shift 3
-ddj31<- ddj_model(out2_df_1_N_CB)
-plot_ddj_1(out2_df_1_N_CB, ddj31, name_3)
-plot_ddj_2(out2_df_1_N_CB, ddj31, name_3)
-
-ddj32<- ddj_model(out2_df_1_N_PB)
-plot_ddj_1(out2_df_1_N_PB, ddj32, name_3)
-plot_ddj_2(out2_df_1_N_PB, ddj32, name_3)
-
-ddj33<- ddj_model(out2_df_1_N_SB)
-plot_ddj_1(out2_df_1_N_SB, ddj33, name_3)
-plot_ddj_2(out2_df_1_N_SB, ddj33, name_3)
-
-#shift 4
-ddj41<- ddj_model(out2_df_2_N_CB)
-plot_ddj_1(out2_df_2_N_CB, ddj41, name_4)
-plot_ddj_2(out2_df_2_N_CB, ddj41, name_4)
-
-ddj42<- ddj_model(out2_df_2_N_PB)
-plot_ddj_1(out2_df_2_N_PB, ddj42, name_4)
-plot_ddj_2(out2_df_2_N_PB, ddj42, name_4)
-
-ddj43<- ddj_model(out2_df_2_N_SB)
-plot_ddj_1(out2_df_2_N_SB, ddj43, name_4)
-plot_ddj_2(out2_df_2_N_SB, ddj43, name_4)
-```

### 2021-04-14 various_stuff/dev_vignettes/within-function-group-diversity.Rmd
reorganisation of vignettes
now need to get them working!

--- 2021-04-13 various_stuff/dev_vignettes/Bush_basic_reproduction_new.Rmd
@@
 ---
-title: "Reproducing some results of Bush et al 2017"
+title: A study of the effects of within functional group diversity on the response
+  to environmental change.
 author: "Owen Petchey"
@@
 vignette: >
-  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
+  %\VignetteIndexEntry{Effects of within functional group diversity}
   %\VignetteEngine{knitr::rmarkdown}
@@
 * Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-* Add an investigation of how changes in tolerances alter the response to environmental change.
 
-We also add:
-
-* Option to include random noise in substrate concentrations.
-* Option to temporally vary oxygen diffusivity.
-* Option to include immigration.
-* Option to include a minimum abundances.
-
-Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups.
-That is, there is only one parameter set per functional group.
-
-Here is the original publication:
-https://www.nature.com/articles/s41467-017-00912-x
-
-And here is the supplementary information, including the table of parameter values:
-https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf
-
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
-(There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
+Please see the User manual vignette for further introduction and general methods.
 
@@
 
-## Model dimensions
+### Default initial state
 
-* Time: hours
-* Volume: litres
-* Substrate quantity: micromoles
-* Organism quantity: cells
-
-## Modelling framework
-
-The framework used for running a simulation, and for varying aspects of a simulation run is:
-
-* Define the default simulation conditions.
-* Use these defaults unless other conditions are specified.
-* The function call `run_simulation()` would therefore run a simulation with default conditions.
-* Whereas, for example, the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values.
-
-### Dynamic model
-
-The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model_strains()`.
-This is an augemented version that can handle multiple strains within each of the three functional groups.
-The basic version that has only one strain per functional group is `bushplus_dynamic_model()`.
+No default.
 
@@
 
-```{r}
-default_parameter_values <- c(
-
-  ## maximum specific growth rates
-  g_max_CB = 0.05,
-  g_max_PB = 0.07,
-  g_max_SB = 0.1,
-
-  ## half-saturation constants
-  k_PB_SR = 10,
-  k_SB_SO = 5,
-  k_CB_P = 0.2,
-  k_PB_P = 0.5,
-  k_SB_P = 0.5,
-
-  ## half-inhibition constant
-  h_SR_CB = 300,
-  h_O_PB = 100,
-  h_O_SB = 100,
-
-  ## yield (cells per micromole)
-  Y_SO_SB = 3.33e7,  
-  Y_SR_PB = 1.25e7, ## Y_S_PB in the main text
-  y_P_CB = 1.67e8,
-  y_P_PB = 1.67e8,
-  y_P_SB = 1.67e8,
-
-  ## oxygen production per microbial cell
-  Pr_CB = 6e-9,
-
-  ## mortality rate
-  m_CB = 0.02,
-  m_PB = 0.028,
-  m_SB = 0.04,
-
-  ## substrate diffusivity
-  a_S = 0.001,
-  a_O = 8e-4,
-  a_P = 0.01,
-
-  ## background substrate concentration
-  back_SR = 300,
-  back_SO = 300,
-  back_O = 300,
-  back_P = 9.5,
-
-  ## immigration rates (cells per time)
-  i_CB <- 0,
-  i_PB <- 0,
-  i_SB <- 0,
-
-  ## oxidisation rate of reduced sulphur
-  c = 4e-5
-  )
-```
+No default.
 
@@
 
-Used to add noise and to ensure minimum abundance.
-Noise is added to the resource concentrations, and not to the organism abundances.
-Minimum organism abundance is set.
-I.e.
-organismal abundance cannot go below some very low level.
+The default definition is in the function `default_event_definition()`
 
-The event definition is made in the form of a function. The default definition is in the function `default_event_definition()`
+This currently ensure abundances do not go below minimum and adds noise (or does not, see below) to an abiotic state variable.
 
@@
 
-Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.
+No forcing -- constant diffusivity at a level that allows alternate stable states.
 
@@
 
-### Default initial state
+# Test results
 
-Specify the default set of initial conditions (here favouring anoxic organisms).
-
-```{r}
-default_initial_state <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-```
-
-## Code test
-
-```{r}
-dynamic_model = bushplus_dynamic_model
-event_definition = default_event_definition
-parameter_values = default_parameter_values
-event_interval = default_event_interval
-noise_sigma = default_noise_sigma
-minimum_abundances = default_minimum_abundances
-sim_duration = default_sim_duration
-sim_sample_interval = default_sim_sample_interval
-log10a_series = default_log10a_series
-initial_state = default_initial_state
-solver_method = "radau"
-
-times <- seq(0,
-             sim_duration,
-             by = sim_sample_interval)
-
-event_times <- seq(min(times),
-                   max(times),
-                   by = event_interval)
-
-log10a_forcing <- matrix(ncol=2,
-                         byrow=F,
-                         data=c(ceiling(seq(0,
-                                            max(times),
-                                            length=length(log10a_series))),
-                                log10a_series))
-
-log10a_forcing <- approxfun(x = log10a_forcing[,1],
-                            y = log10a_forcing[,2],
-                            method = "linear",
-                            rule = 2)
-assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
-
-out <- as.data.frame(ode(y = initial_state,
-                         times = times,
-                         func = dynamic_model,
-                         parms = parameter_values,
-                         method = solver_method,
-                         events = list(func=event_definition,
-                                       time=event_times)))
-
-bushplus_dynamic_model(1, initial_state, default_parameter_values)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-                           event_definition = default_event_definition,
-                           parameter_values = default_parameter_values,
-                           event_interval = default_event_interval,
-                           noise_sigma = default_noise_sigma,
-                           minimum_abundances = default_minimum_abundances,
-                           sim_duration = default_sim_duration,
-                           sim_sample_interval = default_sim_sample_interval,
-                           log10a_series = default_log10a_series,
-                           initial_state = default_initial_state,
-                           solver_method = "radau")
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)
-
-plot_dynamics(sim_res)
-
-```
-
-# Results
-
-## Initial conditions favouring anoxic, no noise, no forcing
-
-Figure 2a, b in Bush et al.
-
-```{r}
-
-init_state_favouring_anoxic <- c(N_CB = 5e1,
-           N_PB = 1e7,
-           N_SB = 1e7,
-           SO = 300,
-           SR = 300,
-           O = 1e1,
-           P = 1e1)
-
-parameters <- default_parameter_values
-state <- init_state_favouring_anoxic
-bushplus_dynamic_model(100, state, parameters)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-  initial_state = init_state_favouring_anoxic)
-plot_dynamics(simulation_result = sim_res)
-
-#ggsave('bacteria_no_noise.png')
-```
-
-Population dynamics seem as in the paper.
-Resource dynamics fluctuate more than in the paper.
-
-## Initial conditions favouring oxic, no noise, no forcing
-
-Figure 2c, d in Bush et al.
-
-```{r} 
-init_state_favouring_oxic <- c(N_CB = 1e7,
-           N_PB = 1e5,
-           N_SB = 1e5,
-           SO = 500,
-           SR = 50,
-           O = 30,
-           P = 10)
-
-sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
-  initial_state = init_state_favouring_oxic)
-
-plot_dynamics(sim_res)
-##ggsave("high_CB.png")
-```
-Population and resource dynamics seen as those in the paper.
-
-## Stable states
-
-```{r eval = run_steadystate_sims}
-ssfind_minimum_abundances <- rep(0, 3)
-names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
-ssfind_simulation_duration <- 10000
-ssfind_simulation_sampling_interval <- ssfind_simulation_duration
-ssfind_event_interval <- ssfind_simulation_duration
-ssfind_parameters <- default_parameter_values
-# ssfind_init_state <- c(N_CB = NA, 
-#                        N_PB = NA,
-#                        N_SB = NA,
-#                        SO = 500,
-#                        SR = 50,
-#                        O = 30,
-#                        P = 10)
-##ssfind_init_state <- default_initial_state
-ssfind_init_state <- init_state_favouring_anoxic
-ssfind_init_state <- c(N_CB = NA, #this one is varied
-           N_PB = 1e8,
-           N_SB = 1e8,
-           SO = 200,
-           SR = 200,
-           O = 10,
-           P = 9.5)
-grid_num_a <- 2
-a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
-grid_num_N <- 2
-
-initial_CBs <- 10^seq(0, 8, length=grid_num_N)
-initial_PBs <- 1e5
-initial_SBs <- 1e5
-expt <- expand.grid(N_CB = initial_CBs,
-                    N_PB = initial_PBs,
-                    N_SB = initial_SBs,
-                    a_O = a_Os)
-
-ss_CB_SB_PB <- ss_by_a_N(expt)
-save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
-```
-
-```{r}
-load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
-ss_CB_SB_PB %>%
-  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species),
-            size=0.8) +
-  ylab(expression(Log[10](Abundance))) +
-  xlab(expression(Log[10](Oxygen~diffusivity))) +
-  theme_light() +
-  ylim(0,10) +
-  theme(plot.title=element_text(size=10, hjust=0.5)) +
-  labs(title="CB, SB, & PB")  %>%
-  print()
-##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)
-
-```
-
-```{r}
-p1 <- ss_CB_SB_PB %>%
-  select(a, SO, SR, O, P) %>%
-  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
-  ggplot() +
-  geom_path(aes(x = a, y = log10(Quantity), col = Species))
-print(p1)
-```
-
-# Repeat the results above using the functions that allow simulation of multiple strains per functional group
-
-## Test with one strain per functional group
+## One strain per functional group
 
@@
 
-### Test with N strains per functional group
+## Three strains per functional group
 

### 2021-04-14 R/microxanox.R

@@ -1,0 +1,25 @@
+#' R/microxanox: Microbial oxic and anoxic ecosystem simulations
+#'
+#' Ecosystems containing microbes can be oxic (oxyen present) or
+#' anoxic (oxygen absent). These states can be both stable even with
+#' identical biotic and abiotic circumstances, i.e. they can be
+#' alternate stable states, implying history alone can determine the
+#' present state. The biological reason is that the microbes can alter
+#' the environmental conditions in a way that favour themselves, while
+#' making them less suitable for other organisms. This creates mutual
+#' inhibition. Which state occurs also depends on numerous
+#' abiotic and biotic factors. This package facilitates simulation studies
+#' of the influence of several of these factors.
+#'
+#' Vignettes include:
+#'
+#' - A user guide (maybe).
+#' - A reproduction of some results of Bush et al 2017, on which the simulation is based. It shows the alternate stable states. It also extends the original study by adding temporal forcing and therefore switching between states, and also effects of biotic composition on the response to environmental change.
+#' - A study of the effect of weaker or strong inbihibtion on the response to environmental change, include the extent of environmental conditions for which alternate stable states are present.
+#' - A study of the effects of within functional group diversity on the response to environmental change.
+#'
+#' 
+#' @name microxanox-package
+#' @aliases microxanox
+#' @docType package
+NULL

### 2021-04-14 various_stuff/dumping_ground/dumping_ground1.R

@@ -1,0 +1,315 @@
+## Code test
+
+```{r eval = FALSE}
+dynamic_model = bushplus_dynamic_model
+event_definition = default_event_definition
+parameter_values = default_parameter_values
+event_interval = default_event_interval
+noise_sigma = default_noise_sigma
+minimum_abundances = default_minimum_abundances
+sim_duration = default_sim_duration
+sim_sample_interval = default_sim_sample_interval
+log10a_series = default_log10a_series
+initial_state = default_initial_state
+solver_method = "radau"
+
+times <- seq(0,
+             sim_duration,
+             by = sim_sample_interval)
+
+event_times <- seq(min(times),
+                   max(times),
+                   by = event_interval)
+
+log10a_forcing <- matrix(ncol=2,
+                         byrow=F,
+                         data=c(ceiling(seq(0,
+                                            max(times),
+                                            length=length(log10a_series))),
+                                log10a_series))
+
+
+log10a_forcing <- approxfun(x = log10a_forcing[,1],
+                            y = log10a_forcing[,2],
+                            method = "linear",
+                            rule = 2)
+assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
+
+out <- as.data.frame(ode(y = initial_state,
+                         times = times,
+                         func = dynamic_model,
+                         parms = parameter_values,
+                         method = solver_method,
+                         events = list(func=event_definition,
+                                       time=event_times)))
+
+bushplus_dynamic_model(1, initial_state, default_parameter_values)
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+                          event_definition = default_event_definition,
+                          parameter_values = default_parameter_values,
+                          event_interval = default_event_interval,
+                          noise_sigma = default_noise_sigma,
+                          minimum_abundances = default_minimum_abundances,
+                          sim_duration = default_sim_duration,
[…]

### 2021-04-15 various_stuff/dev_vignettes/User-guide.Rmd
partial reproduction vignette code and data added

--- 2021-04-14 various_stuff/dev_vignettes/User manual.Rmd
@@
 ---
-title: "User manual -- microxanox R Package for microbial oxic and anoxic ecosystem simulations"
+title: "User guide -- microxanox R Package for microbial oxic and anoxic ecosystem simulations"
 author: "Owen Petchey"
@@
 vignette: >
-  %\VignetteIndexEntry{User manual -- microxanox R Package for microbial oxic and anoxic ecosystem simulations}
+  %\VignetteIndexEntry{User guide -- microxanox R Package for microbial oxic and anoxic ecosystem simulations}
   %\VignetteEngine{knitr::rmarkdown}

### 2021-04-16 R/plot_dynamics.R
Some errors in vignette and code

@@ -13,7 +13,7 @@
 plot_dynamics <- function(simulation_result, every_n = 1) {
   
   simulation_result$result %>%
-    filter(row_number() %% every_n == 0) %>%
+    dplyr::filter(row_number() %% every_n == 0) %>%
     mutate(a = 10^a) %>%
     gather(species, quantity, 2:ncol(.)) %>% 
     mutate(var_type=ifelse(str_sub(species, 1, 1)=="N", "Organism", "Substrate"), 

### 2021-04-16 various_stuff/dev_vignettes/within-function-group-diversity.Rmd
various

@@
 ```{r}
+default_parameter_values <- return_default_parameter_values()
 default_log10a_series <- c(log10(default_parameter_values["a_O"]),
@@
 
-simulation_result <- run_simulation(dynamic_model =
-                                      bushplus_dynamic_model_strains,
+simulation_result <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
                                     parameter_values = one_strain_params,
-                                    initial_state = one_strain_initial_state)
+                                    initial_state = one_strain_initial_state,
+                                    event_definition = default_event_definition_strains)
 
@@
                                     SB_strain_params,
-                                    all_other_parms)
+                                    all_other_params)
 
 multi_strain_initial_state <- c(
-  CB_1 = 5e1,
-  CB_2 = 0,
+  CB_1 = 5e3,
+  CB_2 = 1e3,
   CB_3 = 0,
@@
                parameter_values = multi_strain_params,
-               initial_state = multi_strain_initial_state)
+               initial_state = multi_strain_initial_state,
+               event_definition = default_event_definition_strains)
+
 plot_dynamics_strains(simulation_result)

### 2021-04-16 various_stuff/dev_vignettes/within-function-group-diversity.Rmd
fixed the n() error
by dplyr::filter , rather than just filter !!!!!!!!

@@
 rm(list=ls())
-knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = FALSE,
+                      fig.align = "centre", fig.width = 7, fig.height = 7)
 
@@
                                     event_definition = default_event_definition_strains)
-
+every_n <- 10
 plot_dynamics_strains(simulation_result)

### 2021-04-16 R/plot_dynamics_strains.R

@@ -11,11 +11,9 @@
 #' 
 #' @export
 plot_dynamics_strains <- function(simulation_result, every_n = 1) {
-  
-  row_nums <- 1:nrow(simulation_result$result)
-  
+
   simulation_result$result %>%
-    filter(row_nums %% every_n == 0) %>%
+    dplyr::filter(row_number() %% every_n == 0 ) %>%
     mutate(a = 10^a) %>%
     gather(species, quantity, 2:ncol(.)) %>% 
     mutate(var_type=ifelse(grepl("B_", species), "Organism", "Substrate"), 

### 2021-04-16 various_stuff/dev_vignettes/User-guide.Rmd
changed framework
defaults now only refer to the values of the run_simultion arguments used if none are given in the function call

@@
 
+```{r setup, echo = FALSE, output = FALSE, message = FALSE}
+
+knitr::opts_chunk$set(cache = FALSE,
+                      fig.align = "centre", fig.width = 7, fig.height = 7)
+
+library(tidyverse)
+library(deSolve)
+library(rootSolve)
+library(here)
+library(microxanox)
+library(patchwork)
+```
+
+```{r echo = TRUE}
+library(microxanox)
+```
+
 # Introduction
 
-The aims here are:
+The aims of the `microxanox` package are:
 
-* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-* Add an investigation of how changes in tolerances alter the response to environmental change.
+* Facilitate reproduction of the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
+* Include new functionality that allows new research questions to be answered.
 
-We also add:
+We added functionality for:
 
-* Option to include random noise in substrate concentrations.
-* Option to temporally vary oxygen diffusivity.
-* Option to include immigration.
-* Option to include a minimum abundances.
+* adding random noise in substrate concentrations.
+* adding temporally vary oxygen diffusivity.
+* including immigration.
+* setting minimum population abundances
+* changing organismal tolerances to environmental conditions.
+* Changing the number of strains per functional group.
 
@@
 That is, there is only one parameter set per functional group.
+Please see the vignette for multi-strain simulations for information about implementing those.
 
-Here is the original publication:
+Here (again) is the original publication:
 https://www.nature.com/articles/s41467-017-00912-x
@@
 
-Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text.
+Note that om this package we use the notation from the table S1 of the supplementary information, and not the notation in the main text.
 (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
-
-# Methods
-
-## Modelling framework
-
-The framework used for running a simulation, and for varying aspects of a simulation run is:
-
-* Define the default simulation conditions.
-* Use these defaults unless other conditions are specified.
-* The function call `run_simulation()` would therefore run a simulation with default conditions.
-* Whereas, for example, the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values.
-
-## Dynamic model
-
-There are two versions...
-
-The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model_strains()`.
-This is an augmented version that can handle multiple strains within each of the three functional groups.
-The basic version that has only one strain per functional group is `bushplus_dynamic_model()`.
 
@@
 * Organism quantity: cells
+
+# Modelling framework
+
+The framework used for running a simulation, and for varying aspects of a simulation run is:
+
+* Define the default simulation conditions.
+* Use these defaults unless other conditions are specified.
+* The function call `run_simulation()` would therefore run a simulation with default conditions.
+* The function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values.
+
+# (Default) conditions that must be defined
+
+Conditions that must be set, with an example of setting them:
+
+## The dynamic model
+
+The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`.
+An augmented version `bushplus_dynamic_model_strains()` can handle multiple strains within each of the three functional groups.
+The multiple strain version can be run with one strain per functional group, and then gives the same results as the basic version.
+The multiple strain versions is, however, quite a bit slower, even for one strain.
+
+Here we define the default dynamic model to be the basic, one strain per functional group model
+
+```{r}
+default_dynamic_model <- bushplus_dynamic_model
+```
+
+## The parameter values
+
+The parameter values:
+
+```{r}
+default_parameter_values <- return_bushetal2017_parameter_values()
+```
+
+## The initial values of state variables:
+
+```{r}
+default_initial_state <- return_anoxic_1strain_initial_state()
+```
+
+## The duration of the simulation:
+
+```{r}
+default_sim_duration <- 2000
+```
+
+## The sampling interval:
+
+```{r}
+default_sim_sample_interval <- 1
+```
+
+## The oxygen diffusivity dynamics
+
+Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.
+
+```{r}
+default_log10a_series <- c(log10(default_parameter_values["a_O"]),
+                           log10(default_parameter_values["a_O"]))
+```
+
+## Events
+
+These are "things" that happen during the simulation. One has to specify what happens, and when it happens.
+
+#### What happens
+
+Default is to ensure abundances are not below minimum, and to add the amount of noise specified below (none in this vignette).
+
+At present there are two functions available, one for single strain simulations, and one for multi-strain simulations.
+Both can add noise to a state variable, and/or assure minimum abundance is not exceeded.
+
+```{r}
+default_event_definition <- event_definition_1
+```
+
+#### Default event interval
+
+Here set to an event every 10 time units.
+
+```{r}
+default_event_interval <- 10
+```
+
+#### Default noise event options
+
+Default to no noise added to substrate concentrations:
+
+```{r}
+default_noise_sigma <- 0
+```
+
+#### Default minimum abundance event
+
+Default to following minimum abundances:
+
+```{r}
+default_minimum_abundances <- rep(1e2, 3)
+names(default_minimum_abundances) <- c("CB", "PB", "SB") 
+```
+
+# Warning: Assignment to global environment
+
+Please note that the following objects are created during simulations and are assign to the global environment:
+
+* `log10a_forcing_func`
+* `noise_sigma`
+* `minimum_abundances`
+
+# Run a simulation
+
+Once all the conditions are assigned to the default objects, then the `run_simulation` function will pick them up without being explicitly given them:
+
+```{r}
+default_sim_results <- run_simulation()
+```
+
+# Plot the results of a simulation
+
+```{r}
+plot_dynamics(default_sim_results)
+```
+
+# Run simulation with different defaults
+
+```{r}
+initial_state_favouring_oxic <- c(N_CB = 1e7,
+           N_PB = 1e5,
+           N_SB = 1e5,
+           SO = 500,
+           SR = 50,
+           O = 30,
+           P = 10)
+
+new_sim_results <- run_simulation(initial_state = initial_state_favouring_oxic)
+plot_dynamics(new_sim_results)
+```
+
+# Appendix
+
+The code below was useful for debugging.
+
+```{r eval = FALSE}
+dynamic_model = bushplus_dynamic_model
+event_definition = default_event_definition
+parameter_values = default_parameter_values
+event_interval = default_event_interval
+noise_sigma = default_noise_sigma
+minimum_abundances = default_minimum_abundances
+sim_duration = default_sim_duration
+sim_sample_interval = default_sim_sample_interval
+log10a_series = default_log10a_series
+initial_state = default_initial_state
+solver_method = "radau"
+
+times <- seq(0,
+             sim_duration,
+             by = sim_sample_interval)
+
+event_times <- seq(min(times),
+                   max(times),
+                   by = event_interval)
+
+log10a_forcing <- matrix(ncol=2,
+                         byrow=F,
+                         data=c(ceiling(seq(0,
+                                            max(times),
+                                            length=length(log10a_series))),
+                                log10a_series))
+
+log10a_forcing <- approxfun(x = log10a_forcing[,1],
+                            y = log10a_forcing[,2],
+                            method = "linear",
+                            rule = 2)
+assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
+
+out <- as.data.frame(ode(y = initial_state,
+                         times = times,
+                         func = dynamic_model,
+                         parms = parameter_values,
+                         method = solver_method,
+                         events = list(func=event_definition,
+                                       time=event_times)))
+
+bushplus_dynamic_model(1, initial_state, default_parameter_values)
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
+                           event_definition = default_event_definition,
+                           parameter_values = default_parameter_values,
+                           event_interval = default_event_interval,
+                           noise_sigma = default_noise_sigma,
+                           minimum_abundances = default_minimum_abundances,
+                           sim_duration = default_sim_duration,
+                           sim_sample_interval = default_sim_sample_interval,
+                           log10a_series = default_log10a_series,
+                           initial_state = default_initial_state,
+                           solver_method = "radau")
+
+sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)
+
+plot_dynamics(sim_res)
+
+```

### 2021-04-16 R/return_anoxic_1strain_initial_state.R

@@ -5,7 +5,7 @@
 #' @return A named vector of parameter values 
 #' @export
 #'
-return_default_initial_state <- function() {
+return_anoxic_1strain_initial_state <- function() {
     
     default_initial_state <- c(N_CB = 5e1,
                                N_PB = 1e7,

### 2021-04-20 R/bushplus_dynamic_model.R
Running but wrong results

@@ -7,40 +7,40 @@
 #' @export
 bushplus_dynamic_model <- function(t, state, parameters) {
 
-  with(as.list(c(state, parameters)),
+  with(parameters,
        {
 
          # rates of change
-         N_CB_growth_rate <- growth1(P, g_max_CB, k_CB_P) * inhibition(SR, h_SR_CB) * N_CB
+         N_CB_growth_rate <- growth1(state[["P"]], CB$g_max_CB, CB$k_CB_P) * inhibition(state[["SR"]], CB$h_SR_CB) * state[["N_CB"]]
          N_CB_mortality_rate <- m_CB * N_CB
          N_CB_rate <- N_CB_growth_rate - N_CB_mortality_rate + i_CB
 
-         N_PB_growth_rate <- growth2(P, SR, g_max_PB, k_PB_P, k_PB_SR) * inhibition(O, h_O_PB) * N_PB
+         N_PB_growth_rate <- growth2(state[["P"]], state[["SR"]], g_max_PB, k_PB_P, k_PB_SR) * inhibition(state[["O"]], h_O_PB) * state[["N_PB"]]
          N_PB_mortality_rate <- m_PB * N_PB
          N_PB_rate <- N_PB_growth_rate - N_PB_mortality_rate + i_PB
 
-         N_SB_growth_rate <- growth2(P, SO, g_max_SB, k_SB_P, k_SB_SO) * inhibition(O, h_O_SB) * N_SB
+         N_SB_growth_rate <- growth2(state[["P"]], state[["SO"]], g_max_SB, k_SB_P, k_SB_SO) * inhibition(state[["O"]], h_O_SB) * state[["N_SB"]]
          N_SB_mortality_rate <- m_SB * N_SB
          N_SB_rate <- N_SB_growth_rate - N_SB_mortality_rate + i_SB
 
          SO_rate <- 1 / Y_SR_PB * N_PB_growth_rate -
            1 / Y_SO_SB * N_SB_growth_rate +
-           c * O * SR +
-           a_S * (back_SO - SO)
+           c * state[["O"]] * state[["SR"]] +
+           a_S * (back_SO - state[["SO"]])
 
          SR_rate <- - 1 / Y_SR_PB * N_PB_growth_rate +
            1 / Y_SO_SB * N_SB_growth_rate -
-           c * O * SR +
-           a_S * (back_SR - SR)
+           c * state[["O"]] * state[["SR"]] +
+           a_S * (back_SR - state[["SR"]])
 
[…]

### 2021-04-20 R/bushplus_dynamic_model_strains.R

@@ -8,11 +8,7 @@
 #' @return Growth rate
 #' @export
 bushplus_dynamic_model_strains <- function(t, state, parameters) {
-  
-  with(as.list(c(state, parameters)),
-       {
-         
-         
+
          CB <- state[grep("CB", names(state))]
          names_CB <- names(CB)[order(names(CB))]
          CB <- as.numeric(CB[order(names(CB))])
@@ -22,95 +18,97 @@ bushplus_dynamic_model_strains <- function(t, state, parameters) {
          SB <- state[grep("SB", names(state))]
          names_SB <- names(SB)[order(names(SB))]
          SB <- as.numeric(SB[order(names(SB))])
-         #print(c(CB, PB, SB))
+         # print(c(CB, PB, SB))
          
+         ### CB
+         # g_max_CB <- parameters[grep("g_max_CB", names(parameters))]
+         # g_max_CB <- as.numeric(g_max_CB[order(names(g_max_CB))])
+         # k_CB_P <- parameters[grep("k_CB_P", names(parameters))]
+         # k_CB_P <- as.numeric(k_CB_P[order(names(k_CB_P))])
+         # h_SR_CB <- parameters[grep("h_SR_CB", names(parameters))]
+         # h_SR_CB <- as.numeric(h_SR_CB[order(names(h_SR_CB))])
+         # y_P_CB <- parameters[grep("y_P_CB", names(parameters))]
+         # y_P_CB <- as.numeric(y_P_CB[order(names(y_P_CB))])
+         # Pr_CB <- parameters[grep("Pr_CB", names(parameters))]
+         # Pr_CB <- as.numeric(Pr_CB[order(names(Pr_CB))])
+         # m_CB <- parameters[grep("m_CB", names(parameters))]
+         # m_CB <- as.numeric(m_CB[order(names(m_CB))])
+         # i_CB <- parameters[grep("i_CB", names(parameters))]
+         # i_CB <- as.numeric(i_CB[order(names(i_CB))])
          
+         ### PB
+         # g_max_PB <- parameters[grep("g_max_PB", names(parameters))]
+         # g_max_PB <- as.numeric(g_max_PB[order(names(g_max_PB))])
+         # k_PB_SR <- parameters[grep("k_PB_SR", names(parameters))]
+         # k_PB_SR <- as.numeric(k_PB_SR[order(names(k_PB_SR))])
[…]

### 2021-04-20 R/new_strain_parameter.R

@@ -1,0 +1,111 @@
+#' Creates empty parameter
+#'
+#' @param CB_strain_parameter 
+#' @param PB_strain_parameter 
+#' @param SB_strain_parameter 
+#' @param a_S 
+#' @param a_O 
+#' @param a_P 
+#' @param back_SR 
+#' @param back_SO 
+#' @param back_O 
+#' @param back_P 
+#' @param c 
+#'
+#' @return
+#' @export
+#'
+#' @examples
+new_strain_parameter <- function(
+  ## strain parameter
+  CB_strain_parameter = new_CB_strain_parameter(),
+  PB_strain_parameter = new_PB_strain_parameter(),
+  SB_strain_parameter = new_SB_strain_parameter,
+  ## substrate diffusivity
+  a_S = 0.001,
+  a_O = 8e-4,
+  a_P = 0.01,
+  ## background substrate concentration
+  back_SR = 300,
+  back_SO = 300,
+  back_O = 300,
+  back_P = 9.5,
+  ## oxidisation rate of reduced sulphur
+  c = 4e-5
+){
+  parms <- list()
+
+# strain parameter --------------------------------------------------------
+
+  parms$CB <- CB_strain_parameter
+  parms$PB <- PB_strain_parameter
+  parms$SB <- SB_strain_parameter
+  
+  # other parameter ---------------------------------------------------------
+
+  ## substrate diffusivity
+  parms$a_S = as.numeric(NA)
+  parms$a_O = as.numeric(NA)
+  parms$a_P = as.numeric(NA)
+  
+  ## background substrate concentration
+  parms$back_SR = as.numeric(NA)
+  parms$back_SO = as.numeric(NA)
+  parms$back_O = as.numeric(NA)
+  parms$back_P = as.numeric(NA)
+  
+  ## oxidisation rate of reduced sulphur
+  parms$c = as.numeric(NA)
+  
+  class(parms) <- append("strain_parameter", class(parms))
+  
+  return(parms)
+}
+
+
+new_CB_strain_parameter <- function(){
+  result <- data.frame(
+    strain_name = c("CB_1"),
+    g_max_CB = as.numeric(NA),
+    k_CB_P = as.numeric(NA),
+    h_SR_CB = as.numeric(NA),
+    y_P_CB = as.numeric(NA),
+    Pr_CB = as.numeric(NA),
+    m_CB = as.numeric(NA),
[…]

### 2021-04-21 R/new_strain_parameter.R
Combines parameter and initial state in one variable
Vignette within-function-group-diversity.Rmd adapted

The others still need to be checked

@@ -1,13 +1,14 @@
 #' Creates empty parameter
 #' 
 #' The default is to create the parameter set used in Bush et al. (2017)
-#' @param n_CB number of strains 
-#' @param values values_CB to be used for CB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param n_PB number of strains 
-#' @param values values_PB to be used for PB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param n_SB number of strains 
-#' @param values values_SB to be used for SB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
+#' @param n_CB number of CB strains 
+#' @param values_CB to be used for CB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
+#' @param n_PB number of PB strains 
+#' @param values_PB to be used for PB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
+#' @param n_SB number of SB strains 
+#' @param values_SB to be used for SB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
 #' @param values_other values to be used for other parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
+#' @param values_initial_state values to be used for initial values or \code{"bush"}, in which case the default from Bush et al (2017) will be used
 #'
 #' @return \code{list} with additional class \code{strain_parameter}
 #' @export
@@ -19,14 +20,23 @@ new_strain_parameter <- function(
   values_PB = "bush",
   n_SB = 1,
   values_SB = "bush",
-  other = "bush"
+  values_other = "bush",
+  values_initial_state = "bush"
 ){
-  if (is.na(other)) {
-    other <- "NA"
+  if (is.na(values_other)) {
+    values_other <- "NA"
   }
   
-  if (!(other %in% c("NA", "bush"))) {
-    stop("Not supported value for `other`!\n", "Only NA, 'NA' and 'bush' supported!")
[…]

### 2021-04-23 R/get_final_states_a_N.R
partial-reproduction-Bushetal2017 works mainly.

@@ -12,11 +12,11 @@ get_final_states_a_N <- function(x) {
   ssfind_parameters["a_O"] <- x["a_O"]      
   
   ## update forcing series
-  ss_log10a_series <- c(log10(ssfind_parameters["a_O"]),
-                        log10(ssfind_parameters["a_O"]))
+  ss_log10a_series <- c(log10(ssfind_parameters[["a_O"]]),
+                        log10(ssfind_parameters[["a_O"]]))
   
-  simres <- run_simulation(dynamic_model = bushplus_dynamic_model,
-    initial_state = ssfind_init_state,
+  simres <- run_simulation(dynamic_model = bushplus_dynamic_model_strains,
+                           initial_state = ssfind_init_state,
                            parameter_values = ssfind_parameters,
                            sim_duration = ssfind_simulation_duration,
                            sim_sample_interval = ssfind_simulation_sampling_interval,

### 2021-04-30 R/bushplus_dynamic_model_strains.R
basic documentation done

@@ -1,7 +1,7 @@
 
 
-#' The rate equations, as published in the paper, but with forcing of oxygen diffusivity (a_0) potential added.
-#' And multiple strains
+#' The rate equations, as published in the paper, but with forcing of oxygen diffusivity \code{"a_0"} potential added,
+#' and the possibility to simulate multiple strains per functional group
 #'
 #' @param t The current time in the simulation
 #' @param state A vector containing the current (named) values of each state variable

### 2021-05-01 R/initialisation_functions.R
vignettes done for now

@@ -14,7 +14,7 @@
 #' @md
 #' @export
 #'
-new_strain_starter <- function(
+new_starter <- function(
   n_CB = 1,
   values_CB = "bush",
   n_PB = 1,

### 2021-05-05 R/get_final_states_a_N.R
multiple improvements

@@ -5,11 +5,12 @@
 #' @param x A vector of oxygen diffusivity and initial conditions.
 #' @return A vector of steady states
 #' @export
-get_final_states_a_N <- function(x) { 
+get_final_states_a_N <- function(x, ssfind_parameters) { 
   
   #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
   
   ssfind_parameters$a_O <- x["a_O"]
+  print(ssfind_parameters$a_O)
   
   CBs <- unlist(rep(x["N_CB"] / length(grep("CB", names(ssfind_parameters$initial_state))),
       length(grep("CB", names(ssfind_parameters$initial_state)))))

### 2021-05-05 R/plot_dynamics.R

@@ -1,7 +1,7 @@
 #' Convenience function to plot the dynamics of a model run,
-#' with strains within functional groups
+#' with strains within functional groups.
 #'
-#' @param simulation_result Object returned by the run_simulation function.
+#' @param simulation_result Object returned by the run_simulation function
 #' @param every_n Plot data of every other n sample.
 #' @return Nothing.
 #' 
@@ -11,32 +11,60 @@
 #' 
 #' @export
 plot_dynamics <- function(simulation_result, every_n = 1) {
-
-  simulation_result$result %>%
+  
+  colfunc_CB <- colorRampPalette(c("#024F17", "#B5FFC9"))
+  colfunc_SB <- colorRampPalette(c("#7D1402", "#FCBEB3"))
+  colfunc_PB <- colorRampPalette(c("#6E0172", "#F9AEFC"))
+  
+  temp <- simulation_result$result %>%
     dplyr::filter(row_number() %% every_n == 0 ) %>%
     mutate(a = 10^a) %>%
     gather(species, quantity, 2:ncol(.)) %>% 
-    mutate(var_type=ifelse(grepl("B_", species), "Organism", "Substrate"), 
-           trans_quantity=log10(quantity)) %>% 
-    ggplot(aes(x=time, y=trans_quantity, col=species)) +
-    facet_wrap(~var_type, scales = "free", ncol = 1) + 
+    mutate(var_type=ifelse(grepl("B_", species), "Organism", "Substrate"),
+           functional_group = case_when(str_detect(species, "CB_") ~ "CB",
+                                        str_detect(species, "SB_") ~ "SB",
+                                        str_detect(species, "PB_") ~ "PB"),
+           log10_quantity=log10(quantity))
+  
+  num_CB_strains <- nrow(simulation_result$parameter_values$CB)
+  num_SB_strains <- nrow(simulation_result$parameter_values$SB)
+  num_PB_strains <- nrow(simulation_result$parameter_values$PB)
+  
+  p1 <- temp %>%
+    dplyr::filter(functional_group == "CB") %>%
+    ggplot(aes(x=time, y=log10_quantity, col=species)) +
     geom_line() +
-    ylab('log(quantity)') +
+    ylab('log10(quantity [cells])') +
     xlab('time [hours]') +
-    labs(color = "Organism/Substrate")
[…]

### 2021-05-05 various_stuff/reading/exponential_growth_example/Run_EG_GEM_intro.m

@@ -1,60 +1,0 @@
-clear; clc; clf; % clear everything
-tic; % start a time counter
-
-%% pre GEM info
-
-% parameters and replicates
-    b = 1; % birth rate
-    d = 0.5; % death rate
-    num_replicates = 50; % number of times to run GEM
-
-% evolution and solution details
-    h_2 = 1; % narrow-sense heritability
-    cv = 0.3; % coefficient of variation of trait
-    y0 = 5; % starting population size
-    t_max = 6; % time span to run simulations    
-    tspan = [0 t_max]; % start and end times
-
-%% set up and run GEM 
-
-[stand_times, R_data_out, x_data_out, x_var_data_out] = GEMv2_exponential_growth(b, d, cv, h_2, num_replicates, y0, t_max);
-
-%% Solve non-evolutionary ODE for realized initial conditions
-
-realized_initial_b = x_data_out(1,1); % get the average of the realized starting b's
-realized_initial_b_var = x_var_data_out(1,1); % get the average of the realized variances of b
-
-ode = @(t,y) EG_model(t,y,realized_initial_b,d); % compile function and call
-    [t1,y1] = ode45(ode, tspan, y0); % return time and population density vectors
-
-%% plotting
-figure(1); % plot medians and ci's overtop individual lines
-    subplot(3,1,1); box on;
-        jbfill(stand_times,R_data_out(2,:),R_data_out(3,:),[0.5 0.5 0.5],'w',1,0.2); 
-
-% X=[stand_times,fliplr(stand_times)];                %#create continuous x value array for plotting
-% Y=[R_data_out(2,:),fliplr(R_data_out(3,:))];              %#create y values for out and then back
-% fill(X,Y,[0.5 0.5 0.5]); alpha 0.2
-
-        hold on;
-        plot(stand_times,R_data_out(1,:),'-','Color','b','LineWidth',2);
-        plot(t1,y1(:,1),'-k','LineWidth',2);
-        xlim([0 t_max]);
-        ylabel('Population size','FontSize',12);     
-    subplot(3,1,2); box on;
-        jbfill(stand_times,x_data_out(2,:),x_data_out(3,:),[0.5 0.5 0.5],'w',1,0.2); 
-        hold on;
-        plot(stand_times,x_data_out(1,:),'-','Color','b','LineWidth',2);
-        plot([0 t_max],[realized_initial_b realized_initial_b],'--k');
[…]

### 2021-05-13 R/initialisation_functions.R
some small edits

@@ -1,14 +1,14 @@
-#' Creates empty startter values for the simulation
+#' Creates a set of parameters and starting conditions for a simulation
 #' 
-#' The default is to create the parameter set used in Bush et al. (2017) and anoxic initial conditions, 
-#' @param n_CB number of CB strains 
-#' @param values_CB to be used for CB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param n_PB number of PB strains 
-#' @param values_PB to be used for PB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param n_SB number of SB strains 
-#' @param values_SB to be used for SB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param values_other values to be used for other parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used
-#' @param values_initial_state values to be used for initial values or \code{"bush_anoxic"} or \code{"bush_oxic"}, in which case the default from Bush et al (2017) will be used
+#' The default is to create the parameter set used in Bush et al. (2017) and the anoxic favourable initial conditions used in the simulations for Figure 2a&b of that publication. 
+#' @param n_CB number of CB strains.
+#' @param values_CB to be used for CB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used.
+#' @param n_PB number of PB strains. 
+#' @param values_PB to be used for PB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used.
+#' @param n_SB number of SB strains. 
+#' @param values_SB to be used for SB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used.
+#' @param values_other values to be used for other parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used.
+#' @param values_initial_state values to be used for initial values or \code{"bush_anoxic"} or \code{"bush_oxic"}, in which case the default from Bush et al (2017) will be used.
 #'
 #' @return \code{list} with additional class \code{strain_parameter}
 #' @md

### 2021-06-09 R/bushplus_dynamic_model.R
r files have reasonable documentation, user guide ok too

@@ -1,12 +1,12 @@
 
 
-#' The rate equations, as published in the paper, but with forcing of oxygen diffusivity \code{"a_0"} potential added,
+#' The rate equations, as published in the Bush et al 2017 paper, but with forcing of oxygen diffusivity \code{"a_0"} potential added,
 #' and the possibility to simulate multiple strains per functional group
 #'
 #' @param t The current time in the simulation
 #' @param state A vector containing the current (named) values of each state variable
-#' @param parameters A vector containing the current (named) values of each parameter
-#' @return Growth rate
+#' @param parameters A list containing the parameter values
+#' @return A list containing the rate of change of each state variable, and also the current values of oxygen diffusivity.
 #' @export
 bushplus_dynamic_model <- function(t, state, parameters) {
 
@@ -21,57 +21,7 @@ bushplus_dynamic_model <- function(t, state, parameters) {
          SB <- as.numeric(SB[order(names(SB))])
          # print(c(CB, PB, SB))
          
-         ### CB
-         # g_max_CB <- parameters[grep("g_max_CB", names(parameters))]
-         # g_max_CB <- as.numeric(g_max_CB[order(names(g_max_CB))])
-         # k_CB_P <- parameters[grep("k_CB_P", names(parameters))]
-         # k_CB_P <- as.numeric(k_CB_P[order(names(k_CB_P))])
-         # h_SR_CB <- parameters[grep("h_SR_CB", names(parameters))]
-         # h_SR_CB <- as.numeric(h_SR_CB[order(names(h_SR_CB))])
-         # y_P_CB <- parameters[grep("y_P_CB", names(parameters))]
-         # y_P_CB <- as.numeric(y_P_CB[order(names(y_P_CB))])
-         # Pr_CB <- parameters[grep("Pr_CB", names(parameters))]
-         # Pr_CB <- as.numeric(Pr_CB[order(names(Pr_CB))])
-         # m_CB <- parameters[grep("m_CB", names(parameters))]
-         # m_CB <- as.numeric(m_CB[order(names(m_CB))])
-         # i_CB <- parameters[grep("i_CB", names(parameters))]
-         # i_CB <- as.numeric(i_CB[order(names(i_CB))])
-         
-         ### PB
[…]

### 2021-06-09 R/get_nonlinearity.R

@@ -1,5 +1,5 @@
 #' Get the amount of non-linearity in a state-environment relationship.
-#' Emancipator & Knoll (1993) Clinical Chemistry 39, 766-772
+#' Method described in Emancipator & Knoll (1993) Clinical Chemistry 39, 766-772
 #'
 #' @param x Environmental driver state variable
 #' @param y Ecosystem state variable

### 2021-06-09 R/initialisation_functions.R

@@ -1,5 +1,5 @@
 #' Creates a set of parameters and starting conditions for a simulation
-#' 
+#' A function to assist with the initialisation of a simulation, by providing various reference sets of parameter values and initial conditions.
 #' The default is to create the parameter set used in Bush et al. (2017) and the anoxic favourable initial conditions used in the simulations for Figure 2a&b of that publication. 
 #' @param n_CB number of CB strains.
 #' @param values_CB to be used for CB strains parameter or \code{"bush"}, in which case the default from Bush et al (2017) will be used.

### 2021-06-09 R/microxanox-package.R

@@ -9,14 +9,12 @@
 #' making them less suitable for other organisms. This creates mutual
 #' inhibition. Which state occurs also depends on numerous
 #' abiotic and biotic factors. This package facilitates simulation studies
-#' of the influence of several of these factors.
+#' of the influence of several of these factors, and is based on work described in (Bush et al) 2017 (DOI: 10.1038/s41467-017-00912-x)
 #'
 #' Vignettes include:
 #'
-#' - A user guide (maybe).
+#' - A user guide.
 #' - A reproduction of some results of Bush et al 2017, on which the simulation is based. It shows the alternate stable states. It also extends the original study by adding temporal forcing and therefore switching between states, and also effects of biotic composition on the response to environmental change.
-#' - A study of the effect of weaker or strong inbihibtion on the response to environmental change, include the extent of environmental conditions for which alternate stable states are present.
-#' - A study of the effects of within functional group diversity on the response to environmental change.
 #'
 #' 
 #' @name microxanox-package

### 2021-06-09 R/ss_by_a_N.R

@@ -3,7 +3,7 @@
 #' @param ss_expt A data frame of oxygen diffusivities and initial conditions combinations,
 #' for example created with \code{"expand.grid"}.
 #' @param param A set of parameter values and initial states
-#' @return Experimental result.
+#' @return Processed data about steady states
 #' @export
 
 ss_by_a_N <- function(ss_expt, param)

### 2021-06-21 R/get_stability_measures.R
working on the vignettes

@@ -15,8 +15,7 @@ get_stability_measures <- function(ss_object) {
   these <- c(these, which(names(ss_object) %in% c("SO", "SR", "O", "P")))
   temp <- ss_object %>%
     #rbind(ss_object) %>%
-    mutate(direction = c(rep("up", nrow(ss_object)/2),
-                         rep("down", nrow(ss_object)/2))) %>%
+    mutate(direction = ifelse(initial_N_CB == 1, "up", "down")) %>%
     tidyr::gather(key = "Species", value = Quantity, these) %>%
     dplyr::select(-initial_N_CB) %>%
     tidyr::spread(key = direction, value=Quantity, drop=T)
@@ -24,12 +23,13 @@ get_stability_measures <- function(ss_object) {
   res <- temp %>%
     dplyr::group_by(Species) %>%
     dplyr::summarise(hyst_tot = get_hysteresis_total(log10(up+1), log10(down+1)),
-              hyst_range = get_hysteresis_range(log10(up+1), log10(down+1), a),
-              nl_up = get_nonlinearity(a, log10(up+1)),
-              nl_down = get_nonlinearity(a, log10(down+1))
+                     hyst_range = get_hysteresis_range(log10(up+1), log10(down+1), a),
+                     hyst_min = get_hysteresis_min(log10(up+1), log10(down+1), a),
+                     hyst_max = get_hysteresis_max(log10(up+1), log10(down+1), a),
+                     nl_up = get_nonlinearity(a, log10(up+1)),
+                     nl_down = get_nonlinearity(a, log10(down+1))
     )
   res
-  
   
 }
 
@@ -57,18 +57,90 @@ get_hysteresis_total <- function(up, down) {
 #' @param a An environmental driver, here it is usually oxygen diffusivity
 #' @return A numeric value, which is the extent of the range of conditions for which alternate stable states exist.
 #' @export
+get_hysteresis_min <- function(up, down, a)
+{
+  
+  temp1 <- abs(up - down) > 0.1
+  
+  if(sum((up+down)) == 0) {
+    res = NA
+    min_flip = NA
+    max_flip = NA
+  }
+  else(
+    if(sum(temp1)==0) {
+      res <- 0
+      min_flip = 0
+      max_flip = 0
+    }
+      
+    else({
+      res <-  min(a[temp1])
+    }
+    )
+  )
+  res
+}
+
+#' Get the range of environmental conditions for which alternate stable states exist
+#'
+#' @param up State variable values as the environmental condition increases
[…]

### 2021-09-02 R/new_runsim_result.R
In progress

@@ -1,0 +1,26 @@
+new_runsim_results <- function(
+  parameter, result, help = FALSE
+){
+  if (help) {
+    p <- list(
+      dynamic_model = NULL, # default_dynamic_model,
+      event_definition = NULL, # default_event_definition,
+      parameter_values = NULL, # default_parameter_values,
+      event_interval = NULL, # default_event_interval,
+      noise_sigma = NULL, # default_noise_sigma,
+      minimum_abundances = NULL, # default_minimum_abundances,
+      sim_duration = NULL, # default_sim_duration,
+      sim_sample_interval = NULL, # default_sim_sample_interval,
+      log10a_series = NULL, # default_log10a_series,
+      initial_state = NULL, # default_initial_state,
+      solver_method = NULL, # "radau",
+      result = "The result dataframe"
+    )
+  } else {
+    p <- parameter
+    p$result <- result
+    class(p) <- append(class(p), "runsim_result")
+
+  }
+  return(p)
+}

### 2021-09-02 R/ss_by_a_N.R

@@ -1,27 +1,28 @@
 #' Function to get the steady states for combinations of a (oxygen diffusivity) and initial states.
 #'
-#' @param ss_expt A data frame of oxygen diffusivities and initial conditions combinations,
-#' for example created with \code{"expand.grid"}.
-#' @param param A set of parameter values and initial states
+#' @param parameter object of class \code{ss_by_a_N_parameter} as generated by \code{new_ss_by_a_N_parameter()}
 #' @param mc.cores the number of cores to be used. If 0, the old sequential version is used.
 #'   The default is read from the option \code{mc.cores}, i.e. using \code{getOption("mc.cores", 0)}
 #' @return Processed data about steady states
 #' @export
 
-ss_by_a_N <- function(ss_expt, param, mc.cores = getOption("mc.cores", 0)) {
-
+ss_by_a_N <- function(parameter, mc.cores = getOption("mc.cores", 0)) {
+  if (!inherits(parameter, "ss_by_a_N_parameter")) {
+    stop("parameter has to be an object of type `ss_by_a_N_parameter`!")
+  }
+  
   if (Sys.info()['sysname'] == "Window") {
     warning("On windows, mc.cores will always be set to one!")
     mc.cores <- 1
   }
   
   if (mc.cores == 0) {
-    temp_result <- apply(ss_expt, 1, function(x) get_final_states_a_N(x, param))
+    temp_result <- apply(parameter$ss_expt, 1, function(x) get_final_states_a_N(x, parameter))
   } else {
     temp_result <- parallel::mclapply(
-      1:nrow(ss_expt),
+      1:nrow(parameter$ss_expt),
       function(i) {
-        get_final_states_a_N(ss_expt[i, ], param)
+        get_final_states_a_N(parameter$ss_expt[i, ], parameter)
       }, mc.preschedule = FALSE,
       mc.cores = mc.cores
     )
@@ -33,10 +34,10 @@ ss_by_a_N <- function(ss_expt, param, mc.cores = getOption("mc.cores", 0)) {
   result <- temp_result %>%
     tibble::tibble() %>%
     tidyr::unnest(cols = 1) %>%
-    dplyr::mutate(initial_N_CB = ss_expt$N_CB,
-                  initial_N_PB = ss_expt$N_PB,
-                  initial_N_SB = ss_expt$N_SB,
-                  a_O = ss_expt$a_O)
[…]

### 2021-09-06 R/event_definition_1.R
0.2.2: Parameter object implemented for partial-reproduction-Bushetal2017.Rmd vignette

@@ -6,7 +6,7 @@
 #' @param parms Model parameters
 #' @return A vector of the new state variables
 #' @export
-event_definition_1 <- function(times=times, y=state, parms = parameters) {
+event_definition_1 <- function(times=times, y=state, parms = parameters, log10a_forcing_func, noise_sigma, minimum_abundances) {
   with(
     as.list(y),
     {

### 2021-09-06 R/get_final_states_a_N.R

@@ -13,20 +13,20 @@ get_final_states_a_N <- function(x, parameter) {
   parameter$a_O <- x["a_O"]
   
   ## and set all the initial states in the parameter object to the required values
-  CBs <- unlist(rep(x["N_CB"] / length(grep("CB", names(parameter$initial_state))),
-      length(grep("CB", names(parameter$initial_state)))))
+  CBs <- unlist(rep(x["N_CB"] / length(grep("CB", names(parameter$strain_parameter$initial_state))),
+      length(grep("CB", names(parameter$strain_parameter$initial_state)))))
   names(CBs) <- NULL
-  parameter$initial_state[grep("CB", names(parameter$initial_state))] <- CBs
+  parameter$strain_parameter$initial_state[grep("CB", names(parameter$strain_parameter$initial_state))] <- CBs
     
-  PBs <- unlist(rep(x["N_PB"] / length(grep("PB", names(parameter$initial_state))),
-                    length(grep("PB", names(parameter$initial_state)))))
+  PBs <- unlist(rep(x["N_PB"] / length(grep("PB", names(parameter$strain_parameter$initial_state))),
+                    length(grep("PB", names(parameter$strain_parameter$initial_state)))))
   names(PBs) <- NULL
-  parameter$initial_state[grep("PB", names(parameter$initial_state))] <- PBs
+  parameter$strain_parameter$initial_state[grep("PB", names(parameter$strain_parameter$initial_state))] <- PBs
   
-  SBs <- unlist(rep(x["N_SB"] / length(grep("SB", names(parameter$initial_state))),
-                    length(grep("SB", names(parameter$initial_state)))))
+  SBs <- unlist(rep(x["N_SB"] / length(grep("SB", names(parameter$strain_parameter$initial_state))),
+                    length(grep("SB", names(parameter$strain_parameter$initial_state)))))
   names(SBs) <- NULL
-  parameter$initial_state[grep("SB", names(parameter$initial_state))] <- SBs
+  parameter$strain_parameter$initial_state[grep("SB", names(parameter$strain_parameter$initial_state))] <- SBs
   
  ## update forcing series with the required oxygen diffus. value
   parameter$log10a_series <- c(
@@ -34,6 +34,7 @@ get_final_states_a_N <- function(x, parameter) {
     log10(parameter[["a_O"]])
   )
   
+  parameter$initial_state = parameter$strain_parameter$initial_state
   ## run the simulation
   simres <- run_simulation(
[…]

### 2021-09-06 R/new_runsim_parameter.R

@@ -3,31 +3,31 @@ new_runsim_parameter <- function(
 ){
   if (help) {
     p <- list(
-      dynamic_model = NULL, # default_dynamic_model,
-      event_definition = NULL, # default_event_definition,
-      parameter_values = NULL, # default_parameter_values,
-      event_interval = NULL, # default_event_interval,
-      noise_sigma = NULL, # default_noise_sigma,
-      minimum_abundances = NULL, # default_minimum_abundances,
-      sim_duration = NULL, # default_sim_duration,
-      sim_sample_interval = NULL, # default_sim_sample_interval,
-      log10a_series = NULL, # default_log10a_series,
-      initial_state = NULL, # default_initial_state,
-      solver_method = NULL # "radau",
+      dynamic_model = NA, # default_dynamic_model,
+      event_definition = NA, # default_event_definition,
+      strain_parameter = NA, # default_parameter_values,
+      event_interval = NA, # default_event_interval,
+      noise_sigma = NA, # default_noise_sigma,
+      minimum_abundances = NA, # default_minimum_abundances,
+      sim_duration = NA, # default_sim_duration,
+      sim_sample_interval = NA, # default_sim_sample_interval,
+      log10a_series = NA, # default_log10a_series,
+      initial_state = NA, # default_initial_state,
+      solver_method = "radau" # "radau",
     )
   } else {
     p <- list(
-      dynamic_model = NULL, # default_dynamic_model,
-      event_definition = NULL, # default_event_definition,
-      parameter_values = NULL, # default_parameter_values,
-      event_interval = NULL, # default_event_interval,
-      noise_sigma = NULL, # default_noise_sigma,
-      minimum_abundances = NULL, # default_minimum_abundances,
-      sim_duration = NULL, # default_sim_duration,
-      sim_sample_interval = NULL, # default_sim_sample_interval,
-      log10a_series = NULL, # default_log10a_series,
-      initial_state = NULL, # default_initial_state,
-      solver_method = NULL # "radau",
+      dynamic_model = NA, # default_dynamic_model,
+      event_definition = NA, # default_event_definition,
+      strain_parameter = NA, # default_parameter_values,
+      event_interval = NA, # default_event_interval,
+      noise_sigma = NA, # default_noise_sigma,
+      minimum_abundances = NA, # default_minimum_abundances,
+      sim_duration = NA, # default_sim_duration,
+      sim_sample_interval = NA, # default_sim_sample_interval,
[…]

### 2021-09-06 R/new_ssbyaN_result.R

@@ -1,0 +1,19 @@
+new_ss_by_a_N_results <- function(
+  parameter, result, help = FALSE
+){
+  if (help) {
+    p <- list(
+      result = "The result dataframe"
+    )
+  } else {
+    if (!inherits(parameter, "ss_by_a_N_parameter")) {
+      stop("`parameter` has to be of class ss_by_a_N_parameter")
+    }
+    
+    p <- parameter
+    p$result <- result
+    class(p) <- append(class(p), "ss_by_a_N_result")
+
+  }
+  return(p)
+}

### 2021-09-06 R/ss_by_a_N.R

@@ -39,5 +39,5 @@ ss_by_a_N <- function(parameter, mc.cores = getOption("mc.cores", 0)) {
                   initial_N_SB = parameter$ss_expt$N_SB,
                   a_O = parameter$ss_expt$a_O)
 
-   result
+   return(new_ss_by_a_N_results(parameter, result))
 }

### 2021-09-07 R/new_runsim_parameter.R
Removal of initial_state from root of parameter

@@ -22,7 +22,7 @@ new_runsim_parameter <- function(
       sim_duration = NA, # default_sim_duration,
       sim_sample_interval = NA, # default_sim_sample_interval,
       log10a_series = NA, # default_log10a_series,
-      initial_state = NA, # default_initial_state,
+      # initial_state = NA, # default_initial_state,
       solver_method = "radau" # "radau",
     )
   } else {
@@ -36,7 +36,7 @@ new_runsim_parameter <- function(
       sim_duration = NA, # default_sim_duration,
       sim_sample_interval = NA, # default_sim_sample_interval,
       log10a_series = NA, # default_log10a_series,
-      initial_state = NA, # default_initial_state,
+      # initial_state = NA, # default_initial_state,
       solver_method = "radau" # "radau",
     )
     class(p) <- append(class(p), "runsim_parameter")

### 2021-09-07 R/run_simulation.R

@@ -80,7 +80,7 @@ run_simulation <- function(
   
   out <- as.data.frame(
     ode(
-      y = parameter$initial_state,
+      y = parameter$strain_parameter$initial_state,
       times = times,
       func = parameter$dynamic_model,
       parms = parameter$strain_parameter,

### 2021-09-09 R/get_stability_measures.R
Changes in get_stability_measures

@@ -1,36 +1,43 @@
 #' Get various measures of the stability, specifically non-linearity and hysteresis measures, of an ecosystem response to environmental change.
 #' Takes steady state data as the input. 
 #'
-#' @param ss_object An object of the type returned by the ss_by_a_N() function
+#' @param ss_object An object of class \code{ss_by_a_N_result} as returned by the ss_by_a_N() function or
+#'   the \code{result} e;ement of that object, i.e. \code{x$result}.
 #' @return A data frame of stability measures of each state variable
 #' @importFrom tidyr gather spread
 #' @importFrom dplyr summarise
 #' 
 #' @export
 get_stability_measures <- function(ss_object) {
+  if (inherits(ss_object, "ss_by_a_N_result")) {
+    result <- ss_object$result
+  } else {
+    result <- ss_object
+  }
+
   
   ## which initial condition varies?
   init_varying <- NA
-  if(length(unique(ss_object$initial_N_CB)) > 1)
+  if(length(unique(result$initial_N_CB)) > 1)
      init_varying <- "initial_N_CB"
-  if(length(unique(ss_object$initial_N_SB)) > 1)
+  if(length(unique(result$initial_N_SB)) > 1)
     init_varying <- "initial_N_SB"
-  if(length(unique(ss_object$initial_N_PB)) > 1)
+  if(length(unique(result$initial_N_PB)) > 1)
     init_varying <- "initial_N_PB"
   
   if(!is.na(init_varying)) {
   
-  ss_object$init_varying <- pull(ss_object[,init_varying],1)
+  result$init_varying <- pull(result[,init_varying],1)
   
-  min_iniN <- min(ss_object$init_varying)
-  max_iniN <- max(ss_object$init_varying)
+  min_iniN <- min(result$init_varying)
+  max_iniN <- max(result$init_varying)
   
   ## The following is preparing the data
-  these <- grep("B_", names(ss_object))
+  these <- grep("B_", names(result))
   #these <- these[c(-(length(these)-1), -length(these))]
-  these <- c(these, which(names(ss_object) %in% c("SO", "SR", "O", "P")))
-  temp <- ss_object %>%
-    #rbind(ss_object) %>%
+  these <- c(these, which(names(result) %in% c("SO", "SR", "O", "P")))
+  temp <- result %>%
+    #rbind(result) %>%
     mutate(direction = ifelse(init_varying == min_iniN, "up", "down")) %>%
[…]

### 2021-09-16 R/get_stability_measures.R
get_stability_measures: add raw in addition to log and changes required in vignettes.

@@ -49,14 +49,25 @@ get_stability_measures <- function(ss_object) {
   ## then get the stability measures
   res <- temp %>%
     dplyr::group_by(Species) %>%
-    dplyr::summarise(hyst_tot = get_hysteresis_total(log10(up+1), log10(down+1)),
-                     hyst_range = get_hysteresis_range(log10(up+1), log10(down+1), a),
-                     hyst_min = get_hysteresis_min(log10(up+1), log10(down+1), a),
-                     hyst_max = get_hysteresis_max(log10(up+1), log10(down+1), a),
-                     nl_up = get_nonlinearity(a, log10(up+1)),
-                     nl_down = get_nonlinearity(a, log10(down+1))
+    dplyr::summarise(
     )
-  
+ 
+  res <- temp %>%
+    dplyr::group_by(Species) %>%
+    dplyr::summarise(hyst_tot_raw = get_hysteresis_total(up, down),
+                     hyst_range_raw = get_hysteresis_range(up, down, a),
+                     hyst_min_raw = get_hysteresis_min(up, down, a),
+                     hyst_max_raw = get_hysteresis_max(up, down, a),
+                     nl_up_raw = get_nonlinearity(a, up),
+                     nl_down_raw = get_nonlinearity(a, down),
+                     
+                     hyst_tot_log = get_hysteresis_total(log10(up+1), log10(down+1)),
+                     hyst_range_log = get_hysteresis_range(log10(up+1), log10(down+1), a),
+                     hyst_min_log = get_hysteresis_min(log10(up+1), log10(down+1), a),
+                     hyst_max_log = get_hysteresis_max(log10(up+1), log10(down+1), a),
+                     nl_up_log = get_nonlinearity(a, log10(up+1)),
+                     nl_down_log = get_nonlinearity(a, log10(down+1))
+    ) 
   }
   
   if(is.na(init_varying)) {
@@ -71,12 +82,20 @@ get_stability_measures <- function(ss_object) {
   
   res <- temp %>%
     dplyr::group_by(Species) %>%
-    dplyr::summarise(hyst_tot = 0,
-                     hyst_range = 0,
-                     hyst_min = 0,
-                     hyst_max = 0,
-                     nl_up = get_nonlinearity(a, log10(Quantity+1)),
[…]

### 2021-10-04 various_stuff/EWS functions/condhet_analysis.R
small update to repro. vignette and remove various_stuff folder

@@ -1,143 +1,0 @@
-#---
-# title: "Analysis functions"
-# author: "Jennifer Probst"
-# date: "2019 M08 21"
-# output: html_document
-# Functions used for the analysis of EWS
-# ---
-# adjust the paths where the graphs should be saved!! (lines 274, 465, 507, 664 and 681)
-
-#------------------------------------------------------------------------------------#
-
-# Conditional Heteroskedasticity 
-
-#' To Add
-#'
-#' @param dataframe TODO
-#' @param name TODO
-#' @param interpolate TODO
-#'
-#' @return TODO
-#' 
-#' @importFrom graphics abline legend lines mtext par text
-#' @importFrom grDevices png dev.off
-#' @importFrom stats approx ar lm na.omit qchisq ts
-#' 
-#' @export
-#'
-#' @examples
-condhet_analysis <- function(dataframe, name, interpolate=FALSE) {
-  
-  #input: dataframe with observed univariate time series, name of timeseries and 'TRUE' if data needs to be interpolated
-  #output: plot of conditional heteroskedasticity over time
-  
-  #functions from https://github.com/earlywarningtoolbox/earlywarnings-R/blob/master/earlywarnings/R/ch_ews.R
-  ch_ews <- function(timeseries, winsize = 10, alpha = 0.1, optim = TRUE, lags = 4, 
-                     logtransform = FALSE, interpolate = FALSE) {
-    
-    timeseries <- ts(timeseries)  #strict data-types the input data as tseries object for use in later steps
-    if (dim(timeseries)[2] == 1) {
-      ts.in = timeseries
-      timeindex = 1:dim(timeseries)[1]
-    } else if (dim(timeseries)[2] == 2) {
-      ts.in = timeseries[, 2]
-      timeindex = timeseries[, 1]
-    } else {
-      warning("not right format of timeseries input")
-    }
-    
-    # Interpolation
-    if (interpolate) {
-      YY <- approx(timeindex, ts.in, n = length(ts.in), method = "linear")
-      ts.in <- YY$y
-    }
-    
-    # Log-transformation
-    if (logtransform) {
-      ts.in <- log(ts.in + 1)
-    }
-    
-    winSize = round(winsize * length(ts.in)/100)
-    sto <- matrix(nrow = (length(ts.in) - (winSize - 1)), ncol = 5)  # creates a matrix to store output
-    
-    count <- 1  #place holder for writing to the matrix
[…]

### 2021-10-04 various_stuff/EWS functions/ddj_model.R

@@ -1,165 +1,0 @@
-#---
-# title: "Analysis functions"
-# author: "Jennifer Probst"
-# date: "2019 M08 21"
-# output: html_document
-# Functions used for the analysis of EWS
-# ---
-# adjust the paths where the graphs should be saved!! (lines 274, 465, 507, 664 and 681)
-
-#------------------------------------------------------------------------------------#
-
-# Model based indicators: 
-#   Drift Diffusion Jump Nonparametrics Early Warning Signals
-
-#' To Add
-#'
-#' @param dataframe TODO
-#'
-#' @return TODO
-#' 
-#' @importFrom graphics grid lines mtext par
-#' @importFrom grDevices dev.new graphics.off
-#' @importFrom stats approx sd ts
-#' 
-#' @export
-#'
-#' @examples
-ddj_model <- function(dataframe) {
-  
-  #input: dataframe with observed univariate time series
-  #output: data object with various elements, see: ddjnonparam_ews function:  https://cran.r-project.org/web/packages/earlywarnings/earlywarnings.pdf
-
-  #modified functions from the ddjnonparam_ews function:  https://github.com/earlywarningtoolbox/earlywarnings-R/blob/master/earlywarnings/R/ddjnonparam_ews.R
-  # Function to compute Bandi, Johannes etc estimators for time series x
-  Bandi5 <- function(x0, dx, nx, DT, bw, na, avec) {
-    
-    # Set up constants and useful preliminaries
-    SF <- 1/(bw * sqrt(2 * pi))  # scale factor for kernel calculation
-    x02 <- x0 * x0  # second power of x
-    dx2 <- dx * dx  # second power of dx
-    dx4 <- dx2 * dx2  # fourth power of dx
-    dx6 <- dx2 * dx4  # sixth power of dx
-    # Compute matrix of kernel values
-    Kmat <- matrix(0, nrow = na, ncol = nx)
-    for (i in 1:(nx)) {
-      # loop over columns (x0 values)
-      Kmat[, i] <- SF * exp(-0.5 * (x0[i] - avec) * (x0[i] - avec)/(bw * bw))
-    }
[…]

### 2021-10-04 various_stuff/EWS functions/dfa_analysis.R

@@ -1,47 +1,0 @@
-#---
-# title: "Analysis functions"
-# author: "Jennifer Probst"
-# date: "2019 M08 21"
-# output: html_document
-# Functions used for the analysis of EWS
-# ---
-# adjust the paths where the graphs should be saved!! (lines 274, 465, 507, 664 and 681)
-
-#------------------------------------------------------------------------------------#
-
-# Detrending fluctuation analysis
-# https://www.rdocumentation.org/packages/nonlinearTseries/versions/0.2.3/topics/dfa
-
-#' To Add
-#'
-#' @param switch TODO
-#' @param var_col TODO
-#' @param n TODO
-#' @param t TODO
-#' @param name TODO
-#'
-#' @return TODO
-#' 
-#' @importFrom graphics lines
-#' @importFrom grDevices png dev.off
-#' 
-#' @export
-#'
-#' @examples
-dfa_analysis <- function(switch, var_col, n, t, name) {
-  
-  #input: a dataframe with data of respective switch, windowsize n t=totallength-n-1 and the name of timeseries
-  #output: plot of dfa over time
-  
-  dfa_vector <- numeric(t-11)
-  for (i in 1:(t-11)) {
-    x <- dfa(switch[i:n+i, var_col], window.size.range = c(10, 300), npoints = 100, do.plot = FALSE)
-    dfa_vector[i] <- estimate(x, regression.range = NULL, do.plot = FALSE, fit.col = 2, fit.lty = 1, fit.lwd =1)[1]
-  } 
-  
-  #plot
-  mypath <- file.path("C:","Users","probs","Desktop", "Sem 6", "Internship", paste(c(name, colnames(switch)[var_col], '_dfa.png'), collapse = ''))
-  png(filename=mypath)
-  plot(dfa_vector[0:(t-100)]~tottime[n:(length(switch[,var_col])-100)], xlab='timestep', ylab='DFA', main=paste(c('Detrended fluctuation analysis of ',colnames(switch)[var_col], ' of ', name), collapse = ''), type='l')
-  dev.off()
-}

### 2021-10-04 various_stuff/dumping_ground/dumping_ground1.R

@@ -1,321 +1,0 @@
-## Code test
-
-```{r eval = FALSE}
-dynamic_model = bushplus_dynamic_model
-event_definition = default_event_definition
-parameter_values = default_parameter_values
-event_interval = default_event_interval
-noise_sigma = default_noise_sigma
-minimum_abundances = default_minimum_abundances
-sim_duration = default_sim_duration
-sim_sample_interval = default_sim_sample_interval
-log10a_series = default_log10a_series
-initial_state = default_initial_state
-solver_method = "radau"
-
-times <- seq(0,
-             sim_duration,
-             by = sim_sample_interval)
-
-event_times <- seq(min(times),
-                   max(times),
-                   by = event_interval)
-
-log10a_forcing <- matrix(ncol=2,
-                         byrow=F,
-                         data=c(ceiling(seq(0,
-                                            max(times),
-                                            length=length(log10a_series))),
-                                log10a_series))
-
-
-log10a_forcing <- approxfun(x = log10a_forcing[,1],
-                            y = log10a_forcing[,2],
-                            method = "linear",
-                            rule = 2)
-assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)
-
-out <- as.data.frame(ode(y = initial_state,
-                         times = times,
-                         func = dynamic_model,
-                         parms = parameter_values,
-                         method = solver_method,
-                         events = list(func=event_definition,
-                                       time=event_times)))
-
-bushplus_dynamic_model(1, initial_state, default_parameter_values)
-
-sim_res <- run_simulation(
-  parameter
-  # dynamic_model = bushplus_dynamic_model,
-  # event_definition = default_event_definition,
-  # parameter_values = default_parameter_values,
-  # event_interval = default_event_interval,
-  # noise_sigma = default_noise_sigma,
-  # minimum_abundances = default_minimum_abundances,
-  # sim_duration = default_sim_duration,
-  # sim_sample_interval = default_sim_sample_interval,
[…]

### 2021-10-04 various_stuff/matrix_version_foodweb_model/check owen/DefineParameterValues.R

@@ -1,198 +1,0 @@
-## set the working directory
-setwd("~/Dropbox/TMMassie/check")
-
-
-# number of species
-species.number <- 7 
- 
- 
-########################################################################  
-######################################################################## 
-## Define parameter values here!
-
-## Define all time related parameters for integration.
-tperiod = 5      # replacing flask content every 'tperiod' days
-ncycles = 5      # number of replacement events
-t_start = 0		 # simulation starts at t_start
-t_stop  = tperiod*(1+ncycles)	# total simulation time
-t_step  = 0.05	 # time steps for output data
-# define time range for simulation as well as steps in time
-times   <- seq(t_start, t_stop, t_step)				# for the simulation, seq(from, to, by)
-timing  <- seq(tperiod, tperiod*ncycles, tperiod)	# for the discrete events, seq(from, to, by)
-
-## BACTERIAL UPTAKE RATES for resource
-u01 = 2.00     	 # bacterial species 1
-u02 = 2.00     	 # bacterial species 2
-
-## HALF-SATURATION CONSTANTS for bacterial resource uptake
-K01 = 0.50  	 # 1st halfsaturation constant, affinity for medium resource
-K02 = 0.50 		 # 2nd halfsaturation constant, affinity for medium resource
- 
-## SEARCH COEFFICIENTS describing the increase in the instantaneous search rate a
-b13 = 0.70       # species 3 on species 1
-b23 = 0.70       # species 3 on species 2
-b14 = 0.70       # species 4 on species 1
-b24 = 0.70       # species 4 on species 2
-b15 = 0.70       # species 5 on species 1
-b25 = 0.70       # species 5 on species 2
-b36 = 0.80       # species 6 on species 3
-b46 = 0.80       # species 6 on species 4
-b56 = 0.80       # species 6 on species 5
- 
-## HANDLING TIMES
-h13 = 1.50       # time spent of species 3 handling species 1
-h23 = 1.50       # time spent of species 3 handling species 2
-h24 = 1.50       # time spent of species 4 handling species 2
-h15 = 1.50       # time spent of species 5 handling species 1
-h14 = 1.50       # time spent of species 4 handling species 1
[…]

### 2021-10-04 various_stuff/matrix_version_foodweb_model/check owen/nonmatrix.R

@@ -1,309 +1,0 @@
-# Food web model, first attempt.
-# Thomas M. Massie | July 15, 2012
-
-
-#rm(list=ls())
-
-## set the working directory
-
-## source the functions used (none at the moment)
-
-
-## Read parameters values from the shared values file.
-data   <- read.csv("values_all.csv", row.names=1)
-# for (i in 1:length(data[,1])) {
-	# values <- data[i,1]
-	# names  <- rownames(data)[i]
-	# assign(names,values)
-# }
-parms <- c(data[,1])
-names(parms) <- rownames(data)
-
-# parameters <- data[,1]
-
-
-##################################
-## Define parameters.
-dil     <- 0.0
-ext.thresh.Sp1 <- 1e-5
-ext.thresh.Sp2 <- 1e-5
-ext.thresh.Sp3 <- 1e-5
-ext.thresh.Sp4 <- 1e-5
-ext.thresh.Sp5 <- 1e-5
-ext.thresh.Sp6 <- 1e-5
-
-
-
-# declare initial values of state variables S1-S6 (species) and Res (resource)
-Res_ini = 1000  # Resource
-Sp1_ini = 300    # B. subtilus
-Sp2_ini = 200    # S. fonticula
-Sp3_ini = 10    # Collodyction sp.
-Sp4_ini = 20     # Paramecium sp.
-Sp5_ini = 30     # Colpidium sp.
-Sp6_ini = 4     # Dileptus sp.
-Los_ini = 0     # Losses
-
-#Ini     = Sp1_ini + Sp2_ini + Sp3_ini + Sp4_ini + Sp5_ini + Sp6_ini + Res_ini + Los_ini
-
-state <- c(
-  Sp1 = Sp1_ini,    # B. subtilus
-  Sp2 = Sp2_ini,    # S. fonticula
-  Sp3 = Sp3_ini,    # Collodyction sp.
-  Sp4 = Sp4_ini,    # Paramecium sp.
-  Sp5 = Sp5_ini,    # Colpidium sp.
-  Sp6 = Sp6_ini,    # Dileptus sp.
-  Res = Res_ini,    # Resource
-  Los = Los_ini     # Losses due to mortality, inefficiencies...
-)
-
-
[…]

### 2021-10-26 R/event_definition_1.R
Documentation changes

@@ -1,12 +1,20 @@
 #' A function that contains events that should alter the state variables.
 #' (It is named definition "1", in case other definitions are created.)
 #'
-#' @param times The time point in the simulation
-#' @param y Current state variable values
-#' @param parms Model parameters
-#' @return A vector of the new state variables
+#' @param times the time point in the simulation
+#' @param y current state variable values
+#' @param parms object of class `strain_parameter`
+#' @return a vector of the new state variables
+#' @md
 #' @export
-event_definition_1 <- function(times=times, y=state, parms = parameters, log10a_forcing_func, noise_sigma, minimum_abundances) {
+event_definition_1 <- function(
+  times = times, 
+  y = state, 
+  parms = parameters, 
+  log10a_forcing_func, 
+  noise_sigma, 
+  minimum_abundances
+) {
   with(
     as.list(y),
     {

### 2021-10-26 R/run_simulation.R

@@ -1,8 +1,10 @@
-#' Function to run a simulation
+#' Run the simulation
 #'
-#' @param parameter An object of class \code{runsim_parameter} as returned by \code{new_runsim_parameter()}.
-#' @return A list containing the time series of state variables and all arguments passed.
-#' 
+#' @param parameter an object of class `runsim_parameter` as returned by
+#'   `new_runsim_parameter()`.
+#' @return an object of class 'runsim_result`, obtained from running the
+#'   su=imulation as defined in `parameter`.`
+#' @md
 #' @importFrom stats approx approxfun
 #' @importFrom deSolve ode
 #' 
@@ -79,7 +81,7 @@ run_simulation <- function(
 
   
   out <- as.data.frame(
-    ode(
+    deSolve::ode(
       y = parameter$strain_parameter$initial_state,
       times = times,
       func = parameter$dynamic_model,

### 2021-10-29 vignettes/partial-reproduction-Bushetal2017.R
Initial start

@@ -7,15 +7,13 @@ knitr::opts_chunk$set(echo = TRUE,
                       fig.height = 4)
 
 
-
-## -----------------------------------------------------------------------------
 library(tidyverse)
 library(deSolve)
 library(rootSolve)
 library(here)
-
-devtools::load_all(here::here())
-
+#if (!require("microxanoxBeta")) {
+  devtools::load_all(here::here())
+#}
 library(patchwork)
 
 ## Set random number generator seed
@@ -23,7 +21,7 @@ library(patchwork)
 set.seed(13)
 
 ## should some of the more intense simulations be run...
-run_steadystate_sims <- TRUE
+run_steadystate_sims <- FALSE
 options(mc.cores=8)
 
 ## -----------------------------------------------------------------------------
@@ -48,13 +46,12 @@ parameter <- new_runsim_parameter(
 rm(ma)
 
 ## -----------------------------------------------------------------------------
-
-parameter$initial_state  <-  new_initial_state(values = "bush_anoxic_fig2ab")
+parameter$strain_parameter$initial_state  <-  new_initial_state(values = "bush_anoxic_fig2ab")
 sim_res <- run_simulation(parameter)
 plot_dynamics(simulation_result = sim_res)
 
 ## -----------------------------------------------------------------------------
-parameter$initial_state = new_initial_state(values = "bush_oxic_fig2cd")
+parameter$strain_parameter$initial_state = new_initial_state(values = "bush_oxic_fig2cd")
 sim_res <- run_simulation(parameter)
 plot_dynamics(sim_res)
 ##ggsave("high_CB.png")

### 2022-01-06 R/bushplus_dynamic_model.R
CHecked vignettes and some minor changes.

@@ -9,7 +9,8 @@
 #' @param parameters An object of class `runsim_parameter` as returned by
 #'   `new_runsim_parameter()``
 #' @param log10a_forcing_func function to change oxygen diffusivity `a` depending on `t`
-#'   
+#' @param ... not used. Needed to catch additional parameter.
+#' 
 #' @return a list containing two elements, namely the rate of change of the
 #'   strains, and also the current values of oxygen diffusivity `a`.
 #' @md
@@ -19,7 +20,8 @@ bushplus_dynamic_model <- function(
   t, 
   state, 
   parameters, 
-  log10a_forcing_func
+  log10a_forcing_func,
+  ...
 ){
   
   CB <- state[grep("CB", names(state))]

### 2022-01-06 R/run_temporal_ssfind_experiment.R
Commenting of added functions and removal of the plot_amd_save function

@@ -1,10 +1,13 @@
-#' Run a stable state finding experiment (e.g. get the stable states for different levels of diversity).
+#' Run a stable state finding experiment (e.g. get the stable states for
+#' different levels of diversity).
 #'
 #' @param parameter an object of class `runsim_parameter` as returned by
 #'   `new_runsim_parameter()`.
-#' @param var_expt An object that describes different levels of diversity that should be examined. This object is not created in the 'microxanox' package.
-#' @param total_initial_abundances An object containing the total abundance in each functional group.
-#' @return an object ...
+#' @param var_expt An object that describes different levels of diversity that
+#'   should be examined. This object is **not** created in the 'microxanox' package.
+#' @param total_initial_abundances An object containing the total abundance in
+#'   each functional group.
+#' @return an `tibble` object ... TODO
 #' @md
 #' @importFrom multidplyr new_cluster cluster_library partition
 #' @export

### 2022-01-06 R/set_temporal_ssfind_initial_state.R

@@ -1,6 +1,6 @@
 #' Convenience function for setting initial states during the process of finding
 #' stable states using the function `run_temporal_ssfind_experiment`
-#'
+#' Should only be used internally.
 #' @param p parameters for the simulation
 #' @param initial_total_CB Total initial abundance of CB
 #' @param initial_total_PB Total initial abundance of PB
@@ -8,9 +8,6 @@
 #' @return The passed parameter object but with initial states set (overwritten by the new ones)
 #' @md
 #' 
-#' @export
-
-
 set_temporal_ssfind_initial_state <- function(p,
                                               initial_total_CB,
                                               initial_total_PB,

### 2022-01-10 inst/ODEINTR.Rmd

---
title: "odeintr"
author: "Uriah Daugaard"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: true
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# About

In this document I implement the ODEs used by Bush et al. (2017) using the package odeintr. I do this because (1) I'm pretty sure that this is the fastest way of solving ODEs in R and (2) this document can serve as a HOW-TO.

The difference to other approaches is that in ODEINTR the ODEs need to be compiled before they can be used. This takes about 10 seconds (usually, depending on the computer) up to max a minute. This times needs to be accounted when we compare approaches. However, they need to be compiled only once per R-session, so for each run after the first this time does not occur anymore. The running of the ODEs (i.e. solving them) should then be very very veeery fast!

Here I do the following:

* Implement the ODEs
* Reproduce Figure 2a,b of Bush et al 2017 and see how long it takes
* At some point compare with the implemented version in microxanox to see what is faster.

I've mentioned the possibility of using this package back in August: https://github.com/UZH-PEG/microxanox/issues/20

# README

As of now, the two methods (ODEINTR and microxanox) produce different results (when I compare the two datasets at the end of this document). Hence, there is probably a mistake somewhere, so what is shown here should be further tested before any decision is made.

# libraries
```{r}
library(odeintr)
library(tidyverse)
library(patchwork)
library(microbenchmark)
```

# Implementing the ODES with ODEINTR

## The ODEs (in C)

The ODEs like this don't look that nice, but once they're written no need to look at them again...
```{r}
ODEs = '
dxdt[0] = g_max_CB*(x[6]/(K_CB_P + x[6])) * 1/(1 + (x[4]/H_CB_S_R)) * x[0] - m_CB*x[0];
dxdt[1] = g_max_PB*(x[6]/(K_PB_P + x[6]))*(x[4]/(K_PB_S_R + x[4])) * 1/(1 + (x[5]/H_PB_O)) * x[1] - m_PB*x[1];
dxdt[2] = g_max_SB*(x[6]/(K_SB_P + x[6]))*(x[3]/(K_SB_S_O + x[3])) * 1/(1 + (x[5]/H_SB_O)) * x[2] - m_SB*x[2];
dxdt[3] = g_max_PB*(x[6]/(K_PB_P + x[6]))*(x[4]/(K_PB_S_R + x[4])) * 1/(1 + (x[5]/H_PB_O)) * x[1]/gamma_SR_PB - g_max_SB*(x[6]/(K_SB_P + x[6]))*(x[3]/(K_SB_S_O + x[3])) * 1/(1 + (x[5]/H_SB_O)) * x[2]/gamma_SO_SB + c*x[5]*x[4] + alpha_S*(S_O_b-x[3]);
dxdt[4] = - g_max_PB*(x[6]/(K_PB_P + x[6]))*(x[4]/(K_PB_S_R + x[4])) * 1/(1 + (x[5]/H_PB_O)) * x[1]/gamma_SR_PB + g_max_SB*(x[6]/(K_SB_P + x[6]))*(x[3]/(K_SB_S_O + x[3])) * 1/(1 + (x[5]/H_SB_O)) * x[2]/gamma_SO_SB - c*x[5]*x[4] + alpha_S*(S_R_b-x[4]);
dxdt[5] = p_CB * g_max_CB*(x[6]/(K_CB_P + x[6])) * 1/(1 + (x[4]/H_CB_S_R)) * x[0] - c*x[5]*x[4] + alpha_O*(O_b-x[5]);
dxdt[6] = - g_max_CB*(x[6]/(K_CB_P + x[6])) * 1/(1 + (x[4]/H_CB_S_R)) * x[0]/gamma_P_CB - g_max_PB*(x[6]/(K_PB_P + x[6]))*(x[4]/(K_PB_S_R + x[4])) * 1/(1 + (x[5]/H_PB_O)) * x[1]/gamma_P_PB - g_max_SB*(x[6]/(K_SB_P + x[6]))*(x[3]/(K_SB_S_O + x[3]))*1/(1 + (x[5]/H_SB_O))*x[2]/gamma_P_SB + alpha_P*(P_b -x[6]);
'
# x <- c(x0="N_CB",x1="N_PB",x2="N_SB",x3="S_O",x4="S_R",x5="O",x6="P")
```

## Compiling the ODEs

This step takes a little bit of time, but is only needed once per ODE-system and R-session.  
```{r}
start_time <- Sys.time()
compile_sys(
  "ODEs_compiled",
  ODEs,
  pars = c(
    "g_max_CB",
    "g_max_PB",
    "g_max_SB",
    "K_PB_S_R",
    "K_SB_S_O",
    "K_CB_P",
    "K_PB_P",
    "K_SB_P",
    "H_CB_S_R",
    "H_PB_O",
    "H_SB_O",
    "gamma_SO_SB",
    "gamma_SR_PB",
    "gamma_P_CB",
    "gamma_P_PB",
    "gamma_P_SB",
    "p_CB",
    "m_CB",
    "m_PB",
    "m_SB",
    "alpha_S",
    "alpha_O",
    "alpha_P",
    "S_R_b",
    "S_O_b",
    "O_b",
    "P_b",
    "c"
  ),
  method = "rk54"
)
end_time <- Sys.time()

paste("Compiling the ODEs took", end_time-start_time, "seconds, not bad!")
```

# Reproducing Figure 2a,b

## The parameter values

Using the ones for Bush et al Figure 2

```{r}
pars <- c(
  g_max_CB = 0.05,
  g_max_PB = 0.07,
  g_max_SB = 0.1,
  K_PB_S_R = 10,
  K_SB_S_O = 5,
  K_CB_P = 0.2,
  K_PB_P = 0.5,
  K_SB_P = 0.5,
  H_CB_S_R = 300,
  H_PB_O = 100,
  H_SB_O = 100,
  gamma_SO_SB = 3.33 * 10 ^ 7,
  gamma_SR_PB = 1.25 * 10 ^ 7,
  gamma_P_CB = 1.67 * 10 ^ 8,
  gamma_P_PB = 1.67 * 10 ^ 8,
  gamma_P_SB = 1.67 * 10 ^ 8,
  p_CB = 6 * 10 ^ (-9),
  m_CB = 0.02,
  m_PB = 0.028,
  m_SB = 0.04,
  alpha_S = 0.001,
  alpha_O = 8 * 10 ^ (-4),
  alpha_P = 0.01,
  S_R_b = 300,
  S_O_b = 300,
  O_b = 300,
  P_b = 9.5,
  c = 4 * 10 ^ (-5)
)

alpha_O <- c(10^(-6),10^(-2))
P_b <- c(2,10)
```

## Adding the paramter values to the ODEs
```{r}
ODEs_compiled_set_params(
  g_max_CB = pars["g_max_CB"],
  g_max_PB = pars["g_max_PB"],
  g_max_SB = pars["g_max_SB"],
  K_PB_S_R = pars["K_PB_S_R"],
  K_SB_S_O = pars["K_SB_S_O"],
  K_CB_P = pars["K_CB_P"],
  K_PB_P = pars["K_PB_P"],
  K_SB_P = pars["K_SB_P"],
  H_CB_S_R = pars["H_CB_S_R"],
  H_PB_O = pars["H_PB_O"],
  H_SB_O = pars["H_SB_O"],
  gamma_SO_SB = pars["gamma_SO_SB"],
  gamma_SR_PB = pars["gamma_SR_PB"],
  gamma_P_CB = pars["gamma_P_CB"],
  gamma_P_PB = pars["gamma_P_PB"],
  gamma_P_SB = pars["gamma_P_SB"],
  p_CB = pars["p_CB"],
  m_CB = pars["m_CB"],
  m_PB = pars["m_PB"],
  m_SB = pars["m_SB"],
  alpha_S = pars["alpha_S"],
  alpha_O = pars["alpha_O"],
  alpha_P = pars["alpha_P"],
  S_R_b = pars["S_R_b"],
  S_O_b = pars["S_O_b"],
  O_b = pars["O_b"],
  P_b = pars["P_b"],
  c = pars["c"]
)
```

## Starting values
```{r}
starting_densities <-
  c(
    N_CB = 5 * 10 ^ 1,
    N_PB = 1 * 10 ^ 7,
    N_SB = 1 * 10 ^ 7,
    S_O = 300,
    S_R = 300,
    O = 10,
    P = 10
  )
# starting_densities <- c(N_CB = 1*10^8, N_PB = 1*10^2, N_SB = 1*10^2, S_O = 500, S_R = 50, O = 300, P = 4)
```

## Running the model!

I didn't look up exactly how many time steps Bush et al used, but for the moment it doesn't matter. here I just run the ODEs to cover a length of 3840 hours, i.e. 160 days. And I use a time-step of 1/100 (i.e. 100 steps per hour, meaning that I simulate 384000 time-steps). We'll see later how much this values impact the performance, for now i just want to reproduce the figure.

```{r}
start_time <- Sys.time()
run <- ODEs_compiled(starting_densities,160*24,1/100)
end_time <- Sys.time()

paste("Running the model took", end_time-start_time, "seconds, not bad!")
```

## The figure (and putting the data into the right shape)

Finally, here I reproduce the figure.
```{r}
colnames(run) <- c("Time", names(starting_densities))
run_longer <- run %>% pivot_longer(cols = -Time)
run_longer$group <- ifelse(run_longer$name %in% c("N_CB","N_PB","N_SB"),"Bacteria","Substrate")

run_longer %>%
  dplyr::filter(name %in% c("N_CB","N_PB","N_SB")) %>%
  ggplot(aes(Time/24,value,col=name)) +
  geom_line() +
  scale_y_log10(limits=c(10^0,10^9)) +

run_longer %>%
  dplyr::filter(name %in% c("S_O","S_R","O","P")) %>%
  ggplot(aes(Time/24,value,col=name)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2")
```

The figure looks pretty identical! Only P looks different, but I used the starting value they stated, while in their figure they are using something else...
successfully reproduced!

# Stress testing: how long does it take?

For simplicity I just increase length of the simulations
```{r}
# 100 days, 2400 time steps per day
start_time <- Sys.time()
run <- ODEs_compiled(starting_densities,100*24,1/100)
end_time <- Sys.time()
paste("Running the model took", end_time-start_time, "seconds, not bad!")

# 500 days, 2400 time steps per day
start_time <- Sys.time()
run <- ODEs_compiled(starting_densities,500*24,1/100)
end_time <- Sys.time()
paste("Running the model took", end_time-start_time, "seconds, not bad!")

# 1000 days, 2400 time steps per day
start_time <- Sys.time()
run <- ODEs_compiled(starting_densities,1000*24,1/100)
end_time <- Sys.time()
paste("Running the model took", end_time-start_time, "seconds, not bad!")

# # 5000 days, 2400 time steps per day
# start_time <- Sys.time()
# run <- ODEs_compiled(starting_densities,5000*24,1/100)
# end_time <- Sys.time()
# paste("Running the model took", end_time-start_time, "seconds, not bad!")
# 
# # 10000 days, 2400 time steps per day
# start_time <- Sys.time()
# run <- ODEs_compiled(starting_densities,10000*24,1/100)
# end_time <- Sys.time()
# paste("Running the model took", end_time-start_time, "seconds, not bad!")
```

# Comparison with microxanox

Now let's compare with microxanox. Let's copy paste a bit of code from "partial-reproduction-Bushetal2017.Rmd" (i.e. the setup and stuff):

```{r setup_micro, output = FALSE, message = FALSE}
library(tidyverse)
library(deSolve)
library(rootSolve)
library(here)
library(microxanox)
library(patchwork)
set.seed(13)
run_steadystate_sims <- FALSE
options(mc.cores=1) # let's just use 1 core for now!

ma <- rep(1e2, 3)
names(ma) <- c("CB", "PB", "SB")
parameter <- new_runsim_parameter(
  dynamic_model = bushplus_dynamic_model,
  strain_parameter = new_strain_parameter(),
  ## no default_initial_state
  sim_duration = 2000,
  sim_sample_interval = 1,
  log10a_series = c(
    log10(new_strain_parameter()$a_O),
    log10(new_strain_parameter()$a_O)
  ),
  event_definition = event_definition_1,
  event_interval = 10,
  noise_sigma = 0,
  minimum_abundances = ma
)

rm(ma)
# saveRDS(parameter, file = "data/partial_reproduction/parameter_default.RDS")
```

So here microxanox reproduces the figure 2a,b:

```{r}
parameter$strain_parameter$initial_state  <-  new_initial_state(values = "bush_anoxic_fig2ab")
sim_res <- run_simulation(parameter)
# saveRDS(sim_res, file = "data/partial_reproduction/bush_anoxic_fig2ab.RDS")

plot_dynamics(simulation_result = sim_res)
```

The parameters the microxanox used are here: (let's compare them to the ones used above)

```{r}
sim_res$strain_parameter
```

Parameter values and starting values are the same, as they should. Nice!

Ok now the interesting part: let's compare!

Microxanox as above uses 2000 hours and 1 sample per hour as settings. We start with that.

## 2000 hours and 1 sample per hour

```{r}
microbenchmark(
  run_simulation(parameter), # microxanox
  ODEs_compiled(starting_densities,2000,1), # odeintr
  times = 10
)
```

Looks like odeintr is about 1000 times faster! But remember, ODEINTR must be compiled once, which takes some seconds.

Note: the resulting dataset differ quite a bit, so there must be a mistake somewhere...
```{r}
run1 <- run_simulation(parameter)
run1 <- run1$result
run2 <- ODEs_compiled(starting_densities,2000,1)
colnames(run2) <- c("Time", names(starting_densities))
merged  <- cbind(run1,run2)
print.data.frame(merged[1:40,])
```

### 2022-01-17 R/bushplus_dynamic_model.R
final changes

@@ -9,7 +9,7 @@
 #' @param parameters An object of class `runsim_parameter` as returned by
 #'   `new_runsim_parameter()``
 #' @param log10a_forcing_func function to change oxygen diffusivity `a` depending on `t`
-#' @param ... not used. Needed to catch additional parameter.
+#' @param ... not used. Needed to catch additional parameters.
 #' 
 #' @return a list containing two elements, namely the rate of change of the
 #'   strains, and also the current values of oxygen diffusivity `a`.

### 2022-01-17 R/get_final_states_a_N.R

@@ -1,6 +1,6 @@
 #' Get the steady state solutions for a set of oxygen diffusivity and initial states.
 #' 
-#' Probably best to not use this directly, but rather via the function \code{ss_by_a_N()}
+#' Can be used direclty. Also can be used indirectly, via the function \code{ss_by_a_N()}
 #' @param x A vector of oxygen diffusivity and initial conditions.
 #' @param parameter  object of class \code{ss_by_a_N_parameter} as generated by \code{new_ss_by_a_N_parameter()}
 #' 

### 2022-01-17 R/new_CB_strain_parameter.R

@@ -1,4 +1,4 @@
-#' Create CB strain parameter
+#' Create CB strain parameter values
 #'
 #' @param n number of strains
 #' @param values Allowed values are:

### 2022-01-17 R/plot_dynamics.R

@@ -1,7 +1,7 @@
 #' Plot the dynamics of a model run
 #' 
 #' This is a convenience function to plot the dynamics of a model run,
-#' with strains within functional groups. It can be used as a template on how the data can be plotted.
+#' with strains within functional groups displayed.
 #' @param simulation_result Object returned by the run_simulation function
 #' @param every_n Plot data of every other n sample.
 #' @return returns the ggplot object of the plot. If it is assigned to a 

### 2022-01-17 R/run_simulation.R

@@ -1,9 +1,9 @@
 #' Run the simulation
 #' 
-#' This function takes the `parameter` and runs the simulation based on these.
-#' it returns an object of class `runsim_result` which contains an additional
+#' This function takes the `parameter` object and runs a simulation based on these.
+#' It returns an object of class `runsim_result` which contains an additional
 #' entry, i.e. `result` which contains the results of the simulation. The
-#' simulation can be re-run bu using the returned object as input `parameter`.
+#' simulation can be re-run using the returned object as input `parameter`.
 #' @param parameter an object of class `runsim_parameter` as returned by
 #'   `new_runsim_parameter()`.
 #' @return an object of class `runsim_result`, obtained from running the

### 2022-01-18 R/add_strain_var.R
cleaned up
removed commented out code, and some pointless code, and also added comments in code.

@@ -1,29 +1,29 @@
 #' Create variability in strain parameters
 #'
 #' Add variability as defined in the argument `variability` to the variables as follows:
-#'   - `strain_parameter$CB$g_max_CB <- variablility(strain_parameter$CB$g_max_CB, CB_var_gmax)`
-#'   - `strain_parameter$CB$h_SR_CB  <- variablility(strain_parameter$CB$h_SR_CB,  CB_var_h   )`
-#'   - `strain_parameter$SB$g_max_SB <- variablility(strain_parameter$SB$g_max_SB, SB_var_gmax)`
-#'   - `strain_parameter$SB$h_O_SB   <- variablility(strain_parameter$SB$h_O_SB,   SB_var_h   )`
-#'   - `strain_parameter$PB$g_max_PB <- variablility(strain_parameter$PB$g_max_PB, PB_var_gmax)`
-#'   - `strain_parameter$PB$h_O_PB   <- variablility(strain_parameter$PB$h_O_PB,   PB_var_h   )`
+#'   - `strain_parameter$CB$g_max_CB <- variability(strain_parameter$CB$g_max_CB, CB_var_gmax)`
+#'   - `strain_parameter$CB$h_SR_CB  <- variability(strain_parameter$CB$h_SR_CB,  CB_var_h   )`
+#'   - `strain_parameter$SB$g_max_SB <- variability(strain_parameter$SB$g_max_SB, SB_var_gmax)`
+#'   - `strain_parameter$SB$h_O_SB   <- variability(strain_parameter$SB$h_O_SB,   SB_var_h   )`
+#'   - `strain_parameter$PB$g_max_PB <- variability(strain_parameter$PB$g_max_PB, PB_var_gmax)`
+#'   - `strain_parameter$PB$h_O_PB   <- variability(strain_parameter$PB$h_O_PB,   PB_var_h   )`
 #' @param strain_parameter object of class `strain_parameter`
-#' @param CB_var_gmax the argument var for the function `variablility` for the
+#' @param CB_var_gmax the argument var for the function `variability` for the
 #'   variable `strain_parameter$CB$g_max_CB`
-#' @param CB_var_h the argument var for the function `variablility` for the
+#' @param CB_var_h the argument var for the function `variability` for the
 #'   variable `strain_parameter$CB$h_SR_CB`
-#' @param SB_var_gmax the argument var for the function `variablility` for the
+#' @param SB_var_gmax the argument var for the function `variability` for the
[…]

### 2022-01-18 R/get_nonlinearity.R

@@ -12,9 +12,11 @@
 #'
 get_nonlinearity <- function(x, y)
 {
-  
+  ## if all state variable values are zero, set the result value to NA
   if(sum(y)==0)
     L <- NA
+  
+  ## if not all zero, calculate the metric
   if(!(sum(y)==0)) {
     lin_pred <- predict(lm(y ~ x))
     gam_pred <- predict(mgcv::gam(y ~ s(x)))

### 2022-01-18 R/get_ssbyaN_parameter.R

@@ -36,8 +36,8 @@ get_ssbyaN_parameter <- function(result, i) {
     log10(parameter[["a_O"]])
   )
   
+  # set the result to NULL and deal with class
   parameter$result <- NULL
-  
   class(parameter) <- grep( "ss_by_a_N_result", class(parameter), value = TRUE, invert = TRUE)
   
   return(parameter)

### 2022-01-18 R/new_PB_strain_parameter.R

@@ -26,14 +26,16 @@ new_PB_strain_parameter <- function(
   n = 1,
   values = "bush"
 ){
+  
+  # Check for supported values of the argument "values"
   if (is.na(values)) {
     values <- "NA"
   }
-  
   if (!(values %in% c("NA", "bush"))) {
     stop("Not supported value for `values`!\n", "Only NA, 'NA' and 'bush' supported!")
   }
   
+  ## Create object
   x <- rep(as.numeric(NA), n)
   nm <- paste0("PB_", 1:n)
   result <- data.frame(
@@ -47,6 +49,8 @@ new_PB_strain_parameter <- function(
     m_PB = x,
     i_PB = x
   )
+  
+  ## Add values for the case of Bush et al 2017
   if (values == "bush") {
     result$g_max_PB = rep(0.07, n)
     result$k_PB_SR = rep(10, n)

### 2022-01-18 R/new_runsim_parameter.R

@@ -1,4 +1,4 @@
-#' Create parameter to run the simulation by using the function `run_sim()`
+#' Create parameter set with which to run a simulation.
 #'
 #' @param ... named parameter for the simulation to be set. 
 #'   An error will be raised, if they are not part of the parameter set.

### 2022-01-18 R/new_strain_parameter.R

@@ -104,6 +104,7 @@ new_strain_parameter <- function(
     parms$c <- 4e-5
   }
   
+  ## set initial conditions
   parms$initial_state <- new_initial_state(
     n_CB = n_CB, 
     n_PB = n_PB, 

### 2022-01-18 R/run_temporal_ssfind_experiment.R

@@ -21,8 +21,8 @@ run_temporal_ssfind_experiment <- function(parameter,
 
   if(cores == 1)
   {
-    ## Next chunck of code:
-    ## For each line of var_expt, add strain variation, and get stable states.
+    ## For each row of var_expt, add strain variation, set initial state,
+    ## and get stable states.
     result <- var_expt %>%
       mutate(
         ssfind_pars = list({
@@ -40,7 +40,8 @@ run_temporal_ssfind_experiment <- function(parameter,
 
   if(cores > 1)
   {
-
+    ## For each row of var_expt, add strain variation, set initial state,
+    ## and get stable states.
     cluster <- multidplyr::new_cluster(cores)
     multidplyr::cluster_library(cluster, c("microxanox", "dplyr"))
     var_expt <- var_expt %>%

### 2022-01-18 R/ss_by_a_N.R

@@ -15,6 +15,7 @@ ss_by_a_N <- function(
   parameter, 
   mc.cores = getOption("mc.cores", 0)
 ){
+  
   if (!inherits(parameter, "ss_by_a_N_parameter")) {
     stop("parameter has to be an object of type `ss_by_a_N_parameter`!")
   }
@@ -24,6 +25,8 @@ ss_by_a_N <- function(
     mc.cores <- 1
   }
   
+  ## single core:
+  ## apply over rows of ss_expt
   if (mc.cores == 0) {
     temp_result <- apply(
       parameter$ss_expt, 
@@ -32,6 +35,8 @@ ss_by_a_N <- function(
         get_final_states_a_N(x, parameter)
       }
     )
+  ## multi core
+  ## mcapply over rows of ss_expt
   } else {
     temp_result <- parallel::mclapply(
       1:nrow(parameter$ss_expt),
@@ -46,6 +51,7 @@ ss_by_a_N <- function(
     )
   }
 
+  ## Assemble results
   result <- temp_result %>%
     tibble::tibble() %>%
     tidyr::unnest(cols = 1) %>%

### 2022-01-21 R/get_replication_ssfind_parameter.R
0.4.1: Refactoring and fixing of missing vignettes - renamed ss_by_an functions

@@ -1,11 +1,11 @@
-#' Extract \code{runsim_parameter} from \code{ss_by_a_N_parameter} 
+#' Extract \code{runsim_parameter} from \code{replication_ssfind_parameter} 
 #' 
 #' Extract the parameter to run an individual simulation from either a parameter set for a 
-#' steady state determination (\code{ss_by_a_N()}) or the result set returned.
-#' @param result  object of class \code{ss_by_a_N_parameter} as generated by \code{new_ss_by_a_N_parameter()}
+#' steady state determination (\code{run_replication_ssfind_parameter()}) or the result set returned.
+#' @param result  object of class \code{replication_ssfind_parameter} as generated by \code{new_replication_ssfind_parameter()}
 #' @param i Index of the individual simulation to extract the parameter for
 #' 
-#' @return an object of class \code{ss_by_a_N_parameter}
+#' @return an object of class \code{replication_ssfind_parameter}
 #' @export
 get_ssbyaN_parameter <- function(result, i) { 
   
@@ -38,7 +38,7 @@ get_ssbyaN_parameter <- function(result, i) {
   
   # set the result to NULL and deal with class
   parameter$result <- NULL
-  class(parameter) <- grep( "ss_by_a_N_result", class(parameter), value = TRUE, invert = TRUE)
+  class(parameter) <- grep( "replication_ssfind_result", class(parameter), value = TRUE, invert = TRUE)
   
   return(parameter)
 }

### 2022-01-21 R/new_replication_ssfind_result.R

@@ -1,0 +1,30 @@
+#' Create object of type `replication_ssfind_result` which is returned by the
+#' function `run_replication_ssfind_parameter()`.
+#'
+#' @param parameter object of class `replication_ssfind_parameter` which has been used
+#'   to run the simulation
+#' @param result a `data.frame` containing the results of the simulation
+#'
+#' @return object of the class `replication_ssfind_result`. This object of class
+#'   `replication_ssfind_result` is identical to an object of class `replication_ssfind_parameter`
+#'   plus an additional field `result` which contains the result of the simulation.
+#' @md
+#' @export
+#'
+#' @examples
+new_replication_ssfind_results <- function(
+  parameter,
+  result
+  ){
+  if (!inherits(parameter, "replication_ssfind_parameter")) {
+    stop("`parameter` has to be of class replication_ssfind_parameter")
+  }
+
+  p <- parameter
+  p$result <- result
+  if (!inherits(p, "replication_ssfind_result")) {
+    class(p) <- append(class(p), "replication_ssfind_result")
+  }
+
+  return(p)
+}

### 2022-01-21 R/new_ssbyaN_result.R

@@ -1,30 +1,0 @@
-#' Create object of type `ss_by_a_N_result` which is returned by the
-#' function `ss_by_a_N()`.
-#'
-#' @param parameter object of class `ss_by_a_N_parameter` which has been used
-#'   to run the simulation
-#' @param result a `data.frame` containing the results of the simulation
-#'
-#' @return object of the class `ss_by_a_N_result`. This object of class
-#'   `ss_by_a_N_result` is identical to an object of class `ss_by_a_N_parameter` 
-#'   plus an additional field `result` which contains the result of the simulation. 
-#' @md
-#' @export
-#'
-#' @examples
-new_ss_by_a_N_results <- function(
-  parameter, 
-  result 
-  ){
-  if (!inherits(parameter, "ss_by_a_N_parameter")) {
-    stop("`parameter` has to be of class ss_by_a_N_parameter")
-  }
-  
-  p <- parameter
-  p$result <- result
-  if (!inherits(p, "ss_by_a_N_result")) {
-    class(p) <- append(class(p), "ss_by_a_N_result")
-  }
-  
-  return(p)
-}

### 2022-01-21 R/get_stability_measures.R
Intermediate - not completely workig!!!!

@@ -2,223 +2,23 @@
 #' 
 #' The measures include non-linearity and hysteresis measures, of an ecosystem response to environmental change.
 #' Takes steady state data as the input. 
-#' @param ss_object An object of class \code{replication_ssfind_result} as returned by the `run_replication_ssfind_parameter()` function or
-#'   the \code{result} element of that object, i.e. \code{x$result}.
+#' @param x object from which to calculate the stability measures. At the moment `replication_ssfind_result` or `data.frame`.
+#' @param ... additional arguments for methods
+#'
 #' @return A data frame of stability measures of each state variable
-#' 
-#' @global Quantity starts_with direction up down a_O a 
-#' 
-#' @importFrom tidyr gather spread
-#' @importFrom dplyr summarise pull filter across
-#' @importFrom stats na.omit
-#' 
+#'
+#' @md
+#'
+#' @rdname get_stability_measures
 #' @export
+#'
+#' @examples
+#'
 get_stability_measures <- function(
-  ss_object
-){
-  if (inherits(ss_object, "replication_ssfind_result")) {
-    result <- ss_object$result
-  } else {
-    result <- ss_object
-  }
+  x,
+  ...
+) {
   
+  UseMethod("get_stability_measures")
   
-  ## which initial condition varies?
-  init_varying <- NA
-  if(length(unique(result$initial_N_CB)) > 1)
-    init_varying <- "initial_N_CB"
-  if(length(unique(result$initial_N_SB)) > 1)
-    init_varying <- "initial_N_SB"
-  if(length(unique(result$initial_N_PB)) > 1)
-    init_varying <- "initial_N_PB"
-  
-  if(!is.na(init_varying)) {
-    
-    result$init_varying <- dplyr::pull(result[,init_varying],1)
-    
-    min_iniN <- min(result$init_varying)
-    max_iniN <- max(result$init_varying)
-    
-    ## The following is preparing the data
-    these <- grep("B_", names(result))
-    these <- c(these, which(names(result) %in% c("SO", "SR", "O", "P")))
-    temp <- result %>%
-      dplyr::mutate(direction = ifelse(init_varying == min_iniN, "up", "down")) %>%
-      dplyr::filter(dplyr::across(these, ~ .x >-0.001)) %>% ## there are rarely negative abundances greater than -0.001. This line and the na.omit removes them 
-      tidyr::gather(key = "Species", value = Quantity, these) %>%
[…]

### 2022-01-21 R/new_replication_ssfind_result.R

@@ -1,5 +1,5 @@
 #' Create object of type `replication_ssfind_result` which is returned by the
-#' function `run_replication_ssfind_parameter()`.
+#' function `run_replication_ssfind()`.
 #'
 #' @param parameter object of class `replication_ssfind_parameter` which has been used
 #'   to run the simulation

### 2022-01-21 R/new_temporal_ssfind_result.R

@@ -1,0 +1,26 @@
+#' Create object of type `temporal_ssfind_result` which is returned by the
+#' function `run_temporal_ssfind_parameter()`.
+#' 
+#' @param parameter  not used - for consistency.
+#' @param result a `data.frame` containing the results of the simulation
+#'
+#' @return the `result` object.
+#' 
+#' @md
+#' @export
+#'
+#' @examples
+new_temporal_ssfind_results <- function(
+  parameter = NULL,
+  result
+  ){
+  # if (!inherits(parameter, "replication_ssfind_parameter")) {
+  #   stop("`parameter` has to be of class replication_ssfind_parameter")
+  # }
+
+  if (!inherits(result, "temporal_ssfind_result")) {
+    class(result) <- append(class(result), "temporal_ssfind_result")
+  }
+
+  return(result)
+}

### 2022-01-21 R/run_temporal_ssfind.R

@@ -11,47 +11,47 @@
 #' @export
 
 
-run_temporal_ssfind_method <- function(parameter) {
+run_temporal_ssfind <- function(parameter) {
   
   ## recalculate the wait time (length of a step)
-  wait_time <- p$sim_duration / length(p$log10a_series)
+  wait_time <- parameter$sim_duration / length(parameter$log10a_series)
   
   ## make function for the increasing ox diff steps
-  up_l_f_f <- approxfun(x = wait_time * c(0:length(p$log10a_series)), 
-                        y = (c(p$log10a_series, p$log10a_series[length(p$log10a_series)])),
+  up_l_f_f <- approxfun(x = wait_time * c(0:length(parameter$log10a_series)), 
+                        y = (c(parameter$log10a_series, parameter$log10a_series[length(parameter$log10a_series)])),
                         method = "constant", rule = 1)
   
   ## make function for the decreasing ox diff steps
-  down_l_f_f <- approxfun(x = wait_time * c(0:length(p$log10a_series)), 
-                          y = c(rev(p$log10a_series), p$log10a_series[1]),
+  down_l_f_f <- approxfun(x = wait_time * c(0:length(parameter$log10a_series)), 
+                          y = c(rev(parameter$log10a_series), parameter$log10a_series[1]),
                           method = "constant", rule = 1)
   
   ## make times at which observations are made (i.e. at the end of a step)
   times <- c(0,
-             seq(p$sim_sample_interval - 1,
-                 p$sim_duration,
-                 by = p$sim_sample_interval))
+             seq(parameter$sim_sample_interval - 1,
+                 parameter$sim_duration,
+                 by = parameter$sim_sample_interval))
   
   ## make times at which events occur
-  event_times <- c(0, seq(p$event_interval-1,
+  event_times <- c(0, seq(parameter$event_interval-1,
                           max(times),
-                          by = p$event_interval))
+                          by = parameter$event_interval))
   
   ## run a simulation for increasing ox diff
   up_res <- as.data.frame(
     deSolve::ode(
-      y = p$strain_parameter$initial_state,
+      y = parameter$strain_parameter$initial_state,
       times = times,
-      func = p$dynamic_model,
-      parms = p$strain_parameter,
-      method = p$solver_method,
+      func = parameter$dynamic_model,
+      parms = parameter$strain_parameter,
+      method = parameter$solver_method,
       events = list(
[…]

### 2022-01-24 R/get_stability_measures.R
0.4.4: FInalised refactoring
get_stability_measures() now generic function.

@@ -2,9 +2,10 @@
 #' 
 #' The measures include non-linearity and hysteresis measures, of an ecosystem response to environmental change.
 #' Takes steady state data as the input. 
-#' @param x object from which to calculate the stability measures. At the moment `replication_ssfind_result` or `data.frame`.
+#' @param ss_object object from which to calculate the stability measures. At the moment `replication_ssfind_result` or `data.frame`.
+#' @param threshold_diff_log10scale
 #' @param ... additional arguments for methods
-#'
+#' 
 #' @return A data frame of stability measures of each state variable
 #'
 #' @md
@@ -15,7 +16,8 @@
 #' @examples
 #'
 get_stability_measures <- function(
-  x,
+  ss_object,
+  threshold_diff_log10scale,
   ...
 ) {
   

### 2022-01-24 R/new_temporal_ssfind_result.R

@@ -19,7 +19,7 @@ new_temporal_ssfind_results <- function(
   # }
 
   if (!inherits(result, "temporal_ssfind_result")) {
-    class(result) <- append(class(result), "temporal_ssfind_result")
+    class(result) <- append("temporal_ssfind_result", class(result))
   }
 
   return(result)

### 2022-01-26 inst/manuscript/manuscript_microxanox.Rmd
m0.0.1: First outline of manuscript

---
title: "m0.0.1: Microxanox - ..."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
keywords: metadata quality; data curation; archival; long term storage; R package;
abstract: |
  1. set the context for and purpose of the work;
  2. indicate the approach and methods;
  3. outline the main results;
  4. identify the conclusions and the wider implications.

journal: 'Ecology and Evolution - Applications'
bibliography: manuscript_bibliography.bib
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  rmarkdown::html_vignette: default
  pdf_document:
    keep_tex: yes
header-includes:
- \usepackage{lineno}
- \linenumbers
- \usepackage{setspace}
- \doublespacing
- \usepackage{fancyhdr}
- \fancyhead{}
- \fancyfoot{}
- \fancyhead[CO,CE]{Metadata made easy - the dmdScheme approach}
- \pagestyle{fancy}
vignette: |
  %\VignetteIndexEntry{Manuscript - microxanox} 
  %\VignetteEncoding{UTF-8} 
  %\VignetteEngine{knitr::rmarkdown}
nocite: '@bush2017'
linkcolor: blue
link-citations: yes
csl: methods-in-ecology-and-evolution.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(microxanox)
```

# Journal Guidelines
  - Abstract < 350 words
  - Article 3000-4000 words

# Introduction
The aims of the `microxanox` package are:

-   Facilitate reproduction of the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
-   Include new functionality that allows new research questions to be answered.

We added functionality for:

-   Adding random noise in substrate concentrations.
-   Adding temporally varying oxygen diffusivity.
-   Including immigration.
-   Setting minimum population abundances.
-   Multiple strains per functional group.

We also include a function for visualising results, though **please note that it is provided for convenience, and not for production of perfectly formatted print-ready graphs.**

Here (again) is the original publication: <https://www.nature.com/articles/s41467-017-00912-x>

# The Model
Based on [@Bush2017]

The framework used for running a simulation, and for varying aspects of a simulation run is:

-   Define the default simulation conditions (i.e. parameters and initial conditions).
-   Use these defaults unless other conditions are specified.
-   The function call `run_simulation()` would therefore run a simulation with default conditions.
-   The function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values, and otherwise all the specified defaults.

The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`. This augmented version of the model published in Bush et al 2017 can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events (see below for event definition).

# Implementation, Availability and Usage
On github [microxanox github repository](https://github.com/opetchey/microxanox/)

# Use cases

# Conclusions

# References

### 2022-01-26 inst/manuscript/manuscript_microxanox.Rmd
intermediate

@@
   4. identify the conclusions and the wider implications.
-
 journal: 'Ecology and Evolution - Applications'
@@
 
--   Facilitate reproduction of the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
+-   Facilitate reproduction of the results of [@bush2017].
 -   Include new functionality that allows new research questions to be answered.
 
-We added functionality for:
+In addition to the model described in [@bush2017], we added functionality for:
 
@@
 
-We also include a function for visualising results, though **please note that it is provided for convenience, and not for production of perfectly formatted print-ready graphs.**
-
-Here (again) is the original publication: <https://www.nature.com/articles/s41467-017-00912-x>
+The package also ingcludes some functions to viaualise the results which can function as template functions for developing graphs for own rquirements.
 
 # The Model
-Based on [@Bush2017]
+The model is based on [@bush2017].
+We use in this package the notation from the Table S1 of the Supplementary information of [@bush2017]), and not the notation in the main manuscript of [@bush2017].
+(There is only one small difference between these: In the main text the notation is simplified to `y_S_PB` ($y_{S,PB}$) (which is `y_SR_PB` ($y_{SR,PB}$) in Table S1) and `y_S_SB` ($y_{S,SB}$) (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
+We also use `Pr_cB` ($Pr_B$) instead of `P_CB` ($P_{CB}$).
+
+## Model dimensions
+
+-   Time: hours
+-   Volume: litres
+-   Substrate quantity: micromoles
+-   Organism quantity: cells
+
+## Modelling framework
 
@@
 
--   Define the default simulation conditions (i.e. parameters and initial conditions).
--   Use these defaults unless other conditions are specified.
--   The function call `run_simulation()` would therefore run a simulation with default conditions.
--   The function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with the specified initial state values, and otherwise all the specified defaults.
+1. Define a parameter set of class `runsim_parameter` (i.e. parameters and initial conditions) by using the functions `new_runsim_parameter()`, `new_initial_state()` and `new_strain_parameter()`.
+2. The function call `run_simulation(parameter)` which will run a simulation using the parameter as defined in the object `parameter`.
 
 The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`.
-This augmented version of the model published in Bush et al 2017 can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events (see below for event definition).
+This augmented version of the model published in [@bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
 
 # Implementation, Availability and Usage
-On github [microxanox github repository](https://github.com/opetchey/microxanox/)
+The package is available in a github repo at https://github.com/UZH-PEG/microxanox/.
 
 # Use cases
+This package has been used in the production of the paper [@REF_NEEDED].
+It can be used for similar studies in this study systems.
+It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@bush2017] and to extend it to our needs.
+Consequently, any more substantial, are likely to need a change in the source code.
+Nevertheless, the model is structured in a way which builds on a udular structuere, so that e.g. the event defuinition can easily be changed.
 

### 2022-02-03 inst/manuscript/manuscript_microxanox.Rmd
Next Steps

@@
   html_document:
-    df_print: paged
-  rmarkdown::html_vignette: default
+    toc: true
+    toc_float: true
   pdf_document:
@@
 - \pagestyle{fancy}
-vignette: |
-  %\VignetteIndexEntry{Manuscript - microxanox} 
-  %\VignetteEncoding{UTF-8} 
-  %\VignetteEngine{knitr::rmarkdown}
-nocite: '@bush2017'
+
+nocite: '@Bush2017'
 linkcolor: blue
@@
 # Introduction
-The aims of the `microxanox` package are:
 
--   Facilitate reproduction of the results of [@bush2017].
--   Include new functionality that allows new research questions to be answered.
+{>> Biological justification and background <<}
 
-In addition to the model described in [@bush2017], we added functionality for:
+The aims of the `microxanox` package are twofold.
+Firstly, the package aims at reproducing the results shown by [@Bush2017].
+To take one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].
 
+In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.
+
+# The Model
+
+The model is an implementation of the model described by [@Bush2017]. It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.
+
+We extended the model and added functionality for:
+
+- Multiple strains per functional group.
+- Adding temporally varying oxygen diffusivity.
 - Adding random noise in substrate concentrations.
-- Adding temporally varying oxygen diffusivity.
 - Including immigration.
 - Setting minimum population abundances.
-- Multiple strains per functional group.
 
-The package also ingcludes some functions to viaualise the results which can function as template functions for developing graphs for own rquirements.
+We will go through these additions in detail below.
 
-# The Model
-The model is based on [@bush2017].
-We use in this package the notation from the Table S1 of the Supplementary information of [@bush2017]), and not the notation in the main manuscript of [@bush2017].
-(There is only one small difference between these: In the main text the notation is simplified to `y_S_PB` ($y_{S,PB}$) (which is `y_SR_PB` ($y_{SR,PB}$) in Table S1) and `y_S_SB` ($y_{S,SB}$) (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
+We use the notation from the [Table S1](LINK NEEDED) of the Supplementary information of [@Bush2017]), and not the notation in the main manuscript of [@Bush2017].
+In the main manuscript the notation is simplified to `y_S_PB` ($y_{S,PB}$) (which is `y_SR_PB` ($y_{SR,PB}$) in Table S1) and `y_S_SB` ($y_{S,SB}$) (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
 We also use `Pr_cB` ($Pr_B$) instead of `P_CB` ($P_{CB}$).
 
-## Model dimensions
+All parameter and variables follow these base dimensions:
 
--   Time: hours
--   Volume: litres
--   Substrate quantity: micromoles
--   Organism quantity: cells
+- Time: `hours`
+- Volume: `litres`
+- Substrate quantity: `micromoles`
+- Organism quantity: `cells`
 
-## Modelling framework
+{>> Add a more detailed description of the model and the additions <<}
 
-The framework used for running a simulation, and for varying aspects of a simulation run is:
+The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the odeSolve package [@REF_NEEDED]
 
+## The ODEs
+The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
+This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+
+# Availability and Accessability of the Package
+The package is available in a [github repo/](https://github.com/UZH-PEG/microxanox/).
+The newest version is also available under the [DOI](https://DOI) or via RUniverse [@REF_NEEDED].
+This manuscript describes version 1.0 which is available under [DOI](https://DOI) or installable within R by using `remotes::install_github("UZH-PEG", ref = "v1.0")`.
+
+The package will be updated based on further research activities. It will be tested continuously for the last revent, current and devel version of R.
+Issues, questions and suggestions should be communicated through the issue tracker at [https://github.com/UZH-PEG/microxanox/issues](https://github.com/UZH-PEG/microxanox/issues)
+
+# Usage of the Package
+The package includes a full documentetion, which includes
+
+1. Standard R help for each function (`?COMMAND` in R)
+2. Two vignettes acompanying the paclage, also available under RUniverse [Vignette 1](LINK NEEDED)
+3. This paper as a third vignette, also available under RUniverse [Vignette 1](LINK NEEDED)
+
+## Simulation Preparation
 1. Define a parameter set of class `runsim_parameter` (i.e. parameters and initial conditions) by using the functions `new_runsim_parameter()`, `new_initial_state()` and `new_strain_parameter()`.
+
+- Set parameter
+	- `new_xxx_parameter()`
+
+## Simulation steps
 2. The function call `run_simulation(parameter)` which will run a simulation using the parameter as defined in the object `parameter`.
 
-The ordinary differential equations for the rates of change are specified in the function `bushplus_dynamic_model()`.
-This augmented version of the model published in [@bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+- single simulation `run_simulation()`
+- Stable State Finding 
+	- Replication Method `run_replication_ssfind()`
+	- Temporal Method `run_replication_ssfind()`
+- {>>I would leave the `run_temporal_experiment()` out as it has no counterpart for the replication method <<}
 
-# Implementation, Availability and Usage
-The package is available in a github repo at https://github.com/UZH-PEG/microxanox/.
+## Analysis and Graphing
+
+{>> Have to add these aspects here <<}
 
@@
 It can be used for similar studies in this study systems.
-It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@bush2017] and to extend it to our needs.
+It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs.
 Consequently, any more substantial, are likely to need a change in the source code.

### 2022-03-07 inst/manuscript/manuscript_microxanox.Rmd
Next Steps Manuscript

@@
 
+The microxanox package is a package for simulating a three functional group system (CB: cyanobacteria, PB: phototrophic sulfur bacteria, and SB: sulfate-reducing bacteria) with four chemical substrates (P: phosphorus, O: oxygen, SR: reduced sulfur, SO: oxidized sulfur).
+It includes feddback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
+
 The aims of the `microxanox` package are twofold.
 Firstly, the package aims at reproducing the results shown by [@Bush2017].
-To take one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].
-
-In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.
-
-# The Model
-
-The model is an implementation of the model described by [@Bush2017]. It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.
-
+Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].
 We extended the model and added functionality for:
@@
 
-We will go through these additions in detail below.
+In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.
+
+# The Model
+
+The model is an implementation of the model described by [@Bush2017]. It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.
 
@@
 
-The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the odeSolve package [@REF_NEEDED]
+The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the `deSolve` package [@Soetaert2010]
 
@@
 This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+
+For an example how these can be used see [@REF_NEEDED].
 
@@
 
-The package will be updated based on further research activities. It will be tested continuously for the last revent, current and devel version of R.
+The package will be updated based on further research activities.
+It will be tested continuously for the last recent, current and devel version of R.
+{>>If we wnat that, we havt o include some, even rudimentary, GITHUB tests.<<}
+
 Issues, questions and suggestions should be communicated through the issue tracker at [https://github.com/UZH-PEG/microxanox/issues](https://github.com/UZH-PEG/microxanox/issues)
 
-# Usage of the Package
+# The Package
 The package includes a full documentetion, which includes
@@
 1. Standard R help for each function (`?COMMAND` in R)
-2. Two vignettes acompanying the paclage, also available under RUniverse [Vignette 1](LINK NEEDED)
-3. This paper as a third vignette, also available under RUniverse [Vignette 1](LINK NEEDED)
+2. Two vignettes acompanying the package, also available under RUniverse [Vignette 1](LINK NEEDED) and [Vignette 2](LINK NEEDED)
+3. This paper as a third vignette, also available under RUniverse [Vignette 3](LINK NEEDED)
 
-## Simulation Preparation
-1. Define a parameter set of class `runsim_parameter` (i.e. parameters and initial conditions) by using the functions `new_runsim_parameter()`, `new_initial_state()` and `new_strain_parameter()`.
+The simulations can be split into different levels:
 
-- Set parameter
-	- `new_xxx_parameter()`
+## The individuel simulation
+The individual simulation (`run_simulation()` function) is the working horse in this package.
+In this function, the ODEs are solved.
+The function needs only one parameter - an object as created by the function `new_runsim_parameter()`.
+One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
+For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at [User Guide](Link Needed)
 
-## Simulation steps
-2. The function call `run_simulation(parameter)` which will run a simulation using the parameter as defined in the object `parameter`.
+After the have been define, they can be used in the `run_simulation()` function.
+The function returns an object which is identical to the parameter, except of an additional slot, the results.
+This makes it possible to run the simulation again from the results from a previous simulation.
 
-- single simulation `run_simulation()`
-- Stable State Finding 
-	- Replication Method `run_replication_ssfind()`
-	- Temporal Method `run_replication_ssfind()`
-- {>>I would leave the `run_temporal_experiment()` out as it has no counterpart for the replication method <<}
+## Find a Steady State of a model
+There are two methods for finding steady states.
+The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*).
+The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*).
+
+### Replication Method
+The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
+
+### Temporal Method
+The temporal method involves two simulations for a particular system configuration (parameter set).
+In one simulation the oxygen diffusivity is *increased* in a step-wise fashion.
+In the other it is *decreased* in a step-wise fashion.
+That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set.
+Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.
+We hereafter term this "the temporal method".
+This is implemented in the function `run_temporal_ssfind()` while the parameter object is created by the function `new_temporal_ssfind_parameter()()`.'
+
+For a more detailed walk-through of these two approaches and explanation please see the [User Guide](LINK NEEDED).
+
+## Extract Stability Measures
+From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
+These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.
 

### 2022-03-14 inst/manuscript/manuscript_microxanox.Rmd
First draft Manuscript

@@
 
-```{r setup}
-library(microxanox)
+```{r setup, include = FALSE}
 ```
-
-# Journal Guidelines
-  - Abstract < 350 words
-  - Article 3000-4000 words
 
@@
 
-{>> Biological justification and background <<}
+{>> Biological justification and background - one paragraph <<}
 
-The microxanox package is a package for simulating a three functional group system (CB: cyanobacteria, PB: phototrophic sulfur bacteria, and SB: sulfate-reducing bacteria) with four chemical substrates (P: phosphorus, O: oxygen, SR: reduced sulfur, SO: oxidized sulfur).
+The microxanox package is a package for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
 It includes feddback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
@@
 The aims of the `microxanox` package are twofold.
-Firstly, the package aims at reproducing the results shown by [@Bush2017].
+Firstly, the package aims at reproducing the results shown by [@Bush2017], which is acomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED).
 Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].
-We extended the model and added functionality for:
 
-- Multiple strains per functional group.
+For this, we extended the model and added functionality for:
+
+- Multiple strains (effectively unlimited) per functional group.
 - Adding temporally varying oxygen diffusivity.
@@
 
-The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the `deSolve` package [@Soetaert2010]
+The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the `deSolve` package [@Soetaert2010].
 
@@
 
-For an example how these can be used see [@REF_NEEDED].
+For an example how these can be used see [@REF_NEEDED] or the vignettes acompanying this package [@LINK_NEEDED].
 
@@
 # The Package
-The package includes a full documentetion, which includes
+The package contains the full documentetion of the functions, which includes
 
 1. Standard R help for each function (`?COMMAND` in R)
-2. Two vignettes acompanying the package, also available under RUniverse [Vignette 1](LINK NEEDED) and [Vignette 2](LINK NEEDED)
-3. This paper as a third vignette, also available under RUniverse [Vignette 3](LINK NEEDED)
+2. Two vignettes acompanying the package, also available under RUniverse [Vignette 1](@LINK_NEEDED) and [Vignette 2](@LINK_NEEDED)
+3. This paper as a third vignette, also available under RUniverse [Vignette 3](@LINK_NEEDED)
+
+We will discuss the general structure and functiuonalit of the package without going into to mu=ch detail.
 
@@
 One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at [User Guide](Link Needed)
+For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at [User Guide](@LINK_NEEDED)
 
-After the have been define, they can be used in the `run_simulation()` function.
-The function returns an object which is identical to the parameter, except of an additional slot, the results.
-This makes it possible to run the simulation again from the results from a previous simulation.
+After the paramter object has been defined, it can be used in the `run_simulation()` function.
+The function returns an object which is identical to the parameter, except of an additional slot containing the results.
+This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results from.
 
-## Find a Steady State of a model
+## Find a Steady State of the model
 There are two methods for finding steady states.
@@
 We hereafter term this "the temporal method".
-This is implemented in the function `run_temporal_ssfind()` while the parameter object is created by the function `new_temporal_ssfind_parameter()()`.'
+This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .
 
-For a more detailed walk-through of these two approaches and explanation please see the [User Guide](LINK NEEDED).
+For a more detailed walk-through of these two approaches and explanation please see the [User Guide](@LINK_NEEDED).
 
@@
 
-{>> Have to add these aspects here <<}
+The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function.
+This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
+An example plot resulting from this function is shown in Figure 1.
+
+![](./figures/fig1_plot_dynamics.png)
 
@@
 Consequently, any more substantial, are likely to need a change in the source code.
-Nevertheless, the model is structured in a way which builds on a udular structuere, so that e.g. the event defuinition can easily be changed.
+
+Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.
+the event definition can easily be changed.
+or other aspects can be adjusted.
+All values in the parameter object can be changed.
 

### 2022-03-15 inst/manuscript/manuscript_microxanox.Rmd
Update manuscript_microxanox.Rmd
Dear Rainer, I made some suggestions. Please take a look and we should then briefly discuss. Cheers, Owen

@@
 
-## The ODEs
-The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
-This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
-
-For an example how these can be used see [@REF_NEEDED] or the vignettes acompanying this package [@LINK_NEEDED].
-
 # Availability and Accessability of the Package
@@
 
-# The Package
+# Using the package
+
+## Documentation
+
 The package contains the full documentetion of the functions, which includes
@@
 
+## The simulation framework
+
+Here make an overview of how conditions and parameters are specified, stored, and results stored. I think there is text in the user guide vignette that can be used (duplicated) here.
+
 We will discuss the general structure and functiuonalit of the package without going into to mu=ch detail.
 
-The simulations can be split into different levels:
+The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
+This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
 
-## The individuel simulation
+In the following sections we describe the general goals of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.
+
+## Running one simulation
+
 The individual simulation (`run_simulation()` function) is the working horse in this package.
@@
 ## Find a Steady State of the model
+
 There are two methods for finding steady states.
@@
 
-## Analysis and Graphing
+## Visualisation
 
@@
 
+**Please include figures so they can have captions and be cross referenced, as in the Supplement.**
+
 ![](./figures/fig1_plot_dynamics.png)
@@
 # Use cases
+
+The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.
+
+## Regime shifts during temporal environmental change
+
+Take from the user guide.
+
+## The extent of hysteresis depends on community composition
+
+Take from the partial reproduction.
+
+## Effects of functional diversity on regime shifts
+
 This package has been used in the production of the paper [@REF_NEEDED].

### 2022-03-16 inst/manuscript/manuscript_microxanox.Rmd
Some Changes to Manuscript

@@
 output:
-  html_document:
+  bookdown::html_document2:
     toc: true
-    toc_float: true
+    toc_float:
+      collapsed: true
+      smooth_scroll: true
   pdf_document:
@@
 
-Here make an overview of how conditions and parameters are specified, stored, and results stored. I think there is text in the user guide vignette that can be used (duplicated) here.
+The framework used for running simulations aims at reproducibility of the results and id=s therefore build on the following points:
 
-We will discuss the general structure and functiuonalit of the package without going into to mu=ch detail.
+1.
+all parameter needed to run a simulation or find a stable state are contained in a single para,meter object.
+This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`.
+Which one of the `new_..._parameter()` functions has to be used, will be discussed imn the section  \@ref(runsim)  and in the [User Guide](LINK_NEEDED).
+2. The function call `run_...(parameter)` will than run the simulation using the parameter as defined in the object `parameter`.
+3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additionla slot named `results` which contains the results of the run
+4. As this return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.
+
+The fact that the results op=bject contains all perameter needed to run the s=imulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.
+
+A typical simulatiopn would look as followed:
+
+```{r, eval = FALSE}
+parameter <- new_runsim_parameter()
+## manually setting certain parameter
+result <- run_simulatrion(parameter)
+saveRDS(result, "sim1.rds")
+## Do other stuff
+parameter <- loadRDS("sim1.rds")
+## change some parameter
+result <- run_simulatrion(parameter)
+saveRDS(result, "sim2,rds")
+```
+
+## Running a simulation {#runsim}
+
+We will now discuss the general structure and functiuonalit of the package without going into to much detail.
 
@@
 
-## Running one simulation
+### Running one simulation
 
@@
 
-## Find a Steady State of the model
+### Find a Steady State of the model
 
@@
 
-### Replication Method
+#### Replication Method
 The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
 
-### Temporal Method
+#### Temporal Method
 The temporal method involves two simulations for a particular system configuration (parameter set).
@@
 This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in Figure 1.
+An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).
 
-**Please include figures so they can have captions and be cross referenced, as in the Supplement.**
-
-![](./figures/fig1_plot_dynamics.png)
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function `plot_dynamics()`.
+Details can be found in the ["User Guide" section "Three strains per functional group"](LINK_NEEDED_).', fig.align="center", out.width = 500 }
+knitr::include_graphics("figures/fig1_plot_dynamics.png")
+```
 
@@
 
-Take from the user guide.
+{>>Take from the user guide.<<}
 
@@
 
-Take from the partial reproduction.
+{>>Take from the partial reproduction.<<}
 

### 2022-03-21 inst/manuscript/manuscript_microxanox.Rmd
Next edits

--- 2022-03-15 inst/manuscript/manuscript_microxanox.Rmd
@@
 ---
-title: "m0.0.1: Microxanox - ..."
+title: "Microxanox - an R package to simulate ..."
 author:
@@
   4. identify the conclusions and the wider implications.
-journal: 'Ecology and Evolution - Applications'
 bibliography: manuscript_bibliography.bib
 date: "`r Sys.time()`"
+
+# RMarkdown output settings
 output:
-  html_document:
+  bookdown::html_document2:
     toc: true
-    toc_float: true
-  pdf_document:
+    toc_float:
+      collapsed: true
+      smooth_scroll: true
+  bookdown::pdf_document2:
     keep_tex: yes
-header-includes:
-- \usepackage{lineno}
-- \linenumbers
-- \usepackage{setspace}
-- \doublespacing
-- \usepackage{fancyhdr}
-- \fancyhead{}
-- \fancyfoot{}
-- \fancyhead[CO,CE]{Metadata made easy - the dmdScheme approach}
-- \pagestyle{fancy}
 
+# qarto output settings
+format:
+  html:
+    self-contained: true
+    toc: true
+    toc-depth: 3
+    link-external-icon: true
+    title-block-style: default
+
+# filters:
+#   - type: json
+#     path: criticpandoc.py
+
+# Citation settings    
 nocite: '@Bush2017'
@@
 link-citations: yes
+
 csl: methods-in-ecology-and-evolution.csl
@@
 
-The package will be updated based on further research activities.
-It will be tested continuously for the last recent, current and devel version of R.
-{>>If we wnat that, we havt o include some, even rudimentary, GITHUB tests.<<}
+The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
+{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
+{>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}
 
@@
 
-# Using the package
+# The R Package
 
@@
 
-## The simulation framework
+## The modelling framework
 
-Here make an overview of how conditions and parameters are specified, stored, and results stored. I think there is text in the user guide vignette that can be used (duplicated) here.
+The framework used when writing this package aims at reproducibility of the results. It builds on the folowing main considerations:
 
-We will discuss the general structure and functiuonalit of the package without going into to mu=ch detail.
+1.
+all parameter needed to run a simulation or find a stable state are contained in a single parameter object.
+This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`.
+Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim)  and in the [User Guide](LINK_NEEDED).
+2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
+3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additionl slot named `results` which contains the results of the run
+4. As this return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.
+
+The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.
+
+A typical simulation would look as followed:
+
+```{r, eval = FALSE}
+## Create the parameter
+parameter <- new_runsim_parameter()
+# manually setting certain parameter
+
+## Run the simulation and save the result
+result <- run_simulatrion(parameter)
+saveRDS(result, "sim1.rds")
+
+## Do other stuff, e.g. plotting
+
+## Load results, change some parameter, and rerun the simulation and save the result
+parameter <- loadRDS("sim1.rds")
+# change some parameter
+result <- run_simulatrion(parameter)
+saveRDS(result, "sim2,rds")
+```
+
+## Using the package {#runsim}
+
+We will now discuss the general structure and functiuonality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
 
@@
 
-In the following sections we describe the general goals of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.
+In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.
 
-## Running one simulation
+### Running one simulation
 
@@
 
-## Find a Steady State of the model
+The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function.
+This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
+An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).
+
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function `plot_dynamics()`$`.
+Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
+knitr::include_graphics("figures/fig1_plot_dynamics.png")
+```
+
+### Find a Steady State of the model
 
@@
 Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.
-We hereafter term this "the temporal method".
+
 This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .
@@
 
-## Visualisation
-
-The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function.
-This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in Figure 1.
-
-**Please include figures so they can have captions and be cross referenced, as in the Supplement.**
-
-![](./figures/fig1_plot_dynamics.png)
-
 # Use cases
 
-The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.
+The first two use cases come from the User Guide and the Partial Reproduction Vignettes.
+The third comes from a research article that relied on the package.
+All of these use cases can be expanded to larger numbers of strains per functional group.
 
@@
 
-Take from the user guide.
+In the [User Guide](LINK_NEEDED) we ued a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to analyse as an example the determination of the stable states during temporal environmental chan ge (the oxygen diffusivity).
+From these simulatiuons, we extracted measures of nonlinearity and hysteresis.
 
@@
 
-Take from the partial reproduction.
+One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED).
+All aspects in the paper could be reproduced.
 
@@
 
-This package has been used in the production of the paper [@REF_NEEDED].
-It can be used for similar studies in this study systems.
-It is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs.
-Consequently, any more substantial, are likely to need a change in the source code.
+This model (as part of the package) has been used in the production of the paper [@REF_NEEDED]. It can be used for similar studies in this study systems.
+
+{>>More details from the paper.<<}
+
+## General usability and flexibility of the model
+The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
+Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.
 
@@
 or other aspects can be adjusted.
-All values in the parameter object can be changed.
+All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
 
@@
 
+{>>Dependant on journal<<}
+
 # References

### 2022-03-21 inst/manuscript/manuscript_microxanox.tex
Next edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Microxanox - an R package to simulate \ldots{}},
  pdfkeywords={metadata quality; data curation; archival; long term storage; R package;},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Microxanox - an R package to simulate \ldots{}}
\author{true \and true}
\date{2022-03-21 14:42:18}

\begin{document}
\maketitle
\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  set the context for and purpose of the work;
\item
  indicate the approach and methods;
\item
  outline the main results;
\item
  identify the conclusions and the wider implications.
\end{enumerate}
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\{\textgreater\textgreater{} Biological justification and background - one paragraph \textless\textless\}

The microxanox package is a package for simulating a three functional group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced sulfur, \emph{SO}: oxidized sulfur). It includes feddback between biogeochemical processes and is based on (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) (See (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) for a detailed discussion of the model).

The aims of the \texttt{microxanox} package are twofold. Firstly, the package aims at reproducing the results shown by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}), which is acomplished in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}. Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).

For this, we extended the model and added functionality for:

\begin{itemize}
\tightlist
\item
  Multiple strains (effectively unlimited) per functional group.
\item
  Adding temporally varying oxygen diffusivity.
\item
  Adding random noise in substrate concentrations.
\item
  Including immigration.
\item
  Setting minimum population abundances.
\end{itemize}

In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.

\hypertarget{the-model}{%
\section{The Model}\label{the-model}}

The model is an implementation of the model described by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.

We use the notation from the \href{LINK\%20NEEDED}{Table S1} of the Supplementary information of (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017})), and not the notation in the main manuscript of (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). In the main manuscript the notation is simplified to \texttt{y\_S\_PB} (\(y_{S,PB}\)) (which is \texttt{y\_SR\_PB} (\(y_{SR,PB}\)) in Table S1) and \texttt{y\_S\_SB} (\(y_{S,SB}\)) (which is y\_SO\_SB (\(y_{SO,SB}\)) in Table S1). We also use \texttt{Pr\_cB} (\(Pr_B\)) instead of \texttt{P\_CB} (\(P_{CB}\)).

All parameter and variables follow these base dimensions:

\begin{itemize}
\tightlist
\item
  Time: \texttt{hours}
\item
  Volume: \texttt{litres}
\item
  Substrate quantity: \texttt{micromoles}
\item
  Organism quantity: \texttt{cells}
\end{itemize}

\{\textgreater\textgreater{} Add a more detailed description of the model and the additions \textless\textless\}

The ODEs are solved using the ``General Solver for Ordinary Differential Equations'' of the \texttt{deSolve} package (\protect\hyperlink{ref-Soetaert2010}{Soetaert, Petzoldt, \& Setzer, 2010}).

\hypertarget{availability-and-accessability-of-the-package}{%
\section{Availability and Accessability of the Package}\label{availability-and-accessability-of-the-package}}

The package is available in a \href{https://github.com/UZH-PEG/microxanox/}{github repo/}. The newest version is also available under the \href{https://DOI}{DOI} or via RUniverse (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). This manuscript describes version 1.0 which is available under \href{https://DOI}{DOI} or installable within R by using \texttt{remotes::install\_github("UZH-PEG",\ ref\ =\ "v1.0")}.

The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R.
\{\textgreater\textgreater If we want that, we have to include some, even rudimentary, GITHUB tests.\textless\textless\}
\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk about the package paper.\textless\textless\}

Issues, questions and suggestions should be communicated through the issue tracker at \url{https://github.com/UZH-PEG/microxanox/issues}

\hypertarget{the-r-package}{%
\section{The R Package}\label{the-r-package}}

\hypertarget{documentation}{%
\subsection{Documentation}\label{documentation}}

The package contains the full documentetion of the functions, which includes

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Standard R help for each function (\texttt{?COMMAND} in R)
\item
  Two vignettes acompanying the package, also available under RUniverse \href{@LINK_NEEDED}{Vignette 1} and \href{@LINK_NEEDED}{Vignette 2}
\item
  This paper as a third vignette, also available under RUniverse \href{@LINK_NEEDED}{Vignette 3}
\end{enumerate}

\hypertarget{the-modelling-framework}{%
\subsection{The modelling framework}\label{the-modelling-framework}}

The framework used when writing this package aims at reproducibility of the results. It builds on the folowing main considerations:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  all parameter needed to run a simulation or find a stable state are contained in a single parameter object. This object is created by using the functions \texttt{new\_...\_parameter()}, \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}. Which one of the \texttt{new\_...\_parameter()} functions has to be used when, will be discussed in the section \ref{runsim} and in the \href{LINK_NEEDED}{User Guide}.
\item
  The function call \texttt{run\_...(parameter)} will run the simulation using the parameter as defined in the object \texttt{parameter}.
\item
  The return value of the \texttt{run\_...(parameter)} function is identical to the parameter object plus an additionl slot named \texttt{results} which contains the results of the run
\item
  As this return value contains all parameter, it is possible to re-run the simulatuion by simply running \texttt{run\_...(result)}.
\end{enumerate}

The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.

A typical simulation would look as followed:

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# Create the parameter}
\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{new\_runsim\_parameter}\NormalTok{()}
\CommentTok{\# manually setting certain parameter}

\DocumentationTok{\#\# Run the simulation and save the result}
\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulatrion}\NormalTok{(parameter)}
\FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim1.rds"}\NormalTok{)}

\DocumentationTok{\#\# Do other stuff, e.g. plotting}

\DocumentationTok{\#\# Load results, change some parameter, and rerun the simulation and save the result}
\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{loadRDS}\NormalTok{(}\StringTok{"sim1.rds"}\NormalTok{)}
\CommentTok{\# change some parameter}
\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulatrion}\NormalTok{(parameter)}
\FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim2,rds"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{runsim}{%
\subsection{Using the package}\label{runsim}}

We will now discuss the general structure and functiuonality of the package without going into to much detail. A more detailed discussion can be found in the \href{LINK_NEEDED}{User Guide}.

The ODEs for the rates of change are specified in the function \texttt{bushplus\_dynamic\_model()}. This augmented version of the model published in (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.

In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.

\hypertarget{running-one-simulation}{%
\subsubsection{Running one simulation}\label{running-one-simulation}}

The individual simulation (\texttt{run\_simulation()} function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one parameter - an object as created by the function \texttt{new\_runsim\_parameter()}. One parameter of this object is the \texttt{strain\_parameter} which can be created by the function \texttt{new\_strain\_parameter()}. For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at \href{@LINK_NEEDED}{User Guide}

After the paramter object has been defined, it can be used in the \texttt{run\_simulation()} function. The function returns an object which is identical to the parameter, except of an additional slot containing the results. This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the \texttt{run\_simulation()} function to run the simulation again from the parameter used to generate the results from.

The function \texttt{plot\_dynamics()} plots a single simulation run, as returned from the \texttt{run\_simulation()} function. This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=500px]{figures/fig1_plot_dynamics}

\}

\textbackslash caption\{Plot of results of a simulation run using the function \texttt{plot\_dynamics()}\$`. Details can be found in the ``User Guide'' section ``Three strains per functional group''.\}\label{fig:plot-dynamics}
\textbackslash end\{figure\}

\hypertarget{find-a-steady-state-of-the-model}{%
\subsubsection{Find a Steady State of the model}\label{find-a-steady-state-of-the-model}}

There are two methods for finding steady states. The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the \emph{Replication method}). The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the \emph{Temporal method}).

\hypertarget{replication-method}{%
\paragraph{Replication Method}\label{replication-method}}

The replication method is implementad in the function \texttt{run\_replication\_ssfind()} which takes a parameter object as returned by the function \texttt{new\_replication\_ssfind\_parameter()} and the number of cores for multithreading the simulation.

\hypertarget{temporal-method}{%
\paragraph{Temporal Method}\label{temporal-method}}

The temporal method involves two simulations for a particular system configuration (parameter set). In one simulation the oxygen diffusivity is \emph{increased} in a step-wise fashion. In the other it is \emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set. Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.

This is implemented in the function \texttt{run\_temporal\_ssfind()}, which takes a parameter object as created by the function \texttt{new\_temporal\_ssfind\_parameter()} and a number indicating the .

For a more detailed walk-through of these two approaches and explanation please see the \href{@LINK_NEEDED}{User Guide}.

\hypertarget{extract-stability-measures}{%
\subsection{Extract Stability Measures}\label{extract-stability-measures}}

From the raw results returned by these \texttt{run\_...()} functions, the stability measures can be extracted by using the function \texttt{get\_stability\_measures()}. These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.

\hypertarget{use-cases}{%
\section{Use cases}\label{use-cases}}

The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.

\hypertarget{regime-shifts-during-temporal-environmental-change}{%
\subsection{Regime shifts during temporal environmental change}\label{regime-shifts-during-temporal-environmental-change}}

\{\textgreater\textgreater Take from the user guide.\textless\textless\}

\hypertarget{the-extent-of-hysteresis-depends-on-community-composition}{%
\subsection{The extent of hysteresis depends on community composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}

\{\textgreater\textgreater Take from the partial reproduction.\textless\textless\}

\hypertarget{effects-of-functional-diversity-on-regime-shifts}{%
\subsection{Effects of functional diversity on regime shifts}\label{effects-of-functional-diversity-on-regime-shifts}}

This model (as part of the package) has been used in the production of the paper (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). It can be used for similar studies in this study systems.

\{\textgreater\textgreater More details from the paper.\textless\textless\}

\hypertarget{general-usability-and-flexibility-of-the-model}{%
\subsection{General usability and flexibility of the model}\label{general-usability-and-flexibility-of-the-model}}

The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) and to extend it to our needs (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.

Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.~the event definition can easily be changed. or other aspects can be adjusted. All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

\{\textgreater\textgreater Dependant on journal\textless\textless\}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
Bush, T., Diao, M., Allen, R. J., Sinnige, R., Muyzer, G., \& Huisman, J. (2017). Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics. \emph{Nature Communications}, \emph{8}(1), 789. doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}

\leavevmode\vadjust pre{\hypertarget{ref-REF_NEEDED}{}}%
REF NEEDED, A. (2222). {REFERNECE NEEDED}. \emph{Journal of Missing References}.

\leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
Soetaert, K., Petzoldt, T., \& Setzer, R. W. (2010). Solving differential equations in {R}: {Package deSolve}. \emph{Journal of Statistical Software}, \emph{33}(9), 1--25. doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}

\end{CSLReferences}

\end{document}

### 2022-03-22 inst/manuscript/manuscript_microxanox.Rmd
Next step in manuscript

@@
 bibliography: manuscript_bibliography.bib
-date: "`r Sys.time()`"
+
+date: last-modified 
+date-format: "DD MMM YYYY, HH:mm"
 
@@
     title-block-style: default
-
-# filters:
-#   - type: json
-#     path: criticpandoc.py
+  pdf:
+    toc: true
+    toc-depth: 3
+    lof: true
+    include-in-header:
+      - headers.tex
 
@@
 link-citations: yes
-
 csl: methods-in-ecology-and-evolution.csl
@@
 
-{>> Biological justification and background - one paragraph <<}
+{>>Biological justification and background - one paragraph<<}
+{>>Rewrite the abstract for this<<}
+
+> Abstract Romana of paper: The role of functional diversity in ecosystem regime shifts
+
+> Biodiversity can increase the resilience of ecosystems facing environmental change.
+However, we have limited understanding if this holds true for ecosystems that respond to gradual environmental change with abrupt regime shifts.
+Here we investigated if functional diversity influences the resilience of ecosystems that abruptly transition between alternative ecosystem states.
+As a case study, we used a mathematical model developed by Bush et al.
+(2017) which describes anoxic-oxic regime shifts of aquatic ecosystems mediated by cyanobacteria, sulfate-reducing bacteria, and phototrophic sulfur bacteria.
+We extended this model so as to include functional diversity in one, two, or all three functional groups.
+Specifically, we manipulated the amount of variation in environmental tolerance and in maximum growth rate, and explored how trait variation influenced anoxic-oxic regime shifts along a gradient of oxygen influx.
+We found that greater diversity of either cyanobacteria or sulfate-reducing bacteria increased the resilience of the ecosystem state that these groups dominated, while greater diversity of phototrophic sulfur bacteria reduced the resilience of the state they dominated.
+We also found that simultaneous variation in multiple functional groups reduced, reversed, or did not change the diversity effect of individual functional groups.
+Collectively, our findings highlight the importance of considering multiple interacting functional groups when predicting the effect of functional diversity on ecosystem resilience
 
@@
 {>> Add a more detailed description of the model and the additions <<}
+{>>TODO<<}
 
@@
 1. Standard R help for each function (`?COMMAND` in R)
-2. Two vignettes acompanying the package, also available under RUniverse [Vignette 1](@LINK_NEEDED) and [Vignette 2](@LINK_NEEDED)
+2. Two vignettes acompanying the package:
+  - A **User Guide**, also available on the R-Universe [Vignette 1](@LINK_NEEDED).
+This gives an introduction of the package and its functions and provides a walk-through of the process of running, visualizing and extra
+   and [Vignette 2](@LINK_NEEDED)
 3. This paper as a third vignette, also available under RUniverse [Vignette 3](@LINK_NEEDED)
@@
 {>>More details from the paper.<<}
+{>>Romana will provide two or three sentences<<}
 

### 2022-03-23 inst/manuscript/manuscript_microxanox.Rmd
Nearly final draft

@@
 
-# Introduction
-
-{>>Biological justification and background - one paragraph<<}
-{>>Rewrite the abstract for this<<}
-
+<!-- 
 > Abstract Romana of paper: The role of functional diversity in ecosystem regime shifts
@@
 However, we have limited understanding if this holds true for ecosystems that respond to gradual environmental change with abrupt regime shifts.
-Here we investigated if functional diversity influences the resilience of ecosystems that abruptly transition between alternative ecosystem states.
+
+> Here we investigated if functional diversity influences the resilience of ecosystems that abruptly transition between alternative ecosystem states.
 As a case study, we used a mathematical model developed by Bush et al.
@@
 Collectively, our findings highlight the importance of considering multiple interacting functional groups when predicting the effect of functional diversity on ecosystem resilience
+ -->
+
+# Introduction
+
+{>>Biological justification and background - one paragraph<<}   
+{>>Based on abstract from Romana<<}
+
+Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable.
+One example where the gradual change of a single environmental causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
+This system hass been investigated by @Bush2017.
+We wanted to take this investigation one step further, and look at the role of an increased biodiversity plays in these dynamics [@REF_NEEDED].
+For this purpose, we developed this package.
 
@@
 
-This model (as part of the package) has been used in the production of the paper [@REF_NEEDED]. It can be used for similar studies in this study systems.
+As discussed in the paper [@REF_NEEDED], the role biodiversity plays in abrupt regime shifts based on gradual changing environmental parameter is not well understood.
+This model (as part of the package) has been used to investigate these dynamics and the results are available in @REF_NEEDED.
 

### 2022-03-27 inst/manuscript/manuscript_microxanox.Rmd
Update manuscript_microxanox.Rmd

@@
 ---
-title: "Microxanox - an R package to simulate ..."
+title: "Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
 author:
@@
 
+## General usability and flexibility of the model
+The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
+Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.
+
+Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.
+the event definition can easily be changed.
+or other aspects can be adjusted.
+All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
+
 # The Model
@@
 
-# Availability and Accessability of the Package
+# Availability and Accessability of the package
+
 The package is available in a [github repo/](https://github.com/UZH-PEG/microxanox/).
@@
 
-# The R Package
+# Package description
 
@@
 
-## The modelling framework
+## Reproducibility
 
@@
 In this function, the ODEs are solved.
-The function needs only one parameter - an object as created by the function `new_runsim_parameter()`.
+The function needs only one argument - an object as created by the function `new_runsim_parameter()`.
 One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
@@
 
-After the paramter object has been defined, it can be used in the `run_simulation()` function.
+After the parameter object has been defined, it can be used in the `run_simulation()` function.
 The function returns an object which is identical to the parameter, except of an additional slot containing the results.
@@
 
-### Find a Steady State of the model
+### Finding a Steady State of the model
 
@@
 
-## Extract Stability Measures
+## Extracting Stability Measures
 From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
@@
 
-## General usability and flexibility of the model
-The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
-Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.
-
-Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.
-the event definition can easily be changed.
-or other aspects can be adjusted.
-All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
-
 # Conclusions

### 2022-04-11 inst/manuscript/osp-template.tex
Create osp-template.tex

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

\documentclass[preprint,12pt, a4paper]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{hyperref}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%\usepackage{lineno}

\journal{Software Impacts}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Title/Name of your software}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{A. Author}

\address{Your institute, some address}

\begin{abstract}
%% Text of abstract 
Ca. 100 words

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
keyword 1 \sep keyword 2 \sep keyword 3

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%\linenumbers

\noindent
\textbf{Main text}\\
Maximum 3 pages (excluding metadata, tables, figures, references)\\

\noindent

The description of your software  shall include:
\begin{itemize}
\item A short description of the high-level functionality and purpose of the software for a diverse, non-specialist audience
\item An Impact overview that illustrates the purpose of the software and its achieved results:
\begin{itemize}
\item[-] Indicate in what way, and to what extent, the pursuit of existing research questions is improved. 
\item[-] Indicate in what way new research questions can be pursued because of the software.
\item[-] Indicate in what way the software has changed the daily practice of its users.
\item[-] Indicate how widespread the use of the software is within and outside the intended user group.
\item[-] Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).
\end{itemize}
\item Mentions (if applicable) of any ongoing research projects using the software. 
\item A list of all scholarly publications enabled by the software.
\end{itemize}
\section*{Acknowledgements}
\label{}

Optionally thank people and institutes you need to acknowledge. 
\section*{References}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References:
%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
%%  \bibliographystyle{elsarticle-num} 
%%  \bibliography{<your bibdatabase>}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

\bibitem{}

\end{thebibliography}

\section*{Illustrative Examples}
Optional : you may include one explanatory  video that will appear next to your article, in the right hand side panel. (Please upload any video as a single supplementary file with your article. Only one MP4 formatted, with 50MB maximum size, video is possible per article. Recommended video dimensions are 640 x 480 at a maximum of 30 frames / second. Prior to submission please test and validate your .mp4 file at  \url{http://elsevier-apps.sciverse.com/GadgetVideoPodcastPlayerWeb/verification} . This tool will display your video exactly in the same way as it will appear on ScienceDirect. )

\section*{Required Metadata}
\label{}

\section*{Current code version}
\label{}

Ancillary data table required for subversion of the codebase. Kindly replace examples in right column with the correct information about your current code, and leave the left column as it is.

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this column} \\
\hline
C1 & Current code version & For example v42 \\
\hline
C2 & Permanent link to code/repository used for this code version & For example: $https://github.com/mozart/mozart2$ \\
\hline
C3  & Permanent link to Reproducible Capsule & \\
\hline
C4 & Legal Code License   & List one of the approved licenses \\
\hline
C5 & Code versioning system used & For example svn, git, mercurial, etc. put none if none \\
\hline
C6 & Software code languages, tools, and services used & For example C++, python, r, MPI, OpenCL, etc. \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & \\
\hline
C8 & If available Link to developer documentation/manual & For example: $http://mozart.github.io/documentation/$ \\
\hline
C9 & Support email for questions & \\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{} 
\end{table}

\section*{Current executable software version}
\label{}

Ancillary data table required for sub version of the executable software: (x.1, x.2 etc.) kindly replace examples in right column with the correct information about your executables, and leave the left column as it is.

\begin{table}[!h]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{(Executable) software metadata description} & \textbf{Please fill in this column} \\
\hline
S1 & Current software version & For example 1.1, 2.4 etc. \\
\hline
S2 & Permanent link to executables of this version  & For example: $https://github.com/combogenomics/$ $DuctApe/releases/tag/DuctApe-0.16.4$ \\
\hline
S3  & Permanent link to Reproducible Capsule & \\
\hline
S4 & Legal Software License & List one of the approved licenses \\
\hline
S5 & Computing platforms/Operating Systems & For example Android, BSD, iOS, Linux, OS X, Microsoft Windows, Unix-like , IBM z/OS, distributed/web based etc. \\
\hline
S6 & Installation requirements \& dependencies & \\
\hline
S7 & If available, link to user manual - if formally published include a reference to the publication in the reference list & For example: $http://mozart.github.io/documentation/$ \\
\hline
S8 & Support email for questions & \\
\hline
\end{tabular}
\caption{Software metadata (optional)}
\label{} 
\end{table}

\end{document}
\endinput
%%
%% End of file `SoftwareX_article_template.tex'.

### 2022-04-11 inst/manuscript/manuscript_microxanox.Rmd
Use of elsevier article class and beginning of adaptation to SoftwareX Article Structure

@@
     190, 8057 Zurich
+footnote:
+  - code: 1
+    text: "Corresponding Author"
+  - code: 2
+    text: "Equal contribution"  
 keywords: metadata quality; data curation; archival; long term storage; R package;
 abstract: | 
-  1. set the context for and purpose of the work;
-  2. indicate the approach and methods;
-  3. outline the main results;
-  4. identify the conclusions and the wider implications.
+  Ca. 100 words.
 bibliography: manuscript_bibliography.bib
+journal: "SoftwareX"
+date: "`r Sys.Date()`"
 
-date: last-modified 
-date-format: "DD MMM YYYY, HH:mm"
-
-# RMarkdown output settings
+#linenumbers: true
+#numbersections: true
+csl: elsevier-harvard.csl
 output: 
-  bookdown::html_document2:
-    toc: true 
-    toc_float:
-      collapsed: true
-      smooth_scroll: true
-  bookdown::pdf_document2:
-    keep_tex: yes
-
-# qarto output settings
-format:
-  html:
-    self-contained: true
-    toc: true
-    toc-depth: 3
-    link-external-icon: true
-    title-block-style: default
-  pdf:
-    toc: true
-    toc-depth: 3
-    lof: true
-    include-in-header:
-      - headers.tex
+  rticles::elsevier_article:
+    includes: 
+      in_header: headers.tex
 
@@
 link-citations: yes
-csl: methods-in-ecology-and-evolution.csl
 ---
@@
 
-# Introduction
+# Required Metadata
+## Current code version
+Ancillary data table required for subversion of the codebase.
 
-{>>Biological justification and background - one paragraph<<}   
-{>>Based on abstract from Romana<<}
+| **Nr.** | **Code metadata description** | **Please fill in this column** |
+|---|---|---|
+| C1 | Current code version | v0.9.0 |
+| C2 | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) |
+| C3 | Code Ocean compute capsule |  |
+| C4 | Legal Code License   | CC BY 4.0 |
+| C5 | Code versioning system used |git |
+| C6 | Software code languages, tools, and services used | R | 
+| C7 | Compilation requirements, operating environments | R (>= 4.1.0), ADD PACKAGES | 
+| C8 | If available Link to developer documentation/manual |  |
+| C9 | Support email for questions | Rainer.Krug@uzh.ch; Rainer@krugs.de |
 
+# Motivation and significance
 Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable.
@@
 
+# Software description
+Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.
+
+## Software Architecture
+Give a short overview of the overall software architecture; provide a pic- torial component overview or similar (if possible). If necessary provide im- plementation details.
+
+## Software Functionalities
+Present the major functionalities of the software.
+
+# Illustrative Examples
+Provide at least one illustrative example to demonstrate the major functions.
+Optional: you may include one explanatory video that will appear next to your article, in the right hand side panel.
+
+# Impact
+This is the main section of the article and the reviewers weight the description here appropriately
+Indicate in what way new research questions can be pursued as a result of the software (if any).
+Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
+Indicate in what way the software has changed the daily practice of its users (if so).
+Indicate how widespread the use of the software is within and outside the intended user group.
+Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).
+
+# Conclusions
+Set out the conclusion of this original software publication.
+
+# Conflict of Interest
+We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
+
+# Acknowledgements
+Optionally thank people and institutes you need to acknowledge.
+
+# THIS NEEDS TO BE TRANSFERED IN THE FORMAT ABOVE
+
 # The Model
@@
 
-{>> Add a more detailed description of the model and the additions <<}
-{>>TODO<<}
+<span style="color:red">{>> Add a more detailed description of the model and the additions <<}
+{>>TODO<<}</span>
 
@@
 The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
-{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
+<span style="color:green">{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
 {>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}
+</span>
 
@@
 
-```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function `plot_dynamics()`$`.
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function $plot_dynamics()$.
 Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
-knitr::include_graphics("figures/fig1_plot_dynamics.png")
+knitr::include_graphics("figures/ug_three_strains_dynamics.png")
 ```
@@
 
-The first two use cases come from the User Guide and the Partial Reproduction Vignettes.
-The third comes from a research article that relied on the package.
-All of these use cases can be expanded to larger numbers of strains per functional group.
+The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes.
+The third is taken from the @REF_NEEDED for which this R package was designed.
+All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.
 
@@
 
-In the [User Guide](LINK_NEEDED) we ued a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to analyse as an example the determination of the stable states during temporal environmental chan ge (the oxygen diffusivity).
+In the [User Guide](LINK_NEEDED) we used a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to determine as an example the stable states during temporal environmental changes (the oxygen diffusivity).
 From these simulatiuons, we extracted measures of nonlinearity and hysteresis.
+See Fig \@ref(fig:plot-dynamics) as an example plot of the simulations.
+
+```{r uc1_stable_state, echo = FALSE,  include=TRUE, fig.cap='Plot of the stable states of the simulation runs under different oxygen diffusivity.
+The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities.
+Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
+knitr::include_graphics("figures/ug_three_strains_stable_state.png")
+```
 
@@
 One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED).
-All aspects in the paper could be reproduced.
+All aspects in the paper could be reproduced and are shown in the vignette.
+
+```{r user_guide_hysteresis, echo = FALSE,  include=TRUE, fig.cap='Hysteresis of all assessed combinations of variability.', fig.align="center", out.width = 500 }
+knitr::include_graphics("figures/user_guide_hysteresis.png")
+```
 
@@
 
-{>>More details from the paper.<<}  
-{>>Romana will provide two or three sentences<<}
+<span style="color:green">{>>More details from the paper.<<}  
+{>>Romana will provide two or three sentences<<}</span>
 

### 2022-04-11 inst/manuscript/manuscript_microxanox.tex
Use of elsevier article class and beginning of adaptation to SoftwareX Article Structure

@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-]{article}
-\usepackage{amsmath,amssymb}
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+
+  \journal{SoftwareX} % Sets Journal name
+
+\usepackage{lineno} % add
+
+\usepackage{graphicx}
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
   \fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Microxanox - an R package to simulate \ldots{}},
-  pdfkeywords={metadata quality; data curation; archival; long term storage; R package;},
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Microxanox - an R package for simulating an MICRobial ecosystem that can occupy OXic or ANOXic states.},
             colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
-\usepackage[margin=1in]{geometry}
+            urlcolor=blue,
+            linkcolor=blue,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+
+\setcounter{secnumdepth}{0}
+% Pandoc toggle for numbering sections (defaults to be off)
+\setcounter{secnumdepth}{0}
+
+% Pandoc syntax highlighting
 \usepackage{color}
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+% From pandoc table feature
 \usepackage{longtable,booktabs,array}
@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+
+% Pandoc citation processing
 \newlength{\cslhangindent}
@@
 \setlength{\cslentryspacingunit}{\parskip}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
 \newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 
-\title{Microxanox - an R package to simulate \ldots{}}
-\author{true \and true}
-\date{2022-03-21 14:42:18}
+%% \documentclass[preprint,12pt, a4paper]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
+
+\usepackage{float}
+\restylefloat{table}
+
+\journal{SoftwareX}
+
+%\usepackage{lineno}
+%\linenumbers
+%\usepackage{setspace}
+%\doublespacing
+%\usepackage{fancyhdr}
+%\fancyhead{}
+%\fancyfoot{}
+%\fancyhead[CO,CE]{Microxanox R package}
+%\pagestyle{fancy}
+%
 
 \begin{document}
-\maketitle
+
+\begin{frontmatter}
+
+  \title{Microxanox - an R package for simulating an \(MICR\)obial
+ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
+    \author[University of Zürich]{Rainer M Krug\corref{Corresponding
+Author}}
+   \ead{Rainer.Krug@uzh.ch \& Rainer@krugs.de} 
+    \author[University of Zürich]{Owen L. Petchey}
+   \ead{Owen.Petchey@ieu.uzh.ch} 
+      \address[University of Zürich]{Department of Evolutionary Biology
+and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
+      \cortext[1]{Corresponding Author}
+    \cortext[2]{Equal contribution}
+
   \begin{abstract}
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  set the context for and purpose of the work;
-\item
-  indicate the approach and methods;
-\item
-  outline the main results;
-\item
-  identify the conclusions and the wider implications.
-\end{enumerate}
+  Ca. 100 words.
   \end{abstract}
+   \begin{keyword} metadata quality; data curation; archival; long term
+storage; R package;\end{keyword}
+ \end{frontmatter}
 
-{
-\hypersetup{linkcolor=}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
+\hypertarget{required-metadata}{%
+\section{Required Metadata}\label{required-metadata}}
+
+\hypertarget{current-code-version}{%
+\subsection{Current code version}\label{current-code-version}}
+
+Ancillary data table required for subversion of the codebase.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
+\toprule
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Nr.}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Code metadata description}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Please fill in this column}
+\end{minipage} \\
+\midrule
+\endhead
+C1 & Current code version & For example v42 \\
+C2 & Permanent link to code/repository used for this code version &
+\url{https://github.com/UZH-PEG/microxanox} \\
+C3 & Code Ocean compute capsule & \\
+C4 & Legal Code License & CC BY 4.0 \\
+C5 & Code versioning system used & git \\
+C6 & Software code languages, tools, and services used & R \\
+C7 & Compilation requirements, operating environments & R (\textgreater=
+4.1.0), ADD PACKAGES \\
+C8 & If available Link to developer documentation/manual & \\
+C9 & Support email for questions & Rainer.Krug@uzh.ch;
+Rainer@krugs.de \\
+\bottomrule
+\end{longtable}
+
+\hypertarget{motivation-and-significance}{%
+\section{Motivation and
+significance}\label{motivation-and-significance}}
+
+Introduce the scientific background and the motivation for developing
+the software. Explain why the software is important, and describe the
+exact (scientific) problem(s) it solves. Indicate in what way the
+software has contributed (or how it will con- tribute in the future) to
+the process of scientific discovery; if available, this is to be
+supported by citing a research paper using the software. Provide a
+description of the experimental setting (how does the user use the
+software?). Introduce related work in literature (cite or list
+algorithms used, other software etc.).
+
+\hypertarget{software-description}{%
+\section{Software description}\label{software-description}}
+
+Describe the software in as much as is necessary to establish a
+vocabulary needed to explain its impact.
+
+\hypertarget{software-architecture}{%
+\subsection{Software Architecture}\label{software-architecture}}
+
+Give a short overview of the overall software architecture; provide a
+pic- torial component overview or similar (if possible). If necessary
+provide im- plementation details.
+
+\hypertarget{software-functionalities}{%
+\subsection{Software Functionalities}\label{software-functionalities}}
+
+Present the major functionalities of the software.
+
+\hypertarget{illustrative-examples}{%
+\section{Illustrative Examples}\label{illustrative-examples}}
+
+Provide at least one illustrative example to demonstrate the major
+functions. Optional: you may include one explanatory video that will
+appear next to your article, in the right hand side panel.
+
+\hypertarget{impact}{%
+\section{Impact}\label{impact}}
+
+This is the main section of the article and the reviewers weight the
+description here appropriately Indicate in what way new research
+questions can be pursued as a result of the software (if any). Indicate
+in what way, and to what extent, the pursuit of existing research
+questions is improved (if so). Indicate in what way the software has
+changed the daily practice of its users (if so). Indicate how widespread
+the use of the software is within and outside the intended user group.
+Indicate in what way the software is used in commercial settings and/or
+how it led to the creation of spin-off companies (if so).
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+Set out the conclusion of this original software publication.
+
+\hypertarget{conflict-of-interest}{%
+\section{Conflict of Interest}\label{conflict-of-interest}}
+
+We wish to confirm that there are no known conflicts of interest
+associated with this publication and there has been no significant
+financial support for this work that could have influenced its outcome.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+Optionally thank people and institutes you need to acknowledge.
+
 \hypertarget{introduction}{%
@@
 
-\{\textgreater\textgreater{} Biological justification and background - one paragraph \textless\textless\}
+{\{\textgreater\textgreater Biological justification and background -
+one paragraph - DONE\textless\textless\} }
 
-The microxanox package is a package for simulating a three functional group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced sulfur, \emph{SO}: oxidized sulfur).
-It includes feddback between biogeochemical processes and is based on (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) (See (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) for a detailed discussion of the model).
+Many ecosystems are exposed to gradual changes of environmental
+variables, to which the responses are not always as gradual as the
+change of the environmental variable. One example where the gradual
+change of a single environmental causing an abrupt change of the system
+is the switch from an aerobic to anaerobic system. This system hass been
+investigated by Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}). We
+wanted to take this investigation one step further, and look at the role
+of an increased biodiversity plays in these dynamics
+(\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). For this
+purpose, we developed this package.
 
-The aims of the \texttt{microxanox} package are twofold.
-Firstly, the package aims at reproducing the results shown by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}), which is acomplished in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}.
-Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).
+The microxanox package is a package for simulating a three functional
+group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
+bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
+substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
+sulfur, \emph{SO}: oxidized sulfur). It includes feddback between
+biogeochemical processes and is based on
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) (See
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) for a detailed
+discussion of the model).
+
+The aims of the \texttt{microxanox} package are twofold. Firstly, the
+package aims at reproducing the results shown by
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}), which is
+acomplished in the \href{LINK\%20NEEDED}{vignette Partial reproduction
+of Bush et al}. Secondly, to take these results one step further, it
+includes new functionality to adress our research question as presented
+in (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).
 
@@
 
-In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.
+In addition to the model itself, the package includes some functions to
+analyse the results as well as visualise the results to provide a
+starting point for customised visualisations based on own requirements.
+
+\hypertarget{general-usability-and-flexibility-of-the-model}{%
+\subsection{General usability and flexibility of the
+model}\label{general-usability-and-flexibility-of-the-model}}
+
+The package is not intended to provide a modelling framework which can
+be adjusted easily to all needs, but primarily a tool to implement the
+model used by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) and
+to extend it to our needs (\protect\hyperlink{ref-REF_NEEDED}{REF
+NEEDED, 2222}). Consequently, any more substantial changes and
+adaptaitions, are likely to need a change in the source code.
+
+Nevertheless, the model is structured in a way which builds on a modular
+structuere, so that e.g.~the event definition can easily be changed. or
+other aspects can be adjusted. All values in the parameter object can be
+changed as needed and the general structure of the code should make it
+not to difficult to adapt the model to other similar systems.
 
@@
 
-The model is an implementation of the model described by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).
-It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.
+The model is an implementation of the model described by
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). It is an ordinary
+differential equation (ODE) based numerical non-stochastic model which
+was implemented in form of an R package.
 
-We use the notation from the \href{LINK\%20NEEDED}{Table S1} of the Supplementary information of (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017})), and not the notation in the main manuscript of (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).
-In the main manuscript the notation is simplified to \texttt{y\_S\_PB} (\(y_{S,PB}\)) (which is \texttt{y\_SR\_PB} (\(y_{SR,PB}\)) in Table S1) and \texttt{y\_S\_SB} (\(y_{S,SB}\)) (which is y\_SO\_SB (\(y_{SO,SB}\)) in Table S1).
-We also use \texttt{Pr\_cB} (\(Pr_B\)) instead of \texttt{P\_CB} (\(P_{CB}\)).
+We use the notation from the \href{LINK\%20NEEDED}{Table S1} of the
+Supplementary information of (\protect\hyperlink{ref-Bush2017}{Bush et
+al., 2017})), and not the notation in the main manuscript of
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). In the main
+manuscript the notation is simplified to \texttt{y\_S\_PB}
+(\(y_{S,PB}\)) (which is \texttt{y\_SR\_PB} (\(y_{SR,PB}\)) in Table S1)
+and \texttt{y\_S\_SB} (\(y_{S,SB}\)) (which is y\_SO\_SB (\(y_{SO,SB}\))
+in Table S1). We also use \texttt{Pr\_cB} (\(Pr_B\)) instead of
+\texttt{P\_CB} (\(P_{CB}\)).
 
@@
 
-\{\textgreater\textgreater{} Add a more detailed description of the model and the additions \textless\textless\}
+{\{\textgreater\textgreater{} Add a more detailed description of the
+model and the additions \textless\textless\}
+\{\textgreater\textgreater TODO\textless\textless\}}
 
-The ODEs are solved using the ``General Solver for Ordinary Differential Equations'' of the \texttt{deSolve} package (\protect\hyperlink{ref-Soetaert2010}{Soetaert, Petzoldt, \& Setzer, 2010}).
+The ODEs are solved using the ``General Solver for Ordinary Differential
+Equations'' of the \texttt{deSolve} package
+(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}).
 
 \hypertarget{availability-and-accessability-of-the-package}{%
-\section{Availability and Accessability of the Package}\label{availability-and-accessability-of-the-package}}
+\section{Availability and Accessability of the
+package}\label{availability-and-accessability-of-the-package}}
 
-The package is available in a \href{https://github.com/UZH-PEG/microxanox/}{github repo/}.
-The newest version is also available under the \href{https://DOI}{DOI} or via RUniverse (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).
-This manuscript describes version 1.0 which is available under \href{https://DOI}{DOI} or installable within R by using \texttt{remotes::install\_github("UZH-PEG",\ ref\ =\ "v1.0")}.
+The package is available in a
+\href{https://github.com/UZH-PEG/microxanox/}{github repo/}. The newest
+version is also available under the \href{https://DOI}{DOI} or via
+RUniverse (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). This
+manuscript describes version 1.0 which is available under
+\href{https://DOI}{DOI} or installable within R by using
+\texttt{remotes::install\_github("UZH-PEG",\ ref\ =\ "v1.0")}.
 
-The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R.
-\{\textgreater\textgreater If we want that, we have to include some, even rudimentary, GITHUB tests.\textless\textless\}
-\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk about the package paper.\textless\textless\}
+The package will be updated based on further research activities. It
+will be tested continuously for the last recent, current and devel
+version of R. {\{\textgreater\textgreater If we want that, we have to
+include some, even rudimentary, GITHUB tests.\textless\textless\}
+\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk
+about the package paper.\textless\textless\} }
 
-Issues, questions and suggestions should be communicated through the issue tracker at \url{https://github.com/UZH-PEG/microxanox/issues}
+Issues, questions and suggestions should be communicated through the
+issue tracker at \url{https://github.com/UZH-PEG/microxanox/issues}
 
-\hypertarget{the-r-package}{%
-\section{The R Package}\label{the-r-package}}
+\hypertarget{package-description}{%
+\section{Package description}\label{package-description}}
 
@@
 
-The package contains the full documentetion of the functions, which includes
+The package contains the full documentetion of the functions, which
+includes
 
@@
 \item
-  Two vignettes acompanying the package, also available under RUniverse \href{@LINK_NEEDED}{Vignette 1} and \href{@LINK_NEEDED}{Vignette 2}
-\item
-  This paper as a third vignette, also available under RUniverse \href{@LINK_NEEDED}{Vignette 3}
+  Two vignettes acompanying the package:
 \end{enumerate}
 
-\hypertarget{the-modelling-framework}{%
-\subsection{The modelling framework}\label{the-modelling-framework}}
+\begin{itemize}
+\tightlist
+\item
+  A \textbf{User Guide}, also available on the R-Universe
+  \href{@LINK_NEEDED}{Vignette 1}. This gives an introduction of the
+  package and its functions and provides a walk-through of the process
+  of running, visualizing and extra and \href{@LINK_NEEDED}{Vignette 2}
+\end{itemize}
 
-The framework used when writing this package aims at reproducibility of the results. It builds on the folowing main considerations:
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  This paper as a third vignette, also available under RUniverse
+  \href{@LINK_NEEDED}{Vignette 3}
+\end{enumerate}
+
+\hypertarget{reproducibility}{%
+\subsection{Reproducibility}\label{reproducibility}}
+
+The framework used when writing this package aims at reproducibility of
+the results. It builds on the folowing main considerations:
 
@@
 \item
-  all parameter needed to run a simulation or find a stable state are contained in a single parameter object.
-This object is created by using the functions \texttt{new\_...\_parameter()}, \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}.
-Which one of the \texttt{new\_...\_parameter()} functions has to be used when, will be discussed in the section \ref{runsim} and in the \href{LINK_NEEDED}{User Guide}.
+  all parameter needed to run a simulation or find a stable state are
+  contained in a single parameter object. This object is created by
+  using the functions \texttt{new\_...\_parameter()},
+  \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}.
+  Which one of the \texttt{new\_...\_parameter()} functions has to be
+  used when, will be discussed in the section @ref(runsim) and in the
+  \href{LINK_NEEDED}{User Guide}.
 \item
-  The function call \texttt{run\_...(parameter)} will run the simulation using the parameter as defined in the object \texttt{parameter}.
+  The function call \texttt{run\_...(parameter)} will run the simulation
+  using the parameter as defined in the object \texttt{parameter}.
 \item
-  The return value of the \texttt{run\_...(parameter)} function is identical to the parameter object plus an additionl slot named \texttt{results} which contains the results of the run
+  The return value of the \texttt{run\_...(parameter)} function is
+  identical to the parameter object plus an additionl slot named
+  \texttt{results} which contains the results of the run
 \item
-  As this return value contains all parameter, it is possible to re-run the simulatuion by simply running \texttt{run\_...(result)}.
+  As this return value contains all parameter, it is possible to re-run
+  the simulatuion by simply running \texttt{run\_...(result)}.
 \end{enumerate}
 
-The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.
+The point that the results object contains all parameter needed to run
+the simulation, promotes reproducibility and makes incremental changes
+of individual parameters and re-running the simulations much easier.
 
@@
 
-We will now discuss the general structure and functiuonality of the package without going into to much detail. A more detailed discussion can be found in the \href{LINK_NEEDED}{User Guide}.
+We will now discuss the general structure and functiuonality of the
+package without going into to much detail. A more detailed discussion
+can be found in the \href{LINK_NEEDED}{User Guide}.
 
-The ODEs for the rates of change are specified in the function \texttt{bushplus\_dynamic\_model()}.
-This augmented version of the model published in (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+The ODEs for the rates of change are specified in the function
+\texttt{bushplus\_dynamic\_model()}. This augmented version of the model
+published in (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) can
+handle multiple strains within each of the three functional groups,
+temporal variation in oxygen diffusivity, and events.
 
-In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.
+In the following sections we describe the general usage of the package:
+running one simulation, finding steady states across an environmental
+gradient, calculating measures of stability, and visualisation.
 
@@
 
-The individual simulation (\texttt{run\_simulation()} function) is the working horse in this package.
-In this function, the ODEs are solved.
-The function needs only one parameter - an object as created by the function \texttt{new\_runsim\_parameter()}.
-One parameter of this object is the \texttt{strain\_parameter} which can be created by the function \texttt{new\_strain\_parameter()}.
-For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at \href{@LINK_NEEDED}{User Guide}
+The individual simulation (\texttt{run\_simulation()} function) is the
+working horse in this package. In this function, the ODEs are solved.
+The function needs only one argument - an object as created by the
+function \texttt{new\_runsim\_parameter()}. One parameter of this object
+is the \texttt{strain\_parameter} which can be created by the function
+\texttt{new\_strain\_parameter()}. For a detailed description of the
+parameter and how they are created please see the User Guide and which
+acomnpanies the package or is available at \href{@LINK_NEEDED}{User
+Guide}
 
-After the paramter object has been defined, it can be used in the \texttt{run\_simulation()} function.
-The function returns an object which is identical to the parameter, except of an additional slot containing the results.
-This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the \texttt{run\_simulation()} function to run the simulation again from the parameter used to generate the results from.
+After the parameter object has been defined, it can be used in the
+\texttt{run\_simulation()} function. The function returns an object
+which is identical to the parameter, except of an additional slot
+containing the results. This design produces a fully reproducible object
+as it can be used instead of a parameter object to be fed back into the
+\texttt{run\_simulation()} function to run the simulation again from the
+parameter used to generate the results from.
 
-The function \texttt{plot\_dynamics()} plots a single simulation run, as returned from the \texttt{run\_simulation()} function.
-This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.
+The function \texttt{plot\_dynamics()} plots a single simulation run, as
+returned from the \texttt{run\_simulation()} function. This functio is
+only provided as a convenience function to provide a way to easily see
+the results of a simulation run. An example plot resulting from this
+function is shown in @ref(fig:plot-dynamics).
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=500px]{figures/fig1_plot_dynamics}
+{\centering \includegraphics[width=500px]{figures/ug_three_strains_dynamics}
 
-\}
+}
 
-\textbackslash caption\{Plot of results of a simulation run using the function \texttt{plot\_dynamics()}\$`.
-Details can be found in the ``User Guide'' section ``Three strains per functional group''.\}\label{fig:plot-dynamics}
-\textbackslash end\{figure\}
+\caption{Plot of results of a simulation run using the function $plot_dynamics()$. Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:plot-dynamics}
+\end{figure}
 
-\hypertarget{find-a-steady-state-of-the-model}{%
-\subsubsection{Find a Steady State of the model}\label{find-a-steady-state-of-the-model}}
+\hypertarget{finding-a-steady-state-of-the-model}{%
+\subsubsection{Finding a Steady State of the
+model}\label{finding-a-steady-state-of-the-model}}
 
-There are two methods for finding steady states.
-The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the \emph{Replication method}).
-The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the \emph{Temporal method}).
+There are two methods for finding steady states. The first runs a
+separate simulation for each combination of starting conditions and
+oxygen diffusivity (let us term this the \emph{Replication method}). The
+second runs only two simulations, with step-wise and slowly temporally
+increasing or decreasing oxygen diffusivities and recorded of state just
+before change to a new oxygen diffusivity (let us term this the
+\emph{Temporal method}).
 
@@
 
-The replication method is implementad in the function \texttt{run\_replication\_ssfind()} which takes a parameter object as returned by the function \texttt{new\_replication\_ssfind\_parameter()} and the number of cores for multithreading the simulation.
+The replication method is implementad in the function
+\texttt{run\_replication\_ssfind()} which takes a parameter object as
+returned by the function \texttt{new\_replication\_ssfind\_parameter()}
+and the number of cores for multithreading the simulation.
 
@@
 
-The temporal method involves two simulations for a particular system configuration (parameter set).
-In one simulation the oxygen diffusivity is \emph{increased} in a step-wise fashion.
-In the other it is \emph{decreased} in a step-wise fashion.
-That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set.
-Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.
+The temporal method involves two simulations for a particular system
+configuration (parameter set). In one simulation the oxygen diffusivity
+is \emph{increased} in a step-wise fashion. In the other it is
+\emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is
+held at a constant level for long enough for steady state to be reach,
+that state is recorded, and then a slightly higher (or lower) oxygen
+diffusivity value is set. Hence, at that time point, the system is
+effectively started with initial conditions that are the state of the
+system in the previous time step.
 
-This is implemented in the function \texttt{run\_temporal\_ssfind()}, which takes a parameter object as created by the function \texttt{new\_temporal\_ssfind\_parameter()} and a number indicating the .
+This is implemented in the function \texttt{run\_temporal\_ssfind()},
+which takes a parameter object as created by the function
+\texttt{new\_temporal\_ssfind\_parameter()} and a number indicating the
+.
 
-For a more detailed walk-through of these two approaches and explanation please see the \href{@LINK_NEEDED}{User Guide}.
+For a more detailed walk-through of these two approaches and explanation
+please see the \href{@LINK_NEEDED}{User Guide}.
 
-\hypertarget{extract-stability-measures}{%
-\subsection{Extract Stability Measures}\label{extract-stability-measures}}
+\hypertarget{extracting-stability-measures}{%
+\subsection{Extracting Stability
+Measures}\label{extracting-stability-measures}}
 
-From the raw results returned by these \texttt{run\_...()} functions, the stability measures can be extracted by using the function \texttt{get\_stability\_measures()}.
-These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.
+From the raw results returned by these \texttt{run\_...()} functions,
+the stability measures can be extracted by using the function
+\texttt{get\_stability\_measures()}. These measures include
+non-linearity and hysteresis measures, of the response of the simulated
+system to environmental change.
 
@@
 
-The first two use cases come from the User Guide and the Partial Reproduction Vignettes. The third comes from a research article that relied on the package.
+The first two use cases are described in detail in the User Guide and
+the Partial Reproduction Vignettes. The third is taken from the REF
+NEEDED (\protect\hyperlink{ref-REF_NEEDED}{2222}) for which this R
+package was designed. All of these use cases can be expanded to larger
+numbers of strains per functional group and variable values van be
+changed.
 
 \hypertarget{regime-shifts-during-temporal-environmental-change}{%
-\subsection{Regime shifts during temporal environmental change}\label{regime-shifts-during-temporal-environmental-change}}
+\subsection{Regime shifts during temporal environmental
+change}\label{regime-shifts-during-temporal-environmental-change}}
 
-\{\textgreater\textgreater Take from the user guide.\textless\textless\}
+In the \href{LINK_NEEDED}{User Guide} we used a one strain system
+(section ``1 strain per functional group'') and three strain system
+(section ``3 strains per functional group'') to determine as an example
+the stable states during temporal environmental changes (the oxygen
+diffusivity). From these simulatiuons, we extracted measures of
+nonlinearity and hysteresis. See Fig @ref(fig:plot-dynamics) as an
+example plot of the simulations.
+
+\begin{figure}
+
+{\centering \includegraphics[width=500px]{figures/ug_three_strains_stable_state}
+
+}
+
+\caption{Plot of the stable states of the simulation runs under different oxygen diffusivity.
+The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities.
+Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:uc1_stable_state}
+\end{figure}
 
 \hypertarget{the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsection{The extent of hysteresis depends on community composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}
+\subsection{The extent of hysteresis depends on community
+composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}
 
-\{\textgreater\textgreater Take from the partial reproduction.\textless\textless\}
+One of the reasons to develop this package was to reproduce the results
+presented in Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}), this
+was achieved as demonstrated in the \href{LINK_NEEDED}{Partial
+Reproduction supplement}. All aspects in the paper could be reproduced
+and are shown in the vignette.
+
+\begin{figure}
+
+{\centering \includegraphics[width=500px]{figures/user_guide_hysteresis}
+
+}
+
+\caption{Hysteresis of all assessed combinations of variability.}\label{fig:user_guide_hysteresis}
+\end{figure}
 
 \hypertarget{effects-of-functional-diversity-on-regime-shifts}{%
-\subsection{Effects of functional diversity on regime shifts}\label{effects-of-functional-diversity-on-regime-shifts}}
+\subsection{Effects of functional diversity on regime
+shifts}\label{effects-of-functional-diversity-on-regime-shifts}}
 
-This model (as part of the package) has been used in the production of the paper (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). It can be used for similar studies in this study systems.
+As discussed in the paper (\protect\hyperlink{ref-REF_NEEDED}{REF
+NEEDED, 2222}), the role biodiversity plays in abrupt regime shifts
+based on gradual changing environmental parameter is not well
+understood. This model (as part of the package) has been used to
+investigate these dynamics and the results are available in REF NEEDED
+(\protect\hyperlink{ref-REF_NEEDED}{2222}).
 
-\{\textgreater\textgreater More details from the paper.\textless\textless\}
+{\{\textgreater\textgreater More details from the
+paper.\textless\textless\}\\
+\{\textgreater\textgreater Romana will provide two or three
+sentences\textless\textless\}}
 
-\hypertarget{general-usability-and-flexibility-of-the-model}{%
-\subsection{General usability and flexibility of the model}\label{general-usability-and-flexibility-of-the-model}}
-
-The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) and to extend it to our needs (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).
-Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.
-
-Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.~the event definition can easily be changed.
-or other aspects can be adjusted.
-All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\hypertarget{conclusions-1}{%
+\section{Conclusions}\label{conclusions-1}}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
-Bush, T., Diao, M., Allen, R.
-J., Sinnige, R., Muyzer, G., \& Huisman, J.
-(2017).
-Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics.
-\emph{Nature Communications}, \emph{8}(1), 789.
+Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
+2017. Oxic-anoxic regime shifts mediated by feedbacks between
+biogeochemical processes and microbial community dynamics. Nature
+Communications 8, 789.
 doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-REF_NEEDED}{}}%
-REF NEEDED, A. (2222). {REFERNECE NEEDED}. \emph{Journal of Missing References}.
+REF NEEDED, A., 2222. {REFERNECE NEEDED}. Journal of missing references.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
-Soetaert, K., Petzoldt, T., \& Setzer, R.
-W.
-(2010).
-Solving differential equations in {R}: {Package deSolve}.
-\emph{Journal of Statistical Software}, \emph{33}(9), 1--25.
+Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
+equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
+1--25.
 doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}

### 2022-04-13 inst/manuscript/manuscript_microxanox.Rmd
Re-ordered for the format of SoftwareX

@@
 
-#linenumbers: true
-#numbersections: true
+linenumbers: true
+numbersections: true
 csl: elsevier-harvard.csl
@@
 ```{r setup, include = FALSE}
+library(plantuml)
 ```
@@
 | **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |
-|---|---|---|
+|:-------:|:-------------------------------------------------------------|:-------------------------------------------------------------------------------|
 | C1      | Current code version                                         | v0.9.0                                                                         |
@@
 | C4      | Legal Code License                                           | CC BY 4.0                                                                      |
-| C5 | Code versioning system used |git |
+| C5      | Code versioning system used                                  | git                                                                            |
 | C6      | Software code languages, tools, and services used            | R                                                                              |
@@
 
-## General usability and flexibility of the model
+# Software description
+{>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<}
+
+```{plantuml runsim_example, eval = TRUE, plantuml.format = "auto", plantuml.preview = TRUE}
+@startuml
+partition "Run Single Simulation" {
+   start
+  :parameter <- new_runsim_parameter();
+  split
+    #palegreen:saveRDS(FILE, parameter);
+    detach
+  split again
+    #pink:result <- run_simulation(parameter);
+  end split
+  split
+    #palegreen:saveRDS(FILE, results);
+    detach
+  split again
+    :Change some parameter;
+    #pink:result2 <- run_simulation(results);
+  end split
+  split
+    #palegreen:saveRDS(FILE, results2);
+    detach
+  split again
+    end
+  end split
+}
+@enduml
+```
+
 The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
@@
 
-# Software description
-Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.
-
 ## Software Architecture
-Give a short overview of the overall software architecture; provide a pic- torial component overview or similar (if possible). If necessary provide im- plementation details.
-
-## Software Functionalities
-Present the major functionalities of the software.
-
-# Illustrative Examples
-Provide at least one illustrative example to demonstrate the major functions.
-Optional: you may include one explanatory video that will appear next to your article, in the right hand side panel.
-
-# Impact
-This is the main section of the article and the reviewers weight the description here appropriately
-Indicate in what way new research questions can be pursued as a result of the software (if any).
-Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
-Indicate in what way the software has changed the daily practice of its users (if so).
-Indicate how widespread the use of the software is within and outside the intended user group.
-Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).
-
-# Conclusions
-Set out the conclusion of this original software publication.
-
-# Conflict of Interest
-We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
-
-# Acknowledgements
-Optionally thank people and institutes you need to acknowledge.
-
-# THIS NEEDS TO BE TRANSFERED IN THE FORMAT ABOVE
-
-# The Model
-
-The model is an implementation of the model described by [@Bush2017]. It is an ordinary differential equation (ODE) based numerical non-stochastic model which was implemented in form of an R package.
-
-We use the notation from the [Table S1](LINK NEEDED) of the Supplementary information of [@Bush2017]), and not the notation in the main manuscript of [@Bush2017].
-In the main manuscript the notation is simplified to `y_S_PB` ($y_{S,PB}$) (which is `y_SR_PB` ($y_{SR,PB}$) in Table S1) and `y_S_SB` ($y_{S,SB}$) (which is y_SO_SB ($y_{SO,SB}$) in Table S1).
-We also use `Pr_cB` ($Pr_B$) instead of `P_CB` ($P_{CB}$).
-
-All parameter and variables follow these base dimensions:
-
-- Time: `hours`
-- Volume: `litres`
-- Substrate quantity: `micromoles`
-- Organism quantity: `cells`
-
-<span style="color:red">{>> Add a more detailed description of the model and the additions <<}
-{>>TODO<<}</span>
-
-The ODEs are solved using the "General Solver for Ordinary Differential Equations" of the `deSolve` package [@Soetaert2010].
-
-# Availability and Accessability of the package
-
-The package is available in a [github repo/](https://github.com/UZH-PEG/microxanox/).
-The newest version is also available under the [DOI](https://DOI) or via RUniverse [@REF_NEEDED].
-This manuscript describes version 1.0 which is available under [DOI](https://DOI) or installable within R by using `remotes::install_github("UZH-PEG", ref = "v1.0")`.
-
-The package will be updated based on further research activities. It will be tested continuously for the last recent, current and devel version of R. 
-<span style="color:green">{>>If we want that, we have to include some, even rudimentary, GITHUB tests.<<}
-{>>RMK: Agreed - let's discuss this when we talk about the package paper.<<}
-</span>
-
-Issues, questions and suggestions should be communicated through the issue tracker at [https://github.com/UZH-PEG/microxanox/issues](https://github.com/UZH-PEG/microxanox/issues)
-
-# Package description
-
-## Documentation
-
-The package contains the full documentetion of the functions, which includes
-
-1. Standard R help for each function (`?COMMAND` in R)
-2. Two vignettes acompanying the package:
-  - A **User Guide**, also available on the R-Universe [Vignette 1](@LINK_NEEDED).
-This gives an introduction of the package and its functions and provides a walk-through of the process of running, visualizing and extra
-   and [Vignette 2](@LINK_NEEDED)
-3. This paper as a third vignette, also available under RUniverse [Vignette 3](@LINK_NEEDED)
-
-## Reproducibility
+{>>Give a short overview of the overall software architecture; provide a pic- torial component overview or similar (if possible). If necessary provide im- plementation details.<<}
 
@@
 ## Run the simulation and save the result
-result <- run_simulatrion(parameter)
+result <- run_simulation(parameter)
 saveRDS(result, "sim1.rds")
@@
 # change some parameter
-result <- run_simulatrion(parameter)
+result <- run_simulation(parameter)
 saveRDS(result, "sim2,rds")
@@
 
-## Using the package {#runsim}
+## Software Functionalities
+{>>Present the major functionalities of the software.<<}
 
-We will now discuss the general structure and functiuonality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
+We will now discuss the general structure and functionality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
 
@@
 
-In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualisation.
+In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualization.
+
+# Illustrative Examples
+{>>Present the major functionalities of the software.<<}
 
@@
 
-## Extracting Stability Measures
+### Extracting Stability Measures
 From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
@@
 
-# Use cases
+# Impact
+{>>This is the main section of the article and the reviewers weight the description here appropriately
+Indicate in what way new research questions can be pursued as a result of the software (if any).
+Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
+Indicate in what way the software has changed the daily practice of its users (if so).
+Indicate how widespread the use of the software is within and outside the intended user group.
+Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<}
 
@@
 # Conclusions
+Set out the conclusion of this original software publication.
 
-{>>Dependant on journal<<}
+# Conflict of Interest
+We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
+
+# Acknowledgements
+Optionally thank people and institutes you need to acknowledge.
 

### 2022-04-13 inst/manuscript/manuscript_microxanox.tex
Re-ordered for the format of SoftwareX

@@
 \usepackage{lineno} % add
+  \linenumbers % turns line numbering on
 
@@
 \bibliographystyle{elsarticle-harv}
+\usepackage{graphicx}
 \ifxetex
@@
 
-\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{5}
 % Pandoc toggle for numbering sections (defaults to be off)
-\setcounter{secnumdepth}{0}
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3333}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0596}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4106}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5298}}@{}}
 \toprule
-\begin{minipage}[b]{\linewidth}\raggedright
+\begin{minipage}[b]{\linewidth}\centering
 \textbf{Nr.}
@@
 \endhead
-C1 & Current code version & For example v42 \\
+C1 & Current code version & v0.9.0 \\
 C2 & Permanent link to code/repository used for this code version &
@@
 significance}\label{motivation-and-significance}}
-
-Introduce the scientific background and the motivation for developing
-the software. Explain why the software is important, and describe the
-exact (scientific) problem(s) it solves. Indicate in what way the
-software has contributed (or how it will con- tribute in the future) to
-the process of scientific discovery; if available, this is to be
-supported by citing a research paper using the software. Provide a
-description of the experimental setting (how does the user use the
-software?). Introduce related work in literature (cite or list
-algorithms used, other software etc.).
-
-\hypertarget{software-description}{%
-\section{Software description}\label{software-description}}
-
-Describe the software in as much as is necessary to establish a
-vocabulary needed to explain its impact.
-
-\hypertarget{software-architecture}{%
-\subsection{Software Architecture}\label{software-architecture}}
-
-Give a short overview of the overall software architecture; provide a
-pic- torial component overview or similar (if possible). If necessary
-provide im- plementation details.
-
-\hypertarget{software-functionalities}{%
-\subsection{Software Functionalities}\label{software-functionalities}}
-
-Present the major functionalities of the software.
-
-\hypertarget{illustrative-examples}{%
-\section{Illustrative Examples}\label{illustrative-examples}}
-
-Provide at least one illustrative example to demonstrate the major
-functions. Optional: you may include one explanatory video that will
-appear next to your article, in the right hand side panel.
-
-\hypertarget{impact}{%
-\section{Impact}\label{impact}}
-
-This is the main section of the article and the reviewers weight the
-description here appropriately Indicate in what way new research
-questions can be pursued as a result of the software (if any). Indicate
-in what way, and to what extent, the pursuit of existing research
-questions is improved (if so). Indicate in what way the software has
-changed the daily practice of its users (if so). Indicate how widespread
-the use of the software is within and outside the intended user group.
-Indicate in what way the software is used in commercial settings and/or
-how it led to the creation of spin-off companies (if so).
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-Set out the conclusion of this original software publication.
-
-\hypertarget{conflict-of-interest}{%
-\section{Conflict of Interest}\label{conflict-of-interest}}
-
-We wish to confirm that there are no known conflicts of interest
-associated with this publication and there has been no significant
-financial support for this work that could have influenced its outcome.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-Optionally thank people and institutes you need to acknowledge.
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-{\{\textgreater\textgreater Biological justification and background -
-one paragraph - DONE\textless\textless\} }
 
@@
 
-\hypertarget{general-usability-and-flexibility-of-the-model}{%
-\subsection{General usability and flexibility of the
-model}\label{general-usability-and-flexibility-of-the-model}}
+\hypertarget{software-description}{%
+\section{Software description}\label{software-description}}
+
+\{\textgreater\textgreater Describe the software in as much as is
+necessary to establish a vocabulary needed to explain its
+impact.\textless\textless\}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{@startuml}
+\NormalTok{partition "Run Single Simulation" \{}
+\NormalTok{   start}
+\NormalTok{  :parameter \textless{}{-} new\_runsim\_parameter();}
+\NormalTok{  split}
+\NormalTok{    \#palegreen:saveRDS(FILE, parameter);}
+\NormalTok{    detach}
+\NormalTok{  split again}
+\NormalTok{    \#pink:result \textless{}{-} run\_simulation(parameter);}
+\NormalTok{  end split}
+\NormalTok{  split}
+\NormalTok{    \#palegreen:saveRDS(FILE, results);}
+\NormalTok{    detach}
+\NormalTok{  split again}
+\NormalTok{    :Change some parameter;}
+\NormalTok{    \#pink:result2 \textless{}{-} run\_simulation(results);}
+\NormalTok{  end split}
+\NormalTok{  split}
+\NormalTok{    \#palegreen:saveRDS(FILE, results2);}
+\NormalTok{    detach}
+\NormalTok{  split again}
+\NormalTok{    end}
+\NormalTok{  end split}
+\NormalTok{\}}
+\NormalTok{@enduml}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{/var/folders/50/wcr5bjwn75q595n6x82gxj280000gq/T//RtmpRw3cq4/runsim_example.pdf}
 
@@
 
-\hypertarget{the-model}{%
-\section{The Model}\label{the-model}}
+\hypertarget{software-architecture}{%
+\subsection{Software Architecture}\label{software-architecture}}
 
-The model is an implementation of the model described by
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). It is an ordinary
-differential equation (ODE) based numerical non-stochastic model which
-was implemented in form of an R package.
-
-We use the notation from the \href{LINK\%20NEEDED}{Table S1} of the
-Supplementary information of (\protect\hyperlink{ref-Bush2017}{Bush et
-al., 2017})), and not the notation in the main manuscript of
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}). In the main
-manuscript the notation is simplified to \texttt{y\_S\_PB}
-(\(y_{S,PB}\)) (which is \texttt{y\_SR\_PB} (\(y_{SR,PB}\)) in Table S1)
-and \texttt{y\_S\_SB} (\(y_{S,SB}\)) (which is y\_SO\_SB (\(y_{SO,SB}\))
-in Table S1). We also use \texttt{Pr\_cB} (\(Pr_B\)) instead of
-\texttt{P\_CB} (\(P_{CB}\)).
-
-All parameter and variables follow these base dimensions:
-
-\begin{itemize}
-\tightlist
-\item
-  Time: \texttt{hours}
-\item
-  Volume: \texttt{litres}
-\item
-  Substrate quantity: \texttt{micromoles}
-\item
-  Organism quantity: \texttt{cells}
-\end{itemize}
-
-{\{\textgreater\textgreater{} Add a more detailed description of the
-model and the additions \textless\textless\}
-\{\textgreater\textgreater TODO\textless\textless\}}
-
-The ODEs are solved using the ``General Solver for Ordinary Differential
-Equations'' of the \texttt{deSolve} package
-(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}).
-
-\hypertarget{availability-and-accessability-of-the-package}{%
-\section{Availability and Accessability of the
-package}\label{availability-and-accessability-of-the-package}}
-
-The package is available in a
-\href{https://github.com/UZH-PEG/microxanox/}{github repo/}. The newest
-version is also available under the \href{https://DOI}{DOI} or via
-RUniverse (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). This
-manuscript describes version 1.0 which is available under
-\href{https://DOI}{DOI} or installable within R by using
-\texttt{remotes::install\_github("UZH-PEG",\ ref\ =\ "v1.0")}.
-
-The package will be updated based on further research activities. It
-will be tested continuously for the last recent, current and devel
-version of R. {\{\textgreater\textgreater If we want that, we have to
-include some, even rudimentary, GITHUB tests.\textless\textless\}
-\{\textgreater\textgreater RMK: Agreed - let's discuss this when we talk
-about the package paper.\textless\textless\} }
-
-Issues, questions and suggestions should be communicated through the
-issue tracker at \url{https://github.com/UZH-PEG/microxanox/issues}
-
-\hypertarget{package-description}{%
-\section{Package description}\label{package-description}}
-
-\hypertarget{documentation}{%
-\subsection{Documentation}\label{documentation}}
-
-The package contains the full documentetion of the functions, which
-includes
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Standard R help for each function (\texttt{?COMMAND} in R)
-\item
-  Two vignettes acompanying the package:
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  A \textbf{User Guide}, also available on the R-Universe
-  \href{@LINK_NEEDED}{Vignette 1}. This gives an introduction of the
-  package and its functions and provides a walk-through of the process
-  of running, visualizing and extra and \href{@LINK_NEEDED}{Vignette 2}
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{2}
-\tightlist
-\item
-  This paper as a third vignette, also available under RUniverse
-  \href{@LINK_NEEDED}{Vignette 3}
-\end{enumerate}
-
-\hypertarget{reproducibility}{%
-\subsection{Reproducibility}\label{reproducibility}}
+\{\textgreater\textgreater Give a short overview of the overall software
+architecture; provide a pic- torial component overview or similar (if
+possible). If necessary provide im- plementation
+details.\textless\textless\}
 
@@
 \DocumentationTok{\#\# Run the simulation and save the result}
-\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulatrion}\NormalTok{(parameter)}
+\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulation}\NormalTok{(parameter)}
 \FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim1.rds"}\NormalTok{)}
@@
 \CommentTok{\# change some parameter}
-\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulatrion}\NormalTok{(parameter)}
+\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulation}\NormalTok{(parameter)}
 \FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim2,rds"}\NormalTok{)}
@@
 
-\hypertarget{runsim}{%
-\subsection{Using the package}\label{runsim}}
+\hypertarget{software-functionalities}{%
+\subsection{Software Functionalities}\label{software-functionalities}}
 
-We will now discuss the general structure and functiuonality of the
+\{\textgreater\textgreater Present the major functionalities of the
+software.\textless\textless\}
+
+We will now discuss the general structure and functionality of the
 package without going into to much detail. A more detailed discussion
@@
 running one simulation, finding steady states across an environmental
-gradient, calculating measures of stability, and visualisation.
+gradient, calculating measures of stability, and visualization.
+
+\hypertarget{illustrative-examples}{%
+\section{Illustrative Examples}\label{illustrative-examples}}
+
+\{\textgreater\textgreater Present the major functionalities of the
+software.\textless\textless\}
 
@@
 \hypertarget{extracting-stability-measures}{%
-\subsection{Extracting Stability
+\subsubsection{Extracting Stability
 Measures}\label{extracting-stability-measures}}
@@
 
-\hypertarget{use-cases}{%
-\section{Use cases}\label{use-cases}}
+\hypertarget{impact}{%
+\section{Impact}\label{impact}}
+
+\{\textgreater\textgreater This is the main section of the article and
+the reviewers weight the description here appropriately Indicate in what
+way new research questions can be pursued as a result of the software
+(if any). Indicate in what way, and to what extent, the pursuit of
+existing research questions is improved (if so). Indicate in what way
+the software has changed the daily practice of its users (if so).
+Indicate how widespread the use of the software is within and outside
+the intended user group. Indicate in what way the software is used in
+commercial settings and/or how it led to the creation of spin-off
+companies (if so).\textless\textless\}
 
@@
 {\{\textgreater\textgreater More details from the
-paper.\textless\textless\}\\
-\{\textgreater\textgreater Romana will provide two or three
-sentences\textless\textless\}}
+paper.\textless\textless\} \{\textgreater\textgreater Romana will
+provide two or three sentences\textless\textless\}}
 
-\hypertarget{conclusions-1}{%
-\section{Conclusions}\label{conclusions-1}}
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
 
-\{\textgreater\textgreater Dependant on journal\textless\textless\}
+Set out the conclusion of this original software publication.
+
+\hypertarget{conflict-of-interest}{%
+\section{Conflict of Interest}\label{conflict-of-interest}}
+
+We wish to confirm that there are no known conflicts of interest
+associated with this publication and there has been no significant
+financial support for this work that could have influenced its outcome.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+Optionally thank people and institutes you need to acknowledge.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
-Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
-equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
-1--25.
-doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}
-
 \end{CSLReferences}

### 2022-04-26 inst/manuscript/manuscript_microxanox.Rmd
Update manuscript_microxanox.Rmd

@@
 ```{r setup, include = FALSE}
-library(plantuml)
 ```
@@
  -->
+
+# TODO
+- **update links**
+- **update crossref**
 
@@
 One example where the gradual change of a single environmental causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
-This system hass been investigated by @Bush2017.
-We wanted to take this investigation one step further, and look at the role of an increased biodiversity plays in these dynamics [@REF_NEEDED].
+This system has been investigated by @Bush2017.
+We wanted to take this investigation one step further, and look at the role an increased biodiversity plays in these dynamics [@REF_NEEDED].
 For this purpose, we developed this package.
@@
 The microxanox package is a package for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
-It includes feddback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
+It includes feedback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
 
 The aims of the `microxanox` package are twofold.
-Firstly, the package aims at reproducing the results shown by [@Bush2017], which is acomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED).
-Secondly, to take these results one step further, it includes new functionality to adress our research question as presented in [@REF_NEEDED].
+Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED).
+Secondly, to take these results one step further, it includes new functionality to address our research question as presented in [@REF_NEEDED].
 
@@
 
-In addition to the model itself, the package includes some functions to analyse the results as well as visualise the results to provide a starting point for customised visualisations based on own requirements.
+In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements.
 
 # Software description
-{>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<}
-
-```{plantuml runsim_example, eval = TRUE, plantuml.format = "auto", plantuml.preview = TRUE}
-@startuml
-partition "Run Single Simulation" {
-   start
-  :parameter <- new_runsim_parameter();
-  split
-    #palegreen:saveRDS(FILE, parameter);
-    detach
-  split again
-    #pink:result <- run_simulation(parameter);
-  end split
-  split
-    #palegreen:saveRDS(FILE, results);
-    detach
-  split again
-    :Change some parameter;
-    #pink:result2 <- run_simulation(results);
-  end split
-  split
-    #palegreen:saveRDS(FILE, results2);
-    detach
-  split again
-    end
-  end split
-}
-@enduml
-```
+> {>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<}
 
 The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
-Consequently, any more substantial changes and adaptaitions, are likely to need a change in the source code.
+Consequently, any more substantial changes and adaptations, are likely to need a change in the source code.
 
-Nevertheless, the model is structured in a way which builds on a modular structuere, so that e.g.
-the event definition can easily be changed.
-or other aspects can be adjusted.
+Nevertheless, the model is structured in a way which builds on a modular structure, so that e.g.
+the event definition can easily be changed, or other aspects can be adjusted.
 All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
@@
 ## Software Architecture
-{>>Give a short overview of the overall software architecture; provide a pic- torial component overview or similar (if possible). If necessary provide im- plementation details.<<}
+> {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide im- plementation details.<<}
 
-The framework used when writing this package aims at reproducibility of the results. It builds on the folowing main considerations:
+The framework used when writing this package aims at reproducibility of the results. It builds on the following main considerations:
 
@@
 
-A typical simulation would look as followed:
+## Software Functionalities
+> {>>Present the major functionalities of the software.<<}
 
-```{r, eval = FALSE}
-## Create the parameter
-parameter <- new_runsim_parameter()
-# manually setting certain parameter
-
-## Run the simulation and save the result
-result <- run_simulation(parameter)
-saveRDS(result, "sim1.rds")
-
-## Do other stuff, e.g. plotting
-
-## Load results, change some parameter, and rerun the simulation and save the result
-parameter <- loadRDS("sim1.rds")
-# change some parameter
-result <- run_simulation(parameter)
-saveRDS(result, "sim2,rds")
-```
-
-## Software Functionalities
-{>>Present the major functionalities of the software.<<}
-
-We will now discuss the general structure and functionality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
+We will discuss the general structure and functionality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
 
@@
 # Illustrative Examples
-{>>Present the major functionalities of the software.<<}
+> {>>Present the major functionalities of the software.<<}
 
-### Running one simulation
+## Running one simulation
+
+A typical simulation would look as shown in \@ref(fig:runsim_example).
+
+```{R runsim_example, eval = TRUE, echo = FALSE, fig.cap='Typical flow of a simulation.
+Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.', fig.align="center", out.width = 350 }
+fig <- "./figures/simflow.pdf"
+
+if (!file.exists(fig)){
+  library(plantuml)
+  puml <- '
+    @startuml
+    partition "Run Single Simulation" {
+      start
+      #DarkGrey:\"\"library(microxanox)\"\"\n\"\"parameter <- new_runsim_parameter(...)\"\";
+      :Change some parameter
+      in the \"\"parameter\"\" object;      
+      split
+        #LightGrey:\"\"saveRDS(FILE, parameter)\"\";
+        detach
+      split again
+        #DarkGrey:\"\"result <- run_simulation(parameter)\"\";
+      end split
+      split
+        #LightGrey:\"\"saveRDS(FILE, results)\"\";
+        detach
+      split again
+        #DarkGrey:\"\"plot_dynamics(results)\"\";
+        :Change some parameter
+        in the \"\"result\"\" object;
+        #DarkGrey:\"\"result2 <- run_simulation(results)\"\";
+      end split
+      split
+        #LightGrey:\"\"saveRDS(FILE, results2)\"\";
+        detach
+      split again
+        #DarkGrey:\"\"plot_dynamics(results)\"\";
+        end
+      end split
+    }
+    @enduml
+  '
+
+  plantuml::get_graph(puml, fig)
+}
+
+knitr::include_graphics(fig)
+```
 
@@
 One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter and how they are created please see the User Guide and which acomnpanies the package or is available at [User Guide](@LINK_NEEDED)
+For a detailed description of the parameter and how they are created please see the User Guide which acomnpanies the package or is available at [User Guide](@LINK_NEEDED)
 
@@
 ```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function $plot_dynamics()$.
-Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
+Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/ug_three_strains_dynamics.png")
@@
 
-### Finding a Steady State of the model
+## Finding a Steady State of the model
 
@@
 
-#### Replication Method
+### Replication Method
 The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
 
-#### Temporal Method
+### Temporal Method
 The temporal method involves two simulations for a particular system configuration (parameter set).
@@
 
-### Extracting Stability Measures
+## Extracting Stability Measures
 From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
@@
 # Impact
-{>>This is the main section of the article and the reviewers weight the description here appropriately
+> {>>This is the main section of the article and the reviewers weight the description here appropriately
 Indicate in what way new research questions can be pursued as a result of the software (if any).
@@
 The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities.
-Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 500 }
+Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/ug_three_strains_stable_state.png")
@@
 
-```{r user_guide_hysteresis, echo = FALSE,  include=TRUE, fig.cap='Hysteresis of all assessed combinations of variability.', fig.align="center", out.width = 500 }
+```{r user_guide_hysteresis, echo = FALSE,  include=TRUE, fig.cap='Hysteresis of all assessed combinations of variability.', fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/user_guide_hysteresis.png")

### 2022-05-02 inst/manuscript/manuscript_microxanox.Rmd
New Structure next step

@@
 Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED).
-Secondly, to take these results one step further, it includes new functionality to address our research question as presented in [@REF_NEEDED].
+Secondly, to take these results one step further, it includes new functionality to address our research question as published in [@REF_NEEDED].
 
@@
 
+We will discuss the general structure and functionality of the package here without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
+
+The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
+This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+
 The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
-Consequently, any more substantial changes and adaptations, are likely to need a change in the source code.
+Consequently, any more substantial changes and adaptations are likely to need a change in the source code.
 
-Nevertheless, the model is structured in a way which builds on a modular structure, so that e.g.
+Nevertheless, the source code is structured in modular, so that e.g.
 the event definition can easily be changed, or other aspects can be adjusted.
@@
 This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`.
-Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim)  and in the [User Guide](LINK_NEEDED).
+Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim) and in more detail in the [User Guide](LINK_NEEDED).
 2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
-3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additionl slot named `results` which contains the results of the run
-4. As this return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.
+3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additional slot named `results` which contains the results of the run
+4. As the return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.
 
@@
 
-## Software Functionalities
-> {>>Present the major functionalities of the software.<<}
-
-We will discuss the general structure and functionality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
-
-The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
-This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
-
 In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualization.
-
-# Illustrative Examples
-> {>>Present the major functionalities of the software.<<}
 
@@
 One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter and how they are created please see the User Guide which acomnpanies the package or is available at [User Guide](@LINK_NEEDED)
+For a detailed description of the parameter, their meaning and how they are created please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)
 
@@
 The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function.
-This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run.
+This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
 An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).
@@
 
-This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .
+This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
 
@@
 
-# Impact
-> {>>This is the main section of the article and the reviewers weight the description here appropriately
-Indicate in what way new research questions can be pursued as a result of the software (if any).
-Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
-Indicate in what way the software has changed the daily practice of its users (if so).
-Indicate how widespread the use of the software is within and outside the intended user group.
-Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<}
+## Use Cases
+We will now show two use cases which illustrate applications of the package. All use cases can be used as starting points for other investigations of the behavior of the system.
 
 The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes.
-The third is taken from the @REF_NEEDED for which this R package was designed.
+The third is taken from @REF_NEEDED for which this R package was designed.
 All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.
 
-## Regime shifts during temporal environmental change
+### Regime shifts during temporal environmental change
 
@@
 
-## The extent of hysteresis depends on community composition
+We saved all parameter and result objects, so that all results are completely reproducible. The general flow of the experiment is identical to the one shown in \@ref(fig:runsim_example).
 
-One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED).
+### The extent of hysteresis depends on community composition
+
+One of the reasons to develop this package was to reproduce the results presented in @Bush2017.
+This was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED).
 All aspects in the paper could be reproduced and are shown in the vignette.
@@
 
-## Effects of functional diversity on regime shifts
+Ass all objects were saved, results can be easily reproduced and parameter can be changed to assess the impact of these changes (e.g. to conduc a sensitivity analysis).
+
+### Effects of functional diversity on regime shifts
 
@@
 {>>Romana will provide two or three sentences<<}</span>
+
+# Impact
+> {>>This is the main section of the article and the reviewers weight the description here appropriately
+Indicate in what way new research questions can be pursued as a result of the software (if any).
+Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
+Indicate in what way the software has changed the daily practice of its users (if so).
+Indicate how widespread the use of the software is within and outside the intended user group.
+Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<}
+
+The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing investigations which extend the original publication, as done in @REFNEEDED.
+
+The combination of all parameter in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object and adds a results slot provides a relatively easy to use framework to implement reproducible experiments.
 

### 2022-05-02 inst/manuscript/manuscript_microxanox.Rmd
Update manuscript_microxanox.Rmd
Minor Changes

--- 2022-04-26 inst/manuscript/manuscript_microxanox.Rmd
@@
 - name: Rainer M Krug
-  email: Rainer.Krug@uzh.ch & Rainer@krugs.de
+  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
   affiliation: University of Zürich
@@
   Ca. 100 words.
-bibliography: manuscript_bibliography.bib
 journal: "SoftwareX"
@@
 # Citation settings
+bibliography: bibliography.bib
 nocite: '@Bush2017'
@@
 
-# TODO
-- **update links**
-- **update crossref**
+# \color{red}TODO
+- \color{red}**update links** 
+- \color{red}**update cross references**
+\pagebreak
 

### 2022-05-03 inst/manuscript/manuscript_microxanox.Rmd
Update manuscript_microxanox.Rmd

@@
 Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable.
-One example where the gradual change of a single environmental causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
+One example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
 This system has been investigated by @Bush2017.
@@
 
-The ODEs for the rates of change are specified in the function `bushplus_dynamic_model()`.
+The model by @Bush2017 is based on ODEs for the rates of change which are specified in the function `bushplus_dynamic_model()`.
 This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
@@
 3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additional slot named `results` which contains the results of the run
-4. As the return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.
+4. As the return value contains all parameter, it is possible to re-run the simulation by simply running `run_...(result)`.
 
@@
 
-## Finding a Steady State of the model
+### Finding a Steady State of the model
 
@@
 ## Use Cases
-We will now show two use cases which illustrate applications of the package. All use cases can be used as starting points for other investigations of the behavior of the system.
+We will now show three use cases which illustrate applications of the package. All use cases can be used as starting points for other investigations of the behavior of the system.
 
@@
 
-Ass all objects were saved, results can be easily reproduced and parameter can be changed to assess the impact of these changes (e.g. to conduc a sensitivity analysis).
+As all objects were saved, results can be easily reproduced and parameter can be changed to assess the impact of these changes (e.g. to conduct a sensitivity analysis).
 
@@
 As discussed in the paper [@REF_NEEDED], the role biodiversity plays in abrupt regime shifts based on gradual changing environmental parameter is not well understood.
-This model (as part of the package) has been used to investigate these dynamics and the results are available in @REF_NEEDED.
+This model (as part of the package) has been used to investigate these dynamics and the results are available in with the DOI DOI NEEDED @REF_NEEDED.
 
@@
 
-The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing investigations which extend the original publication, as done in @REFNEEDED.
+The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing investigations which extend the original publication, as done in @REF_NEEDED.
 

### 2022-05-03 inst/manuscript/manuscript_microxanox.tex
Add compiled manuscript

@@
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+
 \usepackage[hyphens]{url}
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\bibliographystyle{elsarticle-harv}
-\usepackage{graphicx}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
 \ifxetex
@@
             pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
 
@@
 % Pandoc toggle for numbering sections (defaults to be off)
-
-% Pandoc syntax highlighting
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 \makesavenoteenv{longtable}
-
-% Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
-    \author[University of Zürich]{Rainer M Krug\corref{Corresponding
-Author}}
-   \ead{Rainer.Krug@uzh.ch \& Rainer@krugs.de} 
+    \author[University of Zürich]{Rainer M Krug%
+  %
+  \fnref{Corresponding Author}}
+   \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
     \author[University of Zürich]{Owen L. Petchey}
    \ead{Owen.Petchey@ieu.uzh.ch} 
-      \address[University of Zürich]{Department of Evolutionary Biology
-and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
-      \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
+      \affiliation[University of Zürich]{Department of Evolutionary
+Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
+    \cortext[cor1]{Corresponding author}
+    \fntext[1]{Corresponding Author}
+    \fntext[2]{Equal contribution}
 
@@
   \end{abstract}
-   \begin{keyword} metadata quality; data curation; archival; long term
-storage; R package;\end{keyword}
+    \begin{keyword}
+    metadata quality; data curation; archival; long term storage; R
+package; \sep 
+    metadata quality; data curation; archival; long term storage; R
+package;
+  \end{keyword}
+
  \end{frontmatter}
+
+\hypertarget{todo}{%
+\section{\texorpdfstring{\color{red}TODO}{TODO}}\label{todo}}
+
+\begin{itemize}
+\tightlist
+\item
+  \color{red}\textbf{update links}
+\item
+  \color{red}\textbf{update cross references} \pagebreak
+\end{itemize}
 
@@
 change of the environmental variable. One example where the gradual
-change of a single environmental causing an abrupt change of the system
-is the switch from an aerobic to anaerobic system. This system hass been
-investigated by Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}). We
-wanted to take this investigation one step further, and look at the role
-of an increased biodiversity plays in these dynamics
-(\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}). For this
+change of a single environmental variable is causing an abrupt change of
+the system is the switch from an aerobic to anaerobic system. This
+system has been investigated by \citet{Bush2017}. We wanted to take this
+investigation one step further, and look at the role an increased
+biodiversity plays in these dynamics \citep{REF_NEEDED}. For this
 purpose, we developed this package.
@@
 substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
-sulfur, \emph{SO}: oxidized sulfur). It includes feddback between
-biogeochemical processes and is based on
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) (See
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) for a detailed
-discussion of the model).
+sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
+biogeochemical processes and is based on \citep{Bush2017} (See
+\citep{Bush2017} for a detailed discussion of the model).
 
 The aims of the \texttt{microxanox} package are twofold. Firstly, the
-package aims at reproducing the results shown by
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}), which is
-acomplished in the \href{LINK\%20NEEDED}{vignette Partial reproduction
-of Bush et al}. Secondly, to take these results one step further, it
-includes new functionality to adress our research question as presented
-in (\protect\hyperlink{ref-REF_NEEDED}{REF NEEDED, 2222}).
+package aims at reproducing the results shown by \citep{Bush2017}, which
+is accomplished in the \href{LINK\%20NEEDED}{vignette Partial
+reproduction of Bush et al}. Secondly, to take these results one step
+further, it includes new functionality to address our research question
+as published in \citep{REF_NEEDED}.
 
@@
 In addition to the model itself, the package includes some functions to
-analyse the results as well as visualise the results to provide a
-starting point for customised visualisations based on own requirements.
+analyse the results as well as visualize the results to provide a
+starting point for customized visualizations based on own requirements.
 
@@
 
+\begin{quote}
 \{\textgreater\textgreater Describe the software in as much as is
@@
 impact.\textless\textless\}
+\end{quote}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{@startuml}
-\NormalTok{partition "Run Single Simulation" \{}
-\NormalTok{   start}
-\NormalTok{  :parameter \textless{}{-} new\_runsim\_parameter();}
-\NormalTok{  split}
-\NormalTok{    \#palegreen:saveRDS(FILE, parameter);}
-\NormalTok{    detach}
-\NormalTok{  split again}
-\NormalTok{    \#pink:result \textless{}{-} run\_simulation(parameter);}
-\NormalTok{  end split}
-\NormalTok{  split}
-\NormalTok{    \#palegreen:saveRDS(FILE, results);}
-\NormalTok{    detach}
-\NormalTok{  split again}
-\NormalTok{    :Change some parameter;}
-\NormalTok{    \#pink:result2 \textless{}{-} run\_simulation(results);}
-\NormalTok{  end split}
-\NormalTok{  split}
-\NormalTok{    \#palegreen:saveRDS(FILE, results2);}
-\NormalTok{    detach}
-\NormalTok{  split again}
-\NormalTok{    end}
-\NormalTok{  end split}
-\NormalTok{\}}
-\NormalTok{@enduml}
-\end{Highlighting}
-\end{Shaded}
+We will discuss the general structure and functionality of the package
+here without going into to much detail. A more detailed discussion can
+be found in the \href{LINK_NEEDED}{User Guide}.
 
-\includegraphics{/var/folders/50/wcr5bjwn75q595n6x82gxj280000gq/T//RtmpRw3cq4/runsim_example.pdf}
+The model by \citet{Bush2017} is based on ODEs for the rates of change
+which are specified in the function \texttt{bushplus\_dynamic\_model()}.
+This augmented version of the model published in \citep{Bush2017} can
+handle multiple strains within each of the three functional groups,
+temporal variation in oxygen diffusivity, and events.
 
@@
 be adjusted easily to all needs, but primarily a tool to implement the
-model used by (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) and
-to extend it to our needs (\protect\hyperlink{ref-REF_NEEDED}{REF
-NEEDED, 2222}). Consequently, any more substantial changes and
-adaptaitions, are likely to need a change in the source code.
+model used by \citep{Bush2017} and to extend it to our needs
+\citep{REF_NEEDED}. Consequently, any more substantial changes and
+adaptations are likely to need a change in the source code.
 
-Nevertheless, the model is structured in a way which builds on a modular
-structuere, so that e.g.~the event definition can easily be changed. or
-other aspects can be adjusted. All values in the parameter object can be
-changed as needed and the general structure of the code should make it
-not to difficult to adapt the model to other similar systems.
+Nevertheless, the source code is structured in modular, so that e.g.~the
+event definition can easily be changed, or other aspects can be
+adjusted. All values in the parameter object can be changed as needed
+and the general structure of the code should make it not to difficult to
+adapt the model to other similar systems.
 
@@
 
+\begin{quote}
 \{\textgreater\textgreater Give a short overview of the overall software
-architecture; provide a pic- torial component overview or similar (if
+architecture; provide a pictorial component overview or similar (if
 possible). If necessary provide im- plementation
 details.\textless\textless\}
+\end{quote}
 
 The framework used when writing this package aims at reproducibility of
-the results. It builds on the folowing main considerations:
+the results. It builds on the following main considerations:
 
@@
   Which one of the \texttt{new\_...\_parameter()} functions has to be
-  used when, will be discussed in the section @ref(runsim) and in the
-  \href{LINK_NEEDED}{User Guide}.
+  used when, will be discussed in the section @ref(runsim) and in more
+  detail in the \href{LINK_NEEDED}{User Guide}.
 \item
@@
   The return value of the \texttt{run\_...(parameter)} function is
-  identical to the parameter object plus an additionl slot named
+  identical to the parameter object plus an additional slot named
   \texttt{results} which contains the results of the run
 \item
-  As this return value contains all parameter, it is possible to re-run
-  the simulatuion by simply running \texttt{run\_...(result)}.
+  As the return value contains all parameter, it is possible to re-run
+  the simulation by simply running \texttt{run\_...(result)}.
 \end{enumerate}
@@
 
-A typical simulation would look as followed:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\DocumentationTok{\#\# Create the parameter}
-\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{new\_runsim\_parameter}\NormalTok{()}
-\CommentTok{\# manually setting certain parameter}
-
-\DocumentationTok{\#\# Run the simulation and save the result}
-\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulation}\NormalTok{(parameter)}
-\FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim1.rds"}\NormalTok{)}
-
-\DocumentationTok{\#\# Do other stuff, e.g. plotting}
-
-\DocumentationTok{\#\# Load results, change some parameter, and rerun the simulation and save the result}
-\NormalTok{parameter }\OtherTok{\textless{}{-}} \FunctionTok{loadRDS}\NormalTok{(}\StringTok{"sim1.rds"}\NormalTok{)}
-\CommentTok{\# change some parameter}
-\NormalTok{result }\OtherTok{\textless{}{-}} \FunctionTok{run\_simulation}\NormalTok{(parameter)}
-\FunctionTok{saveRDS}\NormalTok{(result, }\StringTok{"sim2,rds"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{software-functionalities}{%
-\subsection{Software Functionalities}\label{software-functionalities}}
-
-\{\textgreater\textgreater Present the major functionalities of the
-software.\textless\textless\}
-
-We will now discuss the general structure and functionality of the
-package without going into to much detail. A more detailed discussion
-can be found in the \href{LINK_NEEDED}{User Guide}.
-
-The ODEs for the rates of change are specified in the function
-\texttt{bushplus\_dynamic\_model()}. This augmented version of the model
-published in (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}) can
-handle multiple strains within each of the three functional groups,
-temporal variation in oxygen diffusivity, and events.
-
 In the following sections we describe the general usage of the package:
@@
 
-\hypertarget{illustrative-examples}{%
-\section{Illustrative Examples}\label{illustrative-examples}}
-
-\{\textgreater\textgreater Present the major functionalities of the
-software.\textless\textless\}
-
 \hypertarget{running-one-simulation}{%
 \subsubsection{Running one simulation}\label{running-one-simulation}}
+
+A typical simulation would look as shown in @ref(fig:runsim\_example).
+
+\begin{verbatim}
+#> [1] "./figures/simflow.pdf"
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics[width=350px]{./figures/simflow}
+
+}
+
+\caption{Typical flow of a simulation.
+Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
+\end{figure}
 
@@
 \texttt{new\_strain\_parameter()}. For a detailed description of the
-parameter and how they are created please see the User Guide and which
-acomnpanies the package or is available at \href{@LINK_NEEDED}{User
-Guide}
+parameter, their meaning and how they are created please see the User
+Guide which accompanies the package or is available at
+\href{@LINK_NEEDED}{User Guide}
 
@@
 The function \texttt{plot\_dynamics()} plots a single simulation run, as
-returned from the \texttt{run\_simulation()} function. This functio is
+returned from the \texttt{run\_simulation()} function. This function is
 only provided as a convenience function to provide a way to easily see
@@
 
-{\centering \includegraphics[width=500px]{figures/ug_three_strains_dynamics}
+{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}
 
@@
 which takes a parameter object as created by the function
-\texttt{new\_temporal\_ssfind\_parameter()} and a number indicating the
-.
+\texttt{new\_temporal\_ssfind\_parameter()}.
 
@@
 
-\hypertarget{impact}{%
-\section{Impact}\label{impact}}
+\hypertarget{use-cases}{%
+\subsection{Use Cases}\label{use-cases}}
 
-\{\textgreater\textgreater This is the main section of the article and
-the reviewers weight the description here appropriately Indicate in what
-way new research questions can be pursued as a result of the software
-(if any). Indicate in what way, and to what extent, the pursuit of
-existing research questions is improved (if so). Indicate in what way
-the software has changed the daily practice of its users (if so).
-Indicate how widespread the use of the software is within and outside
-the intended user group. Indicate in what way the software is used in
-commercial settings and/or how it led to the creation of spin-off
-companies (if so).\textless\textless\}
+We will now show two use cases which illustrate applications of the
+package. All use cases can be used as starting points for other
+investigations of the behavior of the system.
 
 The first two use cases are described in detail in the User Guide and
-the Partial Reproduction Vignettes. The third is taken from the REF
-NEEDED (\protect\hyperlink{ref-REF_NEEDED}{2222}) for which this R
-package was designed. All of these use cases can be expanded to larger
-numbers of strains per functional group and variable values van be
-changed.
+the Partial Reproduction Vignettes. The third is taken from
+\citet{REF_NEEDED} for which this R package was designed. All of these
+use cases can be expanded to larger numbers of strains per functional
+group and variable values van be changed.
 
 \hypertarget{regime-shifts-during-temporal-environmental-change}{%
-\subsection{Regime shifts during temporal environmental
+\subsubsection{Regime shifts during temporal environmental
 change}\label{regime-shifts-during-temporal-environmental-change}}
@@
 
-{\centering \includegraphics[width=500px]{figures/ug_three_strains_stable_state}
+{\centering \includegraphics[width=350px]{figures/ug_three_strains_stable_state}
 
@@
 
+We saved all parameter and result objects, so that all results are
+completely reproducible. The general flow of the experiment is identical
+to the one shown in @ref(fig:runsim\_example).
+
 \hypertarget{the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsection{The extent of hysteresis depends on community
+\subsubsection{The extent of hysteresis depends on community
 composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}
@@
 One of the reasons to develop this package was to reproduce the results
-presented in Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}), this
-was achieved as demonstrated in the \href{LINK_NEEDED}{Partial
-Reproduction supplement}. All aspects in the paper could be reproduced
-and are shown in the vignette.
+presented in \citet{Bush2017}. This was achieved as demonstrated in the
+\href{LINK_NEEDED}{Partial Reproduction supplement}. All aspects in the
+paper could be reproduced and are shown in the vignette.
 
@@
 
-{\centering \includegraphics[width=500px]{figures/user_guide_hysteresis}
+{\centering \includegraphics[width=350px]{figures/user_guide_hysteresis}
 
@@
 
+Ass all objects were saved, results can be easily reproduced and
+parameter can be changed to assess the impact of these changes (e.g.~to
+conduc a sensitivity analysis).
+
 \hypertarget{effects-of-functional-diversity-on-regime-shifts}{%
-\subsection{Effects of functional diversity on regime
+\subsubsection{Effects of functional diversity on regime
 shifts}\label{effects-of-functional-diversity-on-regime-shifts}}
 
-As discussed in the paper (\protect\hyperlink{ref-REF_NEEDED}{REF
-NEEDED, 2222}), the role biodiversity plays in abrupt regime shifts
-based on gradual changing environmental parameter is not well
-understood. This model (as part of the package) has been used to
-investigate these dynamics and the results are available in REF NEEDED
-(\protect\hyperlink{ref-REF_NEEDED}{2222}).
+As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
+plays in abrupt regime shifts based on gradual changing environmental
+parameter is not well understood. This model (as part of the package)
+has been used to investigate these dynamics and the results are
+available in \citet{REF_NEEDED}.
 
@@
 provide two or three sentences\textless\textless\}}
+
+\hypertarget{impact}{%
+\section{Impact}\label{impact}}
+
+\begin{quote}
+\{\textgreater\textgreater This is the main section of the article and
+the reviewers weight the description here appropriately Indicate in what
+way new research questions can be pursued as a result of the software
+(if any). Indicate in what way, and to what extent, the pursuit of
+existing research questions is improved (if so). Indicate in what way
+the software has changed the daily practice of its users (if so).
+Indicate how widespread the use of the software is within and outside
+the intended user group. Indicate in what way the software is used in
+commercial settings and/or how it led to the creation of spin-off
+companies (if so).\textless\textless\}
+\end{quote}
+
+The free and open source implementation and extension of the model used
+in \citet{Bush2017} provides the means of reproducing the results
+published while at the same time provides the means of doing
+investigations which extend the original publication, as done in
+\citet{REF_NEEDED}.
+
+The combination of all parameter in a single parameter object as well as
+the return of the simulation as a result object which inherits from the
+parameter object and adds a results slot provides a relatively easy to
+use framework to implement reproducible experiments.
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
-Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
-2017. Oxic-anoxic regime shifts mediated by feedbacks between
-biogeochemical processes and microbial community dynamics. Nature
-Communications 8, 789.
-doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-REF_NEEDED}{}}%
-REF NEEDED, A., 2222. {REFERNECE NEEDED}. Journal of missing references.
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{bibliography.bib}
 

### 2022-05-06 inst/manuscript/manuscript_microxanox.Rmd
Main Changes done

@@
 ---
-title: "Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
+title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
 author:
@@
     text: "Equal contribution"
-keywords: metadata quality; data curation; archival; long term storage; R package;
+keywords: reproducibility; regime shift; 
 abstract: |
@@
 
-<!--
-> Abstract Romana of paper: The role of functional diversity in ecosystem regime shifts
-
-> Biodiversity can increase the resilience of ecosystems facing environmental change.
-However, we have limited understanding if this holds true for ecosystems that respond to gradual environmental change with abrupt regime shifts.
-
-> Here we investigated if functional diversity influences the resilience of ecosystems that abruptly transition between alternative ecosystem states.
-As a case study, we used a mathematical model developed by Bush et al.
-(2017) which describes anoxic-oxic regime shifts of aquatic ecosystems mediated by cyanobacteria, sulfate-reducing bacteria, and phototrophic sulfur bacteria.
-We extended this model so as to include functional diversity in one, two, or all three functional groups.
-Specifically, we manipulated the amount of variation in environmental tolerance and in maximum growth rate, and explored how trait variation influenced anoxic-oxic regime shifts along a gradient of oxygen influx.
-We found that greater diversity of either cyanobacteria or sulfate-reducing bacteria increased the resilience of the ecosystem state that these groups dominated, while greater diversity of phototrophic sulfur bacteria reduced the resilience of the state they dominated.
-We also found that simultaneous variation in multiple functional groups reduced, reversed, or did not change the diversity effect of individual functional groups.
-Collectively, our findings highlight the importance of considering multiple interacting functional groups when predicting the effect of functional diversity on ecosystem resilience
- -->
-
-# \color{red}TODO
-- \color{red}**update links** 
-- \color{red}**update cross references**
+# \textcolor{red}{TODO}
+- **\textcolor{red}{update links}**
+- **\textcolor{red}{update cross references}**
+- **\textcolor{red}{check keywords}**
+- **\textcolor{red}{add more references}**
+- **\textcolor{red}{expand on Impact and Conclusion section}**
 \pagebreak
@@
 | **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
-|:-------:|:-------------------------------------------------------------|:-------------------------------------------------------------------------------|
+| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
 |   C1    | Current code version                                         | v0.9.0                                                                         |  
@@
 |   C5    | Code versioning system used                                  | git                                                                            |  
-| C6      | Software code languages, tools, and services used            | R                                                                              |
-| C7      | Compilation requirements, operating environments             | R (>= 4.1.0), ADD PACKAGES                                                     |
-| C8      | If available Link to developer documentation/manual          |                                                                                |
+|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
+|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
+|         |                                                              | magrittr                                                                       |  
+|         |                                                              | tibble                                                                         |  
+|         |                                                              | ggplot2                                                                        |  
+|         |                                                              | patchwork                                                                      |  
+|         |                                                              | grDevices                                                                      |  
+|         |                                                              | stats                                                                          |  
+|         |                                                              | mgcv                                                                           |  
+|         |                                                              | deSolve                                                                        |  
+|         |                                                              | dplyr                                                                          |  
+|         |                                                              | tidyr                                                                          |  
+|         |                                                              | stringr                                                                        |  
+|         |                                                              | multidplyr                                                                     |  
+|         |                                                              |                                                                                |  
+|         |                                                              |                                                                                |  
+|         |                                                              |                                                                                |  
+|         |                                                              |                                                                                |  
+|         |                                                              |                                                                                |  
+|   C8    | If available Link to developer documentation/manual          | \textcolor{red}{TO BE ADDED}                                                   |  
 |   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |
@@
 # Motivation and significance
-Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable.
+Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable [@REF_NEEDED].
 One example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
@@
 
-The aims of the `microxanox` package are twofold.
-Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED).
-Secondly, to take these results one step further, it includes new functionality to address our research question as published in [@REF_NEEDED].
+The aims of the `microxanox` package are twofold: firstly, to reproduce the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED) and secondly, to take these results one step further, to includes new functionality to address our research question as published in [@REF_NEEDED].
 
@@
 
-The framework used when writing this package aims at reproducibility of the results. It builds on the following main considerations:
+The framework used when writing this package aims at reproducibility of the results (see [@REF_NEEDED] for examples). It builds on the following main considerations:
 
@@
 
-The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier.
+The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier, as can be seen in our research and the RMarkdown [@REF_NEEDED] file used to run and analyse the experiments [@REF_NEEDED].
 
@@
 The individual simulation (`run_simulation()` function) is the working horse in this package.
-In this function, the ODEs are solved.
-The function needs only one argument - an object as created by the function `new_runsim_parameter()`.
-One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter, their meaning and how they are created please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)
+In this function, the ODEs are solved using the function `ode` in the package deSolve package [@Soetaert2010].
+The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`.
+The parameter object contains among other parameter the `strain_parameter` which can be created by the function `new_strain_parameter()`.
+For a detailed description of the parameter, their meaning and how they are created and can be set and changed please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)
 
 After the parameter object has been defined, it can be used in the `run_simulation()` function.
-The function returns an object which is identical to the parameter, except of an additional slot containing the results.
-This design produces a fully reproducible object as it can be used instead of a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results from.
+The function returns an object which is identical to the parameter object, except of an additional slot containing the results.
+This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results before.
 
@@
 ### Finding a Steady State of the model
+\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}
 
-There are two methods for finding steady states.
-The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*).
-The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*).
+There are two methods for finding steady states implemented in this package.
+The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*) \textcolor{red}{{>>@Owen - does this approach has a name, and do you have a reference where it has been used or described?<<}.
+The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*)\textcolor{red}{{>>@Owen - Has this been used in other publications?
+I do not have a clue what I could search for to find out...<<}.
 
@@
 The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
+As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac.
+It is planned to move to `parLapply()` [@RCoreTeam2022] in a future release.
 
@@
 That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set.
-Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step.
+Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step \textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.
 
 This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
+It is planned for a later release, to run these two simulations in parallel.
 
@@
 From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
-These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change.
+These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change \textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g.
+used in REF?<<}}.
 
@@
 ### Regime shifts during temporal environmental change
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
 
@@
 ### The extent of hysteresis depends on community composition
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
 
@@
 ### Effects of functional diversity on regime shifts
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
 
@@
 
-<span style="color:green">{>>More details from the paper.<<}
-{>>Romana will provide two or three sentences<<}</span>
+\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}
 
@@
 
+\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}
+
 # Conclusions
-Set out the conclusion of this original software publication.
+> {>>Set out the conclusion of this original software publication.<<}
+
+This package makes an important paper \textcolor{red}{{>>@Owen: is this correct?<<}} [@Bush2017] fully reproducible without having to re-implement the code.
+By doing so, it opens the door for further investigations based on the system as done in @REF_NEEDED and, consequently, increases the reach (and ultimaly) the impact of the original paper by @Bush2017.
+This demonstrates again [@REF_NEEDED] the importance of publishing not only the data but also the code used in a publication.
+
+\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}
+
+In addition to the ecological relevance and importance of this package as discussed above, it can be used as a template for the implementation of other different models.
+By doing so, other models can profit from the reproducibility aspects as well as the flexibility as implemented in this package.
 

### 2022-05-06 inst/manuscript/manuscript_microxanox.tex
Main Changes done

@@
             pdfauthor={},
-            pdftitle={Microxanox - an R package for simulating an MICRobial ecosystem that can occupy OXic or ANOXic states.},
+            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
             colorlinks=true,
@@
 
-  \title{Microxanox - an R package for simulating an \(MICR\)obial
-ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
+  \title{Microxanox: an R package for simulating an aquatic
+\(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
     \author[University of Zürich]{Rainer M Krug%
@@
     \begin{keyword}
-    metadata quality; data curation; archival; long term storage; R
-package; \sep 
-    metadata quality; data curation; archival; long term storage; R
-package;
+    reproducibility; regime shift; \sep 
+    reproducibility; regime shift;
   \end{keyword}
@@
 
-\hypertarget{todo}{%
-\section{\texorpdfstring{\color{red}TODO}{TODO}}\label{todo}}
+\hypertarget{section}{%
+\section{\texorpdfstring{\textcolor{red}{TODO}}{}}\label{section}}
 
@@
 \item
-  \color{red}\textbf{update links}
+  \textbf{\textcolor{red}{update links}}
 \item
-  \color{red}\textbf{update cross references} \pagebreak
+  \textbf{\textcolor{red}{update cross references}}
+\item
+  \textbf{\textcolor{red}{check keywords}}
+\item
+  \textbf{\textcolor{red}{add more references}}
+\item
+  \textbf{\textcolor{red}{expand on Impact and Conclusion section}}
+  \pagebreak
 \end{itemize}
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0596}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4106}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5298}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
 \toprule
@@
 C5 & Code versioning system used & git \\
-C6 & Software code languages, tools, and services used & R \\
-C7 & Compilation requirements, operating environments & R (\textgreater=
-4.1.0), ADD PACKAGES \\
-C8 & If available Link to developer documentation/manual & \\
+C6 & Software code languages, tools, and services used &
+\href{https://cran.r-project.org/index.html}{R} \\
+C7 & Compilation requirements, operating environments &
+\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
+& & magrittr \\
+& & tibble \\
+& & ggplot2 \\
+& & patchwork \\
+& & grDevices \\
+& & stats \\
+& & mgcv \\
+& & deSolve \\
+& & dplyr \\
+& & tidyr \\
+& & stringr \\
+& & multidplyr \\
+& & \\
+& & \\
+& & \\
+& & \\
+& & \\
+C8 & If available Link to developer documentation/manual &
+\textcolor{red}{TO BE ADDED} \\
 C9 & Support email for questions & Rainer.Krug@uzh.ch;
@@
 variables, to which the responses are not always as gradual as the
-change of the environmental variable. One example where the gradual
-change of a single environmental variable is causing an abrupt change of
-the system is the switch from an aerobic to anaerobic system. This
-system has been investigated by \citet{Bush2017}. We wanted to take this
-investigation one step further, and look at the role an increased
-biodiversity plays in these dynamics \citep{REF_NEEDED}. For this
-purpose, we developed this package.
+change of the environmental variable \citep{REF_NEEDED}. One example
+where the gradual change of a single environmental variable is causing
+an abrupt change of the system is the switch from an aerobic to
+anaerobic system. This system has been investigated by \citet{Bush2017}.
+We wanted to take this investigation one step further, and look at the
+role an increased biodiversity plays in these dynamics
+\citep{REF_NEEDED}. For this purpose, we developed this package.
 
@@
 
-The aims of the \texttt{microxanox} package are twofold. Firstly, the
-package aims at reproducing the results shown by \citep{Bush2017}, which
-is accomplished in the \href{LINK\%20NEEDED}{vignette Partial
-reproduction of Bush et al}. Secondly, to take these results one step
-further, it includes new functionality to address our research question
-as published in \citep{REF_NEEDED}.
+The aims of the \texttt{microxanox} package are twofold: firstly, to
+reproduce the results shown by \citep{Bush2017}, which is accomplished
+in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et
+al} and secondly, to take these results one step further, to includes
+new functionality to address our research question as published in
+\citep{REF_NEEDED}.
 
@@
 The framework used when writing this package aims at reproducibility of
-the results. It builds on the following main considerations:
+the results (see \citep{REF_NEEDED} for examples). It builds on the
+following main considerations:
 
@@
 the simulation, promotes reproducibility and makes incremental changes
-of individual parameters and re-running the simulations much easier.
+of individual parameters and re-running the simulations much easier, as
+can be seen in our research and the RMarkdown \citep{REF_NEEDED} file
+used to run and analyse the experiments \citep{REF_NEEDED}.
 
@@
 A typical simulation would look as shown in @ref(fig:runsim\_example).
-
-\begin{verbatim}
-#> [1] "./figures/simflow.pdf"
-\end{verbatim}
 
@@
 The individual simulation (\texttt{run\_simulation()} function) is the
-working horse in this package. In this function, the ODEs are solved.
-The function needs only one argument - an object as created by the
-function \texttt{new\_runsim\_parameter()}. One parameter of this object
-is the \texttt{strain\_parameter} which can be created by the function
-\texttt{new\_strain\_parameter()}. For a detailed description of the
-parameter, their meaning and how they are created please see the User
-Guide which accompanies the package or is available at
-\href{@LINK_NEEDED}{User Guide}
+working horse in this package. In this function, the ODEs are solved
+using the function \texttt{ode} in the package deSolve package
+\citep{Soetaert2010}. The \texttt{run\_simulation()} function needs only
+one argument - an object as created by the function
+\texttt{new\_runsim\_parameter()}. The parameter object contains among
+other parameter the \texttt{strain\_parameter} which can be created by
+the function \texttt{new\_strain\_parameter()}. For a detailed
+description of the parameter, their meaning and how they are created and
+can be set and changed please see the User Guide which accompanies the
+package or is available at \href{@LINK_NEEDED}{User Guide}
 
@@
 \texttt{run\_simulation()} function. The function returns an object
-which is identical to the parameter, except of an additional slot
+which is identical to the parameter object, except of an additional slot
 containing the results. This design produces a fully reproducible object
-as it can be used instead of a parameter object to be fed back into the
+as it can be used as a parameter object to be fed back into the
 \texttt{run\_simulation()} function to run the simulation again from the
-parameter used to generate the results from.
+parameter used to generate the results before.
 
@@
 
-There are two methods for finding steady states. The first runs a
-separate simulation for each combination of starting conditions and
-oxygen diffusivity (let us term this the \emph{Replication method}). The
-second runs only two simulations, with step-wise and slowly temporally
-increasing or decreasing oxygen diffusivities and recorded of state just
-before change to a new oxygen diffusivity (let us term this the
-\emph{Temporal method}).
+\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}
+
+There are two methods for finding steady states implemented in this
+package. The first runs a separate simulation for each combination of
+starting conditions and oxygen diffusivity (let us term this the
+\emph{Replication method})
+\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
+- does this approach has a name, and do you have a reference where it
+has been used or described?\textless\textless\}. The second runs only
+two simulations, with step-wise and slowly temporally increasing or
+decreasing oxygen diffusivities and recorded of state just before change
+to a new oxygen diffusivity (let us term this the \emph{Temporal
+method})\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
+- Has this been used in other publications? I do not have a clue what I
+could search for to find out\ldots\textless\textless\}.
 
@@
 returned by the function \texttt{new\_replication\_ssfind\_parameter()}
-and the number of cores for multithreading the simulation.
+and the number of cores for multithreading the simulation. As the
+multithreading uses the package function \texttt{mclapply()} from the
+package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
+works on Linux and Mac. It is planned to move to \texttt{parLapply()}
+\citep{RCoreTeam2022} in a future release.
 
@@
 effectively started with initial conditions that are the state of the
-system in the previous time step.
+system in the previous time step
+\textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.
 
@@
 which takes a parameter object as created by the function
-\texttt{new\_temporal\_ssfind\_parameter()}.
+\texttt{new\_temporal\_ssfind\_parameter()}. It is planned for a later
+release, to run these two simulations in parallel.
 
@@
 non-linearity and hysteresis measures, of the response of the simulated
-system to environmental change.
+system to environmental change
+\textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g. used in REF?<<}}.
 
@@
 
-We will now show two use cases which illustrate applications of the
+We will now show three use cases which illustrate applications of the
 package. All use cases can be used as starting points for other
@@
 change}\label{regime-shifts-during-temporal-environmental-change}}
+
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
 
@@
 
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
+
 One of the reasons to develop this package was to reproduce the results
@@
 
-Ass all objects were saved, results can be easily reproduced and
+As all objects were saved, results can be easily reproduced and
 parameter can be changed to assess the impact of these changes (e.g.~to
-conduc a sensitivity analysis).
+conduct a sensitivity analysis).
 
@@
 
+\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
+
 As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
@@
 has been used to investigate these dynamics and the results are
-available in \citet{REF_NEEDED}.
+available in with the DOI DOI NEEDED \citet{REF_NEEDED}.
 
-{\{\textgreater\textgreater More details from the
-paper.\textless\textless\} \{\textgreater\textgreater Romana will
-provide two or three sentences\textless\textless\}}
+\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}
 
@@
 
+\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}
+
 \hypertarget{conclusions}{%
@@
 
-Set out the conclusion of this original software publication.
+\begin{quote}
+\{\textgreater\textgreater Set out the conclusion of this original
+software publication.\textless\textless\}
+\end{quote}
+
+This package makes an important paper
+\textcolor{red}{{>>@Owen: is this correct?<<}} \citep{Bush2017} fully
+reproducible without having to re-implement the code. By doing so, it
+opens the door for further investigations based on the system as done in
+\citet{REF_NEEDED} and, consequently, increases the reach (and ultimaly)
+the impact of the original paper by \citet{Bush2017}. This demonstrates
+again \citep{REF_NEEDED} the importance of publishing not only the data
+but also the code used in a publication.
+
+\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}
+
+In addition to the ecological relevance and importance of this package
+as discussed above, it can be used as a template for the implementation
+of other different models. By doing so, other models can profit from the
+reproducibility aspects as well as the flexibility as implemented in
+this package.
 

### 2022-05-09 inst/manuscript/manuscript_microxanox.Rmd
a few hours work on the manuscript

@@
 # Motivation and significance
-Many ecosystems are exposed to gradual changes of environmental variables, to which the responses are not always as gradual as the change of the environmental variable [@REF_NEEDED].
-One example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic to anaerobic system.
-This system has been investigated by @Bush2017.
-We wanted to take this investigation one step further, and look at the role an increased biodiversity plays in these dynamics [@REF_NEEDED].
-For this purpose, we developed this package.
 
-The microxanox package is a package for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
-It includes feedback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
+Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@REF_NEEDED]
+- Chaparro-Pedraza, P.C. (2021). Fast environmental change and eco-evolutionary feedbacks can drive regime shifts in ecosystems before tipping points are crossed. Proc. R. Soc. B., 288, 20211192.
+- Binzer, A., Guill, C., Rall, B.C.
+& Brose, U.
+(2016).
+Interactive effects of warming, eutrophication and size structure: impacts on biodiversity and food-web structure.
+Global Change Biology, 22, 220–227.
+- Vasseur, D. a & McCann, K.S. (2005). A mechanistic approach for modeling temperature-dependent consumer-resource dynamics. The American Naturalist, 166, 184–198.
 
-The aims of the `microxanox` package are twofold: firstly, to reproduce the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED) and secondly, to take these results one step further, to includes new functionality to address our research question as published in [@REF_NEEDED].
+.
+They are also critical for developing hypotheses that can then be tested in experimental studies.
+One area of ecology in which simple models have played a very influential role is the area of how ecosystems respond to gradual environmental change [@REF_NEEDED].
+- Scheffer, M., Carpenter, S., Foley, J. a, Folke, C. & Walker, B. (2001). Catastrophic shifts in ecosystems. Nature, 413, 591–6.
 
-For this, we extended the model and added functionality for:
+It is conceivable that an ecosystem state, such as the total biomass of a particular type of bacteria, may remain unchanged when something about its environment changes, such as the temperature.
+It is possible that the ecosystem state changes gradually itself.
+It is also possible that the ecosystem state may change abruptly into a new state that is difficult to recover from [@REF_NEEDED]
+- Scheffer, M., Carpenter, S., Foley, J. a, Folke, C. & Walker, B. (2001). Catastrophic shifts in ecosystems. Nature, 413, 591–6.
 
-- Multiple strains (effectively unlimited) per functional group.
-- Adding temporally varying oxygen diffusivity.
-- Adding random noise in substrate concentrations.
-- Including immigration.
-- Setting minimum population abundances.
+This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern [@REF_NEEDED]. 
+- Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P. & Ratajczak, Z. (2021). Fire frequency, state change and hysteresis in tallgrass prairie. Ecology Letters, 24, 636–647.
+- Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A.
+& Gotelli, N.J.
+(n.d.).
+Clockwise and counterclockwise hysteresis characterize state changes in the same aquatic ecosystem.
+Ecology Letters, n/a.
+- Vandermeer, J. & Perfecto, I. (2019). Hysteresis and critical transitions in a coffee agroecosystem. PNAS, 116, 15074–15079.
 
+An example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic (oxygen is available for metabolism) to anaerobic (oxygen generally unavailable) system.
+This system has been investigated by @Bush2017 in a simulation study of a mathematical model of the system.
+They showed that gradual change in the environment could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+
+One limitation of the study by Bush et al (2017) was limited biodiversity in the ecosystem model.
+Specifically, there were three functional groups of bacteria, but within each of these groups, there was no biodiversity.
+This then leaves open the question of if and how biodiversity within functional groups, in their model, affects the ecosystem response to environmental change.
+This limitation is not limited to the study of Bush et al (2017): there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
+
+We decided to fill this research gap, by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger_et_al_2022, and to base our work on the work and model of Bush et al (2017).
+It was with this goal in mind that we developed the *microxanox* package.
+The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
+The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](LINK NEEDED).
+
+The second stage was to add functionality that would be necessary to answer our research question.
+Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them.
+We also added functionality that allowed: temporally varying environmental conditions, addition of random noise to state variables, and immigration.
 In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements.
+The basic and additional functionality is described in the package [User Guide](LINK_NEEDED).
 
@@
 
-We will discuss the general structure and functionality of the package here without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).
+The *microxanox* package is for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
+It includes feedback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
+At the core of the simulations is a set of ordinary differential equations (specified in the function `bushplus_dynamic_model()`, though users do not need to call this function directly).
+There are functions for running an individual simulations and for running a set of simulations across, for example, a range of environmental conditions.
 
-The model by @Bush2017 is based on ODEs for the rates of change which are specified in the function `bushplus_dynamic_model()`.
-This augmented version of the model published in [@Bush2017] can handle multiple strains within each of the three functional groups, temporal variation in oxygen diffusivity, and events.
+To make the simulation run with multiple species, we expressed different species characteristics in the elements of vectors and matrices.
+We also adapted the code for the ordinary differential equations to include the vectors and matrices, and to use matrix mathematics.
+In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
 
-The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED].
-Consequently, any more substantial changes and adaptations are likely to need a change in the source code.
-
-Nevertheless, the source code is structured in modular, so that e.g.
+The package functions and code are structured in modular, so that some new functionality can be easily added.
+E.g.
 the event definition can easily be changed, or other aspects can be adjusted.
@@
 
-## Software Architecture
-> {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide im- plementation details.<<}
+## Software architecture
+> {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.<<}
 
-The framework used when writing this package aims at reproducibility of the results (see [@REF_NEEDED] for examples). It builds on the following main considerations:
+The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see [@REF_NEEDED] for examples).
+As such, all the parameters needed to run a simulation or find a stable state are contained in a single object (which can easily be created using included functions).
+This parameter object is given to a function that runs the simulations and returns the results.
+The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the results of the run.
+Thus the returned results object contains the simulation conditions (parameters) as well as the rsults, and can be used to run the simulation again.
+This promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations straightforward.
 
-1.
-all parameter needed to run a simulation or find a stable state are contained in a single parameter object.
-This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`.
-Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim) and in more detail in the [User Guide](LINK_NEEDED).
-2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
-3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additional slot named `results` which contains the results of the run
-4. As the return value contains all parameter, it is possible to re-run the simulation by simply running `run_...(result)`.
+In the following sections we describe how to use the package to run one simulation and to find steady states across an environmental gradient.
 
-The point that the results object contains all parameter needed to run the simulation, promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations much easier, as can be seen in our research and the RMarkdown [@REF_NEEDED] file used to run and analyse the experiments [@REF_NEEDED].
-
-In the following sections we describe the general usage of the package: running one simulation, finding steady states across an environmental gradient, calculating measures of stability, and visualization.
-
-### Running one simulation
+## Running one simulation
 
@@
 
-The individual simulation (`run_simulation()` function) is the working horse in this package.
+A simulation is run using the `run_simulation()` function.
 In this function, the ODEs are solved using the function `ode` in the package deSolve package [@Soetaert2010].
 The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`.
-The parameter object contains among other parameter the `strain_parameter` which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter, their meaning and how they are created and can be set and changed please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)
+The parameter object returned by `new_runsim_parameter()` contains among other things the `strain_parameter` object, which can be created by the function `new_strain_parameter()`.
+For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](@LINK_NEEDED)
 
@@
 This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results before.
+
+## Finding Steady States
+
+The general approach used to find the stable state of the system with a specific parameter set is to run the simulation for a long time and record the final state.
+When one does this across a range of environmental conditions, one discovers how the steady state of the system responds to the environmental conditions.
+The package contains functionality for finding steady states that correspond to values of one environmental variable, namely the value of oxygen diffusivity.
+
+Two methods for finding steady states are implemented.
+The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*).
+This is the method used in the Bush et al (2017) study.
+The second runs only two simulations, with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing or decreasing oxygen diffusivity.
+During this temporal environmental change, the state of the system was recorded just before change to a new oxygen diffusivity (let us term this the *Temporal method*).
+We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method.
+
+The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
+As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac.
+It is planned to move to `parLapply()` [@RCoreTeam2022] in a future release.
+
+This temporal method implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
+It is planned for a later release, to run these two simulations in parallel.
+
+For a more detailed walk-through of these two approaches and explanation please see the [User Guide](@LINK_NEEDED).
+
+## Visualising and analysing results
+
+From the raw results returned can be extracted summary quantity about how the ecoystem stable states respond to environmental change.
+The function `get_stability_measures()` returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
 
@@
 
-### Finding a Steady State of the model
-\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}
+# Impact
 
-There are two methods for finding steady states implemented in this package.
-The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*) \textcolor{red}{{>>@Owen - does this approach has a name, and do you have a reference where it has been used or described?<<}.
-The second runs only two simulations, with step-wise and slowly temporally increasing or decreasing oxygen diffusivities and recorded of state just before change to a new oxygen diffusivity (let us term this the *Temporal method*)\textcolor{red}{{>>@Owen - Has this been used in other publications?
-I do not have a clue what I could search for to find out...<<}.
-
-#### Replication Method
-The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
-As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac.
-It is planned to move to `parLapply()` [@RCoreTeam2022] in a future release.
-
-#### Temporal Method
-The temporal method involves two simulations for a particular system configuration (parameter set).
-In one simulation the oxygen diffusivity is *increased* in a step-wise fashion.
-In the other it is *decreased* in a step-wise fashion.
-That is, oxygen diffusivity is held at a constant level for long enough for steady state to be reach, that state is recorded, and then a slightly higher (or lower) oxygen diffusivity value is set.
-Hence, at that time point, the system is effectively started with initial conditions that are the state of the system in the previous time step \textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.
-
-This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
-It is planned for a later release, to run these two simulations in parallel.
-
-For a more detailed walk-through of these two approaches and explanation please see the [User Guide](@LINK_NEEDED).
-
-### Extracting Stability Measures
-From the raw results returned by these `run_...()` functions, the stability measures can be extracted by using the function `get_stability_measures()`.
-These measures include non-linearity and hysteresis measures, of the response of the simulated system to environmental change \textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g.
-used in REF?<<}}.
-
-## Use Cases
-We will now show three use cases which illustrate applications of the package. All use cases can be used as starting points for other investigations of the behavior of the system.
-
-The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes.
-The third is taken from @REF_NEEDED for which this R package was designed.
-All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.
-
-### Regime shifts during temporal environmental change
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-In the [User Guide](LINK_NEEDED) we used a one strain system (section "1 strain per functional group") and three strain system (section "3 strains per functional group") to determine as an example the stable states during temporal environmental changes (the oxygen diffusivity).
-From these simulatiuons, we extracted measures of nonlinearity and hysteresis.
-See Fig \@ref(fig:plot-dynamics) as an example plot of the simulations.
-
-```{r uc1_stable_state, echo = FALSE,  include=TRUE, fig.cap='Plot of the stable states of the simulation runs under different oxygen diffusivity.
-The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities.
-Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 350 }
-knitr::include_graphics("figures/ug_three_strains_stable_state.png")
-```
-
-We saved all parameter and result objects, so that all results are completely reproducible. The general flow of the experiment is identical to the one shown in \@ref(fig:runsim_example).
-
-### The extent of hysteresis depends on community composition
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-One of the reasons to develop this package was to reproduce the results presented in @Bush2017.
-This was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED).
-All aspects in the paper could be reproduced and are shown in the vignette.
-
-```{r user_guide_hysteresis, echo = FALSE,  include=TRUE, fig.cap='Hysteresis of all assessed combinations of variability.', fig.align="center", out.width = 350 }
-knitr::include_graphics("figures/user_guide_hysteresis.png")
-```
-
-As all objects were saved, results can be easily reproduced and parameter can be changed to assess the impact of these changes (e.g. to conduct a sensitivity analysis).
-
-### Effects of functional diversity on regime shifts
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-As discussed in the paper [@REF_NEEDED], the role biodiversity plays in abrupt regime shifts based on gradual changing environmental parameter is not well understood.
-This model (as part of the package) has been used to investigate these dynamics and the results are available in with the DOI DOI NEEDED @REF_NEEDED.
-
-\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}
-
-# Impact
 > {>>This is the main section of the article and the reviewers weight the description here appropriately
@@
 
-The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing investigations which extend the original publication, as done in @REF_NEEDED.
+The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems respond to environmental change.
+The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
 
-The combination of all parameter in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object and adds a results slot provides a relatively easy to use framework to implement reproducible experiments.
+While the package was not intended to provide a code which can be very easily adapted to other types of ecosystems and types of organism, it seems likely that the overall framework could be used as a basis for models of other ecosystems and organisms.
+For example, all such models would have parameters that differ among species and need to be described in an object.
+All would need to run simulations and sets of simulations across environmental conditions.
+Researchers want to model a new ecosystem do not, therefore, have to start from scratch.
+Nevertheless, depending on how different were the new ecosystem to model, the require code changes could be quite significant.
 
-\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}
+We evidence the impact of the *microxanox* package by describing three use cases, each of which was only possible with the package and that contains new results.
+The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes.
+The third is taken from @REF_NEEDED for which this R package was designed.
+All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.
+
+## Use case 1:  Regime shifts during temporal environmental change
+
+The package contains functionality to make a specific pattern of temporal change in the environmental condition oxygen diffusivity; this functionality forms the basis of the temporal method for finding stable states.
+A simple example of this functionality is given in the Partial Reproduction vignette, which we briefly show here (Figure \@ref(fig:uc1).
+The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
+Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types.
+The oxygen diffusivity is then quickly increased and at around hour 38'000 the system abruptly switches back to the oxic state.
+
+**Rainer... the figure needs beautifying please.**
+
+```{r uc1, echo = FALSE,  include=TRUE, fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
+Plot of the stable states of the simulation runs under different oxygen diffusivity.', fig.align="center", out.width = 350 }
+knitr::include_graphics("figures/ug_temporal_state_switching.png.png")
+```
+
+## Use case 2: The extent of hysteresis depends on community composition
+
+The package also contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
+Hysteresis is a key feature of ecosystem responses to environemtal change, because it is related to how difficult it can be to reserve the effects of environmental change [@REF_NEEDED Scheffer, M., Carpenter, S., Foley, J.
+a, Folke, C.
+& Walker, B.
+(2001).
+Catastrophic shifts in ecosystems.
+Nature, 413, 591–6.
+].
+The amount of hysteresis was measured here as the extent of the environmental condition for which there were two stable states.
+I.e.
+it was the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \@ref(fig:uc2)).
+For example, with only the CB (cyanobacteria) present, there was no hysteresis.
+In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis.
+(These results are also given in the Partial Reproduction vignette.)
+
+**Rainer... the figure needs beautifying please.**
+
+```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present.', fig.align="center", out.width = 350 }
+knitr::include_graphics("figures/user_guide_hysteresis.png")
+```
+
+## Use case 3: Effects of functional diversity on regime shifts
+
+As discussed in the Introduction section, the package was largely motivated by the question of how biodiversity influences ecosystem responses to environmental change.
+Extensive results concerning this question are given in a separate publication @REF_NEEDED.
+Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
+I.e.
+biodiversity can qualitatively change the state of the ecosystem.
+
+Biodiversity was added to the functional groups using the 'new_strain_parameter()' function to create a parameter set with multiple species per functional group (albeit all with identical features) and then to add variability among the species by calling the 'add_strain_var()' function.
+This function takes an already existing parameter set and adds the specified about of variation.
+The new parameter object is then used as before.
+
+Figure \@ref(fig:uc3) shows a simulation with two species (strains) in each of the three functional groups.
+The ecosystem starts in the oxic state, though with relatively high abundance of each functional group.
+The strain of SB that is more tolerant to oxygen (SB_1) initially decreases in abundance, but then increases, and the other (SB_2) strain then becomes abundance and SB_1 declines.
+Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
+Importantly, the switch does not occur if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3).
+
+**Rainer... please beautify the figure as required. It comes from the Limberger et al supplement, Figure 5.4**
+
+```{r uc3, echo = FALSE,  include=TRUE, fig.cap='The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.', fig.align="center", out.width = 350 }
+knitr::include_graphics("figures/uc3_2strain_temporal_switch.png")
+```
 
@@
 
-This package makes an important paper \textcolor{red}{{>>@Owen: is this correct?<<}} [@Bush2017] fully reproducible without having to re-implement the code.
-By doing so, it opens the door for further investigations based on the system as done in @REF_NEEDED and, consequently, increases the reach (and ultimaly) the impact of the original paper by @Bush2017.
-This demonstrates again [@REF_NEEDED] the importance of publishing not only the data but also the code used in a publication.
+*microxanox* allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
+It has been used for the research described in another paper of ours that is one of the first investigations of the effects of diversity on ecosystem resilience @REF_NEEDED.
+In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, within which one can easily manipulate the amount of biodiversity.
+The 'microxanox' package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.
 
-\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}
-
-In addition to the ecological relevance and importance of this package as discussed above, it can be used as a template for the implementation of other different models.
-By doing so, other models can profit from the reproducibility aspects as well as the flexibility as implemented in this package.
+The package greatly lowers the amount of work required in further investigations of the specific ecosystem modelled.
+There has, for example, been quite limited investigation of how biodiversity influences the short-term responses of the modelled ecosystem to environmental change.
+Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change @REF_NEEDED ->>> Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R., Dakos, V., et al.
+(2009).
+Early-warning signals for critical transitions.
+Nature, 461, 53–59.
+].
+In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.
+By doing so, other models can profit from the overall framework used, and the reproducibility aspects as well as the flexibility implemented.
 
 # Conflict of Interest
-We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
+The authors declare no known conflicting or competing interests associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
 
 # Acknowledgements
-Optionally thank people and institutes you need to acknowledge.
+This project was part of SNF Project 310030_188431. The project was also supported by the University of Zurich Research Priority Programme in Global Change and Biodiversity.
 

### 2022-05-10 inst/manuscript/manuscript_microxanox.Rmd
Additional references in bibliograhy and beautified graphs.

@@
 
-Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@REF_NEEDED]
-- Chaparro-Pedraza, P.C. (2021). Fast environmental change and eco-evolutionary feedbacks can drive regime shifts in ecosystems before tipping points are crossed. Proc. R. Soc. B., 288, 20211192.
-- Binzer, A., Guill, C., Rall, B.C.
-& Brose, U.
-(2016).
-Interactive effects of warming, eutrophication and size structure: impacts on biodiversity and food-web structure.
-Global Change Biology, 22, 220–227.
-- Vasseur, D. a & McCann, K.S. (2005). A mechanistic approach for modeling temperature-dependent consumer-resource dynamics. The American Naturalist, 166, 184–198.
-
-.
+Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
 They are also critical for developing hypotheses that can then be tested in experimental studies.
@@
 It is possible that the ecosystem state changes gradually itself.
-It is also possible that the ecosystem state may change abruptly into a new state that is difficult to recover from [@REF_NEEDED]
-- Scheffer, M., Carpenter, S., Foley, J. a, Folke, C. & Walker, B. (2001). Catastrophic shifts in ecosystems. Nature, 413, 591–6.
+It is also possible that the ecosystem state may change abruptly into a new state that is difficult to recover from [@Scheffer2001].
 
-This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern [@REF_NEEDED]. 
-- Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P. & Ratajczak, Z. (2021). Fire frequency, state change and hysteresis in tallgrass prairie. Ecology Letters, 24, 636–647.
-- Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A.
-& Gotelli, N.J.
-(n.d.).
-Clockwise and counterclockwise hysteresis characterize state changes in the same aquatic ecosystem.
-Ecology Letters, n/a.
-- Vandermeer, J. & Perfecto, I. (2019). Hysteresis and critical transitions in a coffee agroecosystem. PNAS, 116, 15074–15079.
+This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern [@Collins2021; @Northrop2021; @Vandermeer2019].
 
@@
 
-One limitation of the study by Bush et al (2017) was limited biodiversity in the ecosystem model.
+One limitation of the study by @Bush2017 was limited biodiversity in the ecosystem model.
 Specifically, there were three functional groups of bacteria, but within each of these groups, there was no biodiversity.
 This then leaves open the question of if and how biodiversity within functional groups, in their model, affects the ecosystem response to environmental change.
-This limitation is not limited to the study of Bush et al (2017): there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
+This limitation is not limited to the study of @Bush2017, there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
 
-We decided to fill this research gap, by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger_et_al_2022, and to base our work on the work and model of Bush et al (2017).
+We decided to fill this research gap, by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
 It was with this goal in mind that we developed the *microxanox* package.
@@
 The *microxanox* package is for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
-It includes feedback between biogeochemical processes and is based on [@Bush2017] (See [@Bush2017] for a detailed discussion of the model).
+It includes feedback between biogeochemical processes and is based on @Bush2017 (See @Bush2017 for a detailed discussion of the model).
 At the core of the simulations is a set of ordinary differential equations (specified in the function `bushplus_dynamic_model()`, though users do not need to call this function directly).
-There are functions for running an individual simulations and for running a set of simulations across, for example, a range of environmental conditions.
+There are functions for running individual simulations and for running a set of simulations across, for example, a range of environmental conditions.
 
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see [@REF_NEEDED] for examples).
+The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement to @Limberger2022 for an example of how this is used).
 As such, all the parameters needed to run a simulation or find a stable state are contained in a single object (which can easily be created using included functions).
@@
 A simulation is run using the `run_simulation()` function.
-In this function, the ODEs are solved using the function `ode` in the package deSolve package [@Soetaert2010].
+In this function, the ODEs are solved using the function `ode()` in the package *deSolve* package [@Soetaert2010].
 The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`.
@@
 The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*).
-This is the method used in the Bush et al (2017) study.
+This is the method used in the @Bush2017 study.
 The second runs only two simulations, with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing or decreasing oxygen diffusivity.
@@
 
-From the raw results returned can be extracted summary quantity about how the ecoystem stable states respond to environmental change.
+From the raw results returned, summary measures about how the ecoystem stable states respond to environmental change can be extracted.
 The function `get_stability_measures()` returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
@@
 
-**Rainer... the figure needs beautifying please.**
-
 ```{r uc1, echo = FALSE,  include=TRUE, fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.', fig.align="center", out.width = 350 }
-knitr::include_graphics("figures/ug_temporal_state_switching.png.png")
+Plot of the stable states of the simulation runs under different oxygen diffusivity.', fig.align="center", out.width = 350}
+fig <- "./figures/gen_uc1_partrep_temporal_state_switching.pdf"
+if (!file.exists(fig)){
+  library(ggplot2)
+  library(microxanox)
+  x <- readRDS("./data/hlh_with_defaults.RDS")
+  p <- plot_dynamics(x)
+  p[[1]] <- p[[1]] + 
+    xlab(NULL) + 
+    ylab(expression(log[10](density)~(cells~L^{-1}))) +
+    theme_bw()
+  p[[2]] <- p[[2]] + 
+    xlab(NULL) + 
+    ylab(expression(log[10](density)~(cells~L^{-1}))) +
+    theme_bw()
+  p[[3]] <- p[[3]] + 
+    xlab(NULL) + 
+    ylab(expression(log[10](density)~(cells~L^{-1}))) +
+    theme_bw()
+  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
+  p[[4]]$data$species <- p[[4]]$data$species %>%
+    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
+  p[[4]] <- p[[4]] +
+    xlab("time (h)") + 
+    ylab(expression(log[10](concentration)~(µM))) + 
+    labs(colour = "substrate") +
+    theme_bw() +
+    xlab(expression(time~(hours)))
+  ggsave(
+    filename = fig,
+    plot = p,
+    device = "pdf",
+    width = 8,
+    height = 8
+  )  
+}
+knitr::include_graphics(fig)
 ```
@@
 The package also contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
-Hysteresis is a key feature of ecosystem responses to environemtal change, because it is related to how difficult it can be to reserve the effects of environmental change [@REF_NEEDED Scheffer, M., Carpenter, S., Foley, J.
-a, Folke, C.
-& Walker, B.
-(2001).
-Catastrophic shifts in ecosystems.
-Nature, 413, 591–6.
-].
+Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change [@Scheffer2001].
 The amount of hysteresis was measured here as the extent of the environmental condition for which there were two stable states.
@@
 
-**Rainer... the figure needs beautifying please.**
+**@Rainer... the figure needs beautifying please.**
+**@Owen: is this OK like this?**
 
 ```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present.', fig.align="center", out.width = 350 }
-knitr::include_graphics("figures/user_guide_hysteresis.png")
+fig <- "./figures/gen_uc2_user_guide_hysteresis.pdf"
+if (!file.exists(fig)){
+  library(ggplot2)
+  x <- readRDS("./data/resp_summ5.rds")
+  x$composition <- as.character(x$composition)
+  x <- x[order(x$mean_mean_hyst),]
+  x$composition <- factor(x$composition, x$composition, ordered = TRUE)
+
+ p <- ggplot(x) +
+  geom_point(aes(x = composition, y = mean_mean_hyst), shape = 95, size = 25) +
+  xlab("Functional groups present") +
+  ylab("Amount of hysteresis") +
+  #ylim(0,2) +
+  theme_light() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ ggsave(
+   filename = fig,
+   plot = p,
+   device = "pdf"
+ )
+
+}
+knitr::include_graphics(fig)
 ```
@@
 As discussed in the Introduction section, the package was largely motivated by the question of how biodiversity influences ecosystem responses to environmental change.
-Extensive results concerning this question are given in a separate publication @REF_NEEDED.
+Extensive results concerning this question are given in a separate publication @Limberger2022.
 Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
@@
 
-Biodiversity was added to the functional groups using the 'new_strain_parameter()' function to create a parameter set with multiple species per functional group (albeit all with identical features) and then to add variability among the species by calling the 'add_strain_var()' function.
+Biodiversity was added to the functional groups using the `new_strain_parameter()` function to create a parameter set with multiple species per functional group (albeit all with identical features) and then to add variability among the species by calling the `add_strain_var()` function.
 This function takes an already existing parameter set and adds the specified about of variation.
@@
 **Rainer... please beautify the figure as required. It comes from the Limberger et al supplement, Figure 5.4**
+**@Owen: is this OK like this? Used the graph from the supplement**
 
 ```{r uc3, echo = FALSE,  include=TRUE, fig.cap='The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.', fig.align="center", out.width = 350 }
-knitr::include_graphics("figures/uc3_2strain_temporal_switch.png")
+knitr::include_graphics("figures/uc3_supplement_5_4.png")
 ```
@@
 *microxanox* allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
-It has been used for the research described in another paper of ours that is one of the first investigations of the effects of diversity on ecosystem resilience @REF_NEEDED.
+It has been used for the research described in another paper of ours that is one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
 In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, within which one can easily manipulate the amount of biodiversity.
-The 'microxanox' package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.
+The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.
 
@@
 There has, for example, been quite limited investigation of how biodiversity influences the short-term responses of the modelled ecosystem to environmental change.
-Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change @REF_NEEDED ->>> Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R., Dakos, V., et al.
-(2009).
-Early-warning signals for critical transitions.
-Nature, 461, 53–59.
-].
+Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change [@Scheffer2009].
 In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.

### 2022-05-10 inst/manuscript/manuscript_microxanox.tex
Additional references in bibliograhy and beautified graphs.

@@
 
-Many ecosystems are exposed to gradual changes of environmental
-variables, to which the responses are not always as gradual as the
-change of the environmental variable \citep{REF_NEEDED}. One example
-where the gradual change of a single environmental variable is causing
-an abrupt change of the system is the switch from an aerobic to
-anaerobic system. This system has been investigated by \citet{Bush2017}.
-We wanted to take this investigation one step further, and look at the
-role an increased biodiversity plays in these dynamics
-\citep{REF_NEEDED}. For this purpose, we developed this package.
+Mathematical models play a key role in the development of understanding
+about how ecosystems work and how they respond to environmental changes
+\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
+critical for developing hypotheses that can then be tested in
+experimental studies. One area of ecology in which simple models have
+played a very influential role is the area of how ecosystems respond to
+gradual environmental change \citep{REF_NEEDED}. - Scheffer, M.,
+Carpenter, S., Foley, J. a, Folke, C. \& Walker, B. (2001). Catastrophic
+shifts in ecosystems. Nature, 413, 591--6.
 
-The microxanox package is a package for simulating a three functional
-group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
-bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
-substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
-sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-biogeochemical processes and is based on \citep{Bush2017} (See
-\citep{Bush2017} for a detailed discussion of the model).
+It is conceivable that an ecosystem state, such as the total biomass of
+a particular type of bacteria, may remain unchanged when something about
+its environment changes, such as the temperature. It is possible that
+the ecosystem state changes gradually itself. It is also possible that
+the ecosystem state may change abruptly into a new state that is
+difficult to recover from \citep{Scheffer2001}.
 
-The aims of the \texttt{microxanox} package are twofold: firstly, to
-reproduce the results shown by \citep{Bush2017}, which is accomplished
-in the \href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et
-al} and secondly, to take these results one step further, to includes
-new functionality to address our research question as published in
-\citep{REF_NEEDED}.
+This possibility for abrupt, perhaps catastrophic changes that are
+difficult to reverse causes considerable concern
+\citep{Collins2021, Northrop2021, Vandermeer2019}.
 
-For this, we extended the model and added functionality for:
+An example where the gradual change of a single environmental variable
+is causing an abrupt change of the system is the switch from an aerobic
+(oxygen is available for metabolism) to anaerobic (oxygen generally
+unavailable) system. This system has been investigated by
+\citet{Bush2017} in a simulation study of a mathematical model of the
+system. They showed that gradual change in the environment could cause
+catastrophic changes in the ecosystem state that would be difficult to
+reverse.
 
-\begin{itemize}
-\tightlist
-\item
-  Multiple strains (effectively unlimited) per functional group.
-\item
-  Adding temporally varying oxygen diffusivity.
-\item
-  Adding random noise in substrate concentrations.
-\item
-  Including immigration.
-\item
-  Setting minimum population abundances.
-\end{itemize}
+One limitation of the study by \citet{Bush2017} was limited biodiversity
+in the ecosystem model. Specifically, there were three functional groups
+of bacteria, but within each of these groups, there was no biodiversity.
+This then leaves open the question of if and how biodiversity within
+functional groups, in their model, affects the ecosystem response to
+environmental change. This limitation is not limited to the study of
+\citet{Bush2017}, there are few if any studies of the effects of
+biodiversity on abrupt transitions between ecosystem states.
 
-In addition to the model itself, the package includes some functions to
-analyse the results as well as visualize the results to provide a
-starting point for customized visualizations based on own requirements.
+We decided to fill this research gap, by making a simulation study of
+how within functional group biodiversity affects ecosystem responses to
+environmental change \citet{Limberger2022}, and to base our work on the
+work and model of \citet{Bush2017}. It was with this goal in mind that
+we developed the \emph{microxanox} package. The first stage of
+development was to write code from scratch (as there was no available
+code to start from) and to confirm that this new implementation would
+reproduce the previously published results. The resulting reproduction
+is available as one of the package vignettes:
+\href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}.
+
+The second stage was to add functionality that would be necessary to
+answer our research question. Most importantly, we made it possible to
+have multiple species of bacteria within each of the three functional
+groups, for the multiple species to differ in their characteristics, and
+to vary the number of species and amount of variability among them. We
+also added functionality that allowed: temporally varying environmental
+conditions, addition of random noise to state variables, and
+immigration. In addition to the model itself, the package includes some
+functions to analyse the results as well as visualize the results to
+provide a starting point for customized visualizations based on own
+requirements. The basic and additional functionality is described in the
+package \href{LINK_NEEDED}{User Guide}.
 
@@
 
-We will discuss the general structure and functionality of the package
-here without going into to much detail. A more detailed discussion can
-be found in the \href{LINK_NEEDED}{User Guide}.
+The \emph{microxanox} package is for simulating a three functional group
+system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
+bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
+substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
+sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
+biogeochemical processes and is based on \citet{Bush2017} (See
+\citet{Bush2017} for a detailed discussion of the model). At the core of
+the simulations is a set of ordinary differential equations (specified
+in the function \texttt{bushplus\_dynamic\_model()}, though users do not
+need to call this function directly). There are functions for running
+individual simulations and for running a set of simulations across, for
+example, a range of environmental conditions.
 
-The model by \citet{Bush2017} is based on ODEs for the rates of change
-which are specified in the function \texttt{bushplus\_dynamic\_model()}.
-This augmented version of the model published in \citep{Bush2017} can
-handle multiple strains within each of the three functional groups,
-temporal variation in oxygen diffusivity, and events.
+To make the simulation run with multiple species, we expressed different
+species characteristics in the elements of vectors and matrices. We also
+adapted the code for the ordinary differential equations to include the
+vectors and matrices, and to use matrix mathematics. In this way, we
+made it possible to run simulations with different numbers of species
+without having to change the underlying code.
 
-The package is not intended to provide a modelling framework which can
-be adjusted easily to all needs, but primarily a tool to implement the
-model used by \citep{Bush2017} and to extend it to our needs
-\citep{REF_NEEDED}. Consequently, any more substantial changes and
-adaptations are likely to need a change in the source code.
-
-Nevertheless, the source code is structured in modular, so that e.g.~the
-event definition can easily be changed, or other aspects can be
-adjusted. All values in the parameter object can be changed as needed
-and the general structure of the code should make it not to difficult to
-adapt the model to other similar systems.
+The package functions and code are structured in modular, so that some
+new functionality can be easily added. E.g. the event definition can
+easily be changed, or other aspects can be adjusted. All values in the
+parameter object can be changed as needed and the general structure of
+the code should make it not to difficult to adapt the model to other
+similar systems.
 
 \hypertarget{software-architecture}{%
-\subsection{Software Architecture}\label{software-architecture}}
+\subsection{Software architecture}\label{software-architecture}}
 
@@
 architecture; provide a pictorial component overview or similar (if
-possible). If necessary provide im- plementation
+possible). If necessary provide implementation
 details.\textless\textless\}
@@
 
-The framework used when writing this package aims at reproducibility of
-the results (see \citep{REF_NEEDED} for examples). It builds on the
-following main considerations:
+The framework used when writing this package aimed to maximise
+simplicity for the user, and to make it easy to reproduce results (see
+the supplement to \citet{Limberger2022} for an example of how this is
+used). As such, all the parameters needed to run a simulation or find a
+stable state are contained in a single object (which can easily be
+created using included functions). This parameter object is given to a
+function that runs the simulations and returns the results. The returned
+results object is identical to the parameter object but with an
+additional slot named \texttt{results}, which contains the results of
+the run. Thus the returned results object contains the simulation
+conditions (parameters) as well as the rsults, and can be used to run
+the simulation again. This promotes reproducibility and makes
+incremental changes of individual parameters and re-running the
+simulations straightforward.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  all parameter needed to run a simulation or find a stable state are
-  contained in a single parameter object. This object is created by
-  using the functions \texttt{new\_...\_parameter()},
-  \texttt{new\_initial\_state()} and \texttt{new\_strain\_parameter()}.
-  Which one of the \texttt{new\_...\_parameter()} functions has to be
-  used when, will be discussed in the section @ref(runsim) and in more
-  detail in the \href{LINK_NEEDED}{User Guide}.
-\item
-  The function call \texttt{run\_...(parameter)} will run the simulation
-  using the parameter as defined in the object \texttt{parameter}.
-\item
-  The return value of the \texttt{run\_...(parameter)} function is
-  identical to the parameter object plus an additional slot named
-  \texttt{results} which contains the results of the run
-\item
-  As the return value contains all parameter, it is possible to re-run
-  the simulation by simply running \texttt{run\_...(result)}.
-\end{enumerate}
-
-The point that the results object contains all parameter needed to run
-the simulation, promotes reproducibility and makes incremental changes
-of individual parameters and re-running the simulations much easier, as
-can be seen in our research and the RMarkdown \citep{REF_NEEDED} file
-used to run and analyse the experiments \citep{REF_NEEDED}.
-
-In the following sections we describe the general usage of the package:
-running one simulation, finding steady states across an environmental
-gradient, calculating measures of stability, and visualization.
+In the following sections we describe how to use the package to run one
+simulation and to find steady states across an environmental gradient.
 
 \hypertarget{running-one-simulation}{%
-\subsubsection{Running one simulation}\label{running-one-simulation}}
+\subsection{Running one simulation}\label{running-one-simulation}}
 
@@
 
-The individual simulation (\texttt{run\_simulation()} function) is the
-working horse in this package. In this function, the ODEs are solved
-using the function \texttt{ode} in the package deSolve package
-\citep{Soetaert2010}. The \texttt{run\_simulation()} function needs only
-one argument - an object as created by the function
-\texttt{new\_runsim\_parameter()}. The parameter object contains among
-other parameter the \texttt{strain\_parameter} which can be created by
-the function \texttt{new\_strain\_parameter()}. For a detailed
-description of the parameter, their meaning and how they are created and
-can be set and changed please see the User Guide which accompanies the
-package or is available at \href{@LINK_NEEDED}{User Guide}
+A simulation is run using the \texttt{run\_simulation()} function. In
+this function, the ODEs are solved using the function \texttt{ode()} in
+the package \emph{deSolve} package \citep{Soetaert2010}. The
+\texttt{run\_simulation()} function needs only one argument - an object
+as created by the function \texttt{new\_runsim\_parameter()}. The
+parameter object returned by \texttt{new\_runsim\_parameter()} contains
+among other things the \texttt{strain\_parameter} object, which can be
+created by the function \texttt{new\_strain\_parameter()}. For a
+detailed description of the parameter objects, their meaning and how
+they are created and have values set and changed please see the
+\emph{User Guide} which accompanies the package or is available at
+\href{@LINK_NEEDED}{User Guide}
 
@@
 parameter used to generate the results before.
+
+\hypertarget{finding-steady-states}{%
+\subsection{Finding Steady States}\label{finding-steady-states}}
+
+The general approach used to find the stable state of the system with a
+specific parameter set is to run the simulation for a long time and
+record the final state. When one does this across a range of
+environmental conditions, one discovers how the steady state of the
+system responds to the environmental conditions. The package contains
+functionality for finding steady states that correspond to values of one
+environmental variable, namely the value of oxygen diffusivity.
+
+Two methods for finding steady states are implemented. The first runs a
+separate simulation for each combination of starting conditions and
+oxygen diffusivity (let us term this the \emph{Replication method}).
+This is the method used in the \citet{Bush2017} study. The second runs
+only two simulations, with step-wise and slowly temporally increasing
+oxygen diffusivity, and the other with step-wise and slowly decreasing
+or decreasing oxygen diffusivity. During this temporal environmental
+change, the state of the system was recorded just before change to a new
+oxygen diffusivity (let us term this the \emph{Temporal method}). We
+implemented two methods since there is no definitive best method, and in
+order to check if results were sensitive to choice of method.
+
+The replication method is implementad in the function
+\texttt{run\_replication\_ssfind()} which takes a parameter object as
+returned by the function \texttt{new\_replication\_ssfind\_parameter()}
+and the number of cores for multithreading the simulation. As the
+multithreading uses the package function \texttt{mclapply()} from the
+package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
+works on Linux and Mac. It is planned to move to \texttt{parLapply()}
+\citep{RCoreTeam2022} in a future release.
+
+This temporal method implemented in the function
+\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
+created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
+is planned for a later release, to run these two simulations in
+parallel.
+
+For a more detailed walk-through of these two approaches and explanation
+please see the \href{@LINK_NEEDED}{User Guide}.
+
+\hypertarget{visualising-and-analysing-results}{%
+\subsection{Visualising and analysing
+results}\label{visualising-and-analysing-results}}
+
+From the raw results returned, summary measures about how the ecoystem
+stable states respond to environmental change can be extracted. The
+function \texttt{get\_stability\_measures()} returns quantities such as
+the amount of environmental change required for the system to abruptly
+change to a different state.
 
@@
 \end{figure}
-
-\hypertarget{finding-a-steady-state-of-the-model}{%
-\subsubsection{Finding a Steady State of the
-model}\label{finding-a-steady-state-of-the-model}}
-
-\textcolor{red}{@Owen: can you please fill in the references to these mathods? Thanks.}
-
-There are two methods for finding steady states implemented in this
-package. The first runs a separate simulation for each combination of
-starting conditions and oxygen diffusivity (let us term this the
-\emph{Replication method})
-\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
-- does this approach has a name, and do you have a reference where it
-has been used or described?\textless\textless\}. The second runs only
-two simulations, with step-wise and slowly temporally increasing or
-decreasing oxygen diffusivities and recorded of state just before change
-to a new oxygen diffusivity (let us term this the \emph{Temporal
-method})\textbackslash textcolor\{red\}\{\{\textgreater\textgreater{}\citet{Owen}
-- Has this been used in other publications? I do not have a clue what I
-could search for to find out\ldots\textless\textless\}.
-
-\hypertarget{replication-method}{%
-\paragraph{Replication Method}\label{replication-method}}
-
-The replication method is implementad in the function
-\texttt{run\_replication\_ssfind()} which takes a parameter object as
-returned by the function \texttt{new\_replication\_ssfind\_parameter()}
-and the number of cores for multithreading the simulation. As the
-multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
-works on Linux and Mac. It is planned to move to \texttt{parLapply()}
-\citep{RCoreTeam2022} in a future release.
-
-\hypertarget{temporal-method}{%
-\paragraph{Temporal Method}\label{temporal-method}}
-
-The temporal method involves two simulations for a particular system
-configuration (parameter set). In one simulation the oxygen diffusivity
-is \emph{increased} in a step-wise fashion. In the other it is
-\emph{decreased} in a step-wise fashion. That is, oxygen diffusivity is
-held at a constant level for long enough for steady state to be reach,
-that state is recorded, and then a slightly higher (or lower) oxygen
-diffusivity value is set. Hence, at that time point, the system is
-effectively started with initial conditions that are the state of the
-system in the previous time step
-\textcolor{red}{{>>@Owen: is there a ref for this approach?<<}}.
-
-This is implemented in the function \texttt{run\_temporal\_ssfind()},
-which takes a parameter object as created by the function
-\texttt{new\_temporal\_ssfind\_parameter()}. It is planned for a later
-release, to run these two simulations in parallel.
-
-For a more detailed walk-through of these two approaches and explanation
-please see the \href{@LINK_NEEDED}{User Guide}.
-
-\hypertarget{extracting-stability-measures}{%
-\subsubsection{Extracting Stability
-Measures}\label{extracting-stability-measures}}
-
-From the raw results returned by these \texttt{run\_...()} functions,
-the stability measures can be extracted by using the function
-\texttt{get\_stability\_measures()}. These measures include
-non-linearity and hysteresis measures, of the response of the simulated
-system to environmental change
-\textcolor{red}{{>>@Owen: is there a paper, so that we can say "as e.g. used in REF?<<}}.
-
-\hypertarget{use-cases}{%
-\subsection{Use Cases}\label{use-cases}}
-
-We will now show three use cases which illustrate applications of the
-package. All use cases can be used as starting points for other
-investigations of the behavior of the system.
-
-The first two use cases are described in detail in the User Guide and
-the Partial Reproduction Vignettes. The third is taken from
-\citet{REF_NEEDED} for which this R package was designed. All of these
-use cases can be expanded to larger numbers of strains per functional
-group and variable values van be changed.
-
-\hypertarget{regime-shifts-during-temporal-environmental-change}{%
-\subsubsection{Regime shifts during temporal environmental
-change}\label{regime-shifts-during-temporal-environmental-change}}
-
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-In the \href{LINK_NEEDED}{User Guide} we used a one strain system
-(section ``1 strain per functional group'') and three strain system
-(section ``3 strains per functional group'') to determine as an example
-the stable states during temporal environmental changes (the oxygen
-diffusivity). From these simulatiuons, we extracted measures of
-nonlinearity and hysteresis. See Fig @ref(fig:plot-dynamics) as an
-example plot of the simulations.
-
-\begin{figure}
-
-{\centering \includegraphics[width=350px]{figures/ug_three_strains_stable_state}
-
-}
-
-\caption{Plot of the stable states of the simulation runs under different oxygen diffusivity.
-The top graph are the Organisms (ech initially with three strains) while the lower graph is the substrate availability under the same oxygen diffusivities.
-Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:uc1_stable_state}
-\end{figure}
-
-We saved all parameter and result objects, so that all results are
-completely reproducible. The general flow of the experiment is identical
-to the one shown in @ref(fig:runsim\_example).
-
-\hypertarget{the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsubsection{The extent of hysteresis depends on community
-composition}\label{the-extent-of-hysteresis-depends-on-community-composition}}
-
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-One of the reasons to develop this package was to reproduce the results
-presented in \citet{Bush2017}. This was achieved as demonstrated in the
-\href{LINK_NEEDED}{Partial Reproduction supplement}. All aspects in the
-paper could be reproduced and are shown in the vignette.
-
-\begin{figure}
-
-{\centering \includegraphics[width=350px]{figures/user_guide_hysteresis}
-
-}
-
-\caption{Hysteresis of all assessed combinations of variability.}\label{fig:user_guide_hysteresis}
-\end{figure}
-
-As all objects were saved, results can be easily reproduced and
-parameter can be changed to assess the impact of these changes (e.g.~to
-conduct a sensitivity analysis).
-
-\hypertarget{effects-of-functional-diversity-on-regime-shifts}{%
-\subsubsection{Effects of functional diversity on regime
-shifts}\label{effects-of-functional-diversity-on-regime-shifts}}
-
-\textcolor{red}{@Owen: Do you have some references which could be added here and probably add some more justifying text to this section?}
-
-As discussed in the paper \citep{REF_NEEDED}, the role biodiversity
-plays in abrupt regime shifts based on gradual changing environmental
-parameter is not well understood. This model (as part of the package)
-has been used to investigate these dynamics and the results are
-available in with the DOI DOI NEEDED \citet{REF_NEEDED}.
-
-\textcolor{red}{{>>@Owen: More details from the paper. I will approach Romana when the first draft is ready so that she can put something in, onless you will add a few sentences here.<<}}
 
@@
 in \citet{Bush2017} provides the means of reproducing the results
-published while at the same time provides the means of doing
-investigations which extend the original publication, as done in
-\citet{REF_NEEDED}.
+published while at the same time provides the means of doing unique,
+innovative, and important investigations of how ecosystems respond to
+environmental change. The design of the package code and functionality
+is with reproducibility in mind: the combination of all parameters being
+in a single parameter object as well as the return of the simulation as
+a result object which inherits from the parameter object provides a
+relatively easy to use framework to implement reproducible experiments.
 
-The combination of all parameter in a single parameter object as well as
-the return of the simulation as a result object which inherits from the
-parameter object and adds a results slot provides a relatively easy to
-use framework to implement reproducible experiments.
+While the package was not intended to provide a code which can be very
+easily adapted to other types of ecosystems and types of organism, it
+seems likely that the overall framework could be used as a basis for
+models of other ecosystems and organisms. For example, all such models
+would have parameters that differ among species and need to be described
+in an object. All would need to run simulations and sets of simulations
+across environmental conditions. Researchers want to model a new
+ecosystem do not, therefore, have to start from scratch. Nevertheless,
+depending on how different were the new ecosystem to model, the require
+code changes could be quite significant.
 
-\textcolor{red}{{>>@Owen: Could you put a few sentence in highlighting the impact this package could have considering the importance of possible research questions?<<}}
+We evidence the impact of the \emph{microxanox} package by describing
+three use cases, each of which was only possible with the package and
+that contains new results. The first two use cases are described in
+detail in the User Guide and the Partial Reproduction Vignettes. The
+third is taken from \citet{REF_NEEDED} for which this R package was
+designed. All of these use cases can be expanded to larger numbers of
+strains per functional group and variable values van be changed.
+
+\hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
+\subsection{Use case 1: Regime shifts during temporal environmental
+change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
+
+The package contains functionality to make a specific pattern of
+temporal change in the environmental condition oxygen diffusivity; this
+functionality forms the basis of the temporal method for finding stable
+states. A simple example of this functionality is given in the Partial
+Reproduction vignette, which we briefly show here (Figure @ref(fig:uc1).
+The example is composed of a single simulation, at the beginning of
+which the system is in the oxic state with high abundance of
+cyanobacteria. Oyxgen diffusivity is then slowly decreased and
+eventually, around hour 30'000 the system switches to the anoxic state,
+with high abundance of both sulfur bacteria types. The oxygen
+diffusivity is then quickly increased and at around hour 38'000 the
+system abruptly switches back to the oxic state.
+
+\begin{figure}
+
+{\centering \includegraphics[width=350px]{./figures/gen_uc1_partrep_temporal_state_switching}
+
+}
+
+\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
+Plot of the stable states of the simulation runs under different oxygen diffusivity.}\label{fig:uc1}
+\end{figure}
+
+\hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
+\subsection{Use case 2: The extent of hysteresis depends on community
+composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
+
+The package also contains a function to extract summary features of
+ecosystem responses to environmental change, such as the amount of
+hysteresis displayed by the ecosystem. Hysteresis is a key feature of
+ecosystem responses to environmental change, because it is related to
+how difficult it can be to reserve the effects of environmental change
+\citep{Scheffer2001}. The amount of hysteresis was measured here as the
+extent of the environmental condition for which there were two stable
+states. I.e. it was the extent of the environmental conditions for which
+historical conditions play an important role in determining the current
+system state (a definition of hysteresis). The results show that the
+amount of hysteresis depends greatly on the combinations of organisms
+present (Figure @ref(fig:uc2)). For example, with only the CB
+(cyanobacteria) present, there was no hysteresis. In contrast, the
+presence of both CB and SB (sulfate reducing bacteria) led to a large
+amount of hysteresis. (These results are also given in the Partial
+Reproduction vignette.)
+
+\textbf{\citet{Rainer}\ldots{} the figure needs beautifying please.}
+\textbf{\citet{Owen}: is this OK like this?}
+
+\begin{figure}
+
+{\centering \includegraphics[width=350px]{./figures/gen_uc2_user_guide_hysteresis}
+
+}
+
+\caption{The amount of hysteresis depends on the combination of types of organisms present.}\label{fig:uc2}
+\end{figure}
+
+\hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
+\subsection{Use case 3: Effects of functional diversity on regime
+shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
+
+As discussed in the Introduction section, the package was largely
+motivated by the question of how biodiversity influences ecosystem
+responses to environmental change. Extensive results concerning this
+question are given in a separate publication \citet{Limberger2022}. Here
+we describe one of the results, which is that having biodiversity in a
+functional group can allow state changes to occur that otherwise would
+not have. I.e. biodiversity can qualitatively change the state of the
+ecosystem.
+
+Biodiversity was added to the functional groups using the
+\texttt{new\_strain\_parameter()} function to create a parameter set
+with multiple species per functional group (albeit all with identical
+features) and then to add variability among the species by calling the
+\texttt{add\_strain\_var()} function. This function takes an already
+existing parameter set and adds the specified about of variation. The
+new parameter object is then used as before.
+
+Figure @ref(fig:uc3) shows a simulation with two species (strains) in
+each of the three functional groups. The ecosystem starts in the oxic
+state, though with relatively high abundance of each functional group.
+The strain of SB that is more tolerant to oxygen (SB\_1) initially
+decreases in abundance, but then increases, and the other (SB\_2) strain
+then becomes abundance and SB\_1 declines. Furthermore, the
+cyanobacteria crash in abundance, and the system switches to the anoxic
+state. Importantly, the switch does not occur if there are two identical
+strains with tolerance half way between those in Figure @ref(fig:uc3).
+
+\textbf{Rainer\ldots{} please beautify the figure as required. It comes
+from the Limberger et al supplement, Figure 5.4} \textbf{\citet{Owen}:
+is this OK like this? Used the graph from the supplement}
+
+\begin{figure}
+
+{\centering \includegraphics[width=350px]{figures/uc3_supplement_5_4}
+
+}
+
+\caption{The dynamics of the ecosystem when there are two species in each functional group, and some variation (diversity) in species parameters.}\label{fig:uc3}
+\end{figure}
 
@@
 
-This package makes an important paper
-\textcolor{red}{{>>@Owen: is this correct?<<}} \citep{Bush2017} fully
-reproducible without having to re-implement the code. By doing so, it
-opens the door for further investigations based on the system as done in
-\citet{REF_NEEDED} and, consequently, increases the reach (and ultimaly)
-the impact of the original paper by \citet{Bush2017}. This demonstrates
-again \citep{REF_NEEDED} the importance of publishing not only the data
-but also the code used in a publication.
+\emph{microxanox} allows the simulation, visualisation, and analysis of
+a model of a microbial ecosystem while allowing variation in the amount
+of diversity containing in each of the functional groups of organisms
+present. It has been used for the research described in another paper of
+ours that is one of the first investigations of the effects of diversity
+on ecosystem resilience \citet{Limberger2022}. In that paper, we show
+that diversity can have large and important effects of ecosystem
+responses, highlighting the need for models such as ours, within which
+one can easily manipulate the amount of biodiversity. The
+\emph{microxanox} package has also been used to reproduce the results of
+the paper that inspired the package development \citet{Bush2017}.
 
-\textcolor{red}{{>>@Owen: could you write a short conclusion paragraph about the ecological relevance of `microxanox`?<<}}
-
-In addition to the ecological relevance and importance of this package
-as discussed above, it can be used as a template for the implementation
-of other different models. By doing so, other models can profit from the
-reproducibility aspects as well as the flexibility as implemented in
-this package.
+The package greatly lowers the amount of work required in further
+investigations of the specific ecosystem modelled. There has, for
+example, been quite limited investigation of how biodiversity influences
+the short-term responses of the modelled ecosystem to environmental
+change. Likewise, the package could be used to power an investigation of
+the effects of biodiversity on the usefulness of early warning signals
+of abrupt ecosystem change \citep{Scheffer2009}. In addition this
+package could be used as a template for the implementation for
+developing models of other types of ecosystems and organism. By doing
+so, other models can profit from the overall framework used, and the
+reproducibility aspects as well as the flexibility implemented.
 
@@
 
-We wish to confirm that there are no known conflicts of interest
+The authors declare no known conflicting or competing interests
 associated with this publication and there has been no significant
@@
 
-Optionally thank people and institutes you need to acknowledge.
+This project was part of SNF Project 310030\_188431. The project was
+also supported by the University of Zurich Research Priority Programme
+in Global Change and Biodiversity.
 

### 2022-05-26 inst/manuscript/manuscript_microxanox.Rmd
work on ms

@@
 Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
-They are also critical for developing hypotheses that can then be tested in experimental studies.
-One area of ecology in which simple models have played a very influential role is the area of how ecosystems respond to gradual environmental change [@REF_NEEDED].
-- Scheffer, M., Carpenter, S., Foley, J. a, Folke, C. & Walker, B. (2001). Catastrophic shifts in ecosystems. Nature, 413, 591–6.
+They are also important for developing hypotheses that can then be tested in empirical studies.
+One area of ecology in which simple models have played a influential role is the area of how ecosystems respond to gradual change in an environmental driver [@Scheffer2001].
+An environmental driver is a environmental condition that affects an ecosystem, but is assumed to not be affected by the ecosystem, such as the rate of nutrient input into a lake.
 
-It is conceivable that an ecosystem state, such as the total biomass of a particular type of bacteria, may remain unchanged when something about its environment changes, such as the temperature.
-It is possible that the ecosystem state changes gradually itself.
-It is also possible that the ecosystem state may change abruptly into a new state that is difficult to recover from [@Scheffer2001].
-
+It is conceivable that an ecosystem state, such as the total biomass of a particular type of bacteria, may remain unchanged when an environmental driver changes.
+It is also possible that the ecosystem state changes gradually.
+It is also possible that the ecosystem state changes abruptly to a new state that is difficult to recover from [@Scheffer2001].
 This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern [@Collins2021; @Northrop2021; @Vandermeer2019].
 
-An example where the gradual change of a single environmental variable is causing an abrupt change of the system is the switch from an aerobic (oxygen is available for metabolism) to anaerobic (oxygen generally unavailable) system.
-This system has been investigated by @Bush2017 in a simulation study of a mathematical model of the system.
-They showed that gradual change in the environment could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+An example where a gradual change of an environmental variable causes an abrupt change of the system is the switch from an aerobic (oxygen is available for metabolism) to anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
+Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
+This system has been investigated by @Bush2017 in a simulation study of a mathematical model.
+They showed that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed the oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
-One limitation of the study by @Bush2017 was limited biodiversity in the ecosystem model.
-Specifically, there were three functional groups of bacteria, but within each of these groups, there was no biodiversity.
-This then leaves open the question of if and how biodiversity within functional groups, in their model, affects the ecosystem response to environmental change.
-This limitation is not limited to the study of @Bush2017, there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
+One feature of the study by @Bush2017 was limited biodiversity.
+Specifically, there was no biodiversity within each of the three types of bacteria.
+This leaves open the question of if and how biodiversity within these types (i.e.
+functional groups) of bacteria, in their model, affects the ecosystem response to environmental change.
+This limitation is not specific to the study of @Bush2017, there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
 
-We decided to fill this research gap, by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
+We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
 It was with this goal in mind that we developed the *microxanox* package.
@@
 The *microxanox* package is for simulating a three functional group system (*CB*: cyanobacteria, *PB*: phototrophic sulfur bacteria, and *SB*: sulfate-reducing bacteria) with four chemical substrates (*P*: phosphorus, *O*: oxygen, *SR*: reduced sulfur, *SO*: oxidized sulfur).
-It includes feedback between biogeochemical processes and is based on @Bush2017 (See @Bush2017 for a detailed discussion of the model).
-At the core of the simulations is a set of ordinary differential equations (specified in the function `bushplus_dynamic_model()`, though users do not need to call this function directly).
+It includes feedback between organisms and biogeochemical processes and is based on @Bush2017 (See @Bush2017 for a detailed discussion of the model).
+At the core of the simulations is a set of ordinary differential equations (specified in the function `bushplus_dynamic_model()`, though this function need not be directly called).
 There are functions for running individual simulations and for running a set of simulations across, for example, a range of environmental conditions.
@@
 To make the simulation run with multiple species, we expressed different species characteristics in the elements of vectors and matrices.
-We also adapted the code for the ordinary differential equations to include the vectors and matrices, and to use matrix mathematics.
+We also coded the ordinary differential equations to include the vectors and matrices, and to use matrix mathematics.
 In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
 
-The package functions and code are structured in modular, so that some new functionality can be easily added.
+The package functions and code have modular structure, so that new functionality can be easily added.
 E.g.
-the event definition can easily be changed, or other aspects can be adjusted.
-All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.
+temporally defined events of any type could be specified.
+Further, all parameter values required to run a simulation are stored in one object.
+Lastly, the general structure of the code should make it straightforward to adapt the model to other similar systems (described in more detail in the Impact section).
 
@@
 The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement to @Limberger2022 for an example of how this is used).
-As such, all the parameters needed to run a simulation or find a stable state are contained in a single object (which can easily be created using included functions).
+As such, all the parameters needed to run a simulation or find a stable state (i.e.
+the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
 This parameter object is given to a function that runs the simulations and returns the results.
 The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the results of the run.
-Thus the returned results object contains the simulation conditions (parameters) as well as the rsults, and can be used to run the simulation again.
+Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again.
 This promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations straightforward.
@@
 
-## Finding Steady States
+## Finding Stable States
 
@@
 When one does this across a range of environmental conditions, one discovers how the steady state of the system responds to the environmental conditions.
-The package contains functionality for finding steady states that correspond to values of one environmental variable, namely the value of oxygen diffusivity.
+The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
 Two methods for finding steady states are implemented.
-The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (let us term this the *Replication method*).
+The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (we term this the *Replication method*).
 This is the method used in the @Bush2017 study.
-The second runs only two simulations, with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing or decreasing oxygen diffusivity.
-During this temporal environmental change, the state of the system was recorded just before change to a new oxygen diffusivity (let us term this the *Temporal method*).
+The second runs two simulations, with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing or decreasing oxygen diffusivity.
+During this temporal environmental change, the state of the system was recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
 We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method.
 
-The replication method is implementad in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
+The replication method is implemented in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
 As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac.
@@
 
-## Visualising and analysing results
+## Analysing and visualising results
 
-From the raw results returned, summary measures about how the ecoystem stable states respond to environmental change can be extracted.
+From the results returned, summary measures about how the ecosystem stable states respond to environmental change can be extracted.
 The function `get_stability_measures()` returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
@@
 
-```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Plot of results of a simulation run using the function $plot_dynamics()$.
-Details can be found in the "User Guide" section "Three strains per functional group".', fig.align="center", out.width = 350 }
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Results of a simulation run shown using the function $plot_dynamics()$.
+In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example.
+CB_1 = cyanobacteria strain 1; SB_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.', fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/ug_three_strains_dynamics.png")
@@
 
-The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems respond to environmental change.
+The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems respond to environmental change, and how biodiversity may modulate this response.
+
 The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
 
-While the package was not intended to provide a code which can be very easily adapted to other types of ecosystems and types of organism, it seems likely that the overall framework could be used as a basis for models of other ecosystems and organisms.
-For example, all such models would have parameters that differ among species and need to be described in an object.
-All would need to run simulations and sets of simulations across environmental conditions.
-Researchers want to model a new ecosystem do not, therefore, have to start from scratch.
-Nevertheless, depending on how different were the new ecosystem to model, the require code changes could be quite significant.
-
-We evidence the impact of the *microxanox* package by describing three use cases, each of which was only possible with the package and that contains new results.
-The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes.
-The third is taken from @REF_NEEDED for which this R package was designed.
-All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.
+Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
+The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes.
+The third is taken from @REF_NEEDED.
 
@@
 
-The package contains functionality to make a specific pattern of temporal change in the environmental condition oxygen diffusivity; this functionality forms the basis of the temporal method for finding stable states.
-A simple example of this functionality is given in the Partial Reproduction vignette, which we briefly show here (Figure \@ref(fig:uc1).
+The study of @Bush2017 includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic).
+The *microxonax* package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding stable states.
+
+An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \@ref(fig:uc1).
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
 Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types.
-The oxygen diffusivity is then quickly increased and at around hour 38'000 the system abruptly switches back to the oxic state.
+The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.
+
+Also visible in the results are thick lines showing abundances of bacteria when abundances are low.
+This is due to the implementation of a function that at regular intervals, increases the abundance to a preset level.
+This prevents abundances reducing to very small numbers.
+The function that implements this increase abundance can also be made to add a certain abundance to each strain at regular intervals, thus simulating immigration in to the system.
 
 ```{r uc1, echo = FALSE,  include=TRUE, fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.', fig.align="center", out.width = 350}
-fig <- "./figures/gen_uc1_partrep_temporal_state_switching.pdf"
-if (!file.exists(fig)){
+Plot of the stable states of the simulation runs under different oxygen diffusivity.
+In this simulation there is only one strain in each functional group.
+CB_1 = cyanobacteria strain 1; SB_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show a figure adapted from the output of the $plot_dynamics()$  function.', fig.align="center", out.width = 350}
+fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
+#if (!file.exists(fig)){
   library(ggplot2)
   library(microxanox)
-  x <- readRDS("./data/hlh_with_defaults.RDS")
+  library(dplyr)
+  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
   p <- plot_dynamics(x)
@@
     xlab(NULL) + 
-    ylab(expression(log[10](density)~(cells~L^{-1}))) +
+    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
     theme_bw()
@@
     xlab(NULL) + 
-    ylab(expression(log[10](density)~(cells~L^{-1}))) +
+    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
     theme_bw()
+  p_a <- p[[4]]
   p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
@@
     xlab("time (h)") + 
-    ylab(expression(log[10](concentration)~(µM))) + 
+    ylab(expression(atop(log[10](concentration),(µM)))) + 
     labs(colour = "substrate") +
@@
     xlab(expression(time~(hours)))
+  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
+  p_a$data$species <- p_a$data$species %>%
+    dplyr::recode(a = "Oxygen diffusivity ")
+  p_a <- p_a +
+    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
+    labs(colour = "Environmental driver") +
+    theme_bw() +
+    xlab(expression(time~(hours)))
+  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
   ggsave(
@@
   )  
-}
+#}
 knitr::include_graphics(fig)
@@
 
-The package also contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
+The package contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
 Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change [@Scheffer2001].
-The amount of hysteresis was measured here as the extent of the environmental condition for which there were two stable states.
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two stable states.
 I.e.
-it was the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
+it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
+
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of stable states across an environmental gradient, running the stable state finding function, and analysing the results with the function that calculates extent of hysteresis.
+The code for this is provided in the *User Guide* vignette.
+
 The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \@ref(fig:uc2)).
@@
 In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis.
-(These results are also given in the Partial Reproduction vignette.)
+(These results are also given in the *Partial Reproduction* vignette.)
 
-**@Rainer... the figure needs beautifying please.**
-**@Owen: is this OK like this?**
-
-```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present.', fig.align="center", out.width = 350 }
-fig <- "./figures/gen_uc2_user_guide_hysteresis.pdf"
-if (!file.exists(fig)){
+```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present.
+The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
+fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
+#if (!file.exists(fig)){
   library(ggplot2)
-  x <- readRDS("./data/resp_summ5.rds")
+  x <- readRDS(here::here("inst/manuscript/data/resp_summ5.rds"))
   x$composition <- as.character(x$composition)
@@
  p <- ggplot(x) +
-  geom_point(aes(x = composition, y = mean_mean_hyst), shape = 95, size = 25) +
+  geom_point(aes(x = composition, y = mean_mean_hyst)) +
   xlab("Functional groups present") +
@@
    plot = p,
-   device = "pdf"
+   device = "pdf",
+   width = 3, height = 3
  )
 
-}
+#}
 knitr::include_graphics(fig)
@@
 
-As discussed in the Introduction section, the package was largely motivated by the question of how biodiversity influences ecosystem responses to environmental change.
+As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change.
 Extensive results concerning this question are given in a separate publication @Limberger2022.
@@
 Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
-Importantly, the switch does not occur if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3).
-
-**Rainer... please beautify the figure as required. It comes from the Limberger et al supplement, Figure 5.4**
-**@Owen: is this OK like this? Used the graph from the supplement**
+In contrast, if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3) the ecosystem remains in the oxic state.
 
@@
 
+## Adapting for other ecosystems and organisms
+
+We anticipate that the package can be a useful starting point for investigating models of other types of ecosystem and how biodiversity in them affects responses to environmental change.
+The overall framework of the package, the purpose of each function, and the objects used for storing parameters and results could be retained.
+E.g., all such models would have parameters that differ among species and need to be described in an object, and studies will often need to run simulations and sets of simulations across environmental conditions.
+
+Researchers wanting to model a new ecosystem do not, therefore, have to start from scratch.
+This will relieve researchers from needing to making software design decisions, and rather focus on appropriately representing their ecosystem, and finding the results that interest them.
+Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and that is relatively proficient in R programming.
+
 # Conclusions
@@
 
-*microxanox* allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
-It has been used for the research described in another paper of ours that is one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
-In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, within which one can easily manipulate the amount of biodiversity.
+The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
+It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
+In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
 The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.

### 2022-05-26 inst/manuscript/manuscript_microxanox.tex
work on ms

@@
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-
 \usepackage[hyphens]{url}
@@
   \linenumbers % turns line numbering on
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
-
+\bibliographystyle{elsarticle-harv}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \ifxetex
@@
             pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
 
@@
 
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-% From pandoc table feature
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-
+% Pandoc header
 %% \documentclass[preprint,12pt, a4paper]{elsarticle}
@@
 \begin{document}
-
 \begin{frontmatter}
@@
 \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
-    \author[University of Zürich]{Rainer M Krug%
-  %
-  \fnref{Corresponding Author}}
+    \author[University of Zürich]{Rainer M Krug\corref{Corresponding
+Author}}
    \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
@@
    \ead{Owen.Petchey@ieu.uzh.ch} 
-      \affiliation[University of Zürich]{Department of Evolutionary
-Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
-    \cortext[cor1]{Corresponding author}
-    \fntext[1]{Corresponding Author}
-    \fntext[2]{Equal contribution}
+      \address[University of Zürich]{Department of Evolutionary Biology
+and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
+      \cortext[1]{Corresponding Author}
+    \cortext[2]{Equal contribution}
 
@@
   \end{abstract}
-    \begin{keyword}
-    reproducibility; regime shift; \sep 
-    reproducibility; regime shift;
-  \end{keyword}
-
+   \begin{keyword} reproducibility; regime shift;\end{keyword}
  \end{frontmatter}
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.05}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.41}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.54}}@{}}
 \toprule
-\begin{minipage}[b]{\linewidth}\centering
-\textbf{Nr.}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Code metadata description}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Please fill in this column}
-\end{minipage} \\
+\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill
+in this column} \\
 \midrule
@@
 about how ecosystems work and how they respond to environmental changes
-\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
-critical for developing hypotheses that can then be tested in
-experimental studies. One area of ecology in which simple models have
-played a very influential role is the area of how ecosystems respond to
-gradual environmental change \citep{REF_NEEDED}. - Scheffer, M.,
-Carpenter, S., Foley, J. a, Folke, C. \& Walker, B. (2001). Catastrophic
-shifts in ecosystems. Nature, 413, 591--6.
+(\protect\hyperlink{ref-Binzer2016a}{Binzer et al., 2016};
+\protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021};
+\protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}). They are
+also important for developing hypotheses that can then be tested in
+empirical studies. One area of ecology in which simple models have
+played a influential role is the area of how ecosystems respond to
+gradual change in an environmental driver
+(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). An
+environmental driver is a environmental condition that affects an
+ecosystem, but is assumed to not be affected by the ecosystem, such as
+the rate of nutrient input into a lake.
 
 It is conceivable that an ecosystem state, such as the total biomass of
-a particular type of bacteria, may remain unchanged when something about
-its environment changes, such as the temperature. It is possible that
-the ecosystem state changes gradually itself. It is also possible that
-the ecosystem state may change abruptly into a new state that is
-difficult to recover from \citep{Scheffer2001}.
+a particular type of bacteria, may remain unchanged when an
+environmental driver changes. It is also possible that the ecosystem
+state changes gradually. It is also possible that the ecosystem state
+changes abruptly to a new state that is difficult to recover from
+(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). This
+possibility for abrupt, perhaps catastrophic changes that are difficult
+to reverse causes considerable concern
+(\protect\hyperlink{ref-Collins2021}{Collins et al., 2021};
+\protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021};
+\protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).
 
-This possibility for abrupt, perhaps catastrophic changes that are
-difficult to reverse causes considerable concern
-\citep{Collins2021, Northrop2021, Vandermeer2019}.
+An example where a gradual change of an environmental variable causes an
+abrupt change of the system is the switch from an aerobic (oxygen is
+available for metabolism) to anaerobic (oxygen generally unavailable)
+state in a microbial ecosystem. Three types of microbes occur in the
+model: cyanobacteria (CB) dominating the oxic state, and two types of
+sulfur bacteria that dominate the anoxic state (sulfate reducing
+bacteria (SB) and phototrophic sulfur bacteria (PB)). This system has
+been investigated by \protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) in a simulation study of a
+mathematical model. They showed that gradual change in the rate at which
+oxygen could diffuse into the ecosystem (termed the oxygen
+diffusitivity) could cause catastrophic changes in the ecosystem state
+that would be difficult to reverse.
 
-An example where the gradual change of a single environmental variable
-is causing an abrupt change of the system is the switch from an aerobic
-(oxygen is available for metabolism) to anaerobic (oxygen generally
-unavailable) system. This system has been investigated by
-\citet{Bush2017} in a simulation study of a mathematical model of the
-system. They showed that gradual change in the environment could cause
-catastrophic changes in the ecosystem state that would be difficult to
-reverse.
+One feature of the study by \protect\hyperlink{ref-Bush2017}{Bush et
+al.} (\protect\hyperlink{ref-Bush2017}{2017}) was limited biodiversity.
+Specifically, there was no biodiversity within each of the three types
+of bacteria. This leaves open the question of if and how biodiversity
+within these types (i.e.~functional groups) of bacteria, in their model,
+affects the ecosystem response to environmental change. This limitation
+is not specific to the study of \protect\hyperlink{ref-Bush2017}{Bush et
+al.} (\protect\hyperlink{ref-Bush2017}{2017}), there are few if any
+studies of the effects of biodiversity on abrupt transitions between
+ecosystem states.
 
-One limitation of the study by \citet{Bush2017} was limited biodiversity
-in the ecosystem model. Specifically, there were three functional groups
-of bacteria, but within each of these groups, there was no biodiversity.
-This then leaves open the question of if and how biodiversity within
-functional groups, in their model, affects the ecosystem response to
-environmental change. This limitation is not limited to the study of
-\citet{Bush2017}, there are few if any studies of the effects of
-biodiversity on abrupt transitions between ecosystem states.
-
-We decided to fill this research gap, by making a simulation study of
-how within functional group biodiversity affects ecosystem responses to
-environmental change \citet{Limberger2022}, and to base our work on the
-work and model of \citet{Bush2017}. It was with this goal in mind that
-we developed the \emph{microxanox} package. The first stage of
+We decided to fill this research gap by making a simulation study of how
+within functional group biodiversity affects ecosystem responses to
+environmental change \protect\hyperlink{ref-Limberger2022}{Limberger}
+(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
+the work and model of \protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}). It was with this goal in mind
+that we developed the \emph{microxanox} package. The first stage of
 development was to write code from scratch (as there was no available
@@
 sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-biogeochemical processes and is based on \citet{Bush2017} (See
-\citet{Bush2017} for a detailed discussion of the model). At the core of
-the simulations is a set of ordinary differential equations (specified
-in the function \texttt{bushplus\_dynamic\_model()}, though users do not
-need to call this function directly). There are functions for running
-individual simulations and for running a set of simulations across, for
-example, a range of environmental conditions.
+organisms and biogeochemical processes and is based on
+\protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) (See
+\protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of
+the model). At the core of the simulations is a set of ordinary
+differential equations (specified in the function
+\texttt{bushplus\_dynamic\_model()}, though this function need not be
+directly called). There are functions for running individual simulations
+and for running a set of simulations across, for example, a range of
+environmental conditions.
 
@@
 species characteristics in the elements of vectors and matrices. We also
-adapted the code for the ordinary differential equations to include the
-vectors and matrices, and to use matrix mathematics. In this way, we
-made it possible to run simulations with different numbers of species
-without having to change the underlying code.
+coded the ordinary differential equations to include the vectors and
+matrices, and to use matrix mathematics. In this way, we made it
+possible to run simulations with different numbers of species without
+having to change the underlying code.
 
-The package functions and code are structured in modular, so that some
-new functionality can be easily added. E.g. the event definition can
-easily be changed, or other aspects can be adjusted. All values in the
-parameter object can be changed as needed and the general structure of
-the code should make it not to difficult to adapt the model to other
-similar systems.
+The package functions and code have modular structure, so that new
+functionality can be easily added. E.g. temporally defined events of any
+type could be specified. Further, all parameter values required to run a
+simulation are stored in one object. Lastly, the general structure of
+the code should make it straightforward to adapt the model to other
+similar systems (described in more detail in the Impact section).
 
@@
 simplicity for the user, and to make it easy to reproduce results (see
-the supplement to \citet{Limberger2022} for an example of how this is
-used). As such, all the parameters needed to run a simulation or find a
-stable state are contained in a single object (which can easily be
-created using included functions). This parameter object is given to a
-function that runs the simulations and returns the results. The returned
-results object is identical to the parameter object but with an
-additional slot named \texttt{results}, which contains the results of
-the run. Thus the returned results object contains the simulation
-conditions (parameters) as well as the rsults, and can be used to run
-the simulation again. This promotes reproducibility and makes
-incremental changes of individual parameters and re-running the
-simulations straightforward.
+the supplement to \protect\hyperlink{ref-Limberger2022}{Limberger}
+(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
+is used). As such, all the parameters needed to run a simulation or find
+a stable state (i.e.~the final state of the ecosystem) are contained in
+a single object (which can easily be created using included functions).
+This parameter object is given to a function that runs the simulations
+and returns the results. The returned results object is identical to the
+parameter object but with an additional slot named \texttt{results},
+which contains the results of the run. Thus the returned results object
+contains the simulation conditions (parameters) as well as the results,
+and can be used to run the simulation again. This promotes
+reproducibility and makes incremental changes of individual parameters
+and re-running the simulations straightforward.
 
@@
 this function, the ODEs are solved using the function \texttt{ode()} in
-the package \emph{deSolve} package \citep{Soetaert2010}. The
+the package \emph{deSolve} package
+(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}). The
 \texttt{run\_simulation()} function needs only one argument - an object
@@
 
-\hypertarget{finding-steady-states}{%
-\subsection{Finding Steady States}\label{finding-steady-states}}
+\hypertarget{finding-stable-states}{%
+\subsection{Finding Stable States}\label{finding-stable-states}}
 
@@
 functionality for finding steady states that correspond to values of one
-environmental variable, namely the value of oxygen diffusivity.
+environmental driver, namely the value of oxygen diffusivity.
 
@@
 separate simulation for each combination of starting conditions and
-oxygen diffusivity (let us term this the \emph{Replication method}).
-This is the method used in the \citet{Bush2017} study. The second runs
-only two simulations, with step-wise and slowly temporally increasing
-oxygen diffusivity, and the other with step-wise and slowly decreasing
-or decreasing oxygen diffusivity. During this temporal environmental
+oxygen diffusivity (we term this the \emph{Replication method}). This is
+the method used in the \protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) study. The second runs two
+simulations, with step-wise and slowly temporally increasing oxygen
+diffusivity, and the other with step-wise and slowly decreasing or
+decreasing oxygen diffusivity. During this temporal environmental
 change, the state of the system was recorded just before change to a new
-oxygen diffusivity (let us term this the \emph{Temporal method}). We
+oxygen diffusivity (we term this the \emph{Temporal method}). We
 implemented two methods since there is no definitive best method, and in
@@
 
-The replication method is implementad in the function
+The replication method is implemented in the function
 \texttt{run\_replication\_ssfind()} which takes a parameter object as
@@
 multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
-works on Linux and Mac. It is planned to move to \texttt{parLapply()}
-\citep{RCoreTeam2022} in a future release.
+package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core
+Team, 2022}), the multthreading only works on Linux and Mac. It is
+planned to move to \texttt{parLapply()}
+(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future
+release.
 
@@
 
-\hypertarget{visualising-and-analysing-results}{%
-\subsection{Visualising and analysing
-results}\label{visualising-and-analysing-results}}
+\hypertarget{analysing-and-visualising-results}{%
+\subsection{Analysing and visualising
+results}\label{analysing-and-visualising-results}}
 
-From the raw results returned, summary measures about how the ecoystem
+From the results returned, summary measures about how the ecosystem
 stable states respond to environmental change can be extracted. The
@@
 
-\begin{figure}
+\textbackslash begin\{figure\}
 
-{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}
+\{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}
 
-}
+\}
 
-\caption{Plot of results of a simulation run using the function $plot_dynamics()$. Details can be found in the "User Guide" section "Three strains per functional group".}\label{fig:plot-dynamics}
-\end{figure}
+\textbackslash caption\{Results of a simulation run shown using the
+function \(plot_dynamics()\). In this case, there were three strains per
+functional group, though strains within functional groups had identical
+properties in this example. CB\_1 = cyanobacteria strain 1; SB\_1 =
+sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria
+strain 1. Here we show the figure produce by \(plot_dynamics()\), which
+produces figures not with all features ideal for
+publication.\}\label{fig:plot-dynamics} \textbackslash end\{figure\}
 
@@
 
-The free and open source implementation and extension of the model used
-in \citet{Bush2017} provides the means of reproducing the results
-published while at the same time provides the means of doing unique,
-innovative, and important investigations of how ecosystems respond to
-environmental change. The design of the package code and functionality
-is with reproducibility in mind: the combination of all parameters being
-in a single parameter object as well as the return of the simulation as
-a result object which inherits from the parameter object provides a
-relatively easy to use framework to implement reproducible experiments.
+The open source implementation and extension of the model used in
+\protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) provides the means of
+reproducing the results published while at the same time provides the
+means of doing unique, innovative, and important investigations of how
+ecosystems respond to environmental change, and how biodiversity may
+modulate this response.
 
-While the package was not intended to provide a code which can be very
-easily adapted to other types of ecosystems and types of organism, it
-seems likely that the overall framework could be used as a basis for
-models of other ecosystems and organisms. For example, all such models
-would have parameters that differ among species and need to be described
-in an object. All would need to run simulations and sets of simulations
-across environmental conditions. Researchers want to model a new
-ecosystem do not, therefore, have to start from scratch. Nevertheless,
-depending on how different were the new ecosystem to model, the require
-code changes could be quite significant.
+The design of the package code and functionality is with reproducibility
+in mind: the combination of all parameters being in a single parameter
+object as well as the return of the simulation as a result object which
+inherits from the parameter object provides a relatively easy to use
+framework to implement reproducible experiments.
 
-We evidence the impact of the \emph{microxanox} package by describing
-three use cases, each of which was only possible with the package and
-that contains new results. The first two use cases are described in
-detail in the User Guide and the Partial Reproduction Vignettes. The
-third is taken from \citet{REF_NEEDED} for which this R package was
-designed. All of these use cases can be expanded to larger numbers of
-strains per functional group and variable values van be changed.
+Here we evidence the impact of the \emph{microxanox} package by
+describing three use cases and then by describing how the package can be
+a starting point for models of other ecosystems. The first two use cases
+are described in detail (including the code for reproducing them) in the
+\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
+third is taken from \protect\hyperlink{ref-REF_NEEDED}{REF NEEDED}
+(\protect\hyperlink{ref-REF_NEEDED}{2222}).
 
@@
 
-The package contains functionality to make a specific pattern of
-temporal change in the environmental condition oxygen diffusivity; this
-functionality forms the basis of the temporal method for finding stable
-states. A simple example of this functionality is given in the Partial
-Reproduction vignette, which we briefly show here (Figure @ref(fig:uc1).
-The example is composed of a single simulation, at the beginning of
-which the system is in the oxic state with high abundance of
-cyanobacteria. Oyxgen diffusivity is then slowly decreased and
+The study of \protect\hyperlink{ref-Bush2017}{Bush et al.}
+(\protect\hyperlink{ref-Bush2017}{2017}) includes simulations of the
+effect of oxygen diffusivity (an environmental driver, in the sense that
+it affects the ecosystem but is not affected by it) on the ecosystem
+state (oxic or anoxic). The \emph{microxonax} package contains
+functionality to make a specific temporal pattern of change in the
+oxygen diffusivity. As well as allowing individual simulations during
+which oxygen diffusivity varies, this functionality forms the basis of
+the temporal method for finding stable states.
+
+An example of this functionality is given in the \emph{Partial
+Reproduction} vignette, which we briefly show here (Figure
+@ref(fig:uc1). The example is composed of a single simulation, at the
+beginning of which the system is in the oxic state with high abundance
+of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
 eventually, around hour 30'000 the system switches to the anoxic state,
 with high abundance of both sulfur bacteria types. The oxygen
-diffusivity is then quickly increased and at around hour 38'000 the
-system abruptly switches back to the oxic state.
+diffusivity is then increased and at around hour 38'000 the system
+abruptly switches back to the oxic state.
 
-\begin{figure}
+Also visible in the results are thick lines showing abundances of
+bacteria when abundances are low. This is due to the implementation of a
+function that at regular intervals, increases the abundance to a preset
+level. This prevents abundances reducing to very small numbers. The
+function that implements this increase abundance can also be made to add
+a certain abundance to each strain at regular intervals, thus simulating
+immigration in to the system.
 
-{\centering \includegraphics[width=350px]{./figures/gen_uc1_partrep_temporal_state_switching}
+\begin{verbatim}
+#> 
+#> Attaching package: 'dplyr'
+#> The following objects are masked from 'package:stats':
+#> 
+#>     filter, lag
+#> The following objects are masked from 'package:base':
+#> 
+#>     intersect, setdiff, setequal, union
+\end{verbatim}
 
-}
+\textbackslash begin\{figure\}
 
-\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.}\label{fig:uc1}
-\end{figure}
+\{\centering \includegraphics[width=350px]{/Users/owenpetchey/Desktop/microxanox/microxanox/inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching}
+
+\}
+
+\textbackslash caption\{The temporal dynamics of the ecosystem model
+when an environmental condition (here parameter \emph{a}, the oxygen
+diffusivity) changes. Plot of the stable states of the simulation runs
+under different oxygen diffusivity. In this simulation there is only one
+strain in each functional group. CB\_1 = cyanobacteria strain 1; SB\_1 =
+sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria
+strain 1. Here we show a figure adapted from the output of the
+\(plot_dynamics()\) function.\}\label{fig:uc1}
+\textbackslash end\{figure\}
 
@@
 
-The package also contains a function to extract summary features of
-ecosystem responses to environmental change, such as the amount of
-hysteresis displayed by the ecosystem. Hysteresis is a key feature of
-ecosystem responses to environmental change, because it is related to
-how difficult it can be to reserve the effects of environmental change
-\citep{Scheffer2001}. The amount of hysteresis was measured here as the
-extent of the environmental condition for which there were two stable
-states. I.e. it was the extent of the environmental conditions for which
+The package contains a function to extract summary features of ecosystem
+responses to environmental change, such as the amount of hysteresis
+displayed by the ecosystem. Hysteresis is a key feature of ecosystem
+responses to environmental change, because it is related to how
+difficult it can be to reserve the effects of environmental change
+(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). The
+amount of hysteresis is measured as the extent of the environmental
+condition (here oxygen diffusitivity) for which there were two stable
+states. I.e. it is the extent of the environmental conditions for which
 historical conditions play an important role in determining the current
-system state (a definition of hysteresis). The results show that the
-amount of hysteresis depends greatly on the combinations of organisms
-present (Figure @ref(fig:uc2)). For example, with only the CB
-(cyanobacteria) present, there was no hysteresis. In contrast, the
-presence of both CB and SB (sulfate reducing bacteria) led to a large
-amount of hysteresis. (These results are also given in the Partial
-Reproduction vignette.)
+system state (a definition of hysteresis).
 
-\textbf{\citet{Rainer}\ldots{} the figure needs beautifying please.}
-\textbf{\citet{Owen}: is this OK like this?}
+Using the package to calculate the extent of hysteresis involves setting
+ecosystem and simulation parameters, including parameters for the
+finding of stable states across an environmental gradient, running the
+stable state finding function, and analysing the results with the
+function that calculates extent of hysteresis. The code for this is
+provided in the \emph{User Guide} vignette.
+
+The results show that the amount of hysteresis depends greatly on the
+combinations of organisms present (Figure @ref(fig:uc2)). For example,
+with only the CB (cyanobacteria) present, there was no hysteresis. In
+contrast, the presence of both CB and SB (sulfate reducing bacteria) led
+to a large amount of hysteresis. (These results are also given in the
+\emph{Partial Reproduction} vignette.)
 
@@
 
-{\centering \includegraphics[width=350px]{./figures/gen_uc2_user_guide_hysteresis}
+{\centering \includegraphics[width=150px]{/Users/owenpetchey/Desktop/microxanox/microxanox/inst/manuscript/figures/gen_uc2_user_guide_hysteresis}
 
@@
 
-\caption{The amount of hysteresis depends on the combination of types of organisms present.}\label{fig:uc2}
+\caption{The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.}\label{fig:uc2}
 \end{figure}
@@
 
-As discussed in the Introduction section, the package was largely
-motivated by the question of how biodiversity influences ecosystem
-responses to environmental change. Extensive results concerning this
-question are given in a separate publication \citet{Limberger2022}. Here
-we describe one of the results, which is that having biodiversity in a
-functional group can allow state changes to occur that otherwise would
-not have. I.e. biodiversity can qualitatively change the state of the
-ecosystem.
+As discussed in the Introduction section, the package was motivated by
+the question of how biodiversity influences ecosystem responses to
+environmental change. Extensive results concerning this question are
+given in a separate publication
+\protect\hyperlink{ref-Limberger2022}{Limberger}
+(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
+the results, which is that having biodiversity in a functional group can
+allow state changes to occur that otherwise would not have. I.e.
+biodiversity can qualitatively change the state of the ecosystem.
 
@@
 cyanobacteria crash in abundance, and the system switches to the anoxic
-state. Importantly, the switch does not occur if there are two identical
-strains with tolerance half way between those in Figure @ref(fig:uc3).
-
-\textbf{Rainer\ldots{} please beautify the figure as required. It comes
-from the Limberger et al supplement, Figure 5.4} \textbf{\citet{Owen}:
-is this OK like this? Used the graph from the supplement}
+state. In contrast, if there are two identical strains with tolerance
+half way between those in Figure @ref(fig:uc3) the ecosystem remains in
+the oxic state.
 
@@
 
+\hypertarget{adapting-for-other-ecosystems-and-organisms}{%
+\subsection{Adapting for other ecosystems and
+organisms}\label{adapting-for-other-ecosystems-and-organisms}}
+
+We anticipate that the package can be a useful starting point for
+investigating models of other types of ecosystem and how biodiversity in
+them affects responses to environmental change. The overall framework of
+the package, the purpose of each function, and the objects used for
+storing parameters and results could be retained. E.g., all such models
+would have parameters that differ among species and need to be described
+in an object, and studies will often need to run simulations and sets of
+simulations across environmental conditions.
+
+Researchers wanting to model a new ecosystem do not, therefore, have to
+start from scratch. This will relieve researchers from needing to making
+software design decisions, and rather focus on appropriately
+representing their ecosystem, and finding the results that interest
+them. Nevertheless, adaptation of the code in the package will require a
+person / persons that can take a conceptual model of an ecosystem and
+then represent that in terms of parameters and rate equations, and that
+is relatively proficient in R programming.
+
 \hypertarget{conclusions}{%
@@
 
-\emph{microxanox} allows the simulation, visualisation, and analysis of
-a model of a microbial ecosystem while allowing variation in the amount
-of diversity containing in each of the functional groups of organisms
-present. It has been used for the research described in another paper of
-ours that is one of the first investigations of the effects of diversity
-on ecosystem resilience \citet{Limberger2022}. In that paper, we show
+The \emph{microxanox} R package allows the simulation, visualisation,
+and analysis of a model of a microbial ecosystem while allowing
+variation in the amount of diversity containing in each of the
+functional groups of organisms present. It has been used for the
+research described in another paper that describes one of the first
+investigations of the effects of diversity on ecosystem resilience
+\protect\hyperlink{ref-Limberger2022}{Limberger}
+(\protect\hyperlink{ref-Limberger2022}{2022}). In that paper, we show
 that diversity can have large and important effects of ecosystem
-responses, highlighting the need for models such as ours, within which
-one can easily manipulate the amount of biodiversity. The
-\emph{microxanox} package has also been used to reproduce the results of
-the paper that inspired the package development \citet{Bush2017}.
+responses, highlighting the need for models such as ours, with which one
+can easily manipulate the amount of biodiversity. The \emph{microxanox}
+package has also been used to reproduce the results of the paper that
+inspired the package development \protect\hyperlink{ref-Bush2017}{Bush
+et al.} (\protect\hyperlink{ref-Bush2017}{2017}).
 
@@
 the effects of biodiversity on the usefulness of early warning signals
-of abrupt ecosystem change \citep{Scheffer2009}. In addition this
-package could be used as a template for the implementation for
-developing models of other types of ecosystems and organism. By doing
-so, other models can profit from the overall framework used, and the
-reproducibility aspects as well as the flexibility implemented.
+of abrupt ecosystem change
+(\protect\hyperlink{ref-Scheffer2009}{Scheffer et al., 2009}). In
+addition this package could be used as a template for the implementation
+for developing models of other types of ecosystems and organism. By
+doing so, other models can profit from the overall framework used, and
+the reproducibility aspects as well as the flexibility implemented.
 
@@
 
-\renewcommand\refname{References}
-\bibliography{bibliography.bib}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Binzer2016a}{}%
+Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016. Interactive effects
+of warming, eutrophication and size structure: Impacts on biodiversity
+and food-web structure. Global Change Biology 22, 220--227.
+doi:\href{https://doi.org/10.1111/gcb.13086}{10.1111/gcb.13086}
+
+\leavevmode\hypertarget{ref-Bush2017}{}%
+Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
+2017. Oxic-anoxic regime shifts mediated by feedbacks between
+biogeochemical processes and microbial community dynamics. Nature
+Communications 8, 789.
+doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
+
+\leavevmode\hypertarget{ref-Chaparro-Pedraza2021}{}%
+Chaparro-Pedraza, P.C., 2021. Fast environmental change and
+eco-evolutionary feedbacks can drive regime shifts in ecosystems before
+tipping points are crossed. Proceedings of the Royal Society B:
+Biological Sciences 288, 20211192.
+doi:\href{https://doi.org/10.1098/rspb.2021.1192}{10.1098/rspb.2021.1192}
+
+\leavevmode\hypertarget{ref-Collins2021}{}%
+Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P.,
+Ratajczak, Z., 2021. Fire frequency, state change and hysteresis in
+tallgrass prairie. Ecology Letters 24, 636--647.
+doi:\href{https://doi.org/10.1111/ele.13676}{10.1111/ele.13676}
+
+\leavevmode\hypertarget{ref-Limberger2022}{}%
+Limberger, R., 2022. \#\#\#{TO BE UPDATED}\#\#\# - {The} role of
+functional diversity in ecosystem regime shifts.
+
+\leavevmode\hypertarget{ref-Northrop2021}{}%
+Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J.,
+2021. Clockwise and counterclockwise hysteresis characterize state
+changes in the same aquatic ecosystem. Ecology Letters 24, 94--101.
+doi:\href{https://doi.org/10.1111/ele.13625}{10.1111/ele.13625}
+
+\leavevmode\hypertarget{ref-RCoreTeam2022}{}%
+R Core Team, 2022. R: {A} language and environment for statistical
+computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.
+
+\leavevmode\hypertarget{ref-REF_NEEDED}{}%
+REF NEEDED, A., 2222. {REFERNECE NEEDED}. Journal of missing references.
+
+\leavevmode\hypertarget{ref-Scheffer2009}{}%
+Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R.,
+Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
+Early-warning signals for critical transitions. Nature 461, 53--59.
+doi:\href{https://doi.org/10.1038/nature08227}{10.1038/nature08227}
+
+\leavevmode\hypertarget{ref-Scheffer2001}{}%
+Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001.
+Catastrophic shifts in ecosystems. Nature 413, 591--596.
+doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}
+
+\leavevmode\hypertarget{ref-Soetaert2010}{}%
+Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
+equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
+1--25.
+doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}
+
+\leavevmode\hypertarget{ref-Vandermeer2019}{}%
+Vandermeer, J., Perfecto, I., 2019. Hysteresis and critical transitions
+in a coffee agroecosystem. Proceedings of the National Academy of
+Sciences 116, 15074--15079.
+doi:\href{https://doi.org/10.1073/pnas.1902773116}{10.1073/pnas.1902773116}
+
+\leavevmode\hypertarget{ref-Vasseur2005}{}%
+Vasseur, D.A., McCann, K.S., 2005. A {Mechanistic Approach} for
+{Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}. The
+American Naturalist 166, 184--198.
+doi:\href{https://doi.org/10.1086/431285}{10.1086/431285}
+
+\end{CSLReferences}
 

### 2022-06-07 inst/manuscript/manuscript_microxanox.Rmd
Fix crossrefs

@@
 - **\textcolor{red}{update links}**
-- **\textcolor{red}{update cross references}**
+- [x] ~~**\textcolor{red}{update cross references}**~~
 - **\textcolor{red}{check keywords}**
-- **\textcolor{red}{add more references}**
+- [x] ~~**\textcolor{red}{add more references}**~~
 - **\textcolor{red}{expand on Impact and Conclusion section}**
@@
 # Software description
-> {>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<}
+<!-- > {>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<} -->
 
@@
 ## Software architecture
-> {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.<<}
+<!-- > {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.<<} -->
 
@@
 
-A typical simulation would look as shown in \@ref(fig:runsim_example).
+A typical simulation would look as shown in \ref{fig:runsim_example}.
 
@@
 This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).
+An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.
 
-```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap='Results of a simulation run shown using the function $plot_dynamics()$.
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap="Results of a simulation run shown using the function $plot_dynamics()$.
 In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example.
-CB_1 = cyanobacteria strain 1; SB_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
-Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.', fig.align="center", out.width = 350 }
+CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.", fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/ug_three_strains_dynamics.png")
@@
 
-> {>>This is the main section of the article and the reviewers weight the description here appropriately
-Indicate in what way new research questions can be pursued as a result of the software (if any).
-Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).
-Indicate in what way the software has changed the daily practice of its users (if so).
-Indicate how widespread the use of the software is within and outside the intended user group.
-Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<}
+<!-- > {>>This is the main section of the article and the reviewers weight the description here appropriately -->
+<!-- Indicate in what way new research questions can be pursued as a result of the software (if any). -->
+<!-- Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so). -->
+<!-- Indicate in what way the software has changed the daily practice of its users (if so). -->
+<!-- Indicate how widespread the use of the software is within and outside the intended user group. -->
+<!-- Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<} -->
 
@@
 
-An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \@ref(fig:uc1).
+An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}.
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
@@
 In this simulation there is only one strain in each functional group.
-CB_1 = cyanobacteria strain 1; SB_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
 Here we show a figure adapted from the output of the $plot_dynamics()$  function.', fig.align="center", out.width = 350}
@@
 
-The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \@ref(fig:uc2)).
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref(figL:uc2)).
 For example, with only the CB (cyanobacteria) present, there was no hysteresis.
@@
 
-Figure \@ref(fig:uc3) shows a simulation with two species (strains) in each of the three functional groups.
+Figure \ref{fig:uc3} shows a simulation with two species (strains) in each of the three functional groups.
 The ecosystem starts in the oxic state, though with relatively high abundance of each functional group.
@@
 Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
-In contrast, if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3) the ecosystem remains in the oxic state.
+In contrast, if there are two identical strains with tolerance half way between those in Figure \ref{fig:uc3} the ecosystem remains in the oxic state.
 
@@
 # Conclusions
-> {>>Set out the conclusion of this original software publication.<<}
+<!-- > {>>Set out the conclusion of this original software publication.<<} -->
 

### 2022-06-07 inst/manuscript/manuscript_microxanox.tex
Fix crossrefs

@@
   \linenumbers % turns line numbering on
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\bibliographystyle{elsarticle-harv}
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
 \ifxetex
@@
             pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\usepackage[normalem]{ulem}
+% avoid problems with \sout in headers with hyperref:
+\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 
@@
 
-% Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-% Pandoc header
+% From pandoc table feature
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+
 %% \documentclass[preprint,12pt, a4paper]{elsarticle}
@@
 \begin{document}
+
 \begin{frontmatter}
@@
 \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
-    \author[University of Zürich]{Rainer M Krug\corref{Corresponding
-Author}}
+    \author[University of Zürich]{Rainer M Krug%
+  %
+  \fnref{Corresponding Author}}
    \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
@@
    \ead{Owen.Petchey@ieu.uzh.ch} 
-      \address[University of Zürich]{Department of Evolutionary Biology
-and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
-      \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
+      \affiliation[University of Zürich]{Department of Evolutionary
+Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
+    \cortext[cor1]{Corresponding author}
+    \fntext[1]{Corresponding Author}
+    \fntext[2]{Equal contribution}
 
@@
   \end{abstract}
-   \begin{keyword} reproducibility; regime shift;\end{keyword}
+    \begin{keyword}
+    reproducibility; regime shift; \sep 
+    reproducibility; regime shift;
+  \end{keyword}
+
  \end{frontmatter}
@@
   \textbf{\textcolor{red}{update links}}
-\item
-  \textbf{\textcolor{red}{update cross references}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{red}{update cross references}}}
 \item
   \textbf{\textcolor{red}{check keywords}}
-\item
-  \textbf{\textcolor{red}{add more references}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{red}{add more references}}}
 \item
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.05}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.41}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.54}}@{}}
-\toprule
-\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill
-in this column} \\
-\midrule
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
+\toprule()
+\begin{minipage}[b]{\linewidth}\centering
+\textbf{Nr.}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Code metadata description}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Please fill in this column}
+\end{minipage} \\
+\midrule()
 \endhead
@@
 Rainer@krugs.de \\
-\bottomrule
+\bottomrule()
 \end{longtable}
@@
 about how ecosystems work and how they respond to environmental changes
-(\protect\hyperlink{ref-Binzer2016a}{Binzer et al., 2016};
-\protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021};
-\protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}). They are
-also important for developing hypotheses that can then be tested in
-empirical studies. One area of ecology in which simple models have
-played a influential role is the area of how ecosystems respond to
-gradual change in an environmental driver
-(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). An
-environmental driver is a environmental condition that affects an
-ecosystem, but is assumed to not be affected by the ecosystem, such as
-the rate of nutrient input into a lake.
+\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
+important for developing hypotheses that can then be tested in empirical
+studies. One area of ecology in which simple models have played a
+influential role is the area of how ecosystems respond to gradual change
+in an environmental driver \citep{Scheffer2001}. An environmental driver
+is a environmental condition that affects an ecosystem, but is assumed
+to not be affected by the ecosystem, such as the rate of nutrient input
+into a lake.
 
@@
 changes abruptly to a new state that is difficult to recover from
-(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). This
-possibility for abrupt, perhaps catastrophic changes that are difficult
-to reverse causes considerable concern
-(\protect\hyperlink{ref-Collins2021}{Collins et al., 2021};
-\protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021};
-\protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).
+\citep{Scheffer2001}. This possibility for abrupt, perhaps catastrophic
+changes that are difficult to reverse causes considerable concern
+\citep{Collins2021, Northrop2021, Vandermeer2019}.
 
@@
 bacteria (SB) and phototrophic sulfur bacteria (PB)). This system has
-been investigated by \protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) in a simulation study of a
+been investigated by \citet{Bush2017} in a simulation study of a
 mathematical model. They showed that gradual change in the rate at which
@@
 
-One feature of the study by \protect\hyperlink{ref-Bush2017}{Bush et
-al.} (\protect\hyperlink{ref-Bush2017}{2017}) was limited biodiversity.
+One feature of the study by \citet{Bush2017} was limited biodiversity.
 Specifically, there was no biodiversity within each of the three types
@@
 affects the ecosystem response to environmental change. This limitation
-is not specific to the study of \protect\hyperlink{ref-Bush2017}{Bush et
-al.} (\protect\hyperlink{ref-Bush2017}{2017}), there are few if any
+is not specific to the study of \citet{Bush2017}, there are few if any
 studies of the effects of biodiversity on abrupt transitions between
@@
 within functional group biodiversity affects ecosystem responses to
-environmental change \protect\hyperlink{ref-Limberger2022}{Limberger}
-(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
-the work and model of \protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}). It was with this goal in mind
-that we developed the \emph{microxanox} package. The first stage of
+environmental change \citet{Limberger2022}, and to base our work on the
+work and model of \citet{Bush2017}. It was with this goal in mind that
+we developed the \emph{microxanox} package. The first stage of
 development was to write code from scratch (as there was no available
@@
 
-\begin{quote}
-\{\textgreater\textgreater Describe the software in as much as is
-necessary to establish a vocabulary needed to explain its
-impact.\textless\textless\}
-\end{quote}
-
 The \emph{microxanox} package is for simulating a three functional group
@@
 sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-organisms and biogeochemical processes and is based on
-\protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) (See
-\protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of
-the model). At the core of the simulations is a set of ordinary
-differential equations (specified in the function
-\texttt{bushplus\_dynamic\_model()}, though this function need not be
-directly called). There are functions for running individual simulations
-and for running a set of simulations across, for example, a range of
-environmental conditions.
+organisms and biogeochemical processes and is based on \citet{Bush2017}
+(See \citet{Bush2017} for a detailed discussion of the model). At the
+core of the simulations is a set of ordinary differential equations
+(specified in the function \texttt{bushplus\_dynamic\_model()}, though
+this function need not be directly called). There are functions for
+running individual simulations and for running a set of simulations
+across, for example, a range of environmental conditions.
 
@@
 
-\begin{quote}
-\{\textgreater\textgreater Give a short overview of the overall software
-architecture; provide a pictorial component overview or similar (if
-possible). If necessary provide implementation
-details.\textless\textless\}
-\end{quote}
-
 The framework used when writing this package aimed to maximise
 simplicity for the user, and to make it easy to reproduce results (see
-the supplement to \protect\hyperlink{ref-Limberger2022}{Limberger}
-(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
-is used). As such, all the parameters needed to run a simulation or find
-a stable state (i.e.~the final state of the ecosystem) are contained in
-a single object (which can easily be created using included functions).
+the supplement to \citet{Limberger2022} for an example of how this is
+used). As such, all the parameters needed to run a simulation or find a
+stable state (i.e.~the final state of the ecosystem) are contained in a
+single object (which can easily be created using included functions).
 This parameter object is given to a function that runs the simulations
@@
 
-A typical simulation would look as shown in @ref(fig:runsim\_example).
+A typical simulation would look as shown in \ref{fig:runsim_example}.
 
@@
 this function, the ODEs are solved using the function \texttt{ode()} in
-the package \emph{deSolve} package
-(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}). The
+the package \emph{deSolve} package \citep{Soetaert2010}. The
 \texttt{run\_simulation()} function needs only one argument - an object
@@
 oxygen diffusivity (we term this the \emph{Replication method}). This is
-the method used in the \protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) study. The second runs two
+the method used in the \citet{Bush2017} study. The second runs two
 simulations, with step-wise and slowly temporally increasing oxygen
@@
 multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core
-Team, 2022}), the multthreading only works on Linux and Mac. It is
-planned to move to \texttt{parLapply()}
-(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future
-release.
+package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
+works on Linux and Mac. It is planned to move to \texttt{parLapply()}
+\citep{RCoreTeam2022} in a future release.
 
@@
 the results of a simulation run. An example plot resulting from this
-function is shown in @ref(fig:plot-dynamics).
+function is shown in \ref{fig:plot-dynamics}.
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}
+{\centering \includegraphics[width=350px]{figures/ug_three_strains_dynamics}
 
-\}
+}
 
-\textbackslash caption\{Results of a simulation run shown using the
-function \(plot_dynamics()\). In this case, there were three strains per
-functional group, though strains within functional groups had identical
-properties in this example. CB\_1 = cyanobacteria strain 1; SB\_1 =
-sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria
-strain 1. Here we show the figure produce by \(plot_dynamics()\), which
-produces figures not with all features ideal for
-publication.\}\label{fig:plot-dynamics} \textbackslash end\{figure\}
+\caption{Results of a simulation run shown using the function $plot_dynamics()$.
+In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example.
+CB  extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.}\label{fig:plot-dynamics}
+\end{figure}
 
@@
 
-\begin{quote}
-\{\textgreater\textgreater This is the main section of the article and
-the reviewers weight the description here appropriately Indicate in what
-way new research questions can be pursued as a result of the software
-(if any). Indicate in what way, and to what extent, the pursuit of
-existing research questions is improved (if so). Indicate in what way
-the software has changed the daily practice of its users (if so).
-Indicate how widespread the use of the software is within and outside
-the intended user group. Indicate in what way the software is used in
-commercial settings and/or how it led to the creation of spin-off
-companies (if so).\textless\textless\}
-\end{quote}
-
 The open source implementation and extension of the model used in
-\protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) provides the means of
-reproducing the results published while at the same time provides the
-means of doing unique, innovative, and important investigations of how
-ecosystems respond to environmental change, and how biodiversity may
-modulate this response.
+\citet{Bush2017} provides the means of reproducing the results published
+while at the same time provides the means of doing unique, innovative,
+and important investigations of how ecosystems respond to environmental
+change, and how biodiversity may modulate this response.
 
@@
 \emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
-third is taken from \protect\hyperlink{ref-REF_NEEDED}{REF NEEDED}
-(\protect\hyperlink{ref-REF_NEEDED}{2222}).
+third is taken from \citet{REF_NEEDED}.
 
@@
 
-The study of \protect\hyperlink{ref-Bush2017}{Bush et al.}
-(\protect\hyperlink{ref-Bush2017}{2017}) includes simulations of the
-effect of oxygen diffusivity (an environmental driver, in the sense that
-it affects the ecosystem but is not affected by it) on the ecosystem
-state (oxic or anoxic). The \emph{microxonax} package contains
-functionality to make a specific temporal pattern of change in the
-oxygen diffusivity. As well as allowing individual simulations during
-which oxygen diffusivity varies, this functionality forms the basis of
-the temporal method for finding stable states.
+The study of \citet{Bush2017} includes simulations of the effect of
+oxygen diffusivity (an environmental driver, in the sense that it
+affects the ecosystem but is not affected by it) on the ecosystem state
+(oxic or anoxic). The \emph{microxonax} package contains functionality
+to make a specific temporal pattern of change in the oxygen diffusivity.
+As well as allowing individual simulations during which oxygen
+diffusivity varies, this functionality forms the basis of the temporal
+method for finding stable states.
 
@@
 Reproduction} vignette, which we briefly show here (Figure
-@ref(fig:uc1). The example is composed of a single simulation, at the
+\ref{fig:uc1}. The example is composed of a single simulation, at the
 beginning of which the system is in the oxic state with high abundance
@@
 #>     intersect, setdiff, setequal, union
+#> Warning in knitr::include_graphics(fig): It is highly recommended
+#> to use relative paths for images. You had absolute paths: "/
+#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
+#> gen_uc1_partrep_temporal_state_switching.pdf"
 \end{verbatim}
 
-\textbackslash begin\{figure\}
+\begin{figure}
 
-\{\centering \includegraphics[width=350px]{/Users/owenpetchey/Desktop/microxanox/microxanox/inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching}
+{\centering \includegraphics[width=350px]{/Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching}
 
-\}
+}
 
-\textbackslash caption\{The temporal dynamics of the ecosystem model
-when an environmental condition (here parameter \emph{a}, the oxygen
-diffusivity) changes. Plot of the stable states of the simulation runs
-under different oxygen diffusivity. In this simulation there is only one
-strain in each functional group. CB\_1 = cyanobacteria strain 1; SB\_1 =
-sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria
-strain 1. Here we show a figure adapted from the output of the
-\(plot_dynamics()\) function.\}\label{fig:uc1}
-\textbackslash end\{figure\}
+\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
+Plot of the stable states of the simulation runs under different oxygen diffusivity.
+In this simulation there is only one strain in each functional group.
+CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show a figure adapted from the output of the $plot_dynamics()$  function.}\label{fig:uc1}
+\end{figure}
 
@@
 difficult it can be to reserve the effects of environmental change
-(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). The
-amount of hysteresis is measured as the extent of the environmental
-condition (here oxygen diffusitivity) for which there were two stable
-states. I.e. it is the extent of the environmental conditions for which
-historical conditions play an important role in determining the current
-system state (a definition of hysteresis).
+\citep{Scheffer2001}. The amount of hysteresis is measured as the extent
+of the environmental condition (here oxygen diffusitivity) for which
+there were two stable states. I.e. it is the extent of the environmental
+conditions for which historical conditions play an important role in
+determining the current system state (a definition of hysteresis).
 
@@
 The results show that the amount of hysteresis depends greatly on the
-combinations of organisms present (Figure @ref(fig:uc2)). For example,
+combinations of organisms present (Figure \ref(figL:uc2)). For example,
 with only the CB (cyanobacteria) present, there was no hysteresis. In
@@
 
+\begin{verbatim}
+#> Warning in knitr::include_graphics(fig): It is highly recommended
+#> to use relative paths for images. You had absolute paths: "/
+#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
+#> gen_uc2_user_guide_hysteresis.pdf"
+\end{verbatim}
+
 \begin{figure}
 
-{\centering \includegraphics[width=150px]{/Users/owenpetchey/Desktop/microxanox/microxanox/inst/manuscript/figures/gen_uc2_user_guide_hysteresis}
+{\centering \includegraphics[width=150px]{/Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/gen_uc2_user_guide_hysteresis}
 
@@
 environmental change. Extensive results concerning this question are
-given in a separate publication
-\protect\hyperlink{ref-Limberger2022}{Limberger}
-(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
-the results, which is that having biodiversity in a functional group can
-allow state changes to occur that otherwise would not have. I.e.
-biodiversity can qualitatively change the state of the ecosystem.
+given in a separate publication \citet{Limberger2022}. Here we describe
+one of the results, which is that having biodiversity in a functional
+group can allow state changes to occur that otherwise would not have.
+I.e. biodiversity can qualitatively change the state of the ecosystem.
 
@@
 
-Figure @ref(fig:uc3) shows a simulation with two species (strains) in
+Figure \ref{fig:uc3} shows a simulation with two species (strains) in
 each of the three functional groups. The ecosystem starts in the oxic
@@
 state. In contrast, if there are two identical strains with tolerance
-half way between those in Figure @ref(fig:uc3) the ecosystem remains in
+half way between those in Figure \ref{fig:uc3} the ecosystem remains in
 the oxic state.
@@
 
-\begin{quote}
-\{\textgreater\textgreater Set out the conclusion of this original
-software publication.\textless\textless\}
-\end{quote}
-
 The \emph{microxanox} R package allows the simulation, visualisation,
@@
 investigations of the effects of diversity on ecosystem resilience
-\protect\hyperlink{ref-Limberger2022}{Limberger}
-(\protect\hyperlink{ref-Limberger2022}{2022}). In that paper, we show
-that diversity can have large and important effects of ecosystem
-responses, highlighting the need for models such as ours, with which one
-can easily manipulate the amount of biodiversity. The \emph{microxanox}
-package has also been used to reproduce the results of the paper that
-inspired the package development \protect\hyperlink{ref-Bush2017}{Bush
-et al.} (\protect\hyperlink{ref-Bush2017}{2017}).
+\citet{Limberger2022}. In that paper, we show that diversity can have
+large and important effects of ecosystem responses, highlighting the
+need for models such as ours, with which one can easily manipulate the
+amount of biodiversity. The \emph{microxanox} package has also been used
+to reproduce the results of the paper that inspired the package
+development \citet{Bush2017}.
 
@@
 the effects of biodiversity on the usefulness of early warning signals
-of abrupt ecosystem change
-(\protect\hyperlink{ref-Scheffer2009}{Scheffer et al., 2009}). In
-addition this package could be used as a template for the implementation
-for developing models of other types of ecosystems and organism. By
-doing so, other models can profit from the overall framework used, and
-the reproducibility aspects as well as the flexibility implemented.
+of abrupt ecosystem change \citep{Scheffer2009}. In addition this
+package could be used as a template for the implementation for
+developing models of other types of ecosystems and organism. By doing
+so, other models can profit from the overall framework used, and the
+reproducibility aspects as well as the flexibility implemented.
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Binzer2016a}{}%
-Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016. Interactive effects
-of warming, eutrophication and size structure: Impacts on biodiversity
-and food-web structure. Global Change Biology 22, 220--227.
-doi:\href{https://doi.org/10.1111/gcb.13086}{10.1111/gcb.13086}
-
-\leavevmode\hypertarget{ref-Bush2017}{}%
-Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
-2017. Oxic-anoxic regime shifts mediated by feedbacks between
-biogeochemical processes and microbial community dynamics. Nature
-Communications 8, 789.
-doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
-
-\leavevmode\hypertarget{ref-Chaparro-Pedraza2021}{}%
-Chaparro-Pedraza, P.C., 2021. Fast environmental change and
-eco-evolutionary feedbacks can drive regime shifts in ecosystems before
-tipping points are crossed. Proceedings of the Royal Society B:
-Biological Sciences 288, 20211192.
-doi:\href{https://doi.org/10.1098/rspb.2021.1192}{10.1098/rspb.2021.1192}
-
-\leavevmode\hypertarget{ref-Collins2021}{}%
-Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P.,
-Ratajczak, Z., 2021. Fire frequency, state change and hysteresis in
-tallgrass prairie. Ecology Letters 24, 636--647.
-doi:\href{https://doi.org/10.1111/ele.13676}{10.1111/ele.13676}
-
-\leavevmode\hypertarget{ref-Limberger2022}{}%
-Limberger, R., 2022. \#\#\#{TO BE UPDATED}\#\#\# - {The} role of
-functional diversity in ecosystem regime shifts.
-
-\leavevmode\hypertarget{ref-Northrop2021}{}%
-Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J.,
-2021. Clockwise and counterclockwise hysteresis characterize state
-changes in the same aquatic ecosystem. Ecology Letters 24, 94--101.
-doi:\href{https://doi.org/10.1111/ele.13625}{10.1111/ele.13625}
-
-\leavevmode\hypertarget{ref-RCoreTeam2022}{}%
-R Core Team, 2022. R: {A} language and environment for statistical
-computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.
-
-\leavevmode\hypertarget{ref-REF_NEEDED}{}%
-REF NEEDED, A., 2222. {REFERNECE NEEDED}. Journal of missing references.
-
-\leavevmode\hypertarget{ref-Scheffer2009}{}%
-Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R.,
-Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
-Early-warning signals for critical transitions. Nature 461, 53--59.
-doi:\href{https://doi.org/10.1038/nature08227}{10.1038/nature08227}
-
-\leavevmode\hypertarget{ref-Scheffer2001}{}%
-Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001.
-Catastrophic shifts in ecosystems. Nature 413, 591--596.
-doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}
-
-\leavevmode\hypertarget{ref-Soetaert2010}{}%
-Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
-equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
-1--25.
-doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}
-
-\leavevmode\hypertarget{ref-Vandermeer2019}{}%
-Vandermeer, J., Perfecto, I., 2019. Hysteresis and critical transitions
-in a coffee agroecosystem. Proceedings of the National Academy of
-Sciences 116, 15074--15079.
-doi:\href{https://doi.org/10.1073/pnas.1902773116}{10.1073/pnas.1902773116}
-
-\leavevmode\hypertarget{ref-Vasseur2005}{}%
-Vasseur, D.A., McCann, K.S., 2005. A {Mechanistic Approach} for
-{Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}. The
-American Naturalist 166, 184--198.
-doi:\href{https://doi.org/10.1086/431285}{10.1086/431285}
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{bibliography.bib}
 

### 2022-06-07 inst/manuscript/manuscript_microxanox.tex
Ready for final read-through

@@
   \begin{abstract}
-  Ca. 100 words.
+  \(Microxanox\) is an R package to simulate a three functional group
+  system (cyanobacteria, phototrophic sulfur bacteria, and
+  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
+  oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary
+  differential equations. Simulations can be run individually or over a
+  parameter range, to find stable states. The model features multiple
+  species per functional group, where the number is only limited by
+  computational constraints. The R package is constructed in such a way,
+  that the results contain the input parameter used, so that a saved
+  results can be loaded again and the simulation be repeated.
   \end{abstract}
     \begin{keyword}
-    reproducibility; regime shift; \sep 
-    reproducibility; regime shift;
+    reproducibility \sep regime shift \sep stable state \sep 
+    ordinary differential equations
   \end{keyword}
@@
 \item[$\boxtimes$]
-  \sout{\textbf{\textcolor{red}{update cross references}}}
-\item
-  \textbf{\textcolor{red}{check keywords}}
+  \sout{\textbf{\textcolor{green}{update cross references}}}
 \item[$\boxtimes$]
-  \sout{\textbf{\textcolor{red}{add more references}}}
-\item
-  \textbf{\textcolor{red}{expand on Impact and Conclusion section}}
+  \sout{\textbf{\textcolor{green}{add abstract}}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{green}{check keywords}}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{green}{add more references}}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{green}{expand on Impact section}}}
+\item[$\boxtimes$]
+  \sout{\textbf{\textcolor{green}{expand on Conclusion section}}}
   \pagebreak
@@
 & & multidplyr \\
-& & \\
-& & \\
-& & \\
-& & \\
-& & \\
 C8 & If available Link to developer documentation/manual &

### 2022-06-08 inst/manuscript/manuscript_microxanox.Rmd

@@
     text: "Equal contribution"
-keywords: reproducibility; regime shift; 
+keywords: 
+  - reproducibility
+  - regime shift 
+  - stable state
+  - ordinary differential equations
 abstract: |
-  Ca. 100 words.
+  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary differential equations.
+Simulations can be run individually or over a parameter range, to find stable states.
+The model features multiple species per functional group, where the number is only limited by computational constraints.
+  The R package is constructed in such a way, that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
 journal: "SoftwareX"
@@
 - **\textcolor{red}{update links}**
-- [x] ~~**\textcolor{red}{update cross references}**~~
-- **\textcolor{red}{check keywords}**
-- [x] ~~**\textcolor{red}{add more references}**~~
-- **\textcolor{red}{expand on Impact and Conclusion section}**
+- [x] ~~**\textcolor{green}{update cross references}**~~
+- [x] ~~**\textcolor{green}{add abstract}**~~
+- [x] ~~**\textcolor{green}{check keywords}**~~
+- [x] ~~**\textcolor{green}{add more references}**~~
+- [x] ~~**\textcolor{green}{expand on Impact section}**~~
+- [x] ~~**\textcolor{green}{expand on Conclusion section}**~~
 \pagebreak
@@
 |         |                                                              | multidplyr                                                                     |  
-|         |                                                              |                                                                                |  
-|         |                                                              |                                                                                |  
-|         |                                                              |                                                                                |  
-|         |                                                              |                                                                                |  
-|         |                                                              |                                                                                |  
 |   C8    | If available Link to developer documentation/manual          | \textcolor{red}{TO BE ADDED}                                                   |  

### 2022-06-08 inst/manuscript/manuscript_microxanox.Rmd
Update bibliography and finalizing manuscript.

@@
   collapse = TRUE,
-  comment = "#>"
+  comment = "#>",
+  warning = FALSE,
+  messages = FALSE
 )
@@
 
-# \textcolor{red}{TODO}
-- **\textcolor{red}{update links}**
-- [x] ~~**\textcolor{green}{update cross references}**~~
-- [x] ~~**\textcolor{green}{add abstract}**~~
-- [x] ~~**\textcolor{green}{check keywords}**~~
-- [x] ~~**\textcolor{green}{add more references}**~~
-- [x] ~~**\textcolor{green}{expand on Impact section}**~~
-- [x] ~~**\textcolor{green}{expand on Conclusion section}**~~
+<!-- # \textcolor{red}{TODO} -->
+<!-- - [x] ~~**\textcolor{green}{update links}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{update cross references}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{add abstract}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{check keywords}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{add more references}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{expand on Impact section}**~~ -->
+<!-- - [x] ~~**\textcolor{green}{expand on Conclusion section}**~~ -->
 \pagebreak
@@
 |   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
+|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
 |   C3    | Code Ocean compute capsule                                   |                                                                                |  
@@
 |         |                                                              | multidplyr                                                                     |  
-|   C8    | If available Link to developer documentation/manual          | \textcolor{red}{TO BE ADDED}                                                   |  
+|   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
 |   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |
@@
 We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
-It was with this goal in mind that we developed the *microxanox* package.
+It was with this goal in mind that we developed the *microxanox* package [@Krug2022].
 The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
-The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](LINK NEEDED).
+The resulting reproduction is available as one of the package vignettes: [vignette Partial reproduction of Bush et al](https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html).
 
@@
 In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements.
-The basic and additional functionality is described in the package [User Guide](LINK_NEEDED).
+The basic and additional functionality is described in the package [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
 
@@
 The parameter object returned by `new_runsim_parameter()` contains among other things the `strain_parameter` object, which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](@LINK_NEEDED)
+For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
 
@@
 
-For a more detailed walk-through of these two approaches and explanation please see the [User Guide](@LINK_NEEDED).
+For a more detailed walk-through of these two approaches and explanation please see the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
 

### 2022-06-08 inst/manuscript/manuscript_microxanox.tex
Update bibliography and finalizing manuscript.

@@
             pdfborder={0 0 0}}
-\usepackage[normalem]{ulem}
-% avoid problems with \sout in headers with hyperref:
-\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
 
@@
 
-\hypertarget{section}{%
-\section{\texorpdfstring{\textcolor{red}{TODO}}{}}\label{section}}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{\textcolor{red}{update links}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{update cross references}}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{add abstract}}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{check keywords}}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{add more references}}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{expand on Impact section}}}
-\item[$\boxtimes$]
-  \sout{\textbf{\textcolor{green}{expand on Conclusion section}}}
 \pagebreak
-\end{itemize}
 
@@
 \url{https://github.com/UZH-PEG/microxanox} \\
+& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
 C3 & Code Ocean compute capsule & \\
@@
 C8 & If available Link to developer documentation/manual &
-\textcolor{red}{TO BE ADDED} \\
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{Usert
+Manual} \\
 C9 & Support email for questions & Rainer.Krug@uzh.ch;
@@
 work and model of \citet{Bush2017}. It was with this goal in mind that
-we developed the \emph{microxanox} package. The first stage of
-development was to write code from scratch (as there was no available
-code to start from) and to confirm that this new implementation would
-reproduce the previously published results. The resulting reproduction
-is available as one of the package vignettes:
-\href{LINK\%20NEEDED}{vignette Partial reproduction of Bush et al}.
+we developed the \emph{microxanox} package \citep{Krug2022}. The first
+stage of development was to write code from scratch (as there was no
+available code to start from) and to confirm that this new
+implementation would reproduce the previously published results. The
+resulting reproduction is available as one of the package vignettes:
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
+Partial reproduction of Bush et al}.
 
@@
 requirements. The basic and additional functionality is described in the
-package \href{LINK_NEEDED}{User Guide}.
+package
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
+Guide}.
 
@@
 \emph{User Guide} which accompanies the package or is available at
-\href{@LINK_NEEDED}{User Guide}
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
+Guide}.
 
@@
 For a more detailed walk-through of these two approaches and explanation
-please see the \href{@LINK_NEEDED}{User Guide}.
+please see the
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
+Guide}.
 
@@
 #>     intersect, setdiff, setequal, union
-#> Warning in knitr::include_graphics(fig): It is highly recommended
-#> to use relative paths for images. You had absolute paths: "/
-#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
-#> gen_uc1_partrep_temporal_state_switching.pdf"
 \end{verbatim}
@@
 \emph{Partial Reproduction} vignette.)
-
-\begin{verbatim}
-#> Warning in knitr::include_graphics(fig): It is highly recommended
-#> to use relative paths for images. You had absolute paths: "/
-#> Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/
-#> gen_uc2_user_guide_hysteresis.pdf"
-\end{verbatim}
 

### 2022-06-16 inst/manuscript/manuscript_microxanox.Rmd
Final draft

@@
 abstract: |
-  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary differential equations.
+  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
 Simulations can be run individually or over a parameter range, to find stable states.
@@
 |   C3    | Code Ocean compute capsule                                   |                                                                                |  
-|   C4    | Legal Code License                                           | CC BY 4.0                                                                      |  
+|   C4    | Legal Code License                                           | MIT                                                                            |  
 |   C5    | Code versioning system used                                  | git                                                                            |  
@@
 |   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
-|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |
+|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |
 
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement to @Limberger2022 for an example of how this is used).
+The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
 As such, all the parameters needed to run a simulation or find a stable state (i.e.
@@
 
-A typical simulation would look as shown in \ref{fig:runsim_example}.
+A typical simulation would look as shown in Figure \ref{fig:runsim_example}.
 
@@
 This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.
+An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.
 
-```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap="Results of a simulation run shown using the function $plot_dynamics()$.
+```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap="Results of a simulation run shown using the function $plot\\_dynamics()$.
 In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example.
-CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
-Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.", fig.align="center", out.width = 350 }
+CB\\_1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB\\_1 = phototrophic sulfur bacteria strain 1.", fig.align="center", out.width = 350 }
 knitr::include_graphics("figures/ug_three_strains_dynamics.png")
@@
 The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes.
-The third is taken from @REF_NEEDED.
+The third is taken from @Limberger2022 and @Petchey2022.
 
@@
 
-An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}.
+An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}).
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
@@
 
-```{r uc1, echo = FALSE,  include=TRUE, fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
+```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
 Plot of the stable states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
-CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
-Here we show a figure adapted from the output of the $plot_dynamics()$  function.', fig.align="center", out.width = 350}
+CB\textunderscore1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
 fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
-#if (!file.exists(fig)){
+if (!file.exists(fig)){
   library(ggplot2)
@@
   )  
-#}
+}
 knitr::include_graphics(fig)
@@
 
-The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref(figL:uc2)).
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}).
 For example, with only the CB (cyanobacteria) present, there was no hysteresis.
@@
 fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
-#if (!file.exists(fig)){
+if (!file.exists(fig)){
   library(ggplot2)
@@
  )  
-
-#}
+}
 knitr::include_graphics(fig)
@@
 In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
-The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.
+The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development [@Bush2017].
 

### 2022-06-16 inst/manuscript/manuscript_microxanox.tex
Final draft

@@
   sulfate-reducing bacteria) with four chemical substrates (phosphorus,
-  oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary
+  oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
   differential equations. Simulations can be run individually or over a
@@
 C3 & Code Ocean compute capsule & \\
-C4 & Legal Code License & CC BY 4.0 \\
+C4 & Legal Code License & MIT \\
 C5 & Code versioning system used & git \\
@@
 C9 & Support email for questions & Rainer.Krug@uzh.ch;
-Rainer@krugs.de \\
+Owen.Petchey@ieu.uzh \\
 \bottomrule()
@@
 simplicity for the user, and to make it easy to reproduce results (see
-the supplement to \citet{Limberger2022} for an example of how this is
-used). As such, all the parameters needed to run a simulation or find a
-stable state (i.e.~the final state of the ecosystem) are contained in a
-single object (which can easily be created using included functions).
-This parameter object is given to a function that runs the simulations
-and returns the results. The returned results object is identical to the
-parameter object but with an additional slot named \texttt{results},
-which contains the results of the run. Thus the returned results object
-contains the simulation conditions (parameters) as well as the results,
-and can be used to run the simulation again. This promotes
-reproducibility and makes incremental changes of individual parameters
-and re-running the simulations straightforward.
+the supplement \citep{Petchey2022} to \citet{Limberger2022} for an
+example of how this is used). As such, all the parameters needed to run
+a simulation or find a stable state (i.e.~the final state of the
+ecosystem) are contained in a single object (which can easily be created
+using included functions). This parameter object is given to a function
+that runs the simulations and returns the results. The returned results
+object is identical to the parameter object but with an additional slot
+named \texttt{results}, which contains the results of the run. Thus the
+returned results object contains the simulation conditions (parameters)
+as well as the results, and can be used to run the simulation again.
+This promotes reproducibility and makes incremental changes of
+individual parameters and re-running the simulations straightforward.
 
@@
 
-A typical simulation would look as shown in \ref{fig:runsim_example}.
+A typical simulation would look as shown in Figure
+\ref{fig:runsim_example}.
 
@@
 the results of a simulation run. An example plot resulting from this
-function is shown in \ref{fig:plot-dynamics}.
+function is shown in Figure \ref{fig:plot-dynamics}.
 
@@
 
-\caption{Results of a simulation run shown using the function $plot_dynamics()$.
+\caption{Results of a simulation run shown using the function $plot\_dynamics()$.
 In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example.
-CB  extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
-Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.}\label{fig:plot-dynamics}
+CB\_1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB\_1 = phototrophic sulfur bacteria strain 1.}\label{fig:plot-dynamics}
 \end{figure}
@@
 \emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
-third is taken from \citet{REF_NEEDED}.
+third is taken from \citet{Limberger2022} and \citet{Petchey2022}.
 
@@
 Reproduction} vignette, which we briefly show here (Figure
-\ref{fig:uc1}. The example is composed of a single simulation, at the
+\ref{fig:uc1}). The example is composed of a single simulation, at the
 beginning of which the system is in the oxic state with high abundance
@@
 
-\begin{verbatim}
-#> 
-#> Attaching package: 'dplyr'
-#> The following objects are masked from 'package:stats':
-#> 
-#>     filter, lag
-#> The following objects are masked from 'package:base':
-#> 
-#>     intersect, setdiff, setequal, union
-\end{verbatim}
-
 \begin{figure}
@@
 
-\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes.
+\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
 Plot of the stable states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
-CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
-Here we show a figure adapted from the output of the $plot_dynamics()$  function.}\label{fig:uc1}
+CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
 \end{figure}
@@
 The results show that the amount of hysteresis depends greatly on the
-combinations of organisms present (Figure \ref(figL:uc2)). For example,
+combinations of organisms present (Figure \ref{fig:uc2}). For example,
 with only the CB (cyanobacteria) present, there was no hysteresis. In
@@
 to reproduce the results of the paper that inspired the package
-development \citet{Bush2017}.
+development \citep{Bush2017}.
 

### 2022-06-16 inst/manuscript/manuscript_microxanox.Rmd
ms revised

@@
 abstract: |
-  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
+  $Microxanox$ is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
 Simulations can be run individually or over a parameter range, to find stable states.
-The model features multiple species per functional group, where the number is only limited by computational constraints.
-  The R package is constructed in such a way, that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
+The model can be implemented with different numbers of species per functional group.
+The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
+Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystem.
 journal: "SoftwareX"
@@
 |         |                                                              | multidplyr                                                                     |  
-|   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
+|   C8    | If available Link to developer documentation/manual          | [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
 |   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |
@@
 Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
-They are also important for developing hypotheses that can then be tested in empirical studies.
-One area of ecology in which simple models have played a influential role is the area of how ecosystems respond to gradual change in an environmental driver [@Scheffer2001].
+They are also important for developing hypotheses to test in empirical studies.
+One area of ecology in which models have played a influential role is how ecosystems respond to gradual change in an environmental driver [@Scheffer2001].
 An environmental driver is a environmental condition that affects an ecosystem, but is assumed to not be affected by the ecosystem, such as the rate of nutrient input into a lake.
@@
 An example where a gradual change of an environmental variable causes an abrupt change of the system is the switch from an aerobic (oxygen is available for metabolism) to anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
+This system has been investigated by @Bush2017 in a simulation study of a mathematical model.
 Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
-This system has been investigated by @Bush2017 in a simulation study of a mathematical model.
-They showed that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed the oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed the oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
@@
 This leaves open the question of if and how biodiversity within these types (i.e.
-functional groups) of bacteria, in their model, affects the ecosystem response to environmental change.
-This limitation is not specific to the study of @Bush2017, there are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
+functional groups) of bacteria affects the ecosystem response to environmental change.
+This limitation is not specific to the study of @Bush2017.
+There are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
 
@@
 We also added functionality that allowed: temporally varying environmental conditions, addition of random noise to state variables, and immigration.
-In addition to the model itself, the package includes some functions to analyse the results as well as visualize the results to provide a starting point for customized visualizations based on own requirements.
+In addition to the model itself, the package includes some functions to analyse the results as well as to visualize the results to provide a starting point for customized visualizations based on own requirements.
 The basic and additional functionality is described in the package [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
@@
 
-To make the simulation run with multiple species, we expressed different species characteristics in the elements of vectors and matrices.
+To make the simulation run with multiple species per functional group, we expressed different species characteristics in the elements of vectors and matrices.
 We also coded the ordinary differential equations to include the vectors and matrices, and to use matrix mathematics.
@@
 
-The package functions and code have modular structure, so that new functionality can be easily added.
+The package functions and code have modular structure so that new functionality can be easily added.
 E.g.
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
+The framework used when writing this package aimed to maximise simplicity for the user, and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
 As such, all the parameters needed to run a simulation or find a stable state (i.e.
@@
 This is the method used in the @Bush2017 study.
-The second runs two simulations, with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing or decreasing oxygen diffusivity.
-During this temporal environmental change, the state of the system was recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
+The second runs two simulations, one with step-wise and slowly temporally *increasing* oxygen diffusivity, and the other with step-wise and slowly *decreasing* oxygen diffusivity.
+During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
 We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method.
@@
 The replication method is implemented in the function `run_replication_ssfind()` which takes a parameter object as returned by the function `new_replication_ssfind_parameter()` and the number of cores for multithreading the simulation.
-As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multthreading only works on Linux and Mac.
+As the multithreading uses the package function `mclapply()` from the package `parallel` [@RCoreTeam2022], the multithreading only works on Linux and Mac.
 It is planned to move to `parLapply()` [@RCoreTeam2022] in a future release.
 
-This temporal method implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
+The temporal method is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.
 It is planned for a later release, to run these two simulations in parallel.
@@
 Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of stable states across an environmental gradient, running the stable state finding function, and analysing the results with the function that calculates extent of hysteresis.
-The code for this is provided in the *User Guide* vignette.
+The code for this is provided in the *User Guide*.
 
@@
 Researchers wanting to model a new ecosystem do not, therefore, have to start from scratch.
-This will relieve researchers from needing to making software design decisions, and rather focus on appropriately representing their ecosystem, and finding the results that interest them.
+This will relieve researchers from needing to make software design decisions, and rather focus on appropriately representing their ecosystem and finding the results that interest them.
 Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and that is relatively proficient in R programming.

### 2022-06-16 inst/manuscript/manuscript_microxanox.tex
Changes Bibliography and re-compiled.

@@
   \(Microxanox\) is an R package to simulate a three functional group
-  system (cyanobacteria, phototrophic sulfur bacteria, and
+  ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
   sulfate-reducing bacteria) with four chemical substrates (phosphorus,
@@
   differential equations. Simulations can be run individually or over a
-  parameter range, to find stable states. The model features multiple
-  species per functional group, where the number is only limited by
-  computational constraints. The R package is constructed in such a way,
-  that the results contain the input parameter used, so that a saved
-  results can be loaded again and the simulation be repeated.
+  parameter range, to find stable states. The model can be implemented
+  with different numbers of species per functional group. The package is
+  constructed in such a way that the results contain the input parameter
+  used, so that a saved results can be loaded again and the simulation
+  be repeated. Furthermore, the package framework and code should serve
+  as a useful starting point for making simulation models of other types
+  of ecosystem.
   \end{abstract}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
-\toprule()
+\toprule
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule()
+\midrule
 \endhead
@@
 C8 & If available Link to developer documentation/manual &
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{Usert
-Manual} \\
+\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
+Guide} \\
 C9 & Support email for questions & Rainer.Krug@uzh.ch;
 Owen.Petchey@ieu.uzh \\
-\bottomrule()
+\bottomrule
 \end{longtable}
@@
 \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
-important for developing hypotheses that can then be tested in empirical
-studies. One area of ecology in which simple models have played a
-influential role is the area of how ecosystems respond to gradual change
-in an environmental driver \citep{Scheffer2001}. An environmental driver
-is a environmental condition that affects an ecosystem, but is assumed
-to not be affected by the ecosystem, such as the rate of nutrient input
-into a lake.
+important for developing hypotheses to test in empirical studies. One
+area of ecology in which models have played a influential role is how
+ecosystems respond to gradual change in an environmental driver
+\citep{Scheffer2001}. An environmental driver is a environmental
+condition that affects an ecosystem, but is assumed to not be affected
+by the ecosystem, such as the rate of nutrient input into a lake.
 
@@
 available for metabolism) to anaerobic (oxygen generally unavailable)
-state in a microbial ecosystem. Three types of microbes occur in the
-model: cyanobacteria (CB) dominating the oxic state, and two types of
-sulfur bacteria that dominate the anoxic state (sulfate reducing
-bacteria (SB) and phototrophic sulfur bacteria (PB)). This system has
-been investigated by \citet{Bush2017} in a simulation study of a
-mathematical model. They showed that gradual change in the rate at which
-oxygen could diffuse into the ecosystem (termed the oxygen
-diffusitivity) could cause catastrophic changes in the ecosystem state
-that would be difficult to reverse.
+state in a microbial ecosystem. This system has been investigated by
+\citet{Bush2017} in a simulation study of a mathematical model. Three
+types of microbes occur in the model: cyanobacteria (CB) dominating the
+oxic state, and two types of sulfur bacteria that dominate the anoxic
+state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria
+(PB)). The model shows that gradual change in the rate at which oxygen
+could diffuse into the ecosystem (termed the oxygen diffusitivity) could
+cause catastrophic changes in the ecosystem state that would be
+difficult to reverse.
 
@@
 of bacteria. This leaves open the question of if and how biodiversity
-within these types (i.e.~functional groups) of bacteria, in their model,
-affects the ecosystem response to environmental change. This limitation
-is not specific to the study of \citet{Bush2017}, there are few if any
-studies of the effects of biodiversity on abrupt transitions between
-ecosystem states.
+within these types (i.e.~functional groups) of bacteria affects the
+ecosystem response to environmental change. This limitation is not
+specific to the study of \citet{Bush2017}. There are few if any studies
+of the effects of biodiversity on abrupt transitions between ecosystem
+states.
 
@@
 immigration. In addition to the model itself, the package includes some
-functions to analyse the results as well as visualize the results to
+functions to analyse the results as well as to visualize the results to
 provide a starting point for customized visualizations based on own
@@
 
-To make the simulation run with multiple species, we expressed different
-species characteristics in the elements of vectors and matrices. We also
-coded the ordinary differential equations to include the vectors and
-matrices, and to use matrix mathematics. In this way, we made it
-possible to run simulations with different numbers of species without
-having to change the underlying code.
+To make the simulation run with multiple species per functional group,
+we expressed different species characteristics in the elements of
+vectors and matrices. We also coded the ordinary differential equations
+to include the vectors and matrices, and to use matrix mathematics. In
+this way, we made it possible to run simulations with different numbers
+of species without having to change the underlying code.
 
-The package functions and code have modular structure, so that new
+The package functions and code have modular structure so that new
 functionality can be easily added. E.g. temporally defined events of any
@@
 The framework used when writing this package aimed to maximise
-simplicity for the user, and to make it easy to reproduce results (see
-the supplement \citep{Petchey2022} to \citet{Limberger2022} for an
-example of how this is used). As such, all the parameters needed to run
-a simulation or find a stable state (i.e.~the final state of the
+simplicity for the user, and to make it straightforward to reproduce
+results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
+for an example of how this is used). As such, all the parameters needed
+to run a simulation or find a stable state (i.e.~the final state of the
 ecosystem) are contained in a single object (which can easily be created
@@
 the method used in the \citet{Bush2017} study. The second runs two
-simulations, with step-wise and slowly temporally increasing oxygen
-diffusivity, and the other with step-wise and slowly decreasing or
-decreasing oxygen diffusivity. During this temporal environmental
-change, the state of the system was recorded just before change to a new
+simulations, one with step-wise and slowly temporally \emph{increasing}
+oxygen diffusivity, and the other with step-wise and slowly
+\emph{decreasing} oxygen diffusivity. During this temporal environmental
+change, the state of the system is recorded just before change to a new
 oxygen diffusivity (we term this the \emph{Temporal method}). We
@@
 multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} \citep{RCoreTeam2022}, the multthreading only
+package \texttt{parallel} \citep{RCoreTeam2022}, the multithreading only
 works on Linux and Mac. It is planned to move to \texttt{parLapply()}
@@
 
-This temporal method implemented in the function
+The temporal method is implemented in the function
 \texttt{run\_temporal\_ssfind()}, which takes a parameter object as
@@
 
-{\centering \includegraphics[width=350px]{/Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching}
+{\centering \includegraphics[width=350px]{figures/gen_uc1_partrep_temporal_state_switching}
 
@@
 function that calculates extent of hysteresis. The code for this is
-provided in the \emph{User Guide} vignette.
+provided in the \emph{User Guide}.
 
@@
 
-{\centering \includegraphics[width=150px]{/Users/rainer/git/0.RPackages/microxanox/inst/manuscript/figures/gen_uc2_user_guide_hysteresis}
+{\centering \includegraphics[width=150px]{figures/gen_uc2_user_guide_hysteresis}
 
@@
 Researchers wanting to model a new ecosystem do not, therefore, have to
-start from scratch. This will relieve researchers from needing to making
+start from scratch. This will relieve researchers from needing to make
 software design decisions, and rather focus on appropriately
-representing their ecosystem, and finding the results that interest
-them. Nevertheless, adaptation of the code in the package will require a
+representing their ecosystem and finding the results that interest them.
+Nevertheless, adaptation of the code in the package will require a
 person / persons that can take a conceptual model of an ecosystem and

### 2023-05-15 inst/manuscript/Response.md
WORKING ON TEX ARTICLE!

CC: nkourkou@uoi.gr

Manuscript Number: SOFTX-D-22-00172

Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic orANOXic states.

Dear Dr. Krug,

Thank you for submitting your manuscript to SoftwareX.

I have completed my evaluation of your manuscript. The reviewers recommend reconsideration of your manuscript following minor revision and modification. I invite you to resubmit your manuscript after addressing the comments below. Please resubmit your revised manuscript by Jun 12, 2023.

When revising your manuscript, please consider all issues mentioned in the reviewers' comments carefully: please outline every change made in response to their comments and provide suitable rebuttals for any comments not addressed. Please note that your revised submission may need to be re-reviewed.

To submit your revised manuscript, please log in as an author at https://www.editorialmanager.com/softx/, and navigate to the "Submissions Needing Revision" folder under the Author Main Menu.

SoftwareX values your contribution and I look forward to receiving your revised manuscript.￼

Kind regards,

Randall Sobie

Editor-in-Chief

SoftwareX

Editor and Reviewer comments:

# Reviewer #1:
- 1. Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.

- 2 The authors in section 3.3 claim several sentences that could confuse readers.
    - 2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
         - a) `The steady state`. Is there a unique steady state? Is there a reported contribution concerning a formal analysis of the model in the sense of stability (stable state existence and uniqueness)? Maybe, authors should refer as `a system's stable state (if it exists)`.
         - b) `The steady state of the system responds`. It should be noted that the steady state does not respond to the environment. Maybe, the authors refer to the numerical result of a steady state as a function of parameter and initial condition (if a steady state is not unique).
    - 2.2 `finding steady states that correspond to values of one environmental driver`. The software does not find a steady state. Maybe the authors refer to the software running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (oxygen diffusivity). Then, the numerical evaluation of steady states could be related to environmental conditions in real systems.
    - 2.3 `Two methods for finding steady states are implemented`. Again, the software does not find a steady state. Maybe the authors refer to two simulations proposed to evaluate an open loop simulation of the dynamic model numerically.
    - 2.4 `The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity`. This sentence is confusing. What the authors mean with `separate simulation` and `starting conditions`. Maybe the authors refer to each initial condition and parameter set value (oxygen diffusivity) as an open loop simulation run of the dynamic system. Are these simulations results focused on obtaining a bifurcation diagram Fig. 3 and 4 from {10.1038/s41467-017-00912-x}? Please clarified this point.
    - 2.5 `The second runs two simulations, one with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing oxygen diffusivity`. This sentence needs to be clarified. Authors refer `temporally` as several changes in the value of oxygen diffusivity during a simulation run time. Maybe the authors mean the following: to explore the parametric sensitivity of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated. Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.
    - 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
         - a) `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
         - b) `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
         - c) `in order to check if results were sensitive to choice of method`. This sentence needs to be clarified. There is no comparison between numerical evaluations in any sense. The numerical simulations or simulation results are not sensitive to choice or method. Indeed, the numerical solution of the dynamical model is sensitive to a set of parameters and initial conditions. Maybe the authors mean the following: to evaluate the performance of the numerical parametric sensitivity analysis of the dynamical model concerning one parametric variation (oxygen diffusivity) under two schemes.2.7 It is not clear why there are two `methods` for numerical parametric sensitivity analysis. The authors should include the arguments of this.
    - 2.8 Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H. K. (2002). Nonlinear Systems. Prentice Hall, New Jersey.)

- 3. Please consider the following suggestions:
    - Line 5 `types of ecosystem` -> `types of ecosystems` <span style="color:green">**Done**</span>
    - Line 10 `Permanent link to` -> `Permanent links to` <span style="color:green">**Done**</span>
    - Line 12 `role in the development of understanding about how ecosystems` -> `role in understanding how ecosystems` <span style="color:green">**Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`**</span>
    - Line 13: `and how they respond` -> `and respond` <span style="color:green">**Done**</span>
    - Line 15 `have played a influential role is 16 how ecosystems respond to gradual change` -> `have been influential in how ecosystems respond to a gradual change` <span style="color:green">**Changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`**</span>
    - Line 16 `is a environmental` -> `is an environmental` <span style="color:green">**Done**</span>
    - Line 17 `affects an ecosystem, but 18 is assumed to not be` -> `affects an ecosystem. Still, it is assumed not to be` <span style="color:green">**Done**</span>
    - Line 27 `change of the system` -> `change in the system` <span style="color:green">**Done**</span>
    - Line 28 `to anaerobic` -> `to an anaerobic`  <span style="color:green">**Done**</span>
    - Line 34 `termed the oxygen diffusitivity` -> `termed oxygen diffusivity` <span style="color:green">**Done**</span>
    - Line 38 `This leaves open the question of if and how biodiversity within these types (i.e. functional groups) of bacteria affects the ecosystem` -> `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem` <span style="color:green">**Done**</span>
    - Line 47 `It was with this goal in mind that we developed the microxanox package` -> `With this goal in mind, we developed the microxanox package` <span style="color:green">**Done**</span>
    - Line 58 `conditions, addition` -> `conditions, the addition` <span style="color:green">**Done**</span>
    - Line 60 `some functions to analyse the results as well as to visualize the results to provide a starting` -> `some functions to analyze the results and visualize them to provide a starting`<span style="color:green">** Changed to `some functions to analyse the results and visualize these to provide a starting point`**</span>
    - Line 71 `for running individual simulations and for running a set of simulations` -> `for running individual simulations and a set of simulations` <span style="color:green">**Done**</span>
    - Line 76 `the vectors and matrices, and to use matrix mathematics` -> `the vectors and matrices and used matrix mathematics` <span style="color:green">**Done**</span>
    - Line 80 `code have modular structure so that new functionality can be easily added. E.g. temporally` -> `code have a modular structure so that new functionality can be easily added. E.g., temporally` <span style="color:green">**Done**</span>
    - Line 87 `to maximise simplicity for the user, and` -> `to maximize simplicity for the user and` <span style="color:green">**Done**</span>
    - Line 95 `results of the run` -> `simulation results` <span style="color:green">**Done**</span>
    - Line 98 `parameters and re-running the simulations straightforward. In the following sections we` -> `parameters, and reruns the simulations straightforwardly. In the following sections, we` <span style="color:green">**Changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`**</span>
    - Line 107 `contains among other things the` -> `contains, among other things, the` <span style="color:green">**Done**</span>
    - Line 109 `their meaning and how they are created and have values set and changed please` -> `their meaning, and how they are created and have values set and changed, please` <span style="color:green">**Done**</span>
    - Line 113 ` an object which is identical to the parameter` -> `an object identical to the parameter` <span style="color:green">**Done**</span>
    - Line 116 `to run the simulation again from` -> `to rerun the simulation from` <span style="color:green">**Done**</span>
    - Line 119 `The general approach used to find the stable state of the system` -> `The general approach to finding the numerical value of a system's stable state (if it exists)` <span style="color:green">**Done**</span>
    - Line …

- 4. Please, consider English Editing Service or Proofread Your Paper.

- 5. Please, consider include two or three references of SoftwareX Journal.

More information and support

FAQ: How do I revise my submission in Editorial Manager?

https://service.elsevier.com/app/answers/detail/a_id/28463/supporthub/publishing/

You will find information relevant for you as an author on Elsevier’s Author Hub: https://www.elsevier.com/authors

FAQ: How can I reset a forgotten password?
https://service.elsevier.com/app/answers/detail/a_id/28452/supporthub/publishing/
For further assistance, please visit our customer service site: https://service.elsevier.com/app/home/supporthub/publishing/
Here you can search for solutions on a range of topics, find answers to frequently asked questions, and learn more about Editorial Manager via interactive tutorials. You can also talk 24/7 to our customer support team by phone and 24/7 by live chat and email

#AU_SOFTX#

To ensure this email reaches the intended recipient, please do not delete the above code

In compliance with data protection regulations, you may request that we remove your personal registration details at any time. (Remove my information/details). Please contact the publication office if you have any questions.

### 2023-05-15 inst/manuscript/manuscript_microxanox.Rmd
WORKING ON TEX ARTICLE!

@@
 The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
-Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystem.
+Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
 journal: "SoftwareX"
@@
 |   C1    | Current code version                                         | v0.9.0                                                                         |  
-|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
+|   C2    | Permanent links to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
 |         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
@@
 
-Mathematical models play a key role in the development of understanding about how ecosystems work and how they respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
+Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work and  respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
 They are also important for developing hypotheses to test in empirical studies.
-One area of ecology in which models have played a influential role is how ecosystems respond to gradual change in an environmental driver [@Scheffer2001].
-An environmental driver is a environmental condition that affects an ecosystem, but is assumed to not be affected by the ecosystem, such as the rate of nutrient input into a lake.
+One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].
+An environmental driver is an environmental condition that affects an ecosystem.
+Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.
 
@@
 
-An example where a gradual change of an environmental variable causes an abrupt change of the system is the switch from an aerobic (oxygen is available for metabolism) to anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
+An example where a gradual change of an environmental variable causes an abrupt change in the system is the switch from an aerobic (oxygen is available for metabolism) to an anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
 This system has been investigated by @Bush2017 in a simulation study of a mathematical model.
 Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
-The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed the oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
@@
 Specifically, there was no biodiversity within each of the three types of bacteria.
-This leaves open the question of if and how biodiversity within these types (i.e.
-functional groups) of bacteria affects the ecosystem response to environmental change.
+This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem response to environmental change.
 This limitation is not specific to the study of @Bush2017.
@@
 We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
-It was with this goal in mind that we developed the *microxanox* package [@Krug2022].
+With this goal in mind, we developed the microxanox package [@Krug2022].
 The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
@@
 Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them.
-We also added functionality that allowed: temporally varying environmental conditions, addition of random noise to state variables, and immigration.
-In addition to the model itself, the package includes some functions to analyse the results as well as to visualize the results to provide a starting point for customized visualizations based on own requirements.
+We also added functionality that allowed: temporally varying environmental conditions, the addition of random noise to state variables, and immigration.
+In addition to the model itself, the package includes some functions to analyse the results and visualize these to provide a starting point for customized visualizations based on own requirements.
 The basic and additional functionality is described in the package [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
@@
 At the core of the simulations is a set of ordinary differential equations (specified in the function `bushplus_dynamic_model()`, though this function need not be directly called).
-There are functions for running individual simulations and for running a set of simulations across, for example, a range of environmental conditions.
+There are functions for running individual simulations and a set of simulations across, for example, a range of environmental conditions.
 
 To make the simulation run with multiple species per functional group, we expressed different species characteristics in the elements of vectors and matrices.
-We also coded the ordinary differential equations to include the vectors and matrices, and to use matrix mathematics.
+We also coded the ordinary differential equations to include the vectors and matrices and used matrix mathematics.
 In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
 
-The package functions and code have modular structure so that new functionality can be easily added.
-E.g.
-temporally defined events of any type could be specified.
+The package functions and code have a modular structure so that new functionality can be easily added.
+E.g., temporally defined events of any type could be specified.
 Further, all parameter values required to run a simulation are stored in one object.
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user, and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
 As such, all the parameters needed to run a simulation or find a stable state (i.e.
@@
 This parameter object is given to a function that runs the simulations and returns the results.
-The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the results of the run.
+The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the simulation results.
 Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again.
-This promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations straightforward.
+This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.
 
@@
 The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`.
-The parameter object returned by `new_runsim_parameter()` contains among other things the `strain_parameter` object, which can be created by the function `new_strain_parameter()`.
-For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed please see the *User Guide* which accompanies the package or is available at [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
+The parameter object returned by `new_runsim_parameter()` contains, among other things, the `strain_parameter` object, which can be created by the function `new_strain_parameter()`.
+For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed, please see the *User Guide* which accompanies the package or is available at [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
 
 After the parameter object has been defined, it can be used in the `run_simulation()` function.
-The function returns an object which is identical to the parameter object, except of an additional slot containing the results.
-This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the `run_simulation()` function to run the simulation again from the parameter used to generate the results before.
+The function returns an object identical to the parameter object, except of an additional slot containing the results.
+This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the `run_simulation()` function to rerun the simulation again from the parameter used to generate the results before.
 
@@
 
-The general approach used to find the stable state of the system with a specific parameter set is to run the simulation for a long time and record the final state.
+The general approach to finding the numerical value of a system's stable state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
 When one does this across a range of environmental conditions, one discovers how the steady state of the system responds to the environmental conditions.

### 2023-05-15 inst/manuscript/manuscript_microxanox.tex
WORKING ON TEX ARTICLE!

@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+ \documentclass[preprint,12pt, a4paper]{elsarticle}
+
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
@@
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
+% \usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
 
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
 \usepackage[unicode=true]{hyperref}
-\fi
+
 \hypersetup{breaklinks=true,
@@
 \makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 
@@
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-\usepackage{lineno}
-
 \usepackage{float}
@@
 \journal{SoftwareX}
-
-%\usepackage{lineno}
-%\linenumbers
-%\usepackage{setspace}
-%\doublespacing
-%\usepackage{fancyhdr}
-%\fancyhead{}
-%\fancyfoot{}
-%\fancyhead[CO,CE]{Microxanox R package}
-%\pagestyle{fancy}
-%
 
@@
    \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
-    \author[University of Zürich]{Owen L. Petchey}
+    \author[University of Zürich]{Owen L. Petchey%
+  %
+  }
    \ead{Owen.Petchey@ieu.uzh.ch} 
@@
   as a useful starting point for making simulation models of other types
-  of ecosystem.
+  of ecosystems.
   \end{abstract}
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
-\toprule
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 C1 & Current code version & v0.9.0 \\
-C2 & Permanent link to code/repository used for this code version &
+C2 & Permanent links to code/repository used for this code version &
 \url{https://github.com/UZH-PEG/microxanox} \\
@@
 Owen.Petchey@ieu.uzh \\
-\bottomrule
 \end{longtable}
@@
 
-Mathematical models play a key role in the development of understanding
-about how ecosystems work and how they respond to environmental changes
+Mathematical models play a key role in the development of ecosystem
+models and the understanding about how ecosystems work and respond to
+environmental changes
 \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
 important for developing hypotheses to test in empirical studies. One
-area of ecology in which models have played a influential role is how
-ecosystems respond to gradual change in an environmental driver
-\citep{Scheffer2001}. An environmental driver is a environmental
-condition that affects an ecosystem, but is assumed to not be affected
-by the ecosystem, such as the rate of nutrient input into a lake.
+area of ecology in which models have been influential is the
+understanding of their response to a gradual change in an environmental
+driver \citep{Scheffer2001}. An environmental driver is an environmental
+condition that affects an ecosystem. Still, it is assumed not to be
+affected by the ecosystem, such as the rate of nutrient input into a
+lake.
 
@@
 An example where a gradual change of an environmental variable causes an
-abrupt change of the system is the switch from an aerobic (oxygen is
-available for metabolism) to anaerobic (oxygen generally unavailable)
+abrupt change in the system is the switch from an aerobic (oxygen is
+available for metabolism) to an anaerobic (oxygen generally unavailable)
 state in a microbial ecosystem. This system has been investigated by
@@
 (PB)). The model shows that gradual change in the rate at which oxygen
-could diffuse into the ecosystem (termed the oxygen diffusitivity) could
+could diffuse into the ecosystem (termed oxygen diffusitivity) could
 cause catastrophic changes in the ecosystem state that would be
@@
 Specifically, there was no biodiversity within each of the three types
-of bacteria. This leaves open the question of if and how biodiversity
-within these types (i.e.~functional groups) of bacteria affects the
+of bacteria. This leaves the question of if and how bacteria's
+biodiversity within these types (i.e., functional groups) affects the
 ecosystem response to environmental change. This limitation is not
@@
 environmental change \citet{Limberger2022}, and to base our work on the
-work and model of \citet{Bush2017}. It was with this goal in mind that
-we developed the \emph{microxanox} package \citep{Krug2022}. The first
-stage of development was to write code from scratch (as there was no
-available code to start from) and to confirm that this new
-implementation would reproduce the previously published results. The
-resulting reproduction is available as one of the package vignettes:
+work and model of \citet{Bush2017}. With this goal in mind, we developed
+the microxanox package \citep{Krug2022}. The first stage of development
+was to write code from scratch (as there was no available code to start
+from) and to confirm that this new implementation would reproduce the
+previously published results. The resulting reproduction is available as
+one of the package vignettes:
 \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
@@
 also added functionality that allowed: temporally varying environmental
-conditions, addition of random noise to state variables, and
+conditions, the addition of random noise to state variables, and
 immigration. In addition to the model itself, the package includes some
-functions to analyse the results as well as to visualize the results to
-provide a starting point for customized visualizations based on own
-requirements. The basic and additional functionality is described in the
-package
+functions to analyse the results and visualize these to provide a
+starting point for customized visualizations based on own requirements.
+The basic and additional functionality is described in the package
 \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
@@
 this function need not be directly called). There are functions for
-running individual simulations and for running a set of simulations
-across, for example, a range of environmental conditions.
+running individual simulations and a set of simulations across, for
+example, a range of environmental conditions.
 
@@
 vectors and matrices. We also coded the ordinary differential equations
-to include the vectors and matrices, and to use matrix mathematics. In
-this way, we made it possible to run simulations with different numbers
-of species without having to change the underlying code.
+to include the vectors and matrices and used matrix mathematics. In this
+way, we made it possible to run simulations with different numbers of
+species without having to change the underlying code.
 
-The package functions and code have modular structure so that new
-functionality can be easily added. E.g. temporally defined events of any
-type could be specified. Further, all parameter values required to run a
-simulation are stored in one object. Lastly, the general structure of
-the code should make it straightforward to adapt the model to other
+The package functions and code have a modular structure so that new
+functionality can be easily added. E.g., temporally defined events of
+any type could be specified. Further, all parameter values required to
+run a simulation are stored in one object. Lastly, the general structure
+of the code should make it straightforward to adapt the model to other
 similar systems (described in more detail in the Impact section).
@@
 The framework used when writing this package aimed to maximise
-simplicity for the user, and to make it straightforward to reproduce
+simplicity for the user and to make it straightforward to reproduce
 results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
@@
 object is identical to the parameter object but with an additional slot
-named \texttt{results}, which contains the results of the run. Thus the
+named \texttt{results}, which contains the simulation results. Thus the
 returned results object contains the simulation conditions (parameters)
@@
 This promotes reproducibility and makes incremental changes of
-individual parameters and re-running the simulations straightforward.
+individual parameters with a consecutive re-running of the simulations
+straightforward.
 
@@
 as created by the function \texttt{new\_runsim\_parameter()}. The
-parameter object returned by \texttt{new\_runsim\_parameter()} contains
-among other things the \texttt{strain\_parameter} object, which can be
+parameter object returned by \texttt{new\_runsim\_parameter()} contains,
+among other things, the \texttt{strain\_parameter} object, which can be
 created by the function \texttt{new\_strain\_parameter()}. For a
 detailed description of the parameter objects, their meaning and how
-they are created and have values set and changed please see the
+they are created and have values set and changed, please see the
 \emph{User Guide} which accompanies the package or is available at
@@
 \texttt{run\_simulation()} function. The function returns an object
-which is identical to the parameter object, except of an additional slot
+identical to the parameter object, except of an additional slot
 containing the results. This design produces a fully reproducible object
 as it can be used as a parameter object to be fed back into the
-\texttt{run\_simulation()} function to run the simulation again from the
-parameter used to generate the results before.
+\texttt{run\_simulation()} function to rerun the simulation again from
+the parameter used to generate the results before.
 
@@
 
-The general approach used to find the stable state of the system with a
-specific parameter set is to run the simulation for a long time and
-record the final state. When one does this across a range of
-environmental conditions, one discovers how the steady state of the
-system responds to the environmental conditions. The package contains
-functionality for finding steady states that correspond to values of one
-environmental driver, namely the value of oxygen diffusivity.
+The general approach to finding the numerical value of a system's stable
+state (if it exists) with a specific parameter set is to run the
+simulation for a long time and record the final state. When one does
+this across a range of environmental conditions, one discovers how the
+steady state of the system responds to the environmental conditions. The
+package contains functionality for finding steady states that correspond
+to values of one environmental driver, namely the value of oxygen
+diffusivity.
 
@@
 \renewcommand\refname{References}
-\bibliography{bibliography.bib}
+\bibliography{bibliography}
+\bibliographystyle{elsarticle-num-names.bst}
 

### 2023-05-31 inst/manuscript/Response.md
first go at revision by Owen

@@
-CC: nkourkou@uoi.gr
+CC: [nkourkou\@uoi.gr](mailto:nkourkou@uoi.gr){.email}
 
@@
 
-To submit your revised manuscript, please log in as an author at https://www.editorialmanager.com/softx/, and navigate to the "Submissions Needing Revision" folder under the Author Main Menu.
+To submit your revised manuscript, please log in as an author at <https://www.editorialmanager.com/softx/>, and navigate to the "Submissions Needing Revision" folder under the Author Main Menu.
 
@@
 
+##Rainer: please would you simplify the formatting of what comes below. Ideally it would be like this:
+
+**Comment text**
+
+*Response text*
+
+**Comment text**
+
+*Response text*
+
+**Comment text**
+
+*Response text*
+
+**Comment text**
+
+*Response text*
+
+**Comment text**
+
+*Response text*
+
 # Reviewer #1:
-- 1. Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
+
+-
+
+    1.  Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
+
+##Rainer: please correct and then make a response here.
 
 -   2 The authors in section 3.3 claim several sentences that could confuse readers.
-    - 2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
-         - a) `The steady state`.
+
+    -
+
+        ## 2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
+
+        ```         
+        a)  `The steady state`.
 Is there a unique steady state?
@@
 Maybe, authors should refer as `a system's stable state (if it exists)`.
-         - b) `The steady state of the system responds`.
+        ```
+
+##Response: We have added the text "The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved." ##Response: Since we do not provide methods for formal analysis of stability, we always use "steady state" and have removed all methods of "stable state".
+
+```         
+    -
+
+        b)  `The steady state of the system responds`.
 It should be noted that the steady state does not respond to the environment.
 Maybe, the authors refer to the numerical result of a steady state as a function of parameter and initial condition (if a steady state is not unique).
+
+```
+
+##Response: Thank you. We have edited the text so that we do not state that the steady state responds to the environment.
+
+```         
 -   2.2 `finding steady states that correspond to values of one environmental driver`.
@@
 Then, the numerical evaluation of steady states could be related to environmental conditions in real systems.
+```
+
+##Response: we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
+
+```         
 -   2.3 `Two methods for finding steady states are implemented`.
@@
 Maybe the authors refer to two simulations proposed to evaluate an open loop simulation of the dynamic model numerically.
+```
+
+##Response: we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
+
+```         
 -   2.4 `The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity`.
@@
 Please clarified this point.
+```
+
+##Response: We have clarified what is meant by "separate simulation" by changing to "independent simulation", and change "starting conditions" to "initial conditions".
+We hope that this more standard terminology is clearer.
+Furthermore, we have added that "This is the method used in the @Bush2017 study and was used to obtain the results in figures 3 and 4 of that article."
+
+```         
 -   2.5 `The second runs two simulations, one with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing oxygen diffusivity`.
@@
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.
-    - 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
-         - a) `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
-         - b) `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
-         - c) `in order to check if results were sensitive to choice of method`.
+```
+
+##Response: We are grateful for this different perspective on how to explain the method and have added it to the manuscript ("Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.")
+
+```         
+-
+
+    ## 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
+
+```
+
+##Response: Thank you for pointing out this source of confusion.
+We change the text to "Two strategies are implemented in order to allow comparison of their results." and moved it to just after we state that two strategies are provided.
+
+````         
+    ```         
+    a)  `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
+    ```
+````
+
+##Response: Thank you for this suggestions. We have changed the text to "Two numerical strategies for finding steady states and their sensitivity to parameters are implemented."
+
+```         
+    -
+
+        b)  `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
+
+```
+
+##Response: We have removed this claim since it is not demonstrated to be true and is unimportant.
+
+```         
+    -
+
+        c)  `in order to check if results were sensitive to choice of method`.
 This sentence needs to be clarified.
@@
 The authors should include the arguments of this.
+
+```
+
+##Response: We removed this text and now simply state that having two methods allows the comparison of their results.
+
+```         
 -   2.8 Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H.
@@
 Prentice Hall, New Jersey.)
+```
 
-- 3. Please consider the following suggestions:
-    - Line 5 `types of ecosystem` -> `types of ecosystems` <span style="color:green">**Done**</span>
-    - Line 10 `Permanent link to` -> `Permanent links to` <span style="color:green">**Done**</span>
-    - Line 12 `role in the development of understanding about how ecosystems` -> `role in understanding how ecosystems` <span style="color:green">**Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`**</span>
-    - Line 13: `and how they respond` -> `and respond` <span style="color:green">**Done**</span>
-    - Line 15 `have played a influential role is 16 how ecosystems respond to gradual change` -> `have been influential in how ecosystems respond to a gradual change` <span style="color:green">**Changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`**</span>
-    - Line 16 `is a environmental` -> `is an environmental` <span style="color:green">**Done**</span>
-    - Line 17 `affects an ecosystem, but 18 is assumed to not be` -> `affects an ecosystem. Still, it is assumed not to be` <span style="color:green">**Done**</span>
-    - Line 27 `change of the system` -> `change in the system` <span style="color:green">**Done**</span>
-    - Line 28 `to anaerobic` -> `to an anaerobic`  <span style="color:green">**Done**</span>
-    - Line 34 `termed the oxygen diffusitivity` -> `termed oxygen diffusivity` <span style="color:green">**Done**</span>
+##Response: we are of the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not completely rewritten.
+
+-
+
+    3.  Please consider the following suggestions:
+
+    -   Line 5 `types of ecosystem` -\> `types of ecosystems` [**Done**]{style="color:green"}
+    -   Line 10 `Permanent link to` -\> `Permanent links to` [**Done**]{style="color:green"}
+    -   Line 12 `role in the development of understanding about how ecosystems` -\> `role in understanding how ecosystems` [**Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`**]{style="color:green"}
+    -   Line 13: `and how they respond` -\> `and respond` [**Done**]{style="color:green"}
+    -   Line 15 `have played a influential role is 16 how ecosystems respond to gradual change` -\> `have been influential in how ecosystems respond to a gradual change` [**Changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`**]{style="color:green"}
+    -   Line 16 `is a environmental` -\> `is an environmental` [**Done**]{style="color:green"}
+    -   Line 17 `affects an ecosystem, but 18 is assumed to not be` -\> `affects an ecosystem. Still, it is assumed not to be` [**Done**]{style="color:green"}
+    -   Line 27 `change of the system` -\> `change in the system` [**Done**]{style="color:green"}
+    -   Line 28 `to anaerobic` -\> `to an anaerobic` [**Done**]{style="color:green"}
+    -   Line 34 `termed the oxygen diffusitivity` -\> `termed oxygen diffusivity` [**Done**]{style="color:green"}
     -   Line 38 `This leaves open the question of if and how biodiversity within these types (i.e.
-functional groups) of bacteria affects the ecosystem` -> `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem` <span style="color:green">**Done**</span>
-    - Line 47 `It was with this goal in mind that we developed the microxanox package` -> `With this goal in mind, we developed the microxanox package` <span style="color:green">**Done**</span>
-    - Line 58 `conditions, addition` -> `conditions, the addition` <span style="color:green">**Done**</span>
-    - Line 60 `some functions to analyse the results as well as to visualize the results to provide a starting` -> `some functions to analyze the results and visualize them to provide a starting`<span style="color:green">** Changed to `some functions to analyse the results and visualize these to provide a starting point`**</span>
-    - Line 71 `for running individual simulations and for running a set of simulations` -> `for running individual simulations and a set of simulations` <span style="color:green">**Done**</span>
-    - Line 76 `the vectors and matrices, and to use matrix mathematics` -> `the vectors and matrices and used matrix mathematics` <span style="color:green">**Done**</span>
+functional groups) of bacteria affects the ecosystem` -\> `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem` [**Done**]{style="color:green"}
+    -   Line 47 `It was with this goal in mind that we developed the microxanox package` -\> `With this goal in mind, we developed the microxanox package` [**Done**]{style="color:green"}
+    -   Line 58 `conditions, addition` -\> `conditions, the addition` [**Done**]{style="color:green"}
+    -   Line 60 `some functions to analyse the results as well as to visualize the results to provide a starting` -\> `some functions to analyze the results and visualize them to provide a starting`[\*\* Changed to `some functions to analyse the results and visualize these to provide a starting point`\*\*]{style="color:green"}
+    -   Line 71 `for running individual simulations and for running a set of simulations` -\> `for running individual simulations and a set of simulations` [**Done**]{style="color:green"}
+    -   Line 76 `the vectors and matrices, and to use matrix mathematics` -\> `the vectors and matrices and used matrix mathematics` [**Done**]{style="color:green"}
     -   Line 80 `code have modular structure so that new functionality can be easily added.
 E.g.
-temporally` -> `code have a modular structure so that new functionality can be easily added.
-E.g., temporally` <span style="color:green">**Done**</span>
-    - Line 87 `to maximise simplicity for the user, and` -> `to maximize simplicity for the user and` <span style="color:green">**Done**</span>
-    - Line 95 `results of the run` -> `simulation results` <span style="color:green">**Done**</span>
+temporally` -\> `code have a modular structure so that new functionality can be easily added.
+E.g., temporally` [**Done**]{style="color:green"}
+    -   Line 87 `to maximise simplicity for the user, and` -\> `to maximize simplicity for the user and` [**Done**]{style="color:green"}
+    -   Line 95 `results of the run` -\> `simulation results` [**Done**]{style="color:green"}
     -   Line 98 `parameters and re-running the simulations straightforward.
-In the following sections we` -> `parameters, and reruns the simulations straightforwardly.
-In the following sections, we` <span style="color:green">**Changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`**</span>
-    - Line 107 `contains among other things the` -> `contains, among other things, the` <span style="color:green">**Done**</span>
-    - Line 109 `their meaning and how they are created and have values set and changed please` -> `their meaning, and how they are created and have values set and changed, please` <span style="color:green">**Done**</span>
-    - Line 113 ` an object which is identical to the parameter` -> `an object identical to the parameter` <span style="color:green">**Done**</span>
-    - Line 116 `to run the simulation again from` -> `to rerun the simulation from` <span style="color:green">**Done**</span>
-    - Line 119 `The general approach used to find the stable state of the system` -> `The general approach to finding the numerical value of a system's stable state (if it exists)` <span style="color:green">**Done**</span>
-    - Line …
+In the following sections we` -\> `parameters, and reruns the simulations straightforwardly.
+In the following sections, we` [**Changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`**]{style="color:green"}
+    -   Line 107 `contains among other things the` -\> `contains, among other things, the` [**Done**]{style="color:green"}
+    -   Line 109 `their meaning and how they are created and have values set and changed please` -\> `their meaning, and how they are created and have values set and changed, please` [**Done**]{style="color:green"}
+    -   Line 113 `an object which is identical to the parameter` -\> `an object identical to the parameter` [**Done**]{style="color:green"}
+    -   Line 116 `to run the simulation again from` -\> `to rerun the simulation from` [**Done**]{style="color:green"}
+    -   Line 119 `The general approach used to find the stable state of the system` -\> `The general approach to finding the numerical value of a system's stable state (if it exists)` [**Done**]{style="color:green"}
+    -   Line ...
 
-- 4. Please, consider English Editing Service or Proofread Your Paper.
+-
 
-- 5. Please, consider include two or three references of SoftwareX Journal.
+    4.  Please, consider English Editing Service or Proofread Your Paper.
+
+## ##Rainer: please would you see to this?
+
+```         
+5.  Please, consider include two or three references of SoftwareX Journal.
+```
+
+##Rainer: please would you see to this?
 
@@
 
-https://service.elsevier.com/app/answers/detail/a_id/28463/supporthub/publishing/
+<https://service.elsevier.com/app/answers/detail/a_id/28463/supporthub/publishing/>
 
-You will find information relevant for you as an author on Elsevier’s Author Hub: https://www.elsevier.com/authors
+You will find information relevant for you as an author on Elsevier's Author Hub: <https://www.elsevier.com/authors>
 
 FAQ: How can I reset a forgotten password?
-https://service.elsevier.com/app/answers/detail/a_id/28452/supporthub/publishing/
-For further assistance, please visit our customer service site: https://service.elsevier.com/app/home/supporthub/publishing/
-Here you can search for solutions on a range of topics, find answers to frequently asked questions, and learn more about Editorial Manager via interactive tutorials.
+<https://service.elsevier.com/app/answers/detail/a_id/28452/supporthub/publishing/> For further assistance, please visit our customer service site: <https://service.elsevier.com/app/home/supporthub/publishing/> Here you can search for solutions on a range of topics, find answers to frequently asked questions, and learn more about Editorial Manager via interactive tutorials.
 You can also talk 24/7 to our customer support team by phone and 24/7 by live chat and email

### 2023-05-31 inst/manuscript/manuscript_microxanox.Rmd
first go at revision by Owen

@@
   - regime shift 
-  - stable state
+  - steady state
   - ordinary differential equations
@@
   $Microxanox$ is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
-Simulations can be run individually or over a parameter range, to find stable states.
+Simulations can be run individually or over a parameter range, to find steady states.
 The model can be implemented with different numbers of species per functional group.
@@
 
-Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work and  respond to environmental changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
+Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
 They are also important for developing hypotheses to test in empirical studies.
@@
 The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
-As such, all the parameters needed to run a simulation or find a stable state (i.e.
+As such, all the parameters needed to run a simulation or find a steady state (i.e.
 the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
@@
 
-## Finding Stable States
+## Finding Steady States
 
-The general approach to finding the numerical value of a system's stable state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
-When one does this across a range of environmental conditions, one discovers how the steady state of the system responds to the environmental conditions.
+The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
+When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions.
 The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-Two methods for finding steady states are implemented.
-The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity (we term this the *Replication method*).
-This is the method used in the @Bush2017 study.
-The second runs two simulations, one with step-wise and slowly temporally *increasing* oxygen diffusivity, and the other with step-wise and slowly *decreasing* oxygen diffusivity.
-During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
-We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method.
+Two numerical strategies for finding steady states and their sensitivity to parameters are implemented.
+Two strategies are implemented in order to allow comparison of their results.
+The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the *Replication method*).
+This is the method used in the @Bush2017 study and was used to obtain the results in figures 3 and 4 of that article.
+
+The second method runs two simulations, one with step-wise and slowly temporally *increasing* oxygen diffusivity, and the other with step-wise and slowly *decreasing* oxygen diffusivity.
+(Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
+
+The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.
 
@@
 
-From the results returned, summary measures about how the ecosystem stable states respond to environmental change can be extracted.
+From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted.
 The function `get_stability_measures()` returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
@@
 
-The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems respond to environmental change, and how biodiversity may modulate this response.
+The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
 
@@
 The *microxonax* package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
-As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding stable states.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
 
@@
 ```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.
+Plot of the steady states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
@@
 Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change [@Scheffer2001].
-The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two stable states.
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
 I.e.
@@
 
-Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of stable states across an environmental gradient, running the stable state finding function, and analysing the results with the function that calculates extent of hysteresis.
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
 The code for this is provided in the *User Guide*.

### 2023-06-02 inst/manuscript/Response.md
Formating of Response.md

@@
 
-##Rainer: please would you simplify the formatting of what comes below. Ideally it would be like this:
+# Reviewer #1:
+> 1.  Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
 
-**Comment text**
+*##Rainer: please correct and then make a response here.*
 
-*Response text*
+> The authors in section 3.3 claim several sentences that could confuse readers.
+2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
 
-**Comment text**
-
-*Response text*
-
-**Comment text**
-
-*Response text*
-
-**Comment text**
-
-*Response text*
-
-**Comment text**
-
-*Response text*
-
-# Reviewer #1:
-
--
-
-    1.  Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
-
-##Rainer: please correct and then make a response here.
-
--   2 The authors in section 3.3 claim several sentences that could confuse readers.
-
-    -
-
-        ## 2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
-
-        ```         
-        a)  `The steady state`.
+> a)  `The steady state`.
 Is there a unique steady state?
@@
 Maybe, authors should refer as `a system's stable state (if it exists)`.
-        ```
 
-##Response: We have added the text "The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved." ##Response: Since we do not provide methods for formal analysis of stability, we always use "steady state" and have removed all methods of "stable state".
+**Response:** We have added the text `The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.`
 
-```         
-    -
+**Response:** Since we do not provide methods for formal analysis of stability, we always use "steady state" and have removed all methods of "stable state".
 
-        b)  `The steady state of the system responds`.
+>`b)  `The steady state of the system responds`.
 It should be noted that the steady state does not respond to the environment.
@@
 
-```
+**Response:** Thank you. We have edited the text so that we do not state that the steady state responds to the environment.
 
-##Response: Thank you. We have edited the text so that we do not state that the steady state responds to the environment.
-
-```         
--   2.2 `finding steady states that correspond to values of one environmental driver`.
+> 2.2 `finding steady states that correspond to values of one environmental driver`.
 The software does not find a steady state.
@@
 Then, the numerical evaluation of steady states could be related to environmental conditions in real systems.
-```
 
-##Response: we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
+**Response:** we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
 
-```         
--   2.3 `Two methods for finding steady states are implemented`.
+> 2.3 `Two methods for finding steady states are implemented`.
 Again, the software does not find a steady state.
 Maybe the authors refer to two simulations proposed to evaluate an open loop simulation of the dynamic model numerically.
-```
 
-##Response: we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
+**Response:** we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
 
-```         
--   2.4 `The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity`.
+> 2.4 `The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity`.
 This sentence is confusing.
@@
 Please clarified this point.
-```
 
-##Response: We have clarified what is meant by "separate simulation" by changing to "independent simulation", and change "starting conditions" to "initial conditions".
+**Response:** We have clarified what is meant by "separate simulation" by changing to "independent simulation", and change "starting conditions" to "initial conditions".
 We hope that this more standard terminology is clearer.
@@
 
-```         
--   2.5 `The second runs two simulations, one with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing oxygen diffusivity`.
+> 2.5 `The second runs two simulations, one with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing oxygen diffusivity`.
 This sentence needs to be clarified.
@@
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.
-```
 
-##Response: We are grateful for this different perspective on how to explain the method and have added it to the manuscript ("Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+**Response:** We are grateful for this different perspective on how to explain the method and have added it to the manuscript ("Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.")
 
-```         
--
+> 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
 
-    ## 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
-
-```
-
-##Response: Thank you for pointing out this source of confusion.
+**Response:** Thank you for pointing out this source of confusion.
 We change the text to "Two strategies are implemented in order to allow comparison of their results." and moved it to just after we state that two strategies are provided.
 
-````         
-    ```         
-    a)  `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
-    ```
-````
+> a) `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
 
-##Response: Thank you for this suggestions. We have changed the text to "Two numerical strategies for finding steady states and their sensitivity to parameters are implemented."
+**Response:** Thank you for this suggestions. We have changed the text to "Two numerical strategies for finding steady states and their sensitivity to parameters are implemented."
 
-```         
-    -
+> b) `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
 
-        b)  `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
+**Response:** We have removed this claim since it is not demonstrated to be true and is unimportant.
 
-```
-
-##Response: We have removed this claim since it is not demonstrated to be true and is unimportant.
-
-```         
-    -
-
-        c)  `in order to check if results were sensitive to choice of method`.
+> c)  `in order to check if results were sensitive to choice of method`.
 This sentence needs to be clarified.
@@
 
-```
+**Response:** We removed this text and now simply state that having two methods allows the comparison of their results.
 
-##Response: We removed this text and now simply state that having two methods allows the comparison of their results.
-
-```         
--   2.8 Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H.
+> 2.8 Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H.
 K.
@@
 Prentice Hall, New Jersey.)
-```
 
-##Response: we are of the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not completely rewritten.
+**Response:** we are of the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not completely rewritten.
 
--
+> 3.  Please consider the following suggestions:
 
-    3.  Please consider the following suggestions:
+> Line 5 `types of ecosystem`
 
-    -   Line 5 `types of ecosystem` -\> `types of ecosystems` [**Done**]{style="color:green"}
-    -   Line 10 `Permanent link to` -\> `Permanent links to` [**Done**]{style="color:green"}
-    -   Line 12 `role in the development of understanding about how ecosystems` -\> `role in understanding how ecosystems` [**Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`**]{style="color:green"}
-    -   Line 13: `and how they respond` -\> `and respond` [**Done**]{style="color:green"}
-    -   Line 15 `have played a influential role is 16 how ecosystems respond to gradual change` -\> `have been influential in how ecosystems respond to a gradual change` [**Changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`**]{style="color:green"}
-    -   Line 16 `is a environmental` -\> `is an environmental` [**Done**]{style="color:green"}
-    -   Line 17 `affects an ecosystem, but 18 is assumed to not be` -\> `affects an ecosystem. Still, it is assumed not to be` [**Done**]{style="color:green"}
-    -   Line 27 `change of the system` -\> `change in the system` [**Done**]{style="color:green"}
-    -   Line 28 `to anaerobic` -\> `to an anaerobic` [**Done**]{style="color:green"}
-    -   Line 34 `termed the oxygen diffusitivity` -\> `termed oxygen diffusivity` [**Done**]{style="color:green"}
-    -   Line 38 `This leaves open the question of if and how biodiversity within these types (i.e.
-functional groups) of bacteria affects the ecosystem` -\> `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem` [**Done**]{style="color:green"}
-    -   Line 47 `It was with this goal in mind that we developed the microxanox package` -\> `With this goal in mind, we developed the microxanox package` [**Done**]{style="color:green"}
-    -   Line 58 `conditions, addition` -\> `conditions, the addition` [**Done**]{style="color:green"}
-    -   Line 60 `some functions to analyse the results as well as to visualize the results to provide a starting` -\> `some functions to analyze the results and visualize them to provide a starting`[\*\* Changed to `some functions to analyse the results and visualize these to provide a starting point`\*\*]{style="color:green"}
-    -   Line 71 `for running individual simulations and for running a set of simulations` -\> `for running individual simulations and a set of simulations` [**Done**]{style="color:green"}
-    -   Line 76 `the vectors and matrices, and to use matrix mathematics` -\> `the vectors and matrices and used matrix mathematics` [**Done**]{style="color:green"}
-    -   Line 80 `code have modular structure so that new functionality can be easily added.
-E.g.
-temporally` -\> `code have a modular structure so that new functionality can be easily added.
-E.g., temporally` [**Done**]{style="color:green"}
-    -   Line 87 `to maximise simplicity for the user, and` -\> `to maximize simplicity for the user and` [**Done**]{style="color:green"}
-    -   Line 95 `results of the run` -\> `simulation results` [**Done**]{style="color:green"}
-    -   Line 98 `parameters and re-running the simulations straightforward.
-In the following sections we` -\> `parameters, and reruns the simulations straightforwardly.
-In the following sections, we` [**Changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`**]{style="color:green"}
-    -   Line 107 `contains among other things the` -\> `contains, among other things, the` [**Done**]{style="color:green"}
-    -   Line 109 `their meaning and how they are created and have values set and changed please` -\> `their meaning, and how they are created and have values set and changed, please` [**Done**]{style="color:green"}
-    -   Line 113 `an object which is identical to the parameter` -\> `an object identical to the parameter` [**Done**]{style="color:green"}
-    -   Line 116 `to run the simulation again from` -\> `to rerun the simulation from` [**Done**]{style="color:green"}
-    -   Line 119 `The general approach used to find the stable state of the system` -\> `The general approach to finding the numerical value of a system's stable state (if it exists)` [**Done**]{style="color:green"}
-    -   Line ...
+Changed to  `types of ecosystems`
 
--
+>  Line 10 `Permanent link to`
 
-    4.  Please, consider English Editing Service or Proofread Your Paper.
+Changed to  `Permanent links to`
+
+>  Line 12 `role in the development of understanding about how ecosystems`
+
+`role in understanding how ecosystems` Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`
+
+>  Line 13: `and how they respond`
+
+Changed to  `and respond`
+
+>  Line 15 `have played a influential role is 16 how ecosystems respond to gradual change`
+
+`have been influential in how ecosystems respond to a gradual change` changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`
+
+>  Line 16 `is a environmental`
+
+Changed to `is an environmental`
+
+>  Line 17 `affects an ecosystem, but 18 is assumed to not be`
+
+Changed to  `affects an ecosystem. Still, it is assumed not to be`
+
+>  Line 27 `change of the system`
+
+Changed to  `change in the system`
+
+>  Line 28 `to anaerobic`
+
+Changed to  `to an anaerobic`
+
+>  Line 34 `termed the oxygen diffusitivity`
+
+Changed to  `termed oxygen diffusivity`
+
+>  Line 38 `This leaves open the question of if and how biodiversity within these types (i.e. functional groups) of bacteria affects the ecosystem`
+
+Changed to  `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem`
+
+>  Line 47 `It was with this goal in mind that we developed the microxanox package`
+
+Changed to  `With this goal in mind, we developed the microxanox package`
+
+>  Line 58 `conditions, addition`
+
+Changed to  `conditions, the addition`
+
+>  Line 60 `some functions to analyse the results as well as to visualize the results to provide a starting`
+
+`some functions to analyze the results and visualize them to provide a starting` changed to `some functions to analyse the results and visualize these to provide a starting point`
+
+>  Line 71 `for running individual simulations and for running a set of simulations`
+
+Changed to  `for running individual simulations and a set of simulations`
+
+>  Line 76 `the vectors and matrices, and to use matrix mathematics`
+
+Changed to  `the vectors and matrices and used matrix mathematics`
+
+>  Line 80 `code have modular structure so that new functionality can be easily added. E.g. temporally`
+
+Changed to  `code have a modular structure so that new functionality can be easily added. E.g., temporally`
+
+>  Line 87 `to maximise simplicity for the user, and`
+
+Changed to  `to maximize simplicity for the user and`
+
+>  Line 95 `results of the run`
+
+Changed to  `simulation results`
+
+>  Line 98 `parameters and re-running the simulations straightforward. In the following sections we`
+
+`parameters, and reruns the simulations straightforwardly.
+In the following sections, we` changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`
+
+>  Line 107 `contains among other things the`
+
+Changed to  `contains, among other things, the`
+
+>  Line 109 `their meaning and how they are created and have values set and changed please`
+
+Changed to  `their meaning, and how they are created and have values set and changed, please`
+
+>  Line 113 `an object which is identical to the parameter`
+
+Changed to  `an object identical to the parameter`
+
+>  Line 116 `to run the simulation again from`
+
+Changed to  `to rerun the simulation from`
+
+>  Line 119 `The general approach used to find the stable state of the system`
+
+Changed to  `The general approach to finding the numerical value of a system's stable state (if it exists)`
+
+>  Line ...
+
+>  4.  Please, consider English Editing Service or Proofread Your Paper.
 
@@
 
-```         
-5.  Please, consider include two or three references of SoftwareX Journal.
-```
+>  5.  Please, consider include two or three references of SoftwareX Journal.
 
 ##Rainer: please would you see to this?
+
+---
+
+---
 

### 2023-06-02 inst/manuscript/Response.md
Nearly all done...

@@
+# Manuscript Number: SOFTX-D-22-00172
+
 CC: [nkourkou\@uoi.gr](mailto:nkourkou@uoi.gr){.email}
-
-Manuscript Number: SOFTX-D-22-00172
 
@@
 
-SoftwareX values your contribution and I look forward to receiving your revised manuscript.￼
+SoftwareX values your contribution and I look forward to receiving your revised manuscript.
 
@@
 
-Editor and Reviewer comments:
+## Editor and Reviewer comments
 
-# Reviewer #1:
+### Reviewer #1
+
 > 1. Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
 
-*##Rainer: please correct and then make a response here.*
+{>>Owen: Rainer - please correct and then make a response here.<<} {>>Rainer: Will do this at the end<<} *
 
@@
 2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
-
+>
 > a)  `The steady state`.
@@
 
->`b)  `The steady state of the system responds`.
+>`b)`The steady state of the system responds`.
 It should be noted that the steady state does not respond to the environment.
@@
 > 3. Please consider the following suggestions:
-
+>
 > Line 5 `types of ecosystem`
@@
 > Line ...
-
+>
 > 4. Please, consider English Editing Service or Proofread Your Paper.
 
-## ##Rainer: please would you see to this?
+{>>Owen: Rainer - please would you see to this?<<}
 
@@
 
-##Rainer: please would you see to this?
+{>>Owen: Rainer - please would you see to this?<<}
 

### 2023-06-02 inst/manuscript/manuscript_microxanox.tex
Nearly all done...

@@
- \documentclass[preprint,12pt, a4paper]{elsarticle}
-
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-
-\usepackage[hyphens]{url}
-
- \journal{SoftwareX} % Sets Journal name
-
-\usepackage{lineno} % add
- \linenumbers % turns line numbering on
-
-\usepackage{graphicx}
-%%%%%%%%%%%%%%%% end my additions to header
-
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+% 
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
 \usepackage{lmodern}
-% \usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-
-\usepackage[unicode=true]{hyperref}
-
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
-            colorlinks=true,
-            urlcolor=blue,
-            linkcolor=blue,
-            pdfborder={0 0 0}}
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% From pandoc table feature
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs,array}
@@
 \makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
+  pdfkeywords={reproducibility, regime shift, steady state, ordinary
+differential equations},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-%% \documentclass[preprint,12pt, a4paper]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-\usepackage{float}
-\restylefloat{table}
-
-\journal{SoftwareX}
+\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial
+ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
+\author{true \and true}
+\date{2023-06-02}
 
 \begin{document}
-
-\begin{frontmatter}
-
-  \title{Microxanox: an R package for simulating an aquatic
-\(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
-    \author[University of Zürich]{Rainer M Krug%
-  %
-  \fnref{Corresponding Author}}
-   \ead{Rainer.Krug@uzh.ch; Rainer@krugs.de} 
-    \author[University of Zürich]{Owen L. Petchey%
-  %
-  }
-   \ead{Owen.Petchey@ieu.uzh.ch} 
-      \affiliation[University of Zürich]{Department of Evolutionary
-Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
-    \cortext[cor1]{Corresponding author}
-    \fntext[1]{Corresponding Author}
-    \fntext[2]{Equal contribution}
-
+\maketitle
 \begin{abstract}
@@
 differential equations. Simulations can be run individually or over a
-  parameter range, to find stable states. The model can be implemented
+parameter range, to find steady states. The model can be implemented
 with different numbers of species per functional group. The package is
 constructed in such a way that the results contain the input parameter
-  used, so that a saved results can be loaded again and the simulation
-  be repeated. Furthermore, the package framework and code should serve
-  as a useful starting point for making simulation models of other types
-  of ecosystems.
+used, so that a saved results can be loaded again and the simulation be
+repeated. Furthermore, the package framework and code should serve as a
+useful starting point for making simulation models of other types of
+ecosystems.
 \end{abstract}
-    \begin{keyword}
-    reproducibility \sep regime shift \sep stable state \sep 
-    ordinary differential equations
-  \end{keyword}
-
- \end{frontmatter}
 
@@
 Guide} \\
-C9 & Support email for questions & Rainer.Krug@uzh.ch;
-Owen.Petchey@ieu.uzh \\
+C9 & Support email for questions &
+\href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}};
+\href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
 \end{longtable}
@@
 Mathematical models play a key role in the development of ecosystem
-models and the understanding about how ecosystems work and respond to
-environmental changes
-\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
-important for developing hypotheses to test in empirical studies. One
-area of ecology in which models have been influential is the
+models and the understanding about how and why ecosystems change when
+the environment changes (\protect\hyperlink{ref-Binzer2016a}{Binzer et
+al., 2016};
+\protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021};
+\protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}). They are
+also important for developing hypotheses to test in empirical studies.
+One area of ecology in which models have been influential is the
 understanding of their response to a gradual change in an environmental
-driver \citep{Scheffer2001}. An environmental driver is an environmental
-condition that affects an ecosystem. Still, it is assumed not to be
-affected by the ecosystem, such as the rate of nutrient input into a
-lake.
+driver (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). An
+environmental driver is an environmental condition that affects an
+ecosystem. Still, it is assumed not to be affected by the ecosystem,
+such as the rate of nutrient input into a lake.
 
@@
 changes abruptly to a new state that is difficult to recover from
-\citep{Scheffer2001}. This possibility for abrupt, perhaps catastrophic
-changes that are difficult to reverse causes considerable concern
-\citep{Collins2021, Northrop2021, Vandermeer2019}.
+(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). This
+possibility for abrupt, perhaps catastrophic changes that are difficult
+to reverse causes considerable concern
+(\protect\hyperlink{ref-Collins2021}{Collins et al., 2021};
+\protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021};
+\protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).
 
@@
 state in a microbial ecosystem. This system has been investigated by
-\citet{Bush2017} in a simulation study of a mathematical model. Three
-types of microbes occur in the model: cyanobacteria (CB) dominating the
-oxic state, and two types of sulfur bacteria that dominate the anoxic
-state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria
-(PB)). The model shows that gradual change in the rate at which oxygen
-could diffuse into the ecosystem (termed oxygen diffusitivity) could
-cause catastrophic changes in the ecosystem state that would be
-difficult to reverse.
+Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}) in a simulation
+study of a mathematical model. Three types of microbes occur in the
+model: cyanobacteria (CB) dominating the oxic state, and two types of
+sulfur bacteria that dominate the anoxic state (sulfate reducing
+bacteria (SB) and phototrophic sulfur bacteria (PB)). The model shows
+that gradual change in the rate at which oxygen could diffuse into the
+ecosystem (termed oxygen diffusitivity) could cause catastrophic changes
+in the ecosystem state that would be difficult to reverse.
 
-One feature of the study by \citet{Bush2017} was limited biodiversity.
+One feature of the study by Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) was limited biodiversity.
 Specifically, there was no biodiversity within each of the three types
@@
 ecosystem response to environmental change. This limitation is not
-specific to the study of \citet{Bush2017}. There are few if any studies
+specific to the study of Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}). There are few if any studies
 of the effects of biodiversity on abrupt transitions between ecosystem
@@
 within functional group biodiversity affects ecosystem responses to
-environmental change \citet{Limberger2022}, and to base our work on the
-work and model of \citet{Bush2017}. With this goal in mind, we developed
-the microxanox package \citep{Krug2022}. The first stage of development
-was to write code from scratch (as there was no available code to start
-from) and to confirm that this new implementation would reproduce the
-previously published results. The resulting reproduction is available as
-one of the package vignettes:
+environmental change Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
+the work and model of Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}). With this goal in mind, we
+developed the microxanox package (\protect\hyperlink{ref-Krug2022}{Krug
+and Petchey, 2022}). The first stage of development was to write code
+from scratch (as there was no available code to start from) and to
+confirm that this new implementation would reproduce the previously
+published results. The resulting reproduction is available as one of the
+package vignettes:
 \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
@@
 sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-organisms and biogeochemical processes and is based on \citet{Bush2017}
-(See \citet{Bush2017} for a detailed discussion of the model). At the
-core of the simulations is a set of ordinary differential equations
-(specified in the function \texttt{bushplus\_dynamic\_model()}, though
-this function need not be directly called). There are functions for
-running individual simulations and a set of simulations across, for
-example, a range of environmental conditions.
+organisms and biogeochemical processes and is based on Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) (See Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of
+the model). At the core of the simulations is a set of ordinary
+differential equations (specified in the function
+\texttt{bushplus\_dynamic\_model()}, though this function need not be
+directly called). There are functions for running individual simulations
+and a set of simulations across, for example, a range of environmental
+conditions.
 
@@
 simplicity for the user and to make it straightforward to reproduce
-results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
-for an example of how this is used). As such, all the parameters needed
-to run a simulation or find a stable state (i.e.~the final state of the
-ecosystem) are contained in a single object (which can easily be created
-using included functions). This parameter object is given to a function
-that runs the simulations and returns the results. The returned results
-object is identical to the parameter object but with an additional slot
-named \texttt{results}, which contains the simulation results. Thus the
-returned results object contains the simulation conditions (parameters)
-as well as the results, and can be used to run the simulation again.
-This promotes reproducibility and makes incremental changes of
-individual parameters with a consecutive re-running of the simulations
-straightforward.
+results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey
+et al., 2022}) to Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
+is used). As such, all the parameters needed to run a simulation or find
+a steady state (i.e.~the final state of the ecosystem) are contained in
+a single object (which can easily be created using included functions).
+This parameter object is given to a function that runs the simulations
+and returns the results. The returned results object is identical to the
+parameter object but with an additional slot named \texttt{results},
+which contains the simulation results. Thus the returned results object
+contains the simulation conditions (parameters) as well as the results,
+and can be used to run the simulation again. This promotes
+reproducibility and makes incremental changes of individual parameters
+with a consecutive re-running of the simulations straightforward.
 
@@
 
-\hypertarget{finding-stable-states}{%
-\subsection{Finding Stable States}\label{finding-stable-states}}
+\hypertarget{finding-steady-states}{%
+\subsection{Finding Steady States}\label{finding-steady-states}}
 
-The general approach to finding the numerical value of a system's stable
+The general approach to finding the numerical value of a system's steady
 state (if it exists) with a specific parameter set is to run the
@@
 this across a range of environmental conditions, one discovers how the
-steady state of the system responds to the environmental conditions. The
-package contains functionality for finding steady states that correspond
-to values of one environmental driver, namely the value of oxygen
-diffusivity.
+steady state of the system changes with the environmental conditions.
+The package contains functionality for finding steady states that
+correspond to values of one environmental driver, namely the value of
+oxygen diffusivity.
 
-Two methods for finding steady states are implemented. The first runs a
-separate simulation for each combination of starting conditions and
+Two numerical strategies for finding steady states and their sensitivity
+to parameters are implemented. Two strategies are implemented in order
+to allow comparison of their results. The first method runs a
+independent simulation for each combination of initial conditions and
 oxygen diffusivity (we term this the \emph{Replication method}). This is
-the method used in the \citet{Bush2017} study. The second runs two
-simulations, one with step-wise and slowly temporally \emph{increasing}
-oxygen diffusivity, and the other with step-wise and slowly
-\emph{decreasing} oxygen diffusivity. During this temporal environmental
-change, the state of the system is recorded just before change to a new
-oxygen diffusivity (we term this the \emph{Temporal method}). We
-implemented two methods since there is no definitive best method, and in
-order to check if results were sensitive to choice of method.
+the method used in the Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) study and was used to obtain
+the results in figures 3 and 4 of that article.
+
+The second method runs two simulations, one with step-wise and slowly
+temporally \emph{increasing} oxygen diffusivity, and the other with
+step-wise and slowly \emph{decreasing} oxygen diffusivity. (Put another
+way, to explore the sensitivity of steady states of the dynamic model
+under oxygen diffusivity variation, subset values of this parameter are
+numerically evaluated. Two value vectors of oxygen diffusivity are
+presented step-wise, gradually increasing and decreasing.) During this
+temporal environmental change, the state of the system is recorded just
+before change to a new oxygen diffusivity (we term this the
+\emph{Temporal method}).
+
+The package does not include methods for a formal analysis of the
+stability of the system and users should take care to assess if steady
+states are unique, and indeed if a steady state has been achieved.
 
@@
 multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} \citep{RCoreTeam2022}, the multithreading only
-works on Linux and Mac. It is planned to move to \texttt{parLapply()}
-\citep{RCoreTeam2022} in a future release.
+package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core
+Team, 2022}), the multithreading only works on Linux and Mac. It is
+planned to move to \texttt{parLapply()}
+(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future
+release.
 
@@
 From the results returned, summary measures about how the ecosystem
-stable states respond to environmental change can be extracted. The
+steady states changes with environmental change can be extracted. The
 function \texttt{get\_stability\_measures()} returns quantities such as
@@
 
-The open source implementation and extension of the model used in
-\citet{Bush2017} provides the means of reproducing the results published
-while at the same time provides the means of doing unique, innovative,
-and important investigations of how ecosystems respond to environmental
-change, and how biodiversity may modulate this response.
+The open source implementation and extension of the model used in Bush
+et al. (\protect\hyperlink{ref-Bush2017}{2017}) provides the means of
+reproducing the results published while at the same time provides the
+means of doing unique, innovative, and important investigations of how
+ecosystems change as the environmental changes, and how biodiversity may
+modulate this.
 
@@
 \emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
-third is taken from \citet{Limberger2022} and \citet{Petchey2022}.
+third is taken from Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2022}) and Petchey et al.
+(\protect\hyperlink{ref-Petchey2022}{2022}).
 
@@
 
-The study of \citet{Bush2017} includes simulations of the effect of
-oxygen diffusivity (an environmental driver, in the sense that it
-affects the ecosystem but is not affected by it) on the ecosystem state
-(oxic or anoxic). The \emph{microxonax} package contains functionality
-to make a specific temporal pattern of change in the oxygen diffusivity.
-As well as allowing individual simulations during which oxygen
-diffusivity varies, this functionality forms the basis of the temporal
-method for finding stable states.
+The study of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017})
+includes simulations of the effect of oxygen diffusivity (an
+environmental driver, in the sense that it affects the ecosystem but is
+not affected by it) on the ecosystem state (oxic or anoxic). The
+\emph{microxonax} package contains functionality to make a specific
+temporal pattern of change in the oxygen diffusivity. As well as
+allowing individual simulations during which oxygen diffusivity varies,
+this functionality forms the basis of the temporal method for finding
+steady states.
 
@@
 \caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.
+Plot of the steady states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
@@
 difficult it can be to reserve the effects of environmental change
-\citep{Scheffer2001}. The amount of hysteresis is measured as the extent
-of the environmental condition (here oxygen diffusitivity) for which
-there were two stable states. I.e. it is the extent of the environmental
-conditions for which historical conditions play an important role in
-determining the current system state (a definition of hysteresis).
+(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). The
+amount of hysteresis is measured as the extent of the environmental
+condition (here oxygen diffusitivity) for which there were two steady
+states. I.e. it is the extent of the environmental conditions for which
+historical conditions play an important role in determining the current
+system state (a definition of hysteresis).
 
@@
 ecosystem and simulation parameters, including parameters for the
-finding of stable states across an environmental gradient, running the
-stable state finding function, and analysing the results with the
+finding of steady states across an environmental gradient, running the
+steady state finding function, and analysing the results with the
 function that calculates extent of hysteresis. The code for this is
@@
 environmental change. Extensive results concerning this question are
-given in a separate publication \citet{Limberger2022}. Here we describe
-one of the results, which is that having biodiversity in a functional
-group can allow state changes to occur that otherwise would not have.
-I.e. biodiversity can qualitatively change the state of the ecosystem.
+given in a separate publication Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
+the results, which is that having biodiversity in a functional group can
+allow state changes to occur that otherwise would not have. I.e.
+biodiversity can qualitatively change the state of the ecosystem.
 
@@
 investigations of the effects of diversity on ecosystem resilience
-\citet{Limberger2022}. In that paper, we show that diversity can have
-large and important effects of ecosystem responses, highlighting the
-need for models such as ours, with which one can easily manipulate the
-amount of biodiversity. The \emph{microxanox} package has also been used
-to reproduce the results of the paper that inspired the package
-development \citep{Bush2017}.
+Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2022}). In that
+paper, we show that diversity can have large and important effects of
+ecosystem responses, highlighting the need for models such as ours, with
+which one can easily manipulate the amount of biodiversity. The
+\emph{microxanox} package has also been used to reproduce the results of
+the paper that inspired the package development
+(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).
 
@@
 
-\renewcommand\refname{References}
-\bibliography{bibliography}
-\bibliographystyle{elsarticle-num-names.bst}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Binzer2016a}{}}%
+Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016. Interactive effects
+of warming, eutrophication and size structure: Impacts on biodiversity
+and food-web structure. Global Change Biology 22, 220--227.
+doi:\href{https://doi.org/10.1111/gcb.13086}{10.1111/gcb.13086}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
+Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
+2017. Oxic-anoxic regime shifts mediated by feedbacks between
+biogeochemical processes and microbial community dynamics. Nature
+Communications 8, 789.
+doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Chaparro-Pedraza2021}{}}%
+Chaparro-Pedraza, P.C., 2021. Fast environmental change and
+eco-evolutionary feedbacks can drive regime shifts in ecosystems before
+tipping points are crossed. Proceedings of the Royal Society B:
+Biological Sciences 288, 20211192.
+doi:\href{https://doi.org/10.1098/rspb.2021.1192}{10.1098/rspb.2021.1192}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Collins2021}{}}%
+Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P.,
+Ratajczak, Z., 2021. Fire frequency, state change and hysteresis in
+tallgrass prairie. Ecology Letters 24, 636--647.
+doi:\href{https://doi.org/10.1111/ele.13676}{10.1111/ele.13676}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Krug2022}{}}%
+Krug, R.M., Petchey, O.L., 2022. Microxanox: {Model} to simulate three
+functional group system.
+doi:\href{https://doi.org/10.5281/ZENODO.6624125}{10.5281/ZENODO.6624125}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Limberger2022}{}}%
+Limberger, R., Daugaard, U., Gupta, A., Krug, R., Lemmen, K., Moorsel,
+S. van, Suleiman, M., Zuppinger-Dingley, D., Petchey, O., 2022.
+Functional diversity can facilitate the collapse of an undesirable
+ecosystem state. Authorea Preprint.
+doi:\href{https://doi.org/10.22541/au.165513066.66785786/v1}{10.22541/au.165513066.66785786/v1}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Northrop2021}{}}%
+Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J.,
+2021. Clockwise and counterclockwise hysteresis characterize state
+changes in the same aquatic ecosystem. Ecology Letters 24, 94--101.
+doi:\href{https://doi.org/10.1111/ele.13625}{10.1111/ele.13625}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Petchey2022}{}}%
+Petchey, O.L., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., van
+Moorsel, Sofia J., Suleiman, M., Zuppinger-Dingley, Debra, Limberger,
+R., 2022. Code package for: "{Functional} diversity can facilitate the
+collapse of an undesirable ecosystem state".
+doi:\href{https://doi.org/10.5281/ZENODO.6334544}{10.5281/ZENODO.6334544}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
+R Core Team, 2022. R: {A} language and environment for statistical
+computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Scheffer2009}{}}%
+Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R.,
+Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
+Early-warning signals for critical transitions. Nature 461, 53--59.
+doi:\href{https://doi.org/10.1038/nature08227}{10.1038/nature08227}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Scheffer2001}{}}%
+Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001.
+Catastrophic shifts in ecosystems. Nature 413, 591--596.
+doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
+Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
+equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
+1--25.
+doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Vandermeer2019}{}}%
+Vandermeer, J., Perfecto, I., 2019. Hysteresis and critical transitions
+in a coffee agroecosystem. Proceedings of the National Academy of
+Sciences 116, 15074--15079.
+doi:\href{https://doi.org/10.1073/pnas.1902773116}{10.1073/pnas.1902773116}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Vasseur2005}{}}%
+Vasseur, D.A., McCann, K.S., 2005. A {Mechanistic Approach} for
+{Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}. The
+American Naturalist 166, 184--198.
+doi:\href{https://doi.org/10.1086/431285}{10.1086/431285}
+
+\end{CSLReferences}
 

### 2023-06-05 inst/manuscript/manuscript_microxanox.Rmd
Neraly Final version

@@
 | :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
-|   C1    | Current code version                                         | v0.9.0                                                                         |  
+|   C1    | Current code version                                         | v0.9.1                                                                         |  
 |   C2    | Permanent links to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
-|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
+|         |                                                              | [10.5281/zenodo.7148667](https://zenodo.org/record/7148667)                    |
 |   C3    | Code Ocean compute capsule                                   |                                                                                |  
 |   C4    | Legal Code License                                           | MIT                                                                            |  
-|   C5    | Code versioning system used                                  | git                                                                            |  
+|   C5    | Code versioning system used                                  | [git](https://git-scm.com)                                                                            |  
 |   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
@@
 We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
-With this goal in mind, we developed the microxanox package [@Krug2022].
+With this goal in mind, we developed the `microxanox` package [@Krug2022].
 The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
@@
 Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
-The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes.
+The first two use cases are described in detail (including the code for reproducing them) in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) and the *Partial Reproduction* vignettes.
 The third is taken from @Limberger2022 and @Petchey2022.
@@
 The study of @Bush2017 includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic).
-The *microxonax* package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
+The `microxonox` package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
 As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
@@
 In this simulation there is only one strain in each functional group.
-CB\textunderscore1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
 Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
@@
 Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
-The code for this is provided in the *User Guide*.
+The code for this is provided in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
 
@@
 
+# TODO in RMD file
+# TODO in LaTeX file
+- [ ] fix citation style
+- [ ] fix references to figures
+
 # References

### 2023-06-05 inst/manuscript/manuscript_microxanox.tex
Neraly Final version

@@
   pdftitle={Microxanox: an R package for simulating an aquatic MICRobial ecosystem that can occupy OXic or ANOXic states.},
-  pdfkeywords={reproducibility, regime shift, steady state, ordinary
-differential equations},
+  pdfkeywords={reproducibility, regime shift, steady state, ordinary differential equations},
   colorlinks=true,
@@
 
-\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial
-ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
+\title{Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
 \author{true \and true}
-\date{2023-06-02}
+\date{2023-06-05}
 
@@
 \begin{abstract}
-\(Microxanox\) is an R package to simulate a three functional group
-ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
-sulfate-reducing bacteria) with four chemical substrates (phosphorus,
-oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
-differential equations. Simulations can be run individually or over a
-parameter range, to find steady states. The model can be implemented
-with different numbers of species per functional group. The package is
-constructed in such a way that the results contain the input parameter
-used, so that a saved results can be loaded again and the simulation be
-repeated. Furthermore, the package framework and code should serve as a
-useful starting point for making simulation models of other types of
-ecosystems.
+\(Microxanox\) is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
+Simulations can be run individually or over a parameter range, to find steady states.
+The model can be implemented with different numbers of species per functional group.
+The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
+Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
 \end{abstract}
 
+{
+\hypersetup{linkcolor=}
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
 \pagebreak
@@
 \endlastfoot
-C1 & Current code version & v0.9.0 \\
-C2 & Permanent links to code/repository used for this code version &
-\url{https://github.com/UZH-PEG/microxanox} \\
-& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
+C1 & Current code version & v0.9.1 \\
+C2 & Permanent links to code/repository used for this code version & \url{https://github.com/UZH-PEG/microxanox} \\
+& & \href{https://zenodo.org/record/7148667}{10.5281/zenodo.7148667} \\
 C3 & Code Ocean compute capsule & \\
 C4 & Legal Code License & MIT \\
-C5 & Code versioning system used & git \\
-C6 & Software code languages, tools, and services used &
-\href{https://cran.r-project.org/index.html}{R} \\
-C7 & Compilation requirements, operating environments &
-\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
+C5 & Code versioning system used & \href{https://git-scm.com}{git} \\
+C6 & Software code languages, tools, and services used & \href{https://cran.r-project.org/index.html}{R} \\
+C7 & Compilation requirements, operating environments & \href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
 & & magrittr \\
@@
 & & multidplyr \\
-C8 & If available Link to developer documentation/manual &
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide} \\
-C9 & Support email for questions &
-\href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}};
-\href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
+C8 & If available Link to developer documentation/manual & \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} \\
+C9 & Support email for questions & \href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}}; \href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
 \end{longtable}
@@
 \hypertarget{motivation-and-significance}{%
-\section{Motivation and
-significance}\label{motivation-and-significance}}
+\section{Motivation and significance}\label{motivation-and-significance}}
 
-Mathematical models play a key role in the development of ecosystem
-models and the understanding about how and why ecosystems change when
-the environment changes (\protect\hyperlink{ref-Binzer2016a}{Binzer et
-al., 2016};
-\protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021};
-\protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}). They are
-also important for developing hypotheses to test in empirical studies.
-One area of ecology in which models have been influential is the
-understanding of their response to a gradual change in an environmental
-driver (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). An
-environmental driver is an environmental condition that affects an
-ecosystem. Still, it is assumed not to be affected by the ecosystem,
-such as the rate of nutrient input into a lake.
+Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes (\protect\hyperlink{ref-Binzer2016a}{Binzer et al., 2016}; \protect\hyperlink{ref-Chaparro-Pedraza2021}{Chaparro-Pedraza, 2021}; \protect\hyperlink{ref-Vasseur2005}{Vasseur and McCann, 2005}).
+They are also important for developing hypotheses to test in empirical studies.
+One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}).
+An environmental driver is an environmental condition that affects an ecosystem.
+Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.
 
-It is conceivable that an ecosystem state, such as the total biomass of
-a particular type of bacteria, may remain unchanged when an
-environmental driver changes. It is also possible that the ecosystem
-state changes gradually. It is also possible that the ecosystem state
-changes abruptly to a new state that is difficult to recover from
-(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). This
-possibility for abrupt, perhaps catastrophic changes that are difficult
-to reverse causes considerable concern
-(\protect\hyperlink{ref-Collins2021}{Collins et al., 2021};
-\protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021};
-\protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).
+It is conceivable that an ecosystem state, such as the total biomass of a particular type of bacteria, may remain unchanged when an environmental driver changes.
+It is also possible that the ecosystem state changes gradually.
+It is also possible that the ecosystem state changes abruptly to a new state that is difficult to recover from (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}).
+This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern (\protect\hyperlink{ref-Collins2021}{Collins et al., 2021}; \protect\hyperlink{ref-Northrop2021}{Northrop et al., 2021}; \protect\hyperlink{ref-Vandermeer2019}{Vandermeer and Perfecto, 2019}).
 
-An example where a gradual change of an environmental variable causes an
-abrupt change in the system is the switch from an aerobic (oxygen is
-available for metabolism) to an anaerobic (oxygen generally unavailable)
-state in a microbial ecosystem. This system has been investigated by
-Bush et al. (\protect\hyperlink{ref-Bush2017}{2017}) in a simulation
-study of a mathematical model. Three types of microbes occur in the
-model: cyanobacteria (CB) dominating the oxic state, and two types of
-sulfur bacteria that dominate the anoxic state (sulfate reducing
-bacteria (SB) and phototrophic sulfur bacteria (PB)). The model shows
-that gradual change in the rate at which oxygen could diffuse into the
-ecosystem (termed oxygen diffusitivity) could cause catastrophic changes
-in the ecosystem state that would be difficult to reverse.
+An example where a gradual change of an environmental variable causes an abrupt change in the system is the switch from an aerobic (oxygen is available for metabolism) to an anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
+This system has been investigated by Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) in a simulation study of a mathematical model.
+Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
+The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
@@
 (\protect\hyperlink{ref-Bush2017}{2017}) was limited biodiversity.
-Specifically, there was no biodiversity within each of the three types
-of bacteria. This leaves the question of if and how bacteria's
-biodiversity within these types (i.e., functional groups) affects the
-ecosystem response to environmental change. This limitation is not
-specific to the study of Bush et al.
-(\protect\hyperlink{ref-Bush2017}{2017}). There are few if any studies
-of the effects of biodiversity on abrupt transitions between ecosystem
-states.
+Specifically, there was no biodiversity within each of the three types of bacteria.
+This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem response to environmental change.
+This limitation is not specific to the study of Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}).
+There are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
 
-We decided to fill this research gap by making a simulation study of how
-within functional group biodiversity affects ecosystem responses to
-environmental change Limberger et al.
-(\protect\hyperlink{ref-Limberger2022}{2022}), and to base our work on
-the work and model of Bush et al.
-(\protect\hyperlink{ref-Bush2017}{2017}). With this goal in mind, we
-developed the microxanox package (\protect\hyperlink{ref-Krug2022}{Krug
-and Petchey, 2022}). The first stage of development was to write code
-from scratch (as there was no available code to start from) and to
-confirm that this new implementation would reproduce the previously
-published results. The resulting reproduction is available as one of the
-package vignettes:
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
-Partial reproduction of Bush et al}.
+We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2023}), and to base our work on the work and model of Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}).
+With this goal in mind, we developed the \texttt{microxanox} package (\protect\hyperlink{ref-Krug2022}{Krug and Petchey, 2022}).
+The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
+The resulting reproduction is available as one of the package vignettes: \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette Partial reproduction of Bush et al}.
 
-The second stage was to add functionality that would be necessary to
-answer our research question. Most importantly, we made it possible to
-have multiple species of bacteria within each of the three functional
-groups, for the multiple species to differ in their characteristics, and
-to vary the number of species and amount of variability among them. We
-also added functionality that allowed: temporally varying environmental
-conditions, the addition of random noise to state variables, and
-immigration. In addition to the model itself, the package includes some
-functions to analyse the results and visualize these to provide a
-starting point for customized visualizations based on own requirements.
-The basic and additional functionality is described in the package
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+The second stage was to add functionality that would be necessary to answer our research question.
+Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them.
+We also added functionality that allowed: temporally varying environmental conditions, the addition of random noise to state variables, and immigration.
+In addition to the model itself, the package includes some functions to analyse the results and visualize these to provide a starting point for customized visualizations based on own requirements.
+The basic and additional functionality is described in the package \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
@@
 
-The \emph{microxanox} package is for simulating a three functional group
-system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
-bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
-substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
-sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-organisms and biogeochemical processes and is based on Bush et al.
+The \emph{microxanox} package is for simulating a three functional group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced sulfur, \emph{SO}: oxidized sulfur).
+It includes feedback between organisms and biogeochemical processes and is based on Bush et al.
 (\protect\hyperlink{ref-Bush2017}{2017}) (See Bush et al.
-(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of
-the model). At the core of the simulations is a set of ordinary
-differential equations (specified in the function
-\texttt{bushplus\_dynamic\_model()}, though this function need not be
-directly called). There are functions for running individual simulations
-and a set of simulations across, for example, a range of environmental
-conditions.
+(\protect\hyperlink{ref-Bush2017}{2017}) for a detailed discussion of the model).
+At the core of the simulations is a set of ordinary differential equations (specified in the function \texttt{bushplus\_dynamic\_model()}, though this function need not be directly called).
+There are functions for running individual simulations and a set of simulations across, for example, a range of environmental conditions.
 
-To make the simulation run with multiple species per functional group,
-we expressed different species characteristics in the elements of
-vectors and matrices. We also coded the ordinary differential equations
-to include the vectors and matrices and used matrix mathematics. In this
-way, we made it possible to run simulations with different numbers of
-species without having to change the underlying code.
+To make the simulation run with multiple species per functional group, we expressed different species characteristics in the elements of vectors and matrices.
+We also coded the ordinary differential equations to include the vectors and matrices and used matrix mathematics.
+In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
 
-The package functions and code have a modular structure so that new
-functionality can be easily added. E.g., temporally defined events of
-any type could be specified. Further, all parameter values required to
-run a simulation are stored in one object. Lastly, the general structure
-of the code should make it straightforward to adapt the model to other
-similar systems (described in more detail in the Impact section).
+The package functions and code have a modular structure so that new functionality can be easily added.
+E.g., temporally defined events of any type could be specified.
+Further, all parameter values required to run a simulation are stored in one object.
+Lastly, the general structure of the code should make it straightforward to adapt the model to other similar systems (described in more detail in the Impact section).
 
@@
 
-The framework used when writing this package aimed to maximise
-simplicity for the user and to make it straightforward to reproduce
-results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey
-et al., 2022}) to Limberger et al.
-(\protect\hyperlink{ref-Limberger2022}{2022}) for an example of how this
-is used). As such, all the parameters needed to run a simulation or find
-a steady state (i.e.~the final state of the ecosystem) are contained in
-a single object (which can easily be created using included functions).
-This parameter object is given to a function that runs the simulations
-and returns the results. The returned results object is identical to the
-parameter object but with an additional slot named \texttt{results},
-which contains the simulation results. Thus the returned results object
-contains the simulation conditions (parameters) as well as the results,
-and can be used to run the simulation again. This promotes
-reproducibility and makes incremental changes of individual parameters
-with a consecutive re-running of the simulations straightforward.
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement (\protect\hyperlink{ref-Petchey2022}{Petchey et al., 2022}) to Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2023}) for an example of how this is used).
+As such, all the parameters needed to run a simulation or find a steady state (i.e.~the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
+This parameter object is given to a function that runs the simulations and returns the results.
+The returned results object is identical to the parameter object but with an additional slot named \texttt{results}, which contains the simulation results.
+Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again.
+This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.
 
-In the following sections we describe how to use the package to run one
-simulation and to find steady states across an environmental gradient.
+In the following sections we describe how to use the package to run one simulation and to find steady states across an environmental gradient.
 
@@
 
-A typical simulation would look as shown in Figure
-\ref{fig:runsim_example}.
+A typical simulation would look as shown in Figure \ref{fig:runsim_example}.
 
@@
 \caption{Typical flow of a simulation.
-Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
+Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}(\#fig:runsim_example)
 \end{figure}
 
-A simulation is run using the \texttt{run\_simulation()} function. In
-this function, the ODEs are solved using the function \texttt{ode()} in
-the package \emph{deSolve} package
-(\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}). The
-\texttt{run\_simulation()} function needs only one argument - an object
-as created by the function \texttt{new\_runsim\_parameter()}. The
-parameter object returned by \texttt{new\_runsim\_parameter()} contains,
-among other things, the \texttt{strain\_parameter} object, which can be
-created by the function \texttt{new\_strain\_parameter()}. For a
-detailed description of the parameter objects, their meaning and how
-they are created and have values set and changed, please see the
-\emph{User Guide} which accompanies the package or is available at
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+A simulation is run using the \texttt{run\_simulation()} function.
+In this function, the ODEs are solved using the function \texttt{ode()} in the package \emph{deSolve} package (\protect\hyperlink{ref-Soetaert2010}{Soetaert et al., 2010}).
+The \texttt{run\_simulation()} function needs only one argument - an object as created by the function \texttt{new\_runsim\_parameter()}.
+The parameter object returned by \texttt{new\_runsim\_parameter()} contains, among other things, the \texttt{strain\_parameter} object, which can be created by the function \texttt{new\_strain\_parameter()}.
+For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed, please see the \emph{User Guide} which accompanies the package or is available at \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
-After the parameter object has been defined, it can be used in the
-\texttt{run\_simulation()} function. The function returns an object
-identical to the parameter object, except of an additional slot
-containing the results. This design produces a fully reproducible object
-as it can be used as a parameter object to be fed back into the
-\texttt{run\_simulation()} function to rerun the simulation again from
-the parameter used to generate the results before.
+After the parameter object has been defined, it can be used in the \texttt{run\_simulation()} function.
+The function returns an object identical to the parameter object, except of an additional slot containing the results.
+This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the \texttt{run\_simulation()} function to rerun the simulation again from the parameter used to generate the results before.
 
@@
 
-The general approach to finding the numerical value of a system's steady
-state (if it exists) with a specific parameter set is to run the
-simulation for a long time and record the final state. When one does
-this across a range of environmental conditions, one discovers how the
-steady state of the system changes with the environmental conditions.
-The package contains functionality for finding steady states that
-correspond to values of one environmental driver, namely the value of
-oxygen diffusivity.
+The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
+When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions.
+The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-Two numerical strategies for finding steady states and their sensitivity
-to parameters are implemented. Two strategies are implemented in order
-to allow comparison of their results. The first method runs a
-independent simulation for each combination of initial conditions and
-oxygen diffusivity (we term this the \emph{Replication method}). This is
-the method used in the Bush et al.
-(\protect\hyperlink{ref-Bush2017}{2017}) study and was used to obtain
-the results in figures 3 and 4 of that article.
+Two numerical strategies for finding steady states and their sensitivity to parameters are implemented.
+Two strategies are implemented in order to allow comparison of their results.
+The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the \emph{Replication method}).
+This is the method used in the Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) study and was used to obtain the results in figures 3 and 4 of that article.
 
-The second method runs two simulations, one with step-wise and slowly
-temporally \emph{increasing} oxygen diffusivity, and the other with
-step-wise and slowly \emph{decreasing} oxygen diffusivity. (Put another
-way, to explore the sensitivity of steady states of the dynamic model
-under oxygen diffusivity variation, subset values of this parameter are
-numerically evaluated. Two value vectors of oxygen diffusivity are
-presented step-wise, gradually increasing and decreasing.) During this
-temporal environmental change, the state of the system is recorded just
-before change to a new oxygen diffusivity (we term this the
-\emph{Temporal method}).
+The second method runs two simulations, one with step-wise and slowly temporally \emph{increasing} oxygen diffusivity, and the other with step-wise and slowly \emph{decreasing} oxygen diffusivity.
+(Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the \emph{Temporal method}).
 
-The package does not include methods for a formal analysis of the
-stability of the system and users should take care to assess if steady
-states are unique, and indeed if a steady state has been achieved.
+The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.
 
-The replication method is implemented in the function
-\texttt{run\_replication\_ssfind()} which takes a parameter object as
-returned by the function \texttt{new\_replication\_ssfind\_parameter()}
-and the number of cores for multithreading the simulation. As the
-multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core
-Team, 2022}), the multithreading only works on Linux and Mac. It is
-planned to move to \texttt{parLapply()}
-(\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future
-release.
+The replication method is implemented in the function \texttt{run\_replication\_ssfind()} which takes a parameter object as returned by the function \texttt{new\_replication\_ssfind\_parameter()} and the number of cores for multithreading the simulation.
+As the multithreading uses the package function \texttt{mclapply()} from the package \texttt{parallel} (\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}), the multithreading only works on Linux and Mac.
+It is planned to move to \texttt{parLapply()} (\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}) in a future release.
 
-The temporal method is implemented in the function
-\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
-created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
-is planned for a later release, to run these two simulations in
-parallel.
+The temporal method is implemented in the function \texttt{run\_temporal\_ssfind()}, which takes a parameter object as created by the function \texttt{new\_temporal\_ssfind\_parameter()}.
+It is planned for a later release, to run these two simulations in parallel.
 
-For a more detailed walk-through of these two approaches and explanation
-please see the
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+For a more detailed walk-through of these two approaches and explanation please see the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
 \hypertarget{analysing-and-visualising-results}{%
-\subsection{Analysing and visualising
-results}\label{analysing-and-visualising-results}}
+\subsection{Analysing and visualising results}\label{analysing-and-visualising-results}}
 
-From the results returned, summary measures about how the ecosystem
-steady states changes with environmental change can be extracted. The
-function \texttt{get\_stability\_measures()} returns quantities such as
-the amount of environmental change required for the system to abruptly
-change to a different state.
+From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted.
+The function \texttt{get\_stability\_measures()} returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
 
-The function \texttt{plot\_dynamics()} plots a single simulation run, as
-returned from the \texttt{run\_simulation()} function. This function is
-only provided as a convenience function to provide a way to easily see
-the results of a simulation run. An example plot resulting from this
-function is shown in Figure \ref{fig:plot-dynamics}.
+The function \texttt{plot\_dynamics()} plots a single simulation run, as returned from the \texttt{run\_simulation()} function.
+This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
+An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.
 
@@
 
-The open source implementation and extension of the model used in Bush
-et al. (\protect\hyperlink{ref-Bush2017}{2017}) provides the means of
-reproducing the results published while at the same time provides the
-means of doing unique, innovative, and important investigations of how
-ecosystems change as the environmental changes, and how biodiversity may
-modulate this.
+The open source implementation and extension of the model used in Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
 
-The design of the package code and functionality is with reproducibility
-in mind: the combination of all parameters being in a single parameter
-object as well as the return of the simulation as a result object which
-inherits from the parameter object provides a relatively easy to use
-framework to implement reproducible experiments.
+The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
 
-Here we evidence the impact of the \emph{microxanox} package by
-describing three use cases and then by describing how the package can be
-a starting point for models of other ecosystems. The first two use cases
-are described in detail (including the code for reproducing them) in the
-\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
-third is taken from Limberger et al.
-(\protect\hyperlink{ref-Limberger2022}{2022}) and Petchey et al.
+Here we evidence the impact of the \emph{microxanox} package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
+The first two use cases are described in detail (including the code for reproducing them) in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} and the \emph{Partial Reproduction} vignettes.
+The third is taken from Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2023}) and Petchey et al.
 (\protect\hyperlink{ref-Petchey2022}{2022}).
@@
 \hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
-\subsection{Use case 1: Regime shifts during temporal environmental
-change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
+\subsection{Use case 1: Regime shifts during temporal environmental change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
 
-The study of Bush et al. (\protect\hyperlink{ref-Bush2017}{2017})
-includes simulations of the effect of oxygen diffusivity (an
-environmental driver, in the sense that it affects the ecosystem but is
-not affected by it) on the ecosystem state (oxic or anoxic). The
-\emph{microxonax} package contains functionality to make a specific
-temporal pattern of change in the oxygen diffusivity. As well as
-allowing individual simulations during which oxygen diffusivity varies,
-this functionality forms the basis of the temporal method for finding
-steady states.
+The study of Bush et al.
+(\protect\hyperlink{ref-Bush2017}{2017}) includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic).
+The \texttt{microxonox} package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
 
-An example of this functionality is given in the \emph{Partial
-Reproduction} vignette, which we briefly show here (Figure
-\ref{fig:uc1}). The example is composed of a single simulation, at the
-beginning of which the system is in the oxic state with high abundance
-of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
-eventually, around hour 30'000 the system switches to the anoxic state,
-with high abundance of both sulfur bacteria types. The oxygen
-diffusivity is then increased and at around hour 38'000 the system
-abruptly switches back to the oxic state.
+An example of this functionality is given in the \emph{Partial Reproduction} vignette, which we briefly show here (Figure \ref{fig:uc1}).
+The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
+Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types.
+The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.
 
-Also visible in the results are thick lines showing abundances of
-bacteria when abundances are low. This is due to the implementation of a
-function that at regular intervals, increases the abundance to a preset
-level. This prevents abundances reducing to very small numbers. The
-function that implements this increase abundance can also be made to add
-a certain abundance to each strain at regular intervals, thus simulating
-immigration in to the system.
+Also visible in the results are thick lines showing abundances of bacteria when abundances are low.
+This is due to the implementation of a function that at regular intervals, increases the abundance to a preset level.
+This prevents abundances reducing to very small numbers.
+The function that implements this increase abundance can also be made to add a certain abundance to each strain at regular intervals, thus simulating immigration in to the system.
 
@@
 In this simulation there is only one strain in each functional group.
-CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
 Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
@@
 \hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsection{Use case 2: The extent of hysteresis depends on community
-composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
+\subsection{Use case 2: The extent of hysteresis depends on community composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
 
-The package contains a function to extract summary features of ecosystem
-responses to environmental change, such as the amount of hysteresis
-displayed by the ecosystem. Hysteresis is a key feature of ecosystem
-responses to environmental change, because it is related to how
-difficult it can be to reserve the effects of environmental change
-(\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}). The
-amount of hysteresis is measured as the extent of the environmental
-condition (here oxygen diffusitivity) for which there were two steady
-states. I.e. it is the extent of the environmental conditions for which
-historical conditions play an important role in determining the current
-system state (a definition of hysteresis).
+The package contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
+Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change (\protect\hyperlink{ref-Scheffer2001}{Scheffer et al., 2001}).
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
+I.e.
+it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
 
-Using the package to calculate the extent of hysteresis involves setting
-ecosystem and simulation parameters, including parameters for the
-finding of steady states across an environmental gradient, running the
-steady state finding function, and analysing the results with the
-function that calculates extent of hysteresis. The code for this is
-provided in the \emph{User Guide}.
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
+The code for this is provided in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
-The results show that the amount of hysteresis depends greatly on the
-combinations of organisms present (Figure \ref{fig:uc2}). For example,
-with only the CB (cyanobacteria) present, there was no hysteresis. In
-contrast, the presence of both CB and SB (sulfate reducing bacteria) led
-to a large amount of hysteresis. (These results are also given in the
-\emph{Partial Reproduction} vignette.)
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}).
+For example, with only the CB (cyanobacteria) present, there was no hysteresis.
+In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis.
+(These results are also given in the \emph{Partial Reproduction} vignette.)
 
@@
 \hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
-\subsection{Use case 3: Effects of functional diversity on regime
-shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
+\subsection{Use case 3: Effects of functional diversity on regime shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
 
-As discussed in the Introduction section, the package was motivated by
-the question of how biodiversity influences ecosystem responses to
-environmental change. Extensive results concerning this question are
-given in a separate publication Limberger et al.
-(\protect\hyperlink{ref-Limberger2022}{2022}). Here we describe one of
-the results, which is that having biodiversity in a functional group can
-allow state changes to occur that otherwise would not have. I.e.
+As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change.
+Extensive results concerning this question are given in a separate publication Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2023}).
+Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
+I.e.
 biodiversity can qualitatively change the state of the ecosystem.
 
-Biodiversity was added to the functional groups using the
-\texttt{new\_strain\_parameter()} function to create a parameter set
-with multiple species per functional group (albeit all with identical
-features) and then to add variability among the species by calling the
-\texttt{add\_strain\_var()} function. This function takes an already
-existing parameter set and adds the specified about of variation. The
-new parameter object is then used as before.
+Biodiversity was added to the functional groups using the \texttt{new\_strain\_parameter()} function to create a parameter set with multiple species per functional group (albeit all with identical features) and then to add variability among the species by calling the \texttt{add\_strain\_var()} function.
+This function takes an already existing parameter set and adds the specified about of variation.
+The new parameter object is then used as before.
 
-Figure \ref{fig:uc3} shows a simulation with two species (strains) in
-each of the three functional groups. The ecosystem starts in the oxic
-state, though with relatively high abundance of each functional group.
-The strain of SB that is more tolerant to oxygen (SB\_1) initially
-decreases in abundance, but then increases, and the other (SB\_2) strain
-then becomes abundance and SB\_1 declines. Furthermore, the
-cyanobacteria crash in abundance, and the system switches to the anoxic
-state. In contrast, if there are two identical strains with tolerance
-half way between those in Figure \ref{fig:uc3} the ecosystem remains in
-the oxic state.
+Figure \ref{fig:uc3} shows a simulation with two species (strains) in each of the three functional groups.
+The ecosystem starts in the oxic state, though with relatively high abundance of each functional group.
+The strain of SB that is more tolerant to oxygen (SB\_1) initially decreases in abundance, but then increases, and the other (SB\_2) strain then becomes abundance and SB\_1 declines.
+Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
+In contrast, if there are two identical strains with tolerance half way between those in Figure \ref{fig:uc3} the ecosystem remains in the oxic state.
 
@@
 \hypertarget{adapting-for-other-ecosystems-and-organisms}{%
-\subsection{Adapting for other ecosystems and
-organisms}\label{adapting-for-other-ecosystems-and-organisms}}
+\subsection{Adapting for other ecosystems and organisms}\label{adapting-for-other-ecosystems-and-organisms}}
 
-We anticipate that the package can be a useful starting point for
-investigating models of other types of ecosystem and how biodiversity in
-them affects responses to environmental change. The overall framework of
-the package, the purpose of each function, and the objects used for
-storing parameters and results could be retained. E.g., all such models
-would have parameters that differ among species and need to be described
-in an object, and studies will often need to run simulations and sets of
-simulations across environmental conditions.
+We anticipate that the package can be a useful starting point for investigating models of other types of ecosystem and how biodiversity in them affects responses to environmental change.
+The overall framework of the package, the purpose of each function, and the objects used for storing parameters and results could be retained.
+E.g., all such models would have parameters that differ among species and need to be described in an object, and studies will often need to run simulations and sets of simulations across environmental conditions.
 
-Researchers wanting to model a new ecosystem do not, therefore, have to
-start from scratch. This will relieve researchers from needing to make
-software design decisions, and rather focus on appropriately
-representing their ecosystem and finding the results that interest them.
-Nevertheless, adaptation of the code in the package will require a
-person / persons that can take a conceptual model of an ecosystem and
-then represent that in terms of parameters and rate equations, and that
-is relatively proficient in R programming.
+Researchers wanting to model a new ecosystem do not, therefore, have to start from scratch.
+This will relieve researchers from needing to make software design decisions, and rather focus on appropriately representing their ecosystem and finding the results that interest them.
+Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and that is relatively proficient in R programming.
 
@@
 
-The \emph{microxanox} R package allows the simulation, visualisation,
-and analysis of a model of a microbial ecosystem while allowing
-variation in the amount of diversity containing in each of the
-functional groups of organisms present. It has been used for the
-research described in another paper that describes one of the first
-investigations of the effects of diversity on ecosystem resilience
-Limberger et al. (\protect\hyperlink{ref-Limberger2022}{2022}). In that
-paper, we show that diversity can have large and important effects of
-ecosystem responses, highlighting the need for models such as ours, with
-which one can easily manipulate the amount of biodiversity. The
-\emph{microxanox} package has also been used to reproduce the results of
-the paper that inspired the package development
-(\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).
+The \emph{microxanox} R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
+It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience Limberger et al.
+(\protect\hyperlink{ref-Limberger2022}{2023}).
+In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
+The \emph{microxanox} package has also been used to reproduce the results of the paper that inspired the package development (\protect\hyperlink{ref-Bush2017}{Bush et al., 2017}).
 
-The package greatly lowers the amount of work required in further
-investigations of the specific ecosystem modelled. There has, for
-example, been quite limited investigation of how biodiversity influences
-the short-term responses of the modelled ecosystem to environmental
-change. Likewise, the package could be used to power an investigation of
-the effects of biodiversity on the usefulness of early warning signals
-of abrupt ecosystem change
-(\protect\hyperlink{ref-Scheffer2009}{Scheffer et al., 2009}). In
-addition this package could be used as a template for the implementation
-for developing models of other types of ecosystems and organism. By
-doing so, other models can profit from the overall framework used, and
-the reproducibility aspects as well as the flexibility implemented.
+The package greatly lowers the amount of work required in further investigations of the specific ecosystem modelled.
+There has, for example, been quite limited investigation of how biodiversity influences the short-term responses of the modelled ecosystem to environmental change.
+Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change (\protect\hyperlink{ref-Scheffer2009}{Scheffer et al., 2009}).
+In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.
+By doing so, other models can profit from the overall framework used, and the reproducibility aspects as well as the flexibility implemented.
 
@@
 
-The authors declare no known conflicting or competing interests
-associated with this publication and there has been no significant
-financial support for this work that could have influenced its outcome.
+The authors declare no known conflicting or competing interests associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
 
@@
 
-This project was part of SNF Project 310030\_188431. The project was
-also supported by the University of Zurich Research Priority Programme
-in Global Change and Biodiversity.
+This project was part of SNF Project 310030\_188431. The project was also supported by the University of Zurich Research Priority Programme in Global Change and Biodiversity.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Binzer2016a}{}}%
-Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016. Interactive effects
-of warming, eutrophication and size structure: Impacts on biodiversity
-and food-web structure. Global Change Biology 22, 220--227.
+Binzer, A., Guill, C., Rall, B.C., Brose, U., 2016.
+Interactive effects of warming, eutrophication and size structure: Impacts on biodiversity and food-web structure.
+Global Change Biology 22, 220--227.
 doi:\href{https://doi.org/10.1111/gcb.13086}{10.1111/gcb.13086}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bush2017}{}}%
-Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J.,
-2017. Oxic-anoxic regime shifts mediated by feedbacks between
-biogeochemical processes and microbial community dynamics. Nature
-Communications 8, 789.
+Bush, T., Diao, M., Allen, R.J., Sinnige, R., Muyzer, G., Huisman, J., 2017.
+Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics.
+Nature Communications 8, 789.
 doi:\href{https://doi.org/10.1038/s41467-017-00912-x}{10.1038/s41467-017-00912-x}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Chaparro-Pedraza2021}{}}%
-Chaparro-Pedraza, P.C., 2021. Fast environmental change and
-eco-evolutionary feedbacks can drive regime shifts in ecosystems before
-tipping points are crossed. Proceedings of the Royal Society B:
-Biological Sciences 288, 20211192.
+Chaparro-Pedraza, P.C., 2021.
+Fast environmental change and eco-evolutionary feedbacks can drive regime shifts in ecosystems before tipping points are crossed.
+Proceedings of the Royal Society B: Biological Sciences 288, 20211192.
 doi:\href{https://doi.org/10.1098/rspb.2021.1192}{10.1098/rspb.2021.1192}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Collins2021}{}}%
-Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P.,
-Ratajczak, Z., 2021. Fire frequency, state change and hysteresis in
-tallgrass prairie. Ecology Letters 24, 636--647.
+Collins, S.L., Nippert, J.B., Blair, J.M., Briggs, J.M., Blackmore, P., Ratajczak, Z., 2021.
+Fire frequency, state change and hysteresis in tallgrass prairie.
+Ecology Letters 24, 636--647.
 doi:\href{https://doi.org/10.1111/ele.13676}{10.1111/ele.13676}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Krug2022}{}}%
-Krug, R.M., Petchey, O.L., 2022. Microxanox: {Model} to simulate three
-functional group system.
-doi:\href{https://doi.org/10.5281/ZENODO.6624125}{10.5281/ZENODO.6624125}
+Krug, R.M., Petchey, O.L., 2022. Microxanox: {Model} to simulate three functional group system. doi:\href{https://doi.org/10.5281/ZENODO.6624125}{10.5281/ZENODO.6624125}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Limberger2022}{}}%
-Limberger, R., Daugaard, U., Gupta, A., Krug, R., Lemmen, K., Moorsel,
-S. van, Suleiman, M., Zuppinger-Dingley, D., Petchey, O., 2022.
-Functional diversity can facilitate the collapse of an undesirable
-ecosystem state. Authorea Preprint.
-doi:\href{https://doi.org/10.22541/au.165513066.66785786/v1}{10.22541/au.165513066.66785786/v1}
+Limberger, R., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., Van Moorsel, S.J., Suleiman, M., Zuppinger-Dingley, D., Petchey, O.L., 2023.
+Functional diversity can facilitate the collapse of an undesirable ecosystem state.
+Ecology Letters 26, 883--895.
+doi:\href{https://doi.org/10.1111/ele.14217}{10.1111/ele.14217}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Northrop2021}{}}%
-Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J.,
-2021. Clockwise and counterclockwise hysteresis characterize state
-changes in the same aquatic ecosystem. Ecology Letters 24, 94--101.
+Northrop, A.C., Avalone, V., Ellison, A.M., Ballif, B.A., Gotelli, N.J., 2021.
+Clockwise and counterclockwise hysteresis characterize state changes in the same aquatic ecosystem.
+Ecology Letters 24, 94--101.
 doi:\href{https://doi.org/10.1111/ele.13625}{10.1111/ele.13625}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Petchey2022}{}}%
-Petchey, O.L., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., van
-Moorsel, Sofia J., Suleiman, M., Zuppinger-Dingley, Debra, Limberger,
-R., 2022. Code package for: "{Functional} diversity can facilitate the
-collapse of an undesirable ecosystem state".
+Petchey, O.L., Daugaard, U., Gupta, A., Krug, R.M., Lemmen, K.D., van Moorsel, Sofia J., Suleiman, M., Zuppinger-Dingley, Debra, Limberger, R., 2022.
+Code package for: "{Functional} diversity can facilitate the collapse of an undesirable ecosystem state".
 doi:\href{https://doi.org/10.5281/ZENODO.6334544}{10.5281/ZENODO.6334544}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
-R Core Team, 2022. R: {A} language and environment for statistical
-computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.
+R Core Team, 2022. R: {A} language and environment for statistical computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Scheffer2009}{}}%
-Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R.,
-Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
-Early-warning signals for critical transitions. Nature 461, 53--59.
+Scheffer, M., Bascompte, J., Brock, W.A., Brovkin, V., Carpenter, S.R., Dakos, V., Held, H., van Nes, E.H., Rietkerk, M., Sugihara, G., 2009.
+Early-warning signals for critical transitions.
+Nature 461, 53--59.
 doi:\href{https://doi.org/10.1038/nature08227}{10.1038/nature08227}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Scheffer2001}{}}%
-Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001.
-Catastrophic shifts in ecosystems. Nature 413, 591--596.
-doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}
+Scheffer, M., Carpenter, S., Foley, J.A., Folke, C., Walker, B., 2001. Catastrophic shifts in ecosystems. Nature 413, 591--596. doi:\href{https://doi.org/10.1038/35098000}{10.1038/35098000}
 
 \leavevmode\vadjust pre{\hypertarget{ref-Soetaert2010}{}}%
-Soetaert, K., Petzoldt, T., Setzer, R.W., 2010. Solving differential
-equations in {R}: {Package deSolve}. Journal of Statistical Software 33,
-1--25.
+Soetaert, K., Petzoldt, T., Setzer, R.W., 2010.
+Solving differential equations in {R}: {Package deSolve}.
+Journal of Statistical Software 33, 1--25.
 doi:\href{https://doi.org/10.18637/jss.v033.i09}{10.18637/jss.v033.i09}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Vandermeer2019}{}}%
-Vandermeer, J., Perfecto, I., 2019. Hysteresis and critical transitions
-in a coffee agroecosystem. Proceedings of the National Academy of
-Sciences 116, 15074--15079.
+Vandermeer, J., Perfecto, I., 2019.
+Hysteresis and critical transitions in a coffee agroecosystem.
+Proceedings of the National Academy of Sciences 116, 15074--15079.
 doi:\href{https://doi.org/10.1073/pnas.1902773116}{10.1073/pnas.1902773116}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Vasseur2005}{}}%
-Vasseur, D.A., McCann, K.S., 2005. A {Mechanistic Approach} for
-{Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}. The
-American Naturalist 166, 184--198.
+Vasseur, D.A., McCann, K.S., 2005.
+A {Mechanistic Approach} for {Modeling Temperature}-{Dependent Consumer}-{Resource Dynamics}.
+The American Naturalist 166, 184--198.
 doi:\href{https://doi.org/10.1086/431285}{10.1086/431285}

### 2023-06-05 inst/results_sym_bushplus.Rmd
plot label changed

---
title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
author: "Owen Petchey, Rainer M Krug"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### import libraries and source files

```{r library_sourve, message=FALSE, include=FALSE}

# library import

library(tidyverse)
library(microxanox) # version 0.9.1
library(ggplot2)

```

### set up parameterset

A wait time were set short to be able to simulate many simulation steps and not having long simulation durations at the same time. Short wait times mean fast environmental change.
Number of strains stays one, means no investigations in trait variation in this work.
```{r param}

wait_time <- 10 # time spent at each step
num_strains <- 1 # number of strains per group
event_interval <- 10

## oxygen diffusivity equal to sulfate diffusivity means perfect symmetry
## oxygen diffusivity time series: rep(8e-2, 4), , 10 ^ c(rep(-8, 5),rep(0,5),rep(-8, 5))

# a_S at 0.125

log10a_series <- seq(3, -5, length = 10) # symmetry axis at -1, +- 4 units in both direction

num_CB_strains <- num_strains
num_SB_strains <- num_strains
num_PB_strains <- num_strains

sp <- new_strain_parameter(
  n_CB = num_CB_strains,
  values_CB = "symmetric",
  n_PB = num_SB_strains,
  values_PB = "symmetric",
  n_SB = num_PB_strains,
  values_SB = "symmetric",
  values_other  = "symmetric",
  values_initial_state = "symmetric"

)

parameter <- new_runsim_parameter(
  dynamic_model = bushplus_dynamic_model,  
  event_definition = event_definition_2,       
  event_interval = event_interval,
  noise_sigma = 0,
  minimum_abundances = c(1, 0 , 1),       # PB stays 0
  strain_parameter = sp,
  log10a_series =  log10a_series
)

names(parameter$minimum_abundances) <- c("CB", "PB", "SB")
rm(sp)
parameter$sim_duration <- wait_time * length(parameter$log10a_series)
parameter$sim_sample_interval <- wait_time  # to avoid having negitive ODE results

```

### scenarios from result sections

Simulation duration in each simulation is different to emphasize important dynamics in plots
To investigate in alternative stable states the CB = SB case as well as the differing in initial abundance and favouring a_O scenarios were simulated as described in the report.
To investigate in bistability the bistabilty scenario has been created, as well as adding gradual environmental change is used to induce catatrophic shifts.
All scenario dynamics are saved as RDS files.

```{r expl}
## simulate scenarios

parameter$strain_parameter$a_S <- 0.1

# case CB = SB (no events)

length <- 1e2

parameter$event_interval <- 1e6

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_unstable_eq_noevent.RDS")

# case CB = SB (events)

length <- 6e2

parameter$event_definition <- event_definition_2
parameter$event_interval <- 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_unstable_eq_event.RDS")

# different initial abundance

length <- 4e2

parameter$strain_parameter$initial_state["CB_1"] <- 11
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_different_initial_abundance_CB.RDS")

length <- 4e2

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 11

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_different_initial_abundance_SB.RDS")

# favouring oxygen div

length <- 1.5e2

parameter$strain_parameter$initial_state["CB_1"] <- 1e5
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.001), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_favouring_a_O_SB.RDS")

length <- 1.5e2

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 1e5

parameter$log10a_series <- rep(log10(10), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_favouring_a_O_CB.RDS")

# bistability

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 1e5
parameter$log10a_series <- rep(log10(0.15), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_bistabilty_SB.RDS")

parameter$strain_parameter$initial_state["CB_1"] <- 1e5 
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- rep(log10(0.15), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_bistabilty_CB.RDS")

# gradual environmental change to inestigate in catastrophic shifts

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(-5, times = times), seq(-5, 3, length = length), rep(3, times = times), seq(3, -5, length = length), rep(-5, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_both_collapse.RDS")

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(-5, times = times), seq(-5, 3, length = length), rep(3, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_SB_collapse.RDS")

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(3, times = times), seq(3, -5, length = length), rep(-5, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_CB_collapse.RDS")

```

### plotting scenarios

Scenarios described above are plotted with the plot_dynamics function defined in functions.R after read in from RDS file.

```{r plot_1, include=FALSE}
## ggplot

eq_noevent_results <- readRDS(file = "RDS_files/sym_dynamics_unstable_eq_noevent.RDS")
p1.1 <- plot_dynamics_symmetric(eq_noevent_results, plot_a = FALSE)

eq_event_results <- readRDS(file = "RDS_files/sym_dynamics_unstable_eq_event.RDS")
p1.2 <- plot_dynamics_symmetric(eq_event_results, plot_a = FALSE)

init_ab_CB_results <- readRDS(file = "RDS_files/sym_dynamics_different_initial_abundance_CB.RDS")
p2.1 <- plot_dynamics_symmetric(init_ab_CB_results, plot_a = FALSE)

init_ab_SB_results <- readRDS(file = "RDS_files/sym_dynamics_different_initial_abundance_SB.RDS")
p2.2 <- plot_dynamics_symmetric(init_ab_SB_results, plot_a = FALSE)

favouring_CB_results <- readRDS(file = "RDS_files/sym_dynamics_favouring_a_O_CB.RDS")
p3.1 <- plot_dynamics_symmetric(favouring_CB_results, plot_a = FALSE)

favouring_SB_results <- readRDS(file = "RDS_files/sym_dynamics_favouring_a_O_SB.RDS")
p3.2 <- plot_dynamics_symmetric(favouring_SB_results, plot_a = FALSE)

bistability_CB_results <- readRDS(file = "RDS_files/sym_dynamics_bistabilty_CB.RDS")
p4.1 <- plot_dynamics_symmetric(bistability_CB_results, plot_a = FALSE)

bistability_SB_results <- readRDS(file = "RDS_files/sym_dynamics_bistabilty_SB.RDS")
p4.2 <- plot_dynamics_symmetric(bistability_SB_results, plot_a = FALSE)

a_O_vary_SB_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_SB_collapse.RDS")
p5.1 <- plot_dynamics_symmetric(a_O_vary_SB_results, plot_a = TRUE)

a_O_vary_CB_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_CB_collapse.RDS")
p5.2 <- plot_dynamics_symmetric(a_O_vary_CB_results, plot_a = TRUE)

a_O_vary_both_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_both_collapse.RDS") # add collapse next time run sim above
p5.3 <- plot_dynamics_symmetric(a_O_vary_both_results, plot_a = TRUE)

# chnagel labels smaller or bigger fizre

# discussion pssouble further ideas
```

### saving plots

Plots are saved with ggplot

```{r ggsave_1}
width = 30
height = 15
units = "cm"
## p1
ggsave("GG_figures/eq_noevent_figure.pdf", plot = p1.1, width = width, height = height, units = units)
ggsave("GG_figures/eq_event_figure.pdf", plot = p1.2, width = width, height = height, units = units)
## p2
ggsave("GG_figures/diff_init_ab_CB_figure.pdf", plot = p2.1, width = width, height = height, units = units)
ggsave("GG_figures/diff_init_ab_SB_figure.pdf", plot = p2.2, width = width, height = height, units = units)
## p3
ggsave("GG_figures/favouring_a_O_CB_figure.pdf", plot = p3.1, width = width, height = height, units = units)
ggsave("GG_figures/favouring_a_O_SB_figure.pdf", plot = p3.2, width = width, height = height, units = units)
## p4
ggsave("GG_figures/bistabilty_CB_figure.pdf", plot = p4.1, width = width, height = height, units = units)
ggsave("GG_figures/bistabilty_SB_figure.pdf", plot = p4.2, width = width, height = height, units = units)
## p5
ggsave("GG_figures/a_O_gradual_vary_SB_collapse_figure.pdf", plot = p5.1, width = width, height = height, units = units)
ggsave("GG_figures/a_O_gradual_vary_CB_collapse_figure.pdf", plot = p5.2, width = width, height = height, units = units)
ggsave("GG_figures/a_O_gradual_vary__collapse_figure.pdf", plot = p5.3, width = width, height = height, units = units)

rm(width, height, units)
```

Display plots in file:

```{r plot_2}
p1.1
p1.2
p2.1
p2.2
p3.1
p3.2
p4.1
p4.2
p5.1
p5.2
p5.3

# split ylab \n
```

### temporal results

Gradual envrionmental change is simulated in both direction with the temporal simulation method. Two simulation durations of 1000 respectively 10'000 hours to emphasize problem of bistability with slower environmental change. Simulations saved as RDS files.

```{r temporal_2}
length <- 1e4 # modify sim duration

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- log10a_series <- seq(-5, 3, length = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

temporal_results <- run_temporal_ssfind(parameter) ## simulate
saveRDS(temporal_results, file = "RDS_files/temporal_ss_10000h.RDS")

length <- 1e3 # modify sim duration

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- log10a_series <- seq(-5, 3, length = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time

temporal_results <- run_temporal_ssfind(parameter) ## simulate
saveRDS(temporal_results, file = "RDS_files/temporal_ss_1000h.RDS")
```

Plotting of temporal simulations with plot_temporal_ss from functions.R

```{r plot_3, include=FALSE}
temporal_results_1000h <- readRDS(file = "RDS_files/temporal_ss_1000h.RDS")
p6.1 <- plot_temporal_ss(temporal_results_1000h, zoom = FALSE) ## plot

temporal_results_10000h <- readRDS(file = "RDS_files/temporal_ss_10000h.RDS") # slower environmental change
p6.2 <- plot_temporal_ss(temporal_results_10000h, zoom = FALSE) ## plot

# H: because only a_O chnages
# discusion: vary a_O & a_S, longer simulations, optimazation of symmetry 
# save as GG
```

Displaying plots

```{r plot_4}
p6.1
p6.2

```
GGsave the plots as above if required

```{r ggsave_4}
height = 30
width = 15
units = "cm"

ggsave("GG_figures/temporal_1000h_figure.pdf", plot = p6.1, width = width, height = height, units = units)

ggsave("GG_figures/temporal_10000h_figure.pdf", plot = p6.2, width = width, height = height, units = units)

```

### 2023-06-06 inst/manuscript/manuscript_microxanox.Rmd
References adjusted

@@
 numbersections: true
-csl: elsevier-harvard.csl
 output:
-  rticles::elsevier_article:
+  bookdown::pdf_book:
+    base_format: rticles::elsevier_article
+    keep_tex: true
     includes:
@@
 bibliography: bibliography.bib
+csl: 'softwarex.csl'
 nocite: '@Bush2017'
@@
 
-A typical simulation would look as shown in Figure \ref{fig:runsim_example}.
+A typical simulation would look as shown in Figure \@ref(fig:runsimexample).
 
-```{R runsim_example, eval = TRUE, echo = FALSE, fig.cap='Typical flow of a simulation.
+```{R runsimexample, eval = TRUE, echo = FALSE, fig.cap='Typical flow of a simulation.
 Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.', fig.align="center", out.width = 350 }
@@
 This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
-An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.
+An example plot resulting from this function is shown in Figure \@ref(fig:plot-dynamics).
 
@@
 
-An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}).
+An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \@ref(fig:uc1)).
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
@@
 
-The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}).
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \@ref(fig:uc2)).
 For example, with only the CB (cyanobacteria) present, there was no hysteresis.
@@
 
-Figure \ref{fig:uc3} shows a simulation with two species (strains) in each of the three functional groups.
+Figure \@ref(fig:uc3) shows a simulation with two species (strains) in each of the three functional groups.
 The ecosystem starts in the oxic state, though with relatively high abundance of each functional group.
@@
 Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
-In contrast, if there are two identical strains with tolerance half way between those in Figure \ref{fig:uc3} the ecosystem remains in the oxic state.
+In contrast, if there are two identical strains with tolerance half way between those in Figure \@ref(fig:uc3) the ecosystem remains in the oxic state.
 
@@
 
-# TODO in RMD file
-# TODO in LaTeX file
-- [ ] fix citation style
-- [ ] fix references to figures
-
 # References

### 2023-06-06 inst/manuscript/manuscript_microxanox.tex
References adjusted

--- 2023-05-15 inst/manuscript/manuscript_microxanox.tex
@@
- \documentclass[preprint,12pt, a4paper]{elsarticle}
-
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
@@
 \usepackage{lmodern}
-% \usepackage{amssymb,amsmath}
+\usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
   \usepackage[unicode=true]{hyperref}
-
+\fi
 \hypersetup{breaklinks=true,
@@
 \makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 
@@
 
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+\usepackage{lineno}
+
 \usepackage{float}
@@
 
+%\usepackage{lineno}
+%\linenumbers
+%\usepackage{setspace}
+%\doublespacing
+%\usepackage{fancyhdr}
+%\fancyhead{}
+%\fancyfoot{}
+%\fancyhead[CO,CE]{Microxanox R package}
+%\pagestyle{fancy}
+%
+
 \begin{document}
@@
 
-  \title{Microxanox: an R package for simulating an aquatic
-\(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
+  \title{Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.}
     \author[University of Zürich]{Rainer M Krug%
@@
    \ead{Owen.Petchey@ieu.uzh.ch} 
-      \affiliation[University of Zürich]{Department of Evolutionary
-Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
+      \affiliation[University of Zürich]{Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, 8057 Zurich}
     \cortext[cor1]{Corresponding author}
@@
   \begin{abstract}
-  \(Microxanox\) is an R package to simulate a three functional group
-  ecosystem (cyanobacteria, phototrophic sulfur bacteria, and
-  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
-  oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
-  differential equations. Simulations can be run individually or over a
-  parameter range, to find stable states. The model can be implemented
-  with different numbers of species per functional group. The package is
-  constructed in such a way that the results contain the input parameter
-  used, so that a saved results can be loaded again and the simulation
-  be repeated. Furthermore, the package framework and code should serve
-  as a useful starting point for making simulation models of other types
-  of ecosystems.
+  \(Microxanox\) is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
+Simulations can be run individually or over a parameter range, to find steady states.
+The model can be implemented with different numbers of species per functional group.
+The package is constructed in such a way that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
+Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
   \end{abstract}
     \begin{keyword}
-    reproducibility \sep regime shift \sep stable state \sep 
+    reproducibility \sep regime shift \sep steady state \sep 
     ordinary differential equations
@@
 & & multidplyr \\
-C8 & If available Link to developer documentation/manual &
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide} \\
-C9 & Support email for questions & Rainer.Krug@uzh.ch;
-Owen.Petchey@ieu.uzh \\
+C8 & If available Link to developer documentation/manual & \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} \\
+C9 & Support email for questions & Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh \\
 \end{longtable}
@@
 \hypertarget{motivation-and-significance}{%
-\section{Motivation and
-significance}\label{motivation-and-significance}}
+\section{Motivation and significance}\label{motivation-and-significance}}
 
-Mathematical models play a key role in the development of ecosystem
-models and the understanding about how ecosystems work and respond to
-environmental changes
-\citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}. They are also
-important for developing hypotheses to test in empirical studies. One
-area of ecology in which models have been influential is the
-understanding of their response to a gradual change in an environmental
-driver \citep{Scheffer2001}. An environmental driver is an environmental
-condition that affects an ecosystem. Still, it is assumed not to be
-affected by the ecosystem, such as the rate of nutrient input into a
-lake.
+Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}.
+They are also important for developing hypotheses to test in empirical studies.
+One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver \citep{Scheffer2001}.
+An environmental driver is an environmental condition that affects an ecosystem.
+Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.
 
-It is conceivable that an ecosystem state, such as the total biomass of
-a particular type of bacteria, may remain unchanged when an
-environmental driver changes. It is also possible that the ecosystem
-state changes gradually. It is also possible that the ecosystem state
-changes abruptly to a new state that is difficult to recover from
-\citep{Scheffer2001}. This possibility for abrupt, perhaps catastrophic
-changes that are difficult to reverse causes considerable concern
-\citep{Collins2021, Northrop2021, Vandermeer2019}.
+It is conceivable that an ecosystem state, such as the total biomass of a particular type of bacteria, may remain unchanged when an environmental driver changes.
+It is also possible that the ecosystem state changes gradually.
+It is also possible that the ecosystem state changes abruptly to a new state that is difficult to recover from \citep{Scheffer2001}.
+This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern \citep{Collins2021, Northrop2021, Vandermeer2019}.
 
-An example where a gradual change of an environmental variable causes an
-abrupt change in the system is the switch from an aerobic (oxygen is
-available for metabolism) to an anaerobic (oxygen generally unavailable)
-state in a microbial ecosystem. This system has been investigated by
-\citet{Bush2017} in a simulation study of a mathematical model. Three
-types of microbes occur in the model: cyanobacteria (CB) dominating the
-oxic state, and two types of sulfur bacteria that dominate the anoxic
-state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria
-(PB)). The model shows that gradual change in the rate at which oxygen
-could diffuse into the ecosystem (termed oxygen diffusitivity) could
-cause catastrophic changes in the ecosystem state that would be
-difficult to reverse.
+An example where a gradual change of an environmental variable causes an abrupt change in the system is the switch from an aerobic (oxygen is available for metabolism) to an anaerobic (oxygen generally unavailable) state in a microbial ecosystem.
+This system has been investigated by \citet{Bush2017} in a simulation study of a mathematical model.
+Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
+The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
 One feature of the study by \citet{Bush2017} was limited biodiversity.
-Specifically, there was no biodiversity within each of the three types
-of bacteria. This leaves the question of if and how bacteria's
-biodiversity within these types (i.e., functional groups) affects the
-ecosystem response to environmental change. This limitation is not
-specific to the study of \citet{Bush2017}. There are few if any studies
-of the effects of biodiversity on abrupt transitions between ecosystem
-states.
+Specifically, there was no biodiversity within each of the three types of bacteria.
+This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem response to environmental change.
+This limitation is not specific to the study of \citet{Bush2017}.
+There are few if any studies of the effects of biodiversity on abrupt transitions between ecosystem states.
 
-We decided to fill this research gap by making a simulation study of how
-within functional group biodiversity affects ecosystem responses to
-environmental change \citet{Limberger2022}, and to base our work on the
-work and model of \citet{Bush2017}. With this goal in mind, we developed
-the microxanox package \citep{Krug2022}. The first stage of development
-was to write code from scratch (as there was no available code to start
-from) and to confirm that this new implementation would reproduce the
-previously published results. The resulting reproduction is available as
-one of the package vignettes:
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette
-Partial reproduction of Bush et al}.
+We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change \citet{Limberger2022}, and to base our work on the work and model of \citet{Bush2017}.
+With this goal in mind, we developed the \texttt{microxanox} package \citep{Krug2022}.
+The first stage of development was to write code from scratch (as there was no available code to start from) and to confirm that this new implementation would reproduce the previously published results.
+The resulting reproduction is available as one of the package vignettes: \href{https://uzh-peg.r-universe.dev/articles/microxanox/partial-reproduction-Bushetal2017.html}{vignette Partial reproduction of Bush et al}.
 
-The second stage was to add functionality that would be necessary to
-answer our research question. Most importantly, we made it possible to
-have multiple species of bacteria within each of the three functional
-groups, for the multiple species to differ in their characteristics, and
-to vary the number of species and amount of variability among them. We
-also added functionality that allowed: temporally varying environmental
-conditions, the addition of random noise to state variables, and
-immigration. In addition to the model itself, the package includes some
-functions to analyse the results and visualize these to provide a
-starting point for customized visualizations based on own requirements.
-The basic and additional functionality is described in the package
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+The second stage was to add functionality that would be necessary to answer our research question.
+Most importantly, we made it possible to have multiple species of bacteria within each of the three functional groups, for the multiple species to differ in their characteristics, and to vary the number of species and amount of variability among them.
+We also added functionality that allowed: temporally varying environmental conditions, the addition of random noise to state variables, and immigration.
+In addition to the model itself, the package includes some functions to analyse the results and visualize these to provide a starting point for customized visualizations based on own requirements.
+The basic and additional functionality is described in the package \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
@@
 
-The \emph{microxanox} package is for simulating a three functional group
-system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur
-bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical
-substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced
-sulfur, \emph{SO}: oxidized sulfur). It includes feedback between
-organisms and biogeochemical processes and is based on \citet{Bush2017}
-(See \citet{Bush2017} for a detailed discussion of the model). At the
-core of the simulations is a set of ordinary differential equations
-(specified in the function \texttt{bushplus\_dynamic\_model()}, though
-this function need not be directly called). There are functions for
-running individual simulations and a set of simulations across, for
-example, a range of environmental conditions.
+The \emph{microxanox} package is for simulating a three functional group system (\emph{CB}: cyanobacteria, \emph{PB}: phototrophic sulfur bacteria, and \emph{SB}: sulfate-reducing bacteria) with four chemical substrates (\emph{P}: phosphorus, \emph{O}: oxygen, \emph{SR}: reduced sulfur, \emph{SO}: oxidized sulfur).
+It includes feedback between organisms and biogeochemical processes and is based on \citet{Bush2017} (See \citet{Bush2017} for a detailed discussion of the model).
+At the core of the simulations is a set of ordinary differential equations (specified in the function \texttt{bushplus\_dynamic\_model()}, though this function need not be directly called).
+There are functions for running individual simulations and a set of simulations across, for example, a range of environmental conditions.
 
-To make the simulation run with multiple species per functional group,
-we expressed different species characteristics in the elements of
-vectors and matrices. We also coded the ordinary differential equations
-to include the vectors and matrices and used matrix mathematics. In this
-way, we made it possible to run simulations with different numbers of
-species without having to change the underlying code.
+To make the simulation run with multiple species per functional group, we expressed different species characteristics in the elements of vectors and matrices.
+We also coded the ordinary differential equations to include the vectors and matrices and used matrix mathematics.
+In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
 
-The package functions and code have a modular structure so that new
-functionality can be easily added. E.g., temporally defined events of
-any type could be specified. Further, all parameter values required to
-run a simulation are stored in one object. Lastly, the general structure
-of the code should make it straightforward to adapt the model to other
-similar systems (described in more detail in the Impact section).
+The package functions and code have a modular structure so that new functionality can be easily added.
+E.g., temporally defined events of any type could be specified.
+Further, all parameter values required to run a simulation are stored in one object.
+Lastly, the general structure of the code should make it straightforward to adapt the model to other similar systems (described in more detail in the Impact section).
 
@@
 
-The framework used when writing this package aimed to maximise
-simplicity for the user and to make it straightforward to reproduce
-results (see the supplement \citep{Petchey2022} to \citet{Limberger2022}
-for an example of how this is used). As such, all the parameters needed
-to run a simulation or find a stable state (i.e.~the final state of the
-ecosystem) are contained in a single object (which can easily be created
-using included functions). This parameter object is given to a function
-that runs the simulations and returns the results. The returned results
-object is identical to the parameter object but with an additional slot
-named \texttt{results}, which contains the simulation results. Thus the
-returned results object contains the simulation conditions (parameters)
-as well as the results, and can be used to run the simulation again.
-This promotes reproducibility and makes incremental changes of
-individual parameters with a consecutive re-running of the simulations
-straightforward.
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2022} for an example of how this is used).
+As such, all the parameters needed to run a simulation or find a steady state (i.e.~the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
+This parameter object is given to a function that runs the simulations and returns the results.
+The returned results object is identical to the parameter object but with an additional slot named \texttt{results}, which contains the simulation results.
+Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again.
+This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.
 
-In the following sections we describe how to use the package to run one
-simulation and to find steady states across an environmental gradient.
+In the following sections we describe how to use the package to run one simulation and to find steady states across an environmental gradient.
 
@@
 
-A typical simulation would look as shown in Figure
-\ref{fig:runsim_example}.
+A typical simulation would look as shown in Figure \ref{fig:runsimexample}.
 
@@
 \caption{Typical flow of a simulation.
-Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsim_example}
+Dark Grey boxes: commands necessary for simulation; Light Grey:Saving of parameter and results; Lightest Grey: Different non specified commands.}\label{fig:runsimexample}
 \end{figure}
 
-A simulation is run using the \texttt{run\_simulation()} function. In
-this function, the ODEs are solved using the function \texttt{ode()} in
-the package \emph{deSolve} package \citep{Soetaert2010}. The
-\texttt{run\_simulation()} function needs only one argument - an object
-as created by the function \texttt{new\_runsim\_parameter()}. The
-parameter object returned by \texttt{new\_runsim\_parameter()} contains,
-among other things, the \texttt{strain\_parameter} object, which can be
-created by the function \texttt{new\_strain\_parameter()}. For a
-detailed description of the parameter objects, their meaning and how
-they are created and have values set and changed, please see the
-\emph{User Guide} which accompanies the package or is available at
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+A simulation is run using the \texttt{run\_simulation()} function.
+In this function, the ODEs are solved using the function \texttt{ode()} in the package \emph{deSolve} package \citep{Soetaert2010}.
+The \texttt{run\_simulation()} function needs only one argument - an object as created by the function \texttt{new\_runsim\_parameter()}.
+The parameter object returned by \texttt{new\_runsim\_parameter()} contains, among other things, the \texttt{strain\_parameter} object, which can be created by the function \texttt{new\_strain\_parameter()}.
+For a detailed description of the parameter objects, their meaning and how they are created and have values set and changed, please see the \emph{User Guide} which accompanies the package or is available at \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
-After the parameter object has been defined, it can be used in the
-\texttt{run\_simulation()} function. The function returns an object
-identical to the parameter object, except of an additional slot
-containing the results. This design produces a fully reproducible object
-as it can be used as a parameter object to be fed back into the
-\texttt{run\_simulation()} function to rerun the simulation again from
-the parameter used to generate the results before.
+After the parameter object has been defined, it can be used in the \texttt{run\_simulation()} function.
+The function returns an object identical to the parameter object, except of an additional slot containing the results.
+This design produces a fully reproducible object as it can be used as a parameter object to be fed back into the \texttt{run\_simulation()} function to rerun the simulation again from the parameter used to generate the results before.
 
-\hypertarget{finding-stable-states}{%
-\subsection{Finding Stable States}\label{finding-stable-states}}
+\hypertarget{finding-steady-states}{%
+\subsection{Finding Steady States}\label{finding-steady-states}}
 
-The general approach to finding the numerical value of a system's stable
-state (if it exists) with a specific parameter set is to run the
-simulation for a long time and record the final state. When one does
-this across a range of environmental conditions, one discovers how the
-steady state of the system responds to the environmental conditions. The
-package contains functionality for finding steady states that correspond
-to values of one environmental driver, namely the value of oxygen
-diffusivity.
+The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
+When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions.
+The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-Two methods for finding steady states are implemented. The first runs a
-separate simulation for each combination of starting conditions and
-oxygen diffusivity (we term this the \emph{Replication method}). This is
-the method used in the \citet{Bush2017} study. The second runs two
-simulations, one with step-wise and slowly temporally \emph{increasing}
-oxygen diffusivity, and the other with step-wise and slowly
-\emph{decreasing} oxygen diffusivity. During this temporal environmental
-change, the state of the system is recorded just before change to a new
-oxygen diffusivity (we term this the \emph{Temporal method}). We
-implemented two methods since there is no definitive best method, and in
-order to check if results were sensitive to choice of method.
+Two numerical strategies for finding steady states and their sensitivity to parameters are implemented.
+Two strategies are implemented in order to allow comparison of their results.
+The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the \emph{Replication method}).
+This is the method used in the \citet{Bush2017} study and was used to obtain the results in figures 3 and 4 of that article.
 
-The replication method is implemented in the function
-\texttt{run\_replication\_ssfind()} which takes a parameter object as
-returned by the function \texttt{new\_replication\_ssfind\_parameter()}
-and the number of cores for multithreading the simulation. As the
-multithreading uses the package function \texttt{mclapply()} from the
-package \texttt{parallel} \citep{RCoreTeam2022}, the multithreading only
-works on Linux and Mac. It is planned to move to \texttt{parLapply()}
-\citep{RCoreTeam2022} in a future release.
+The second method runs two simulations, one with step-wise and slowly temporally \emph{increasing} oxygen diffusivity, and the other with step-wise and slowly \emph{decreasing} oxygen diffusivity.
+(Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the \emph{Temporal method}).
 
-The temporal method is implemented in the function
-\texttt{run\_temporal\_ssfind()}, which takes a parameter object as
-created by the function \texttt{new\_temporal\_ssfind\_parameter()}. It
-is planned for a later release, to run these two simulations in
-parallel.
+The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.
 
-For a more detailed walk-through of these two approaches and explanation
-please see the
-\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User
-Guide}.
+The replication method is implemented in the function \texttt{run\_replication\_ssfind()} which takes a parameter object as returned by the function \texttt{new\_replication\_ssfind\_parameter()} and the number of cores for multithreading the simulation.
+As the multithreading uses the package function \texttt{mclapply()} from the package \texttt{parallel} \citep{RCoreTeam2022}, the multithreading only works on Linux and Mac.
+It is planned to move to \texttt{parLapply()} \citep{RCoreTeam2022} in a future release.
+
+The temporal method is implemented in the function \texttt{run\_temporal\_ssfind()}, which takes a parameter object as created by the function \texttt{new\_temporal\_ssfind\_parameter()}.
+It is planned for a later release, to run these two simulations in parallel.
+
+For a more detailed walk-through of these two approaches and explanation please see the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
 \hypertarget{analysing-and-visualising-results}{%
-\subsection{Analysing and visualising
-results}\label{analysing-and-visualising-results}}
+\subsection{Analysing and visualising results}\label{analysing-and-visualising-results}}
 
-From the results returned, summary measures about how the ecosystem
-stable states respond to environmental change can be extracted. The
-function \texttt{get\_stability\_measures()} returns quantities such as
-the amount of environmental change required for the system to abruptly
-change to a different state.
+From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted.
+The function \texttt{get\_stability\_measures()} returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
 
-The function \texttt{plot\_dynamics()} plots a single simulation run, as
-returned from the \texttt{run\_simulation()} function. This function is
-only provided as a convenience function to provide a way to easily see
-the results of a simulation run. An example plot resulting from this
-function is shown in Figure \ref{fig:plot-dynamics}.
+The function \texttt{plot\_dynamics()} plots a single simulation run, as returned from the \texttt{run\_simulation()} function.
+This function is only provided as a convenience function to provide a way to easily see the results of a simulation run.
+An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.
 
@@
 
-The open source implementation and extension of the model used in
-\citet{Bush2017} provides the means of reproducing the results published
-while at the same time provides the means of doing unique, innovative,
-and important investigations of how ecosystems respond to environmental
-change, and how biodiversity may modulate this response.
+The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
 
-The design of the package code and functionality is with reproducibility
-in mind: the combination of all parameters being in a single parameter
-object as well as the return of the simulation as a result object which
-inherits from the parameter object provides a relatively easy to use
-framework to implement reproducible experiments.
+The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
 
-Here we evidence the impact of the \emph{microxanox} package by
-describing three use cases and then by describing how the package can be
-a starting point for models of other ecosystems. The first two use cases
-are described in detail (including the code for reproducing them) in the
-\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
-third is taken from \citet{Limberger2022} and \citet{Petchey2022}.
+Here we evidence the impact of the \emph{microxanox} package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
+The first two use cases are described in detail (including the code for reproducing them) in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} and the \emph{Partial Reproduction} vignettes.
+The third is taken from \citet{Limberger2022} and \citet{Petchey2022}.
 
 \hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
-\subsection{Use case 1: Regime shifts during temporal environmental
-change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
+\subsection{Use case 1: Regime shifts during temporal environmental change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
 
-The study of \citet{Bush2017} includes simulations of the effect of
-oxygen diffusivity (an environmental driver, in the sense that it
-affects the ecosystem but is not affected by it) on the ecosystem state
-(oxic or anoxic). The \emph{microxonax} package contains functionality
-to make a specific temporal pattern of change in the oxygen diffusivity.
-As well as allowing individual simulations during which oxygen
-diffusivity varies, this functionality forms the basis of the temporal
-method for finding stable states.
+The study of \citet{Bush2017} includes simulations of the effect of oxygen diffusivity (an environmental driver, in the sense that it affects the ecosystem but is not affected by it) on the ecosystem state (oxic or anoxic).
+The \texttt{microxonox} package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
 
-An example of this functionality is given in the \emph{Partial
-Reproduction} vignette, which we briefly show here (Figure
-\ref{fig:uc1}). The example is composed of a single simulation, at the
-beginning of which the system is in the oxic state with high abundance
-of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
-eventually, around hour 30'000 the system switches to the anoxic state,
-with high abundance of both sulfur bacteria types. The oxygen
-diffusivity is then increased and at around hour 38'000 the system
-abruptly switches back to the oxic state.
+An example of this functionality is given in the \emph{Partial Reproduction} vignette, which we briefly show here (Figure \ref{fig:uc1}).
+The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
+Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types.
+The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.
 
-Also visible in the results are thick lines showing abundances of
-bacteria when abundances are low. This is due to the implementation of a
-function that at regular intervals, increases the abundance to a preset
-level. This prevents abundances reducing to very small numbers. The
-function that implements this increase abundance can also be made to add
-a certain abundance to each strain at regular intervals, thus simulating
-immigration in to the system.
+Also visible in the results are thick lines showing abundances of bacteria when abundances are low.
+This is due to the implementation of a function that at regular intervals, increases the abundance to a preset level.
+This prevents abundances reducing to very small numbers.
+The function that implements this increase abundance can also be made to add a certain abundance to each strain at regular intervals, thus simulating immigration in to the system.
 
@@
 \caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
-Plot of the stable states of the simulation runs under different oxygen diffusivity.
+Plot of the steady states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
-CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
+CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1.
 Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
@@
 \hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsection{Use case 2: The extent of hysteresis depends on community
-composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
+\subsection{Use case 2: The extent of hysteresis depends on community composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
 
-The package contains a function to extract summary features of ecosystem
-responses to environmental change, such as the amount of hysteresis
-displayed by the ecosystem. Hysteresis is a key feature of ecosystem
-responses to environmental change, because it is related to how
-difficult it can be to reserve the effects of environmental change
-\citep{Scheffer2001}. The amount of hysteresis is measured as the extent
-of the environmental condition (here oxygen diffusitivity) for which
-there were two stable states. I.e. it is the extent of the environmental
-conditions for which historical conditions play an important role in
-determining the current system state (a definition of hysteresis).
+The package contains a function to extract summary features of ecosystem responses to environmental change, such as the amount of hysteresis displayed by the ecosystem.
+Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change \citep{Scheffer2001}.
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
+I.e.
+it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
 
-Using the package to calculate the extent of hysteresis involves setting
-ecosystem and simulation parameters, including parameters for the
-finding of stable states across an environmental gradient, running the
-stable state finding function, and analysing the results with the
-function that calculates extent of hysteresis. The code for this is
-provided in the \emph{User Guide}.
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
+The code for this is provided in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
 
-The results show that the amount of hysteresis depends greatly on the
-combinations of organisms present (Figure \ref{fig:uc2}). For example,
-with only the CB (cyanobacteria) present, there was no hysteresis. In
-contrast, the presence of both CB and SB (sulfate reducing bacteria) led
-to a large amount of hysteresis. (These results are also given in the
-\emph{Partial Reproduction} vignette.)
+The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}).
+For example, with only the CB (cyanobacteria) present, there was no hysteresis.
+In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis.
+(These results are also given in the \emph{Partial Reproduction} vignette.)
 
@@
 \hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
-\subsection{Use case 3: Effects of functional diversity on regime
-shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
+\subsection{Use case 3: Effects of functional diversity on regime shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
 
-As discussed in the Introduction section, the package was motivated by
-the question of how biodiversity influences ecosystem responses to
-environmental change. Extensive results concerning this question are
-given in a separate publication \citet{Limberger2022}. Here we describe
-one of the results, which is that having biodiversity in a functional
-group can allow state changes to occur that otherwise would not have.
-I.e. biodiversity can qualitatively change the state of the ecosystem.
+As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change.
+Extensive results concerning this question are given in a separate publication \citet{Limberger2022}.
+Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
+I.e.
+biodiversity can qualitatively change the state of the ecosystem.
 
-Biodiversity was added to the functional groups using the
-\texttt{new\_strain\_parameter()} function to create a parameter set
-with multiple species per functional group (albeit all with identical
-features) and then to add variability among the species by calling the
-\texttt{add\_strain\_var()} function. This function takes an already
-existing parameter set and adds the specified about of variation. The
-new parameter object is then used as before.
+Biodiversity was added to the functional groups using the \texttt{new\_strain\_parameter()} function to create a parameter set with multiple species per functional group (albeit all with identical features) and then to add variability among the species by calling the \texttt{add\_strain\_var()} function.
+This function takes an already existing parameter set and adds the specified about of variation.
+The new parameter object is then used as before.
 
-Figure \ref{fig:uc3} shows a simulation with two species (strains) in
-each of the three functional groups. The ecosystem starts in the oxic
-state, though with relatively high abundance of each functional group.
-The strain of SB that is more tolerant to oxygen (SB\_1) initially
-decreases in abundance, but then increases, and the other (SB\_2) strain
-then becomes abundance and SB\_1 declines. Furthermore, the
-cyanobacteria crash in abundance, and the system switches to the anoxic
-state. In contrast, if there are two identical strains with tolerance
-half way between those in Figure \ref{fig:uc3} the ecosystem remains in
-the oxic state.
+Figure \ref{fig:uc3} shows a simulation with two species (strains) in each of the three functional groups.
+The ecosystem starts in the oxic state, though with relatively high abundance of each functional group.
+The strain of SB that is more tolerant to oxygen (SB\_1) initially decreases in abundance, but then increases, and the other (SB\_2) strain then becomes abundance and SB\_1 declines.
+Furthermore, the cyanobacteria crash in abundance, and the system switches to the anoxic state.
+In contrast, if there are two identical strains with tolerance half way between those in Figure \ref{fig:uc3} the ecosystem remains in the oxic state.
 
@@
 
-The \emph{microxanox} R package allows the simulation, visualisation,
-and analysis of a model of a microbial ecosystem while allowing
-variation in the amount of diversity containing in each of the
-functional groups of organisms present. It has been used for the
-research described in another paper that describes one of the first
-investigations of the effects of diversity on ecosystem resilience
-\citet{Limberger2022}. In that paper, we show that diversity can have
-large and important effects of ecosystem responses, highlighting the
-need for models such as ours, with which one can easily manipulate the
-amount of biodiversity. The \emph{microxanox} package has also been used
-to reproduce the results of the paper that inspired the package
-development \citep{Bush2017}.
+The \emph{microxanox} R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
+It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience \citet{Limberger2022}.
+In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
+The \emph{microxanox} package has also been used to reproduce the results of the paper that inspired the package development \citep{Bush2017}.
 
-The package greatly lowers the amount of work required in further
-investigations of the specific ecosystem modelled. There has, for
-example, been quite limited investigation of how biodiversity influences
-the short-term responses of the modelled ecosystem to environmental
-change. Likewise, the package could be used to power an investigation of
-the effects of biodiversity on the usefulness of early warning signals
-of abrupt ecosystem change \citep{Scheffer2009}. In addition this
-package could be used as a template for the implementation for
-developing models of other types of ecosystems and organism. By doing
-so, other models can profit from the overall framework used, and the
-reproducibility aspects as well as the flexibility implemented.
+The package greatly lowers the amount of work required in further investigations of the specific ecosystem modelled.
+There has, for example, been quite limited investigation of how biodiversity influences the short-term responses of the modelled ecosystem to environmental change.
+Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change \citep{Scheffer2009}.
+In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.
+By doing so, other models can profit from the overall framework used, and the reproducibility aspects as well as the flexibility implemented.
 
@@
 
-This project was part of SNF Project 310030\_188431. The project was
-also supported by the University of Zurich Research Priority Programme
-in Global Change and Biodiversity.
+This project was part of SNF Project 310030\_188431. The project was also supported by the University of Zurich Research Priority Programme in Global Change and Biodiversity.
 
 \renewcommand\refname{References}
-\bibliography{bibliography}
-\bibliographystyle{elsarticle-num-names.bst}
+\bibliography{bibliography.bib}
 

### 2023-06-09 inst/manuscript/Response.md
Further work on the revision

@@
 
-{>>Owen: Rainer - please correct and then make a response here.<<} {>>Rainer: Will do this at the end<<} *
+{\>\>Owen: Rainer - please correct and then make a response here.\<\<} {\>\>Rainer: Will do this at the end\<\<} \*
 
 > The authors in section 3.3 claim several sentences that could confuse readers.
-2.1 `one discovers how the steady state of the system responds to the environmental conditions`. There are several concerns that the authors should clarify:
+2.1 `one discovers how the steady state of the system responds to the environmental conditions`.
+There are several concerns that the authors should clarify:
 >
@@
 
-**Response:** We have added the text `The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.`
+**Response:** Thank you for raising this important issue, which we had overlooked.
+A number of the reviewer comments below are closely related, hence we sometimes refer to this response and do so using the label "Response R1".
+In order to clarify about the existence and uniqueness of steady states we have added the text `When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the *final state* (found by the simulation) is not guaranteed to be a *steady state*.
+The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In the results presented here, and in the paper @Limberger2022, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
+Furthermore, the package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique.`
 
-**Response:** Since we do not provide methods for formal analysis of stability, we always use "steady state" and have removed all methods of "stable state".
+**Response:** In addition, we notice that we sometimes used the term "stable state" and sometimes "steady state".
+We now use only the term "steady state".
+Furthermore, we now use the term "final state" to refer to the state of the system at the end of a simulation.
 
->`b)`The steady state of the system responds`.
+> `b)`The steady state of the system responds\`.
 It should be noted that the steady state does not respond to the environment.
@@
 
-**Response:** we do not understand why the reviewer states that the software does not find a steady state. We are very sorry that we are missing some key point of understanding here.
+**Response:** We envisage and hope that Response R1 sufficiently clarifies steady state finding.
 
@@
 
-**Response:** we are of the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not completely rewritten.
+**Response:** We have the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not more substatially rewritten.
 
@@
 
-{>>Owen: Rainer - please would you see to this?<<}
+The paper has been proofread and edited.
+
+{\>\>Owen: Rainer - please would you see to this?\<\<}
 
@@
 
-{>>Owen: Rainer - please would you see to this?<<}
+{\>\>Owen: Rainer - please would you see to this?\<\<}
 
----
+------------------------------------------------------------------------
 
----
+------------------------------------------------------------------------
 

### 2023-06-09 inst/manuscript/manuscript_microxanox.Rmd
Further work on the revision

@@
   - steady state
+  - final state
   - ordinary differential equations
@@
   $Microxanox$ is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
-Simulations can be run individually or over a parameter range, to find steady states.
+Simulations can be run individually or over a parameter range.
 The model can be implemented with different numbers of species per functional group.
@@
 <!-- # \textcolor{red}{TODO} -->
+
 <!-- - [x] ~~**\textcolor{green}{update links}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{update cross references}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{add abstract}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{check keywords}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{add more references}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{expand on Impact section}**~~ -->
+
 <!-- - [x] ~~**\textcolor{green}{expand on Conclusion section}**~~ -->
+
 \pagebreak
@@
 # Required Metadata
+
 ## Current code version
+
 Ancillary data table required for subversion of the codebase.
@@
 | **Nr.** | **Code metadata description**                                 | **Please fill in this column**                                                                                         |
-| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
+|:------------:|:------------------------|:--------------------------------|
 |   C1    | Current code version                                          | v0.9.1                                                                                                                 |
-|   C2    | Permanent links to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
+|   C2    | Permanent links to code/repository used for this code version | <https://github.com/UZH-PEG/microxanox>                                                                                |
 |         |                                                               | [10.5281/zenodo.7148667](https://zenodo.org/record/7148667)                                                            |
@@
 |   C6    | Software code languages, tools, and services used             | [R](https://cran.r-project.org/index.html)                                                                             |
-|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
+|   C7    | Compilation requirements, operating environments              | [R (\>= 4.1.0)](https://cran.r-project.org/index.html)                                                                 |
 |         |                                                               | magrittr                                                                                                               |
@@
 |   C8    | If available Link to developer documentation/manual           | [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html)                                       |
-|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |
+|   C9    | Support email for questions                                   | [Rainer.Krug\@uzh.ch](mailto:Rainer.Krug@uzh.ch){.email}; [Owen.Petchey\@ieu.uzh](mailto:Owen.Petchey@ieu.uzh){.email} |
 
@@
 
-Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
+Mathematical models play a key role in the development of ecosystem models and in increasing the understanding of how and why ecosystems change when the environment changes [@Binzer2016a; @Chaparro-Pedraza2021; @Vasseur2005].
 They are also important for developing hypotheses to test in empirical studies.
-One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].
-An environmental driver is an environmental condition that affects an ecosystem.
-Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.
+One area of ecology in which models have been influential is the understanding of ecosystem responses to gradual change in an environmental driver [@Scheffer2001].
+An environmental driver is an environmental condition that affects an ecosystem, but which is not itself affected by the ecosystem.
+An example driver would be the amount of nutrients entering a lake from run-off from surrounding agricultural land.
 
@@
 It is also possible that the ecosystem state changes abruptly to a new state that is difficult to recover from [@Scheffer2001].
-This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern [@Collins2021; @Northrop2021; @Vandermeer2019].
+This possibility for abrupt, perhaps catastrophic changes, which are difficult to reverse, causes considerable concern [@Collins2021; @Northrop2021; @Vandermeer2019].
 
@@
 Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
-The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+The model shows that gradual change in the rate at which oxygen can diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
 One feature of the study by @Bush2017 was limited biodiversity.
-Specifically, there was no biodiversity within each of the three types of bacteria.
-This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem response to environmental change.
+Specifically, there was no biodiversity within each of the three types (i.e., functional groups) of bacteria.
+This leaves the question of if and how biodiversity within these functional groups might affect the ecosystem response to environmental change.
 This limitation is not specific to the study of @Bush2017.
@@
 # Software description
+
 <!-- > {>>Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.<<} -->
@@
 
-To make the simulation run with multiple species per functional group, we expressed different species characteristics in the elements of vectors and matrices.
+To make the simulation run with multiple species per functional group, we expressed different species' characteristics in the elements of vectors and matrices.
 We also coded the ordinary differential equations to include the vectors and matrices and used matrix mathematics.
@@
 ## Software architecture
+
 <!-- > {>>Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.<<} -->
@@
 The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
-As such, all the parameters needed to run a simulation or find a steady state (i.e.
-the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
+As such, all the parameters needed to run a simulation or a set of simulations are contained in a single object (which can easily be created using included functions).
 This parameter object is given to a function that runs the simulations and returns the results.
@@
 
-In the following sections we describe how to use the package to run one simulation and to find steady states across an environmental gradient.
+In the following sections we describe how to use the package to run one simulation and to find final states across an environmental gradient.
 
@@
 
-## Finding Steady States
+## Examining ecosystem responses to environmental change
 
-The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
-When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions.
-The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
+One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's steady state (if it exists).
+The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
+When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
+The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-Two numerical strategies for finding steady states and their sensitivity to parameters are implemented.
+When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the *final state* (found by the simulation) is not guaranteed to be a *steady state*.
+The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In the results presented here, and in the paper @Limberger2022, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
+Furthermore, the package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique.
+
+Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
 Two strategies are implemented in order to allow comparison of their results.
@@
 The second method runs two simulations, one with step-wise and slowly temporally *increasing* oxygen diffusivity, and the other with step-wise and slowly *decreasing* oxygen diffusivity.
-(Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+(Put another way, to explore the sensitivity of final states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
-
-The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.
 
@@
 
-From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted.
+From the results returned, summary measures about how the ecosystem final states changes with environmental change can be extracted.
 The function `get_stability_measures()` returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
@@
 <!-- > {>>This is the main section of the article and the reviewers weight the description here appropriately -->
+
 <!-- Indicate in what way new research questions can be pursued as a result of the software (if any). -->
+
 <!-- Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so). -->
+
 <!-- Indicate in what way the software has changed the daily practice of its users (if so). -->
+
 <!-- Indicate how widespread the use of the software is within and outside the intended user group. -->
+
 <!-- Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).<<} -->
 
-The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
+The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
 
-The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
+The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 
-Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
+Here we evidence the impact of the *microxanox* package by describing three use-cases and then by describing how the package can be a starting point for models of other ecosystems.
 The first two use cases are described in detail (including the code for reproducing them) in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) and the *Partial Reproduction* vignettes.
@@
 The `microxonox` package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
-As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding final states.
 
-An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \@ref(fig:uc1)).
+An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly describe and show here (Figure \@ref(fig:uc1)).
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
@@
 ```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
-Plot of the steady states of the simulation runs under different oxygen diffusivity.
+Plot of the final states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
@@
 The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
-I.e.
-it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
+I.e., it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state.
 
-Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of final states across an environmental gradient, running the final state finding function, and analysing the results with the function that calculates extent of hysteresis.
 The code for this is provided in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html).
@@
 This will relieve researchers from needing to make software design decisions, and rather focus on appropriately representing their ecosystem and finding the results that interest them.
-Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and that is relatively proficient in R programming.
+Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and a person or persons  relatively proficient in R programming.
 
 # Conclusions
+
 <!-- > {>>Set out the conclusion of this original software publication.<<} -->
 
-The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
-It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
-In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
+The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity present in each of the functional groups of organisms present.
+It has been used for the research described in another paper that provides one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
+In that paper, we show that diversity can have large and important effects on ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
 The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development [@Bush2017].
@@
 Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change [@Scheffer2009].
-In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.
+In addition this package could be used as a template for developing models of other types of ecosystems and organism.
 By doing so, other models can profit from the overall framework used, and the reproducibility aspects as well as the flexibility implemented.
@@
 # Conflict of Interest
+
 The authors declare no known conflicting or competing interests associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
@@
 # Acknowledgements
+
 This project was part of SNF Project 310030_188431. The project was also supported by the University of Zurich Research Priority Programme in Global Change and Biodiversity.

### 2023-06-09 inst/manuscript/Response.md
Final Draft Manuscript

@@
 
-CC: [nkourkou\@uoi.gr](mailto:nkourkou@uoi.gr){.email}
+<!-- CC: [nkourkou\@uoi.gr](mailto:nkourkou@uoi.gr){.email}
 
@@
 
-SoftwareX
+SoftwareX -->
 
-## Editor and Reviewer comments
+## Response from Authors to Editor and Reviewer comments
 
@@
 
-{\>\>Owen: Rainer - please correct and then make a response here.\<\<} {\>\>Rainer: Will do this at the end\<\<} \*
+**Response:**: Done
 
@@
 
-{\>\>Owen: Rainer - please would you see to this?\<\<}
+The manuscript has been proofread by a native speaker with extensive experience in publishing scientific articles.
 
@@
 
-{\>\>Owen: Rainer - please would you see to this?\<\<}
+Added three references from SoftwareX.
 
@@
 
-------------------------------------------------------------------------
+I hope we adressed all the comments to your satisfaction.
+
+Sincerely,
+
+Rainer M Krug
+
+<!-- ------------------------------------------------------------------------
 
@@
 Please contact the publication office if you have any questions.
+-->

### 2023-06-09 inst/manuscript/manuscript_microxanox.Rmd
Final Draft Manuscript

@@
 
-We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2022, and to base our work on the work and model of @Bush2017.
+We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change @Limberger2023, and to base our work on the work and model of @Bush2017.
 With this goal in mind, we developed the `microxanox` package [@Krug2022].
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used).
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2023 for an example of how this is used, and XXXXX).
 As such, all the parameters needed to run a simulation or a set of simulations are contained in a single object (which can easily be created using included functions).
@@
 The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
-In the results presented here, and in the paper @Limberger2022, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
+In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
 Furthermore, the package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique.
@@
 
-The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
+The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
+@Correndo2023, @Spillner2022or @Fox2020): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 
@@
 The first two use cases are described in detail (including the code for reproducing them) in the [User Guide](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) and the *Partial Reproduction* vignettes.
-The third is taken from @Limberger2022 and @Petchey2022.
+The third is taken from @Limberger2023 and @Petchey2022.
 
@@
 As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change.
-Extensive results concerning this question are given in a separate publication @Limberger2022.
+Extensive results concerning this question are given in a separate publication @Limberger2023.
 Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
@@
 The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity present in each of the functional groups of organisms present.
-It has been used for the research described in another paper that provides one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022.
+It has been used for the research described in another paper that provides one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2023.
 In that paper, we show that diversity can have large and important effects on ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.

### 2023-06-09 inst/manuscript/manuscript_microxanox.tex
Final Draft Manuscript

@@
   \(Microxanox\) is an R package to simulate a three functional group ecosystem (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations.
-Simulations can be run individually or over a parameter range, to find steady states.
+Simulations can be run individually or over a parameter range.
 The model can be implemented with different numbers of species per functional group.
@@
     \begin{keyword}
-    reproducibility \sep regime shift \sep steady state \sep 
+    reproducibility \sep regime shift \sep steady state \sep final state \sep 
     ordinary differential equations
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1944}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3472}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4583}}@{}}
 \toprule\noalign{}
@@
 C8 & If available Link to developer documentation/manual & \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} \\
-C9 & Support email for questions & Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh \\
+C9 & Support email for questions & \href{mailto:Rainer.Krug@uzh.ch}{\nolinkurl{Rainer.Krug@uzh.ch}}; \href{mailto:Owen.Petchey@ieu.uzh}{\nolinkurl{Owen.Petchey@ieu.uzh}} \\
 \end{longtable}
@@
 
-Mathematical models play a key role in the development of ecosystem models and the understanding about how and why ecosystems change when the environment changes \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}.
+Mathematical models play a key role in the development of ecosystem models and in increasing the understanding of how and why ecosystems change when the environment changes \citep{Binzer2016a, Chaparro-Pedraza2021, Vasseur2005}.
 They are also important for developing hypotheses to test in empirical studies.
-One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver \citep{Scheffer2001}.
-An environmental driver is an environmental condition that affects an ecosystem.
-Still, it is assumed not to be affected by the ecosystem, such as the rate of nutrient input into a lake.
+One area of ecology in which models have been influential is the understanding of ecosystem responses to gradual change in an environmental driver \citep{Scheffer2001}.
+An environmental driver is an environmental condition that affects an ecosystem, but which is not itself affected by the ecosystem.
+An example driver would be the amount of nutrients entering a lake from run-off from surrounding agricultural land.
 
@@
 It is also possible that the ecosystem state changes abruptly to a new state that is difficult to recover from \citep{Scheffer2001}.
-This possibility for abrupt, perhaps catastrophic changes that are difficult to reverse causes considerable concern \citep{Collins2021, Northrop2021, Vandermeer2019}.
+This possibility for abrupt, perhaps catastrophic changes, which are difficult to reverse, causes considerable concern \citep{Collins2021, Northrop2021, Vandermeer2019}.
 
@@
 Three types of microbes occur in the model: cyanobacteria (CB) dominating the oxic state, and two types of sulfur bacteria that dominate the anoxic state (sulfate reducing bacteria (SB) and phototrophic sulfur bacteria (PB)).
-The model shows that gradual change in the rate at which oxygen could diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
+The model shows that gradual change in the rate at which oxygen can diffuse into the ecosystem (termed oxygen diffusitivity) could cause catastrophic changes in the ecosystem state that would be difficult to reverse.
 
 One feature of the study by \citet{Bush2017} was limited biodiversity.
-Specifically, there was no biodiversity within each of the three types of bacteria.
-This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem response to environmental change.
+Specifically, there was no biodiversity within each of the three types (i.e., functional groups) of bacteria.
+This leaves the question of if and how biodiversity within these functional groups might affect the ecosystem response to environmental change.
 This limitation is not specific to the study of \citet{Bush2017}.
@@
 
-We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change \citet{Limberger2022}, and to base our work on the work and model of \citet{Bush2017}.
+We decided to fill this research gap by making a simulation study of how within functional group biodiversity affects ecosystem responses to environmental change \citet{Limberger2023}, and to base our work on the work and model of \citet{Bush2017}.
 With this goal in mind, we developed the \texttt{microxanox} package \citep{Krug2022}.
@@
 
-The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2022} for an example of how this is used).
-As such, all the parameters needed to run a simulation or find a steady state (i.e.~the final state of the ecosystem) are contained in a single object (which can easily be created using included functions).
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2023} for an example of how this is used, and XXXXX).
+As such, all the parameters needed to run a simulation or a set of simulations are contained in a single object (which can easily be created using included functions).
 This parameter object is given to a function that runs the simulations and returns the results.
@@
 
-\hypertarget{finding-steady-states}{%
-\subsection{Finding Steady States}\label{finding-steady-states}}
+\hypertarget{examining-ecosystem-responses-to-environmental-change}{%
+\subsection{Examining ecosystem responses to environmental change}\label{examining-ecosystem-responses-to-environmental-change}}
 
-The general approach to finding the numerical value of a system's steady state (if it exists) with a specific parameter set is to run the simulation for a long time and record the final state.
-When one does this across a range of environmental conditions, one discovers how the steady state of the system changes with the environmental conditions.
-The package contains functionality for finding steady states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
+One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's steady state (if it exists).
+The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
+When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
+The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-Two numerical strategies for finding steady states and their sensitivity to parameters are implemented.
+When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the \emph{final state} (found by the simulation) is not guaranteed to be a \emph{steady state}.
+The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In the results presented here, and in the paper \citet{Limberger2023}, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
+Furthermore, the package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique.
+
+Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
 Two strategies are implemented in order to allow comparison of their results.
@@
 The second method runs two simulations, one with step-wise and slowly temporally \emph{increasing} oxygen diffusivity, and the other with step-wise and slowly \emph{decreasing} oxygen diffusivity.
-(Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
+(Put another way, to explore the sensitivity of final states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the \emph{Temporal method}).
-
-The package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique, and indeed if a steady state has been achieved.
 
@@
 
-From the results returned, summary measures about how the ecosystem steady states changes with environmental change can be extracted.
+From the results returned, summary measures about how the ecosystem final states changes with environmental change can be extracted.
 The function \texttt{get\_stability\_measures()} returns quantities such as the amount of environmental change required for the system to abruptly change to a different state.
@@
 
-The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of doing unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
+The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
 
-The design of the package code and functionality is with reproducibility in mind: the combination of all parameters being in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object provides a relatively easy to use framework to implement reproducible experiments.
+The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
+\citet{Correndo2023}, \citet{Spillner2022or} \citet{Fox2020}): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 
-Here we evidence the impact of the \emph{microxanox} package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems.
+Here we evidence the impact of the \emph{microxanox} package by describing three use-cases and then by describing how the package can be a starting point for models of other ecosystems.
 The first two use cases are described in detail (including the code for reproducing them) in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide} and the \emph{Partial Reproduction} vignettes.
-The third is taken from \citet{Limberger2022} and \citet{Petchey2022}.
+The third is taken from \citet{Limberger2023} and \citet{Petchey2022}.
 
@@
 The \texttt{microxonox} package contains functionality to make a specific temporal pattern of change in the oxygen diffusivity.
-As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding steady states.
+As well as allowing individual simulations during which oxygen diffusivity varies, this functionality forms the basis of the temporal method for finding final states.
 
-An example of this functionality is given in the \emph{Partial Reproduction} vignette, which we briefly show here (Figure \ref{fig:uc1}).
+An example of this functionality is given in the \emph{Partial Reproduction} vignette, which we briefly describe and show here (Figure \ref{fig:uc1}).
 The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria.
@@
 \caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes.
-Plot of the steady states of the simulation runs under different oxygen diffusivity.
+Plot of the final states of the simulation runs under different oxygen diffusivity.
 In this simulation there is only one strain in each functional group.
@@
 The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
-I.e.
-it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state (a definition of hysteresis).
+I.e., it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state.
 
-Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of steady states across an environmental gradient, running the steady state finding function, and analysing the results with the function that calculates extent of hysteresis.
+Using the package to calculate the extent of hysteresis involves setting ecosystem and simulation parameters, including parameters for the finding of final states across an environmental gradient, running the final state finding function, and analysing the results with the function that calculates extent of hysteresis.
 The code for this is provided in the \href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{User Guide}.
@@
 As discussed in the Introduction section, the package was motivated by the question of how biodiversity influences ecosystem responses to environmental change.
-Extensive results concerning this question are given in a separate publication \citet{Limberger2022}.
+Extensive results concerning this question are given in a separate publication \citet{Limberger2023}.
 Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
@@
 This will relieve researchers from needing to make software design decisions, and rather focus on appropriately representing their ecosystem and finding the results that interest them.
-Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and that is relatively proficient in R programming.
+Nevertheless, adaptation of the code in the package will require a person / persons that can take a conceptual model of an ecosystem and then represent that in terms of parameters and rate equations, and a person or persons relatively proficient in R programming.
 
@@
 
-The \emph{microxanox} R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present.
-It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience \citet{Limberger2022}.
-In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
+The \emph{microxanox} R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity present in each of the functional groups of organisms present.
+It has been used for the research described in another paper that provides one of the first investigations of the effects of diversity on ecosystem resilience \citet{Limberger2023}.
+In that paper, we show that diversity can have large and important effects on ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity.
 The \emph{microxanox} package has also been used to reproduce the results of the paper that inspired the package development \citep{Bush2017}.
@@
 Likewise, the package could be used to power an investigation of the effects of biodiversity on the usefulness of early warning signals of abrupt ecosystem change \citep{Scheffer2009}.
-In addition this package could be used as a template for the implementation for developing models of other types of ecosystems and organism.
+In addition this package could be used as a template for developing models of other types of ecosystems and organism.
 By doing so, other models can profit from the overall framework used, and the reproducibility aspects as well as the flexibility implemented.

### 2023-06-12 inst/manuscript/manuscript_microxanox.Rmd
Revised Submission

@@
 
-The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2023 for an example of how this is used, and XXXXX).
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement [@Petchey2022] to @Limberger2023 for an example of how this is used).
 As such, all the parameters needed to run a simulation or a set of simulations are contained in a single object (which can easily be created using included functions).
@@
 The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
-@Correndo2023, @Spillner2022or @Fox2020): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
+@Correndo2023, @Spillner2022 or @Fox2020): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 
@@
 Extensive results concerning this question are given in a separate publication @Limberger2023.
-Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
-I.e.
+Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have, i.e.
 biodiversity can qualitatively change the state of the ecosystem.

### 2023-06-12 inst/manuscript/manuscript_microxanox.tex
Revised Submission

@@
 
-The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2023} for an example of how this is used, and XXXXX).
+The framework used when writing this package aimed to maximise simplicity for the user and to make it straightforward to reproduce results (see the supplement \citep{Petchey2022} to \citet{Limberger2023} for an example of how this is used).
 As such, all the parameters needed to run a simulation or a set of simulations are contained in a single object (which can easily be created using included functions).
@@
 The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
-\citet{Correndo2023}, \citet{Spillner2022or} \citet{Fox2020}): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
+\citet{Correndo2023}, \citet{Spillner2022} or \citet{Fox2020}): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 
@@
 Extensive results concerning this question are given in a separate publication \citet{Limberger2023}.
-Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have.
-I.e.
-biodiversity can qualitatively change the state of the ecosystem.
+Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have, i.e.~biodiversity can qualitatively change the state of the ecosystem.
 

### 2023-06-29 inst/manuscript/Response.md
Update Response.md

@@
 
+> 0.
+It was difficult to follow the [Responses and Suggestions] in the new version.
+Those sections/paragraphs that have been rewritten, added, or modified in the paper should be included in the reply letter and explicitly highlight the parts that correspond precisely to each remark.
+The highlighted paper and the clean version should be shown at the end of the reply letter.
+
+!!! Needs a response.
+
 ### Reviewer #1
@@
 
-> The authors in section 3.3 claim several sentences that could confuse readers.
+> 2.1 The authors in section 3.3 claim several sentences that could confuse readers.
 2.1 `one discovers how the steady state of the system responds to the environmental conditions`.
@@
 
+!!! 2.2 The comment has not been addressed.
+
 > 2.3 `Two methods for finding steady states are implemented`.
@@
 **Response:** We envisage and hope that Response R1 sufficiently clarifies steady state finding.
+
+!!! 2.3 The comment has not been addressed.
 
@@
 Indeed, the numerical solution of the dynamical model is sensitive to a set of parameters and initial conditions.
-Maybe the authors mean the following: to evaluate the performance of the numerical parametric sensitivity analysis of the dynamical model concerning one parametric variation (oxygen diffusivity) under two schemes.2.7 It is not clear why there are two `methods` for numerical parametric sensitivity analysis.
-The authors should include the arguments of this.
+Maybe the authors mean the following: to evaluate the performance of the numerical parametric sensitivity analysis of the dynamical model concerning one parametric variation (oxygen diffusivity) under two schemes.
+
+!!! 2.6c The comment has not been addressed.
+
+> 2.7 It is not clear why there are two `methods` for numerical parametric sensitivity analysis. The authors should include the arguments of this.
 
 **Response:** We removed this text and now simply state that having two methods allows the comparison of their results.
+
+!!! 2.7 The comment has not been addressed.
 
@@
 
+!!! 2.8 The comment has not been addressed.
+
 > 3.  Please consider the following suggestions:
->
+
+!!! 3 It wasn't easy to follow the [Suggestion] in the new version.
+
 > Line 5 `types of ecosystem`

### 2023-06-30 inst/manuscript/revision2/manuscript_microxanox_rev2.Rmd
working in revision 2

--- 2023-06-12 inst/manuscript/manuscript_microxanox.Rmd
@@
   - regime shift 
-  - steady state
   - final state
@@
 
-One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's steady state (if it exists).
+One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's state.
 The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
-When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
+When one does this across a range of environmental conditions, one can discovers how the final state of the system changes with the environmental conditions.
 The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
 
-When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the *final state* (found by the simulation) is not guaranteed to be a *steady state*.
-The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+When one wishes to be able to make conclusions about how the *steady state* of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.
+The software does not provide the user with a steady state.
+In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
 In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
@@
 Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow comparison of their results.
+Two strategies are implemented in order to allow users to compare of their results.
 The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the *Replication method*).
@@
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the *Temporal method*).
+
+An potentially important difference between the two methods is in the system state when a new value of oxygen diffusivity is set.
+In the replication method, the system state when a new value of oxygen diffusivity is set is always the same.
+Whereas in the temporal method, the system state when a new value of oxygen diffusivity is set is the final state of the system for the previously set value of oxygen diffusivity.
+Since some modellers prefer one approach and others another, we decided to implement both.
 
@@
 Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change [@Scheffer2001].
-The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two distinct final states.
 I.e., it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state.

### 2023-09-06 inst/manuscript/manuscript_microxanox.Rmd
Draft Revision 2

@@
 Extensive results concerning this question are given in a separate publication @Limberger2023.
-Here we describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have, i.e.
+Here describe one of the results, which is that having biodiversity in a functional group can allow state changes to occur that otherwise would not have, i.e.
 biodiversity can qualitatively change the state of the ecosystem.

### 2023-09-07 inst/manuscript/revision2/manuscript_microxanox_rev2.Rmd
revised response and ms

@@
 The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
+In terms of non-linear systems analysis, this would be termed a *parameteric sensitivity analysis* being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
 
-When one wishes to be able to make conclusions about how the *steady state* of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.
+When one wishes to be able to make conclusions about how the *steady state* (or an equilibrium point) of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.
 The software does not provide the user with a steady state.

### 2023-09-08 inst/manuscript/revision_2/Response_rev2.md
Reorganisation of folder

--- 2023-06-29 inst/manuscript/Response.md
@@
 
-> 0.
-It was difficult to follow the [Responses and Suggestions] in the new version.
+### Introduction
+
+Thank you for the opportunity to revise the manuscript again. We apologise that our first revision was found to be insufficient and that it was not as easy to follow as it should have been.
+
+We provide a guide to the materials we resubmit:
+
+First is this document.
+In it we use this typeface to give our responses and we use the following typeface when text from the manuscript is being quoted: `this typeface shows text being quoted from the manuscript`.
+In this document we use the following type of box to indicate comments from the reviewer.
+
+> This kind of box contains comments from the reviewer.
+
+The second document is the revised and clean manuscript, and the third is the revised manuscript with changes shown.
+
+Below, we cover each of the latest reviewer comments in turn.
+
+> It was difficult to follow the [Responses and Suggestions] in the new version.
 Those sections/paragraphs that have been rewritten, added, or modified in the paper should be included in the reply letter and explicitly highlight the parts that correspond precisely to each remark.
@@
 
-!!! Needs a response.
+Apologies that we provided materials that were difficult to follow.
+With the current revision we provide two versions of the revised manuscript, as requested, one a clean version, and one with changes highlight.
+*Rainer, please do this.* We also, as requested, include all revised text in this response document and then reference the location in the revised document by line number.
+*Rainer, please check that this is so, before resubmission.
+Wait for our final version of the text before inserting the correct line numbers below.*
 
-### Reviewer #1
+> 2.1a-b The comment has been addressed.
 
-> 1.  Please consider reference style: Number the references (numbers in square brackets) in the list in the order in which they appear in the text.
+Thank you.
 
-**Response:**: Done
+> 2.2 The comment has not been addressed.
 
-> 2.1 The authors in section 3.3 claim several sentences that could confuse readers.
-2.1 `one discovers how the steady state of the system responds to the environmental conditions`.
-There are several concerns that the authors should clarify:
->
-> a)  `The steady state`.
-Is there a unique steady state?
-Is there a reported contribution concerning a formal analysis of the model in the sense of stability (stable state existence and uniqueness)?
-Maybe, authors should refer as `a system's stable state (if it exists)`.
-
-**Response:** Thank you for raising this important issue, which we had overlooked.
-A number of the reviewer comments below are closely related, hence we sometimes refer to this response and do so using the label "Response R1".
-In order to clarify about the existence and uniqueness of steady states we have added the text `When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the *final state* (found by the simulation) is not guaranteed to be a *steady state*.
-The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
-In the results presented here, and in the paper @Limberger2022, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
-Furthermore, the package does not include methods for a formal analysis of the stability of the system and users should take care to assess if steady states are unique.`
-
-**Response:** In addition, we notice that we sometimes used the term "stable state" and sometimes "steady state".
-We now use only the term "steady state".
-Furthermore, we now use the term "final state" to refer to the state of the system at the end of a simulation.
-
-> `b)`The steady state of the system responds\`.
-It should be noted that the steady state does not respond to the environment.
-Maybe, the authors refer to the numerical result of a steady state as a function of parameter and initial condition (if a steady state is not unique).
-
-**Response:** Thank you. We have edited the text so that we do not state that the steady state responds to the environment.
-
-> 2.2 `finding steady states that correspond to values of one environmental driver`.
+> Here is the original comment 2.2: `finding steady states that correspond to values of one environmental driver`.
 The software does not find a steady state.
@@
 
-**Response:** We envisage and hope that Response R1 sufficiently clarifies steady state finding.
+**Response**: We agree--the software does not find a steady state of the system.
+We no longer claim that it does, and we have added `The software does not provide the user with a steady state.` (lines XXX-YYY).
+Furthermore, in lines XXX-YYY of the newly revised manuscript we explain that the simulations provide a *final state* of the system, and that this is not guaranteed to be a steady state of the system: `When one wishes to be able to make conclusions about how the *steady state* of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.` (lines XXX-YYY).
+We also then list the actions that users need to take in order to somewhat safely assume that the provided final state is a steady state of the system: `In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.` (lines XXX-YYY)
 
-!!! 2.2 The comment has not been addressed.
+> 2.3 The comment has not been addressed.
 
-> 2.3 `Two methods for finding steady states are implemented`.
+> Here is the original comment 2.3: `Two methods for finding steady states are implemented`.
 Again, the software does not find a steady state.
@@
 
-**Response:** We envisage and hope that Response R1 sufficiently clarifies steady state finding.
+**Response**: We agree--the software does not find a steady state of the system.
+We no longer claim that it does, and we have added `The software does not provide the user with a steady state.` (lines XXX-YYY).
 
-!!! 2.3 The comment has not been addressed.
+> 2.4 The comment has been addressed.
 
-> 2.4 `The first runs a separate simulation for each combination of starting conditions and oxygen diffusivity`.
-This sentence is confusing.
-What the authors mean with `separate simulation` and `starting conditions`.
-Maybe the authors refer to each initial condition and parameter set value (oxygen diffusivity) as an open loop simulation run of the dynamic system.
-Are these simulations results focused on obtaining a bifurcation diagram Fig.
-3 and 4 from {10.1038/s41467-017-00912-x}?
-Please clarified this point.
+Thank you.
 
-**Response:** We have clarified what is meant by "separate simulation" by changing to "independent simulation", and change "starting conditions" to "initial conditions".
-We hope that this more standard terminology is clearer.
-Furthermore, we have added that "This is the method used in the @Bush2017 study and was used to obtain the results in figures 3 and 4 of that article."
+> 2.5 The comment has been addressed.
 
-> 2.5 `The second runs two simulations, one with step-wise and slowly temporally increasing oxygen diffusivity, and the other with step-wise and slowly decreasing oxygen diffusivity`.
-This sentence needs to be clarified.
-Authors refer `temporally` as several changes in the value of oxygen diffusivity during a simulation run time.
-Maybe the authors mean the following: to explore the parametric sensitivity of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
-Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.
+Thank you.
 
-**Response:** We are grateful for this different perspective on how to explain the method and have added it to the manuscript ("Put another way, to explore the sensitivity of steady states of the dynamic model under oxygen diffusivity variation, subset values of this parameter are numerically evaluated.
-Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.")
+> 2.6a-b The comment has been addressed.
 
-> 2.6 `We implemented two methods since there is no definitive best method, and in order to check if results were sensitive to choice of method`. This sentence is confusing.
+Thank you.
 
-**Response:** Thank you for pointing out this source of confusion.
-We change the text to "Two strategies are implemented in order to allow comparison of their results." and moved it to just after we state that two strategies are provided.
+> 2.6c The comment has not been addressed.
 
-> a)  `two methods` Maybe the authors mean the following: two numerical strategies to explore a parametric sensitivity numerical evaluation.
-
-**Response:** Thank you for this suggestions. We have changed the text to "Two numerical strategies for finding steady states and their sensitivity to parameters are implemented."
-
-> b)  `there is no definitive best method`. Under which criteria? The authors present only two numerical evaluations. Please clarified this point.
-
-**Response:** We have removed this claim since it is not demonstrated to be true and is unimportant.
-
-> c)  `in order to check if results were sensitive to choice of method`.
+> Here is the original comment 2.6c:  `in order to check if results were sensitive to choice of method`.
 This sentence needs to be clarified.
@@
 
-!!! 2.6c The comment has not been addressed.
+**Response:** We agree that the software does not itself make a comparison.
+It only provides users information that they could use to make a comparison.
+This and the surround text have been revised and the points clarified: `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
+Two strategies are implemented in order to allow users to compare of their results.` (lines XXX-YYY).
 
-> 2.7 It is not clear why there are two `methods` for numerical parametric sensitivity analysis. The authors should include the arguments of this.
+> 2.7 The comment has not been addressed.
 
-**Response:** We removed this text and now simply state that having two methods allows the comparison of their results.
+> Here is the original comment 2.7: It is not clear why there are two `methods` for numerical parametric sensitivity analysis. The authors should include the arguments of this.
 
-!!! 2.7 The comment has not been addressed.
+**Response:** We now state that `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
+Two strategies are implemented in order to allow users to compare of their results.` (lines XXX-YYY).
+Furthermore, we have added the following text: `An potentially important difference between the two methods is in the system state when a new value of oxygen diffusivity is set.
+In the replication method, the system state when a new value of oxygen diffusivity is set is always the same.
+Whereas in the temporal method, the system state when a new value of oxygen diffusivity is set is the final state of the system for the previously set value of oxygen diffusivity.
+Since some modellers prefer one approach and others another, we decided to implement both.` (lines XXX-YYY)
 
-> 2.8 Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H.
+> 2.8 The comment has not been addressed.
+
+> Here is the original comment 2.8: Please consider rewrite the section 3.3 (see previous 2.1 - 2.7 points) in the sense of Parametric sensitivity analysis, Open loop dynamical behavior, Equilibrium points (see Khalil, H.
 K.
@@
 
-**Response:** We have the opinion that the with the revisions already made and detailed above, and those below, that the terminology is now more in line with this text, and will be clearly understandable by the target audience (theoretical ecologists) even when the section is not more substatially rewritten.
+**Response:** in our previous response we attempted to argue that the language and terms used in the text were appropriate and understandable by the target audience (quantitative ecologists with some knowledge of dynamical systems modelling).
+*Rainer, you could ask an experienced quantitative ecologist for testimony on this and then include that here as evidence that that language is suitable.*
 
-!!! 2.8 The comment has not been addressed.
+> 3 It wasn't easy to follow the [Suggestion] in the new version.
+
+We interpret this as the reviewer asking for clearer accounts of what was changed, and now provide below, with line numbers corresponding to those in the newly revised document, and the corresponding text in green font.
+*Rainer, please do this.*
 
 > 3.  Please consider the following suggestions:
-
-!!! 3 It wasn't easy to follow the [Suggestion] in the new version.
 
@@
 
-> Line ...
->
-> 4.  Please, consider English Editing Service or Proofread Your Paper.
+> 4 The suggestion seems to have been heeded.
 
-The paper has been proofread and edited.
+Thank you.
 
-The manuscript has been proofread by a native speaker with extensive experience in publishing scientific articles.
+> 5 The suggestion seems to have been heeded.
 
-> 5.  Please, consider include two or three references of SoftwareX Journal.
-
-Added three references from SoftwareX.
-
-------------------------------------------------------------------------
-
-I hope we adressed all the comments to your satisfaction.
-
-Sincerely,
-
-Rainer M Krug
-
-<!-- ------------------------------------------------------------------------
-
-More information and support
-
-FAQ: How do I revise my submission in Editorial Manager?
-
-<https://service.elsevier.com/app/answers/detail/a_id/28463/supporthub/publishing/>
-
-You will find information relevant for you as an author on Elsevier's Author Hub: <https://www.elsevier.com/authors>
-
-FAQ: How can I reset a forgotten password?
-<https://service.elsevier.com/app/answers/detail/a_id/28452/supporthub/publishing/> For further assistance, please visit our customer service site: <https://service.elsevier.com/app/home/supporthub/publishing/> Here you can search for solutions on a range of topics, find answers to frequently asked questions, and learn more about Editorial Manager via interactive tutorials.
-You can also talk 24/7 to our customer support team by phone and 24/7 by live chat and email
-
-#AU_SOFTX#
-
-To ensure this email reaches the intended recipient, please do not delete the above code
-
-In compliance with data protection regulations, you may request that we remove your personal registration details at any time.
-(Remove my information/details).
-Please contact the publication office if you have any questions.
--->
+Thank you.

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.tex
Reorganisation of folder

--- 2023-06-12 inst/manuscript/manuscript_microxanox.tex
@@
     \begin{keyword}
-    reproducibility \sep regime shift \sep steady state \sep final state \sep 
+    reproducibility \sep regime shift \sep final state \sep 
     ordinary differential equations
@@
 
-\hypertarget{required-metadata}{%
-\section{Required Metadata}\label{required-metadata}}
+\section{Required Metadata}\label{required-metadata}
 
-\hypertarget{current-code-version}{%
-\subsection{Current code version}\label{current-code-version}}
+\subsection{Current code version}\label{current-code-version}
 
@@
 
-\hypertarget{motivation-and-significance}{%
-\section{Motivation and significance}\label{motivation-and-significance}}
+\section{Motivation and significance}\label{motivation-and-significance}
 
@@
 
-\hypertarget{software-description}{%
-\section{Software description}\label{software-description}}
+\section{Software description}\label{software-description}
 
@@
 
-\hypertarget{software-architecture}{%
-\subsection{Software architecture}\label{software-architecture}}
+\subsection{Software architecture}\label{software-architecture}
 
@@
 
-\hypertarget{running-one-simulation}{%
-\subsection{Running one simulation}\label{running-one-simulation}}
+\subsection{Running one simulation}\label{running-one-simulation}
 
@@
 
-\hypertarget{examining-ecosystem-responses-to-environmental-change}{%
-\subsection{Examining ecosystem responses to environmental change}\label{examining-ecosystem-responses-to-environmental-change}}
+\subsection{Examining ecosystem responses to environmental change}\label{examining-ecosystem-responses-to-environmental-change}
 
-One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's steady state (if it exists).
+One approach to finding how the ecosystem responds to environmental driver change is to examine the relationship between the environmental driver value and the numerical value of a system's state.
 The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
-When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
+When one does this across a range of environmental conditions, one can discovers how the final state of the system changes with the environmental conditions.
 The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
+In terms of non-linear systems analysis, this would be termed a \emph{parameteric sensitivity analysis} being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
 
-When one wishes to be able to make conclusions about how the steady state of the system is affected by the environmental driver, it is very important to note that the \emph{final state} (found by the simulation) is not guaranteed to be a \emph{steady state}.
-The user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+When one wishes to be able to make conclusions about how the \emph{steady state} (or an equilibrium point) of the system is affected by the environmental driver, it is very important to note that the \emph{final state} (provided by the simulation) is not guaranteed to be a \emph{steady state}.
+The software does not provide the user with a steady state.
+In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
 In the results presented here, and in the paper \citet{Limberger2023}, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
@@
 Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow comparison of their results.
+Two strategies are implemented in order to allow users to compare of their results.
 The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the \emph{Replication method}).
@@
 Two value vectors of oxygen diffusivity are presented step-wise, gradually increasing and decreasing.) During this temporal environmental change, the state of the system is recorded just before change to a new oxygen diffusivity (we term this the \emph{Temporal method}).
+
+An potentially important difference between the two methods is in the system state when a new value of oxygen diffusivity is set.
+In the replication method, the system state when a new value of oxygen diffusivity is set is always the same.
+Whereas in the temporal method, the system state when a new value of oxygen diffusivity is set is the final state of the system for the previously set value of oxygen diffusivity.
+Since some modellers prefer one approach and others another, we decided to implement both.
 
@@
 
-\hypertarget{analysing-and-visualising-results}{%
-\subsection{Analysing and visualising results}\label{analysing-and-visualising-results}}
+\subsection{Analysing and visualising results}\label{analysing-and-visualising-results}
 
@@
 
-\hypertarget{impact}{%
-\section{Impact}\label{impact}}
+\section{Impact}\label{impact}
 
@@
 
-\hypertarget{use-case-1-regime-shifts-during-temporal-environmental-change}{%
-\subsection{Use case 1: Regime shifts during temporal environmental change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}}
+\subsection{Use case 1: Regime shifts during temporal environmental change}\label{use-case-1-regime-shifts-during-temporal-environmental-change}
 
@@
 
-{\centering \includegraphics[width=350px]{figures/gen_uc1_partrep_temporal_state_switching}
+{\centering \includegraphics[width=350px]{../figures/gen_uc1_partrep_temporal_state_switching}
 
@@
 
-\hypertarget{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}{%
-\subsection{Use case 2: The extent of hysteresis depends on community composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}}
+\subsection{Use case 2: The extent of hysteresis depends on community composition}\label{use-case-2-the-extent-of-hysteresis-depends-on-community-composition}
 
@@
 Hysteresis is a key feature of ecosystem responses to environmental change, because it is related to how difficult it can be to reserve the effects of environmental change \citep{Scheffer2001}.
-The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two steady states.
+The amount of hysteresis is measured as the extent of the environmental condition (here oxygen diffusitivity) for which there were two distinct final states.
 I.e., it is the extent of the environmental conditions for which historical conditions play an important role in determining the current system state.
@@
 
-{\centering \includegraphics[width=150px]{figures/gen_uc2_user_guide_hysteresis}
+{\centering \includegraphics[width=150px]{../figures/gen_uc2_user_guide_hysteresis}
 
@@
 
-\hypertarget{use-case-3-effects-of-functional-diversity-on-regime-shifts}{%
-\subsection{Use case 3: Effects of functional diversity on regime shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}}
+\subsection{Use case 3: Effects of functional diversity on regime shifts}\label{use-case-3-effects-of-functional-diversity-on-regime-shifts}
 
@@
 
-\hypertarget{adapting-for-other-ecosystems-and-organisms}{%
-\subsection{Adapting for other ecosystems and organisms}\label{adapting-for-other-ecosystems-and-organisms}}
+\subsection{Adapting for other ecosystems and organisms}\label{adapting-for-other-ecosystems-and-organisms}
 
@@
 
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+\section{Conclusions}\label{conclusions}
 
@@
 
-\hypertarget{conflict-of-interest}{%
-\section{Conflict of Interest}\label{conflict-of-interest}}
+\section{Conflict of Interest}\label{conflict-of-interest}
 
@@
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.Rmd
Reorganisation of folder
Reorganisation step 2 and recompilation of rev 2

--- 2023-09-07 inst/manuscript/revision2/manuscript_microxanox_rev2.Rmd
@@
 Furthermore, the package framework and code should serve as a useful starting point for making simulation models of other types of ecosystems.
-journal: "SoftwareX"
+journal: SoftwareX
 date: "`r Sys.Date()`"
@@
 bibliography: bibliography.bib
-csl: 'softwarex.csl'
+csl: './../softwarex.csl'
 nocite: '@Bush2017'
 linkcolor: blue
-link-citations: yes
+link-citations: true
 ---
@@
 Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
-fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
+fig <- "./figures/gen_uc1_partrep_temporal_state_switching.pdf"
 if (!file.exists(fig)){
@@
 The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
-fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
+fig <- "./figures/gen_uc2_user_guide_hysteresis.pdf"
 if (!file.exists(fig)){

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.tex
Reorganisation step 2 and recompilation of rev 2

@@
 
-{\centering \includegraphics[width=350px]{../figures/gen_uc1_partrep_temporal_state_switching}
+{\centering \includegraphics[width=350px]{./figures/gen_uc1_partrep_temporal_state_switching}
 
@@
 
-{\centering \includegraphics[width=150px]{../figures/gen_uc2_user_guide_hysteresis}
+{\centering \includegraphics[width=150px]{./figures/gen_uc2_user_guide_hysteresis}
 

### 2023-09-08 inst/manuscript/revision_2/Response_rev2.md
Change in Response

@@
 With the current revision we provide two versions of the revised manuscript, as requested, one a clean version, and one with changes highlight.
-*Rainer, please do this.* We also, as requested, include all revised text in this response document and then reference the location in the revised document by line number.
+We also, as requested, include all revised text in this response document and then reference the location in the revised document by line number.
 *Rainer, please check that this is so, before resubmission.

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.Rmd
Changes concerning term inology

@@
 To make the simulation run with multiple species per functional group, we expressed different species' characteristics in the elements of vectors and matrices.
-We also coded the ordinary differential equations to include the vectors and matrices and used matrix mathematics.
+We also coded the ordinary differential equations in matrix form to use matrix mathematics.
 In this way, we made it possible to run simulations with different numbers of species without having to change the underlying code.
@@
 A simulation is run using the `run_simulation()` function.
-In this function, the ODEs are solved using the function `ode()` in the package *deSolve* package [@Soetaert2010].
+In this function, the ODEs are solved using the function `ode()` in the package *deSolve* [@Soetaert2010].
 The `run_simulation()` function needs only one argument - an object as created by the function `new_runsim_parameter()`.
@@
 The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
-When one does this across a range of environmental conditions, one can discovers how the final state of the system changes with the environmental conditions.
+When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
 The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
-In terms of non-linear systems analysis, this would be termed a *parameteric sensitivity analysis* being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
+In terms of non-linear systems analysis, this would be termed a *parametric sensitivity analysis* being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
 
@@
 The software does not provide the user with a steady state.
-In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappear, and must also check the type of long-term dynamics occurring.
 In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
@@
 Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow users to compare of their results.
 The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the *Replication method*).
@@
 
-The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
+The open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environment changes, and how biodiversity may modulate this.
 
 The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
-@Correndo2023, @Spillner2022 or @Fox2020): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
+@Correndo2023, @Spillner2022 or @Fox2020): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits the parameter values from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 

### 2023-09-08 inst/manuscript/revision_2/manuscript_microxanox_rev2.tex
Changes concerning term inology

@@
 A simulation is run using the \texttt{run\_simulation()} function.
-In this function, the ODEs are solved using the function \texttt{ode()} in the package \emph{deSolve} package \citep{Soetaert2010}.
+In this function, the ODEs are solved using the function \texttt{ode()} in the package \emph{deSolve} \citep{Soetaert2010}.
 The \texttt{run\_simulation()} function needs only one argument - an object as created by the function \texttt{new\_runsim\_parameter()}.
@@
 The first component of this approach is to run the simulation for a long time and record the final state (i.e., the state of the system at the end of a simulation).
-When one does this across a range of environmental conditions, one can discovers how the final state of the system changes with the environmental conditions.
+When one does this across a range of environmental conditions, one discovers how the final state of the system changes with the environmental conditions.
 The package contains functionality for finding final states that correspond to values of one environmental driver, namely the value of oxygen diffusivity.
-In terms of non-linear systems analysis, this would be termed a \emph{parameteric sensitivity analysis} being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
+In terms of non-linear systems analysis, this would be termed a \emph{parametric sensitivity analysis} being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).
 
@@
 The software does not provide the user with a steady state.
-In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
+In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappear, and must also check the type of long-term dynamics occurring.
 In the results presented here, and in the paper \citet{Limberger2023}, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.
@@
 Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow users to compare of their results.
 The first method runs a independent simulation for each combination of initial conditions and oxygen diffusivity (we term this the \emph{Replication method}).
@@
 
-The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environmental changes, and how biodiversity may modulate this.
+The open source implementation and extension of the model used in \citet{Bush2017} provides the means of reproducing the results published while at the same time provides the means of making unique, innovative, and important investigations of how ecosystems change as the environment changes, and how biodiversity may modulate this.
 
 The design of the package code and functionality is with reproducibility in mind (for other example of reproducibility directed software, see e.g.
-\citet{Correndo2023}, \citet{Spillner2022} or \citet{Fox2020}): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
+\citet{Correndo2023}, \citet{Spillner2022} or \citet{Fox2020}): the combination of all parameters being in a single parameter object and the return of the simulation as a result object which inherits the parameter values from the parameter object, provides a relatively easy-to-use framework to implement reproducible experiments.
 

### 2023-09-11 inst/manuscript/revision_2/Response_rev2.md
Now finally
update response doc

@@
 SoftwareX -->
+
+Dear Editor,
+
+Thank you for the opportunity to make a second revision.
+The latest comments from the reviewer included that we had made it difficult to follow our previous revisions, and so in this revision we take even more care and provide a difference file that shows the changes from the original ms to the current revision.
+
+The reviewer also pointed out which of their original comments they felt were inadequately addressed.
+We revisited all of these, made additional changes in some cases, while in others we clarify the change already made.
+In all cases, we provide the  make a response that includes the text that was revised.
+
+We would like to thank the reviewer for their insights, which have helped to improve the manuscript again.
+We would also like to that you for allowing us additional time to revise the manuscript.
+We hope that you will find the changes acceptable now, and hope for a positive outcome.
+
+Sincerely,
+
+Rainer M Krug
 
@@
 
-**Response**: We agree--the software does not find a steady state of the system.
-We no longer claim that it does, and we have added `The software does not provide the user with a steady state.` (lines XXX-YYY).
+**Response:** We acknowledge now that the software does find a final state: e.g., `The software does not provide the user with a steady state." and do in the surrounding text carefully describe what the software does do.`
+
+**Response:** We have added `In terms of non-linear systems analysis, this would be termed a *parameteric sensitivity analysis* being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).` to the first paragraph of section 3.3.
 
@@
 
-**Response:** in our previous response we attempted to argue that the language and terms used in the text were appropriate and understandable by the target audience (quantitative ecologists with some knowledge of dynamical systems modelling).
-*Rainer, you could ask an experienced quantitative ecologist for testimony on this and then include that here as evidence that that language is suitable.*
+**Response:** We have considered this request carefully.
+Doing so included reviewing the content of the book *Khalil (2002) Nonlinear Systems*.
+As a result, we have made the following two changes to section 3.3.
+We have added the sentence `In terms of non-linear systems analysis, this would be termed a *parameteric sensitivity analysis* being conducted by running an open loop of the dynamic system under a set of initial conditions and a subset of parameter values (here oxygen diffusivity).` in the first paragraph of section 3.3.
+And we have added `(or an equilibrium point)` to the sentence `When one wishes to be able to make conclusions about how the *steady state* (or an equilibrium point) of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.` in the second paragraph of section 3.3.
+
+**Response:** We have, however, refrained from a "rewrite" of section 3.3.
+The text as written uses terms and language that will be familiar to the intended audience: quantitative ecologists with some interest in simulating ecological systems.
+Changing the terms and language to that in *Khalil, H.
+K.
+(2002).
+Nonlinear Systems* would make it more appropriate for the intended readers of that book, namely engineers and applied mathematicians.
 

### 2023-09-11 inst/manuscript/revision_2/Response_rev2.md
Rev 2

@@
 The second document is the revised and clean manuscript, and the third is the revised manuscript with changes shown.
+Thease take note, that the revised manuscript with changes shows "?" instead of slinks to references and notes.
+We did not find a way of changing this and left it as it does not concern the changes made.
 
@@
 We also, as requested, include all revised text in this response document and then reference the location in the revised document by line number.
-*Rainer, please check that this is so, before resubmission.
-Wait for our final version of the text before inserting the correct line numbers below.*
 
@@
 **Response**: We agree--the software does not find a steady state of the system.
-We no longer claim that it does, and we have added `The software does not provide the user with a steady state.` (lines XXX-YYY).
-Furthermore, in lines XXX-YYY of the newly revised manuscript we explain that the simulations provide a *final state* of the system, and that this is not guaranteed to be a steady state of the system: `When one wishes to be able to make conclusions about how the *steady state* of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.` (lines XXX-YYY).
+We no longer claim that it does, and we have added `The software does not provide the user with a steady state.` (lines 143 in track changes document).
+Furthermore, in lines 139-151 of the newly revised manuscript we explain that the simulations provide a *final state* of the system, and that this is not guaranteed to be a steady state of the system: `When one wishes to be able to make conclusions about how the *steady state* of the system is affected by the environmental driver, it is very important to note that the *final state* (provided by the simulation) is not guaranteed to be a *steady state*.` (lines 139-142).
 We also then list the actions that users need to take in order to somewhat safely assume that the provided final state is a steady state of the system: `In order to somewhat safely assume that the final state is a steady state, the user must ensure that the simulation is run for sufficiently long time for any transient dynamics to disappeared, and must also check the type of long-term dynamics occurring.
-In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.` (lines XXX-YYY)
+In the results presented here, and in the paper @Limberger2023, this was performed by visual inspection, and by checking the sensitivity of conclusions to the length of the simulation.` (lines 143-149)
 
@@
 It only provides users information that they could use to make a comparison.
-This and the surround text have been revised and the points clarified: `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow users to compare of their results.` (lines XXX-YYY).
+This and the surround text have been revised and the points clarified: `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.` (lines 152-153).
 
@@
 
-**Response:** We now state that `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.
-Two strategies are implemented in order to allow users to compare of their results.` (lines XXX-YYY).
+**Response:** We now state that `Two numerical strategies for finding final states and their sensitivity to parameters are implemented.` (lines 152-153).
 Furthermore, we have added the following text: `An potentially important difference between the two methods is in the system state when a new value of oxygen diffusivity is set.
@@
 Whereas in the temporal method, the system state when a new value of oxygen diffusivity is set is the final state of the system for the previously set value of oxygen diffusivity.
-Since some modellers prefer one approach and others another, we decided to implement both.` (lines XXX-YYY)
+Since some modellers prefer one approach and others another, we decided to implement both.` (lines 168-174)
 
@@
 
-We interpret this as the reviewer asking for clearer accounts of what was changed, and now provide below, with line numbers corresponding to those in the newly revised document, and the corresponding text in green font.
-*Rainer, please do this.*
+We interpret this as the reviewer asking for clearer accounts of what was changed, and now provide below, with line numbers corresponding to those in the newly revised document including track changes.
 
@@
 
-Changed to `types of ecosystems`
+Changed to `types of ecosystems` (L5, Abstract)
 
@@
 
-Changed to `Permanent links to`
+Changed to `Permanent links to` (C2 in Ancillary data table)
 
@@
 
-`role in understanding how ecosystems` Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work`
+`role in understanding how ecosystems` Changed to `Mathematical models play a key role in the development of ecosystem models and the understanding about how ecosystems work` (L12)
 
@@
 
-Changed to `and respond`
+Changed to `and respond` (L13)
 
@@
 
-`have been influential in how ecosystems respond to a gradual change` changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`
+`have been influential in how ecosystems respond to a gradual change` changed to `One area of ecology in which models have been influential is the understanding of their response to a gradual change in an environmental driver [@Scheffer2001].`  (L17)
 
@@
 
-Changed to `is an environmental`
+Changed to `is an environmental` (L19)
 
@@
 
-Changed to `affects an ecosystem. Still, it is assumed not to be`
+Changed to `affects an ecosystem. Still, it is assumed not to be` (L20)
 
@@
 
-Changed to `change in the system`
+Changed to `change in the system` (L31)
 
@@
 
-Changed to `to an anaerobic`
+Changed to `to an anaerobic` (L32)
 
@@
 
-Changed to `termed oxygen diffusivity`
+Changed to `termed oxygen diffusivity` (L38)
 
@@
 
-Changed to `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem`
+Changed to `This leaves the question of if and how bacteria's biodiversity within these types (i.e., functional groups) affects the ecosystem` (L43 - L44)
 
@@
 
-Changed to `With this goal in mind, we developed the microxanox package`
+Changed to `With this goal in mind, we developed the microxanox package` (L51)
 
@@
 
-Changed to `conditions, the addition`
+Changed to `conditions, the addition` (L62)
 
@@
 
-`some functions to analyze the results and visualize them to provide a starting` changed to `some functions to analyse the results and visualize these to provide a starting point`
+`some functions to analyze the results and visualize these to provide a starting` changed to `some functions to analyse the results and visualize these to provide a starting point` (L64 - L65)
 
@@
 
-Changed to `for running individual simulations and a set of simulations`
+Changed to `for running individual simulations and a set of simulations` (L77)
 
@@
 
-Changed to `the vectors and matrices and used matrix mathematics`
+Changed to `We also coded the ordinary differential equations in matrix form to use matrix mathematics.` (L81 - L82)
 
@@
 
-Changed to `code have a modular structure so that new functionality can be easily added. E.g., temporally`
+Changed to `code have a modular structure so that new functionality can be easily added. E.g., temporally` (L85)
 
@@
 
-Changed to `to maximize simplicity for the user and`
+Changed to `to maximize simplicity for the user and` (L92)
 
@@
 
-Changed to `simulation results`
+Changed to `simulation results` (L100)
 
@@
 `parameters, and reruns the simulations straightforwardly.
-In the following sections, we` changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive and re-running of the simulations straightforward.`
+In the following sections, we` changed to `This promotes reproducibility and makes incremental changes of individual parameters with a consecutive re-running of the simulations straightforward.` (L102 * L104)
 
@@
 
-Changed to `contains, among other things, the`
+Changed to `contains, among other things, the` (L113)
 
@@
 
-Changed to `their meaning, and how they are created and have values set and changed, please`
+Changed to `their meaning, and how they are created and have values set and changed, please` (L116)
 
@@
 
-Changed to `an object identical to the parameter`
+Changed to `an object identical to the parameter` (L119)
 
@@
 
-Changed to `to rerun the simulation from`
+Changed to `to rerun the simulation from` (L122)
 
@@
 
-Changed to `The general approach to finding the numerical value of a system's stable state (if it exists)`
+Changed to `The general approach to finding the numerical value of a system's stable state (if it exists)` (L125)
 

### 2023-09-27 R/plot_symmetry_measurements.R
vignettes working and checked

@@ -9,8 +9,7 @@
 #' @importFrom dplyr rename mutate select filter
 #' @importFrom ggplot2 geom_path geom_point geom_segment geom_ribbon scale_color_manual labs
 #' 
-
-
+#' @export
 plot_symmetry_measures <- function(res, 
                                    species){
   

### 2023-09-29 R/plot_trajectory_symmetry.R
bug fix v0.9.2

@@ -136,7 +136,7 @@ plot_trajectory_symmetry <-  function(res,
     # scale_color_manual(values = c("#00BD54", "#FF0000")) +
     labs(title = paste(trajectory, "trajectories"),
          x = expression(Log[10](Oxygen~diffusivity)~"["*hr^-1*"]"),
-         y = expression(Log[10](Concentration)~"[μM]"),
+         y = expression(Log[10](Concentration)~"[mM]"),
          color = "type")
     # xlim(-1.5, -0.5)
 
@@ -153,7 +153,7 @@ plot_trajectory_symmetry <-  function(res,
     }
     p <- p + scale_x_continuous(expand = c(0,0), 
                                 sec.axis = sec_axis(trans = . ~ t_func(.),
-                                                    name = expression(Log[10](Sulfide~Diffusivity)~"["*hr^-1*"]")))
+                                                    name = expression(Log[10](Sulfide~Diffusivity)~"["*hr^-1*"]"))
                                 # limits = c(-1.5, -0.5)
                                 )
     }

### 2023-09-29 R/plot_trajectory_symmetry.R
small changes

@@ -153,7 +153,7 @@ plot_trajectory_symmetry <-  function(res,
     }
     p <- p + scale_x_continuous(expand = c(0,0), 
                                 sec.axis = sec_axis(trans = . ~ t_func(.),
-                                                    name = expression(Log[10](Sulfide~Diffusivity)~"["*hr^-1*"]"))
+                                                    name = expression(Log[10](Sulfide~Diffusivity)~"["*hr^-1*"]")))
                                 # limits = c(-1.5, -0.5)
                                 
     }

### 2023-11-13 R/0.microxanox-package.R
Finalises for merge Fixes:  - missing global variable definitiuons  - missing global functions in import  - consistent formating  - other non-functioning related edits

@@ -16,7 +16,7 @@
 #' - A user guide.
 #' - A reproduction of some results of Bush et al 2017, on which the simulation is based. It shows the alternate stable states. It also extends the original study by adding temporal forcing and therefore temporal switching between states, and also effects of biotic composition on the response to environmental change.
 #'
-#' 
+#'
 #' @name microxanox-package
 #' @aliases microxanox
 #' @docType package

### 2023-11-13 R/bushplus_dynamic_model.R

@@ -1,5 +1,5 @@
 #' The rate equations, as published in the Bush et al 2017
-#' \url{https://doi.org/10.1093/clinchem/39.5.766} paper, but with forcing of
+#' \doi{10.1093/clinchem/39.5.766} paper, but with forcing of
 #' oxygen diffusivity `a_0` potential added, and the possibility to simulate
 #' multiple strains per functional group
 #'
@@ -10,20 +10,20 @@
 #'   `new_runsim_parameter()``
 #' @param log10a_forcing_func function to change oxygen diffusivity `a` depending on `t`
 #' @param ... not used. Needed to catch additional parameters.
-#' 
+#'
 #' @return a list containing two elements, namely the rate of change of the
 #'   strains, and also the current values of oxygen diffusivity `a`.
 #' @md
-#' 
+#'
+#' @autoglobal
+#'
 #' @export
 bushplus_dynamic_model <- function(
-  t, 
-  state, 
-  parameters, 
-  log10a_forcing_func,
-  ...
-){
-  
+    t,
+    state,
+    parameters,
+    log10a_forcing_func,
+    ...) {
   ## unpack state variables from the state object
   CB <- state[grep("CB", names(state))]
   names_CB <- names(CB)[order(names(CB))]
@@ -35,8 +35,8 @@ bushplus_dynamic_model <- function(
   names_SB <- names(SB)[order(names(SB))]
   SB <- as.numeric(SB[order(names(SB))])
   # print(c(CB, PB, SB))
-  
-  
+
+
   # CB rates of change
   CB_growth_rate <- growth1(state["P"], parameters$CB$g_max_CB, parameters$CB$k_CB_P) * inhibition(state["SR"], parameters$CB$h_SR_CB) * CB
   CB_mortality_rate <- parameters$CB$m_CB * CB
@@ -46,32 +46,32 @@ bushplus_dynamic_model <- function(
   PB_growth_rate <- growth2(state["P"], state["SR"], parameters$PB$g_max_PB, parameters$PB$k_PB_P, parameters$PB$k_PB_SR) * inhibition(state["O"], parameters$PB$h_O_PB) * PB
   PB_mortality_rate <- parameters$PB$m_PB * PB
   PB_rate <- PB_growth_rate - PB_mortality_rate + parameters$PB$i_PB
-  
+
   # SB rates of change
[…]

### 2023-11-13 R/get_stability_measures.temporal_ssfind_result.R

@@ -1,77 +1,79 @@
 #' #' The function `get_stability_measures.temporal_ssfind_result` extracts the `result`
-#' `data.frame` or `tibble` from the object `x` and processes it. If one wants to extract. If 
+#' `data.frame` or `tibble` from the object `x` and processes it. If one wants to extract. If
 #' one extracts the `results` manually, the function `get_stability_measures_temporal_ssfind_result`
 #' needs to be used.
 #' @rdname get_stability_measures
-#' 
-#' @global Quantity starts_with direction up down a_O a Species time
-#' 
-#' 
+#'
+#' @autoglobal
+#'
+#'
 #' @importFrom tidyr gather spread
-#' @importFrom dplyr summarise pull filter across select 
+#' @importFrom dplyr summarise pull filter across select
 #' @importFrom stats na.omit
-#' 
+#'
+#' @autoglobal
+#'
 #' @export
 get_stability_measures.temporal_ssfind_result <- function(
-  ss_object,
-  threshold_diff_log10scale = 3,
-  ...
-) {
+    ss_object,
+    threshold_diff_log10scale = 3,
+    ...) {
   if (inherits(ss_object, "replication_ssfind_result")) {
     result <- ss_object$result
   } else {
     result <- ss_object
   }
-  
+
   # check whether simulation result from symmetric or bush/trait
   if ("recovery" %in% colnames(result)) {
     # reformat the data.table so it fits the function
     result <- result %>%
-      mutate(direction = ifelse(recovery == "oxic", "up", "down"), 
-             a_O = aO) %>%
+      mutate(
+        direction = ifelse(recovery == "oxic", "up", "down"),
+        a_O = aO
+      ) %>%
       select(-recovery, -aS, -aO)
   }
-  
+
   ## The following is preparing the data
   these <- grep("B_", names(result))
   these <- c(these, which(names(result) %in% c("SO", "SR", "O", "P")))
   temp <- result %>%
-    dplyr::filter(dplyr::across(these, ~ .x >-0.001)) %>% ## there are rarely negative abundances greater than -0.001. This line and the na.omit removes them 
+    dplyr::filter(dplyr::across(these, ~ .x > -0.001)) %>% ## there are rarely negative abundances greater than -0.001. This line and the na.omit removes them
     tidyr::gather(key = "Species", value = Quantity, these) %>%
     select(-time) %>%
[…]

### 2023-11-13 R/growth2.R

@@ -5,14 +5,16 @@
 #' @param g_max Maximum growth rate
 #' @param k_X Half saturation constant for resource X
 #' @param k_Y Half saturation constant for resource Y
-#' 
+#'
 #' @return Growth rate
+#'
+#' @autoglobal
+#'
 #' @export
 growth2 <- function(
-  X, Y, 
-  g_max, 
-  k_X, 
-  k_Y
-) {
-  g_max * ( X / (k_X + X) ) * ( Y / (k_Y + Y))
+    X, Y,
+    g_max,
+    k_X,
+    k_Y) {
+  g_max * (X / (k_X + X)) * (Y / (k_Y + Y))
 }

### 2023-11-13 R/new_CB_strain_parameter.R

@@ -2,13 +2,13 @@
 #'
 #' @param n number of strains
 #' @param values Allowed values are:
-#' - `"bush"`: default values from Bush et al (2017) \url{https://doi.org/10.1038/s41467-017-00912-x} will be used
+#' - `"bush"`: default values from Bush et al (2017) \doi{10.1038/s41467-017-00912-x} will be used
 #' - `"NA"`: all parameter will be set to `NA`. Usable for e.g. setting own parameter
-#'  
+#'
 #' @return object of class \code{CB_strain_parameter}. The object contains a
 #'   `data.frame` with 8 columns and `n` rows. The columns are:
-#' 
-#' columns: 
+#'
+#' columns:
 #' - strain_name: the name of the strain
 #' - g_max_CB:
 #' - k_CB_P:
@@ -17,15 +17,16 @@
 #' - Pr_CB:
 #' - m_CB:
 #' - i_CB:
-#' 
+#'
 #' @md
+#'
+#' @autoglobal
+#'
 #' @export
 #'
 new_CB_strain_parameter <- function(
-  n = 1,
-  values = "bush"
-){
-  
+    n = 1,
+    values = "bush") {
   # Check for supported values of the argument "values"
   if (is.na(values)) {
     values <- "NA"
@@ -33,7 +34,7 @@ new_CB_strain_parameter <- function(
   if (!(values %in% c("NA", "bush", "symmetric"))) {
     stop("Not supported value for `values`!\n", "Only NA, 'NA', 'bush' and 'symmetric' supported!")
   }
-  
+
   ## Create object
   x <- rep(as.numeric(NA), n)
   nm <- paste0("CB_", 1:n)
@@ -47,36 +48,38 @@ new_CB_strain_parameter <- function(
     m_CB = x,
     i_CB = x
   )
-  
+
   ## Add values for the case of Bush et al 2017
   if (values == "bush") {
-    result$g_max_CB = rep(0.05, n)
-    result$k_CB_P = rep(0.2, n)
-    result$h_SR_CB = rep(300, n)
-    result$y_P_CB = rep(1.67e8, n)
-    result$Pr_CB = rep(6e-9, n)
-    result$m_CB = rep(0.02, n)
-    result$i_CB = rep(0, n)
+    result$g_max_CB <- rep(0.05, n)
+    result$k_CB_P <- rep(0.2, n)
+    result$h_SR_CB <- rep(300, n)
+    result$y_P_CB <- rep(1.67e8, n)
[…]

### 2023-11-13 R/new_initial_state.R

@@ -4,26 +4,27 @@
 #' @param n_PB number of PB strains
 #' @param n_SB number of SB strains
 #' @param values Allowed values are:
-#' - `"bush_anoxic_fig2ab"`: initial conditions for Bush et al 2017 figure 2 a & b 
-#' - `"bush_oxic_fig2cd"`: initial conditions for Bush et al 2017 figure 2 c & d 
+#' - `"bush_anoxic_fig2ab"`: initial conditions for Bush et al 2017 figure 2 a & b
+#' - `"bush_oxic_fig2cd"`: initial conditions for Bush et al 2017 figure 2 c & d
 #' - `"bush_ssfig3"`: initial conditions for Bush et al 2017 figure e
 #' - `"NA"`: all initial values set to `NA`. Usable for e.g. setting own initial state
-#'   
+#'
 #' @return initial state \code{vector} aof the system s part of the strain_starter
 #' @md
+#'
+#' @autoglobal
+#'
 #' @export
 #'
 new_initial_state <- function(
-  n_CB = 1,
-  n_PB = 1,
-  n_SB = 1,
-  values = "bush_anoxic_fig2ab"
-){
-  
+    n_CB = 1,
+    n_PB = 1,
+    n_SB = 1,
+    values = "bush_anoxic_fig2ab") {
   if (is.na(values)) {
     values <- "NA"
   }
-  
+
   switch(
     EXPR = values,
     "bush_oxic_fig2cd" = {
@@ -32,48 +33,48 @@ new_initial_state <- function(
       SB <- 1e2
       SO <- 500
       SR <- 50
-      O  <- 300
-      P  <- 4
+      O <- 300
+      P <- 4
     },
     "bush_anoxic_fig2ab" = {
-      CB <-  5e1
-      PB <-  1e7
-      SB <-  1e7
-      SO <-  300
-      SR <-  300
-      O  <-  1e1
-      P  <-  1e1
+      CB <- 5e1
+      PB <- 1e7
+      SB <- 1e7
+      SO <- 300
+      SR <- 300
+      O <- 1e1
+      P <- 1e1
     },
     "bush_ssfig3" = {
-      CB <-  NA
-      PB <-  1e8
-      SB <-  1e8
-      SO <-  250
-      SR <-  350
-      O  <-  150
-      P  <-  9.5
+      CB <- NA
+      PB <- 1e8
+      SB <- 1e8
+      SO <- 250
+      SR <- 350
+      O <- 150
[…]

### 2023-11-13 R/new_replication_ssfind_parameter.R

@@ -16,11 +16,11 @@
 #' @md
 #' @export
 #'
+#' @autoglobal
+#'
 #' @examples
 new_replication_ssfind_parameter <- function(
-  ...
-){
-
+    ...) {
   p <- new_runsim_parameter()
   p$ss_expt <- NA
   if (!inherits(p, "replication_ssfind_parameter")) {

### 2023-11-13 R/plot_temporal_ss.R

@@ -1,82 +1,90 @@
 #' Plot the temporal dynamics of a model run of a symmetric parameter set
-#' 
+#'
 #' This is a convenience function to plot the dynamics of a model run particularly
 #' for a symmetric model
 #' with strains within functional groups displayed.
 #' @param temporal_results Results from \code{run_temporal_ssfind()} or from \code{run_temporal_ssfind_symmetric()}
-#' @return returns the ggplot object of the plot. If it is assigned to a 
+#' @return returns the ggplot object of the plot. If it is assigned to a
 #'    variable, the plot needs to be plotted, otherwise it is plotted.
-#' 
-#' @global
-#' 
-#' @importFrom dplyr filter mutate row_number case_when
-#' @importFrom ggplot2 ggplot aes facet_wrap xlab ylab geom_line scale_colour_manual guides guide_legend
+#'
+#' @autoglobal
+#'
+#' @importFrom dplyr filter mutate row_number case_when starts_with
+#' @importFrom ggplot2 ggplot aes facet_wrap xlab ylab geom_line scale_colour_manual guides guide_legend scale_x_continuous sec_axis ylim
 #' @importFrom tidyr gather
 #' @importFrom grDevices colorRampPalette
 #' @importFrom stringr str_detect
 #' @import patchwork
-#' 
+#'
 #' @export
-plot_temporal_ss <- function(temporal_results){
-
-  # wrangling if non symmetric simulation is put in 
+plot_temporal_ss <- function(temporal_results) {
+  # wrangling if non symmetric simulation is put in
   if ("direction" %in% colnames(temporal_results)) {
-    
-    temporal_results <- temporal_results %>% 
-      mutate(aO = a_O) %>% 
+    temporal_results <- temporal_results %>%
+      mutate(aO = a_O) %>%
       mutate(recovery = ifelse(direction == "up", "oxic", "anoxic"))
   }
-  
+
   p_organisms <- temporal_results %>%
-    select(aO, recovery,
-           starts_with("CB"),
-           starts_with("SB")) %>%
+    select(
+      aO, recovery,
+      starts_with("CB"),
+      starts_with("SB")
+    ) %>%
     gather(key = species, value = Quantity, 3:ncol(.)) %>%
-    ggplot() + 
-    geom_path(aes(x = aO, y = log10(Quantity + 1), col = species, linetype = recovery), linewidth = 1) + 
-    scale_color_manual(values = c("#38761d", "#cc0000" )) + 
+    ggplot() +
[…]

### 2023-11-13 R/run_simulation.R

@@ -1,5 +1,5 @@
 #' Run the simulation
-#' 
+#'
 #' This function takes the `parameter` object and runs a simulation based on these.
 #' It returns an object of class `runsim_result` which contains an additional
 #' entry, i.e. `result` which contains the results of the simulation. The
@@ -11,33 +11,35 @@
 #' @md
 #' @importFrom stats approx approxfun
 #' @importFrom deSolve ode
-#' 
+#'
+#' @autoglobal
+#'
 #' @export
 
 
 run_simulation <- function(
-  parameter
-){
+    parameter) {
   if (!inherits(parameter, "runsim_parameter")) {
     stop("parameter has to be an object of type `runsim_parameter`!")
   }
-  
-  if(parameter$sim_sample_interval > parameter$sim_duration){
+
+  if (parameter$sim_sample_interval > parameter$sim_duration) {
     stop("Simulation sample interval is greater than simulation duration... it should be shorter.")
-  } 
-  
+  }
+
   ## make the times at which observations will be recorded
   times <- seq(
     0,
     parameter$sim_duration,
     by = parameter$sim_sample_interval
   )
-  
+
   ## make the times at which events will occur
   event_times <- seq(min(times),
-                     max(times),
-                     by = parameter$event_interval)  
-  
+    max(times),
+    by = parameter$event_interval
+  )
+
   ## create the series of oxygen diffusivity values
   log10a_forcing <- matrix(
     ncol = 2,
@@ -47,22 +49,22 @@ run_simulation <- function(
         seq(
           0,
           max(times),
-          length=length(parameter$log10a_series)
+          length = length(parameter$log10a_series)
         )
       ),
       parameter$log10a_series
     )
   )
-       
-  ## Make the function to give the oxygen diffusivity at a particular time      
+
+  ## Make the function to give the oxygen diffusivity at a particular time
   l_f_f <- approxfun(
-    x = log10a_forcing[,1],
-    y = log10a_forcing[,2],
+    x = log10a_forcing[, 1],
+    y = log10a_forcing[, 2],
     method = "linear",
     rule = 2
   )
-  
-  if (is.null(parameter$event_definition)){
+
+  if (is.null(parameter$event_definition)) {
     events <- NULL
   } else {
     events <- list(
@@ -84,7 +86,7 @@ run_simulation <- function(
       minimum_abundances = parameter$minimum_abundances
     )
   )
-  
+
   result <- new_runsim_results(parameter, out)
 
   return(result)

### 2023-11-13 R/run_temporal_ssfind_symmetric.R

@@ -10,48 +10,61 @@
 #' @importFrom stats approx approxfun
 #' @importFrom deSolve ode
 #' @importFrom dplyr rename slice
-#' 
-#' @global time
-#' 
+#'
+#' @autoglobal
+#'
 #' @export
 
 run_temporal_ssfind_symmetric <- function(parameter) {
-  
   ## recalculate the wait time (length of a step)
-  wait_time <- parameter$sim_duration / mean(length(parameter$log10aO_series), 
-                                             length(parameter$log10aS_series))
-  
+  wait_time <- parameter$sim_duration / mean(
+    length(parameter$log10aO_series),
+    length(parameter$log10aS_series)
+  )
+
   ## sanity check
-  
+
   ## make function for the increasing oxygen diff steps
-  aO.up_l_f_f <- approxfun(x = wait_time * c(0:length(parameter$log10aO_series)), 
-                           y = (c(parameter$log10aO_series, parameter$log10aO_series[length(parameter$log10aO_series)])),
-                           method = "constant", rule = 1)
+  aO.up_l_f_f <- approxfun(
+    x = wait_time * c(0:length(parameter$log10aO_series)),
+    y = (c(parameter$log10aO_series, parameter$log10aO_series[length(parameter$log10aO_series)])),
+    method = "constant", rule = 1
+  )
   ## make function for the decreasing oxygen diff steps
-  aO.down_l_f_f <- approxfun(x = wait_time * c(0:length(parameter$log10aO_series)), 
-                             y = (c(rev(parameter$log10aO_series), parameter$log10aO_series[1])),
-                             method = "constant", rule = 1)
-  
+  aO.down_l_f_f <- approxfun(
+    x = wait_time * c(0:length(parameter$log10aO_series)),
+    y = (c(rev(parameter$log10aO_series), parameter$log10aO_series[1])),
+    method = "constant", rule = 1
+  )
+
   ## make function for the decreasing sulfide diff steps
-  aS.down_l_f_f <- approxfun(x = wait_time * c(0:length(parameter$log10aS_series)), 
-                             y = c(parameter$log10aS_series, parameter$log10aS_series[length(parameter$log10aS_series)]),
-                             method = "constant", rule = 1)
+  aS.down_l_f_f <- approxfun(
+    x = wait_time * c(0:length(parameter$log10aS_series)),
+    y = c(parameter$log10aS_series, parameter$log10aS_series[length(parameter$log10aS_series)]),
[…]

### 2023-11-13 R/set_diffusivities.R

@@ -1,40 +1,41 @@
 #' Used to update the asymmetry factor with logaO_series & log10aS_series accordingly
-#' 
-#' The `set_diffusivities` enables manipulating `aS` forcing in asymmetric manner to decrease (<1) or increase (>1) stress on cyanobacteria. If the vector in 
+#'
+#' The `set_diffusivities` enables manipulating `aS` forcing in asymmetric manner to decrease (<1) or increase (>1) stress on cyanobacteria. If the vector in
 #' `parameter$log10a_series` has multiple local minima, maxima respectively, it  is directly mirrored. ATTENTION: Mean value of the series has to be the axis at which the
 #' vector is mirrored!
-#' 
+#'
 #' @param param         :parameter set of type `runsim_parameter`
 #' @param asym_factor   :asym_factor to update `runsim_parameter`
-#' 
+#'
 #' @return updated parameterset
-#' 
+#'
 #' @md
+#'
+#' @autoglobal
+#'
 #' @export
 
 
-set_diffusivities <- function(param, 
-                              asym_factor = 1
-) {
-  if (!inherits(param, "runsim_parameter")){
+set_diffusivities <- function(param,
+                              asym_factor = 1) {
+  if (!inherits(param, "runsim_parameter")) {
     stop("Provide parameter object of class `runsim_parameter`")
   }
   # check whether vector is used for time dynamics
-  if (length(which(param$log10a_series == max(param$log10a_series))) > 1 | length(which(param$log10a_series == min(param$log10a_series))) > 1){
+  if (length(which(param$log10a_series == max(param$log10a_series))) > 1 | length(which(param$log10a_series == min(param$log10a_series))) > 1) {
     param$log10aO_series <- param$log10a_series
     param$log10aS_series <- 2 * mean(param$log10a_series) - param$log10a_series
-  } 
-  else {
+  } else {
     param$asym_factor <- asym_factor
     ## extract diffusivity series
     sym.axis <- mean(param$log10a_series) # symmetry axis is between log10a_series limits
     param$log10aO_series <- param$log10a_series
     # mirror the log10_series to obtain symmetry between aO <--> aS(x) at sym.axis (y)
-    param$log10aS_series <- 2*sym.axis - (seq(sym.axis - (abs(sym.axis - min(param$log10a_series)) * param$asym_factor),
-                                              sym.axis + (abs(sym.axis - min(param$log10a_series)) * param$asym_factor),
[…]

### 2023-11-13 R/set_temporal_ssfind_initial_state.R

@@ -9,30 +9,35 @@
 #' @return The passed parameter object but with initial states set (overwritten by the new ones)
 #' @md
 #'
+#' @autoglobal
+#'
 #' @export
 set_temporal_ssfind_initial_state <- function(p,
                                               initial_total_CB,
                                               initial_total_PB,
                                               initial_total_SB) {
-
-  CBs <- unlist(rep(initial_total_CB /
-                      length(grep("CB", names(p$strain_parameter$initial_state))),
-                    length(grep("CB", names(p$strain_parameter$initial_state)))))
+  CBs <- unlist(rep(
+    initial_total_CB /
+      length(grep("CB", names(p$strain_parameter$initial_state))),
+    length(grep("CB", names(p$strain_parameter$initial_state)))
+  ))
   names(CBs) <- NULL
   p$strain_parameter$initial_state[grep("CB", names(p$strain_parameter$initial_state))] <- CBs
 
 
-  PBs <- unlist(rep(initial_total_PB/length(grep("PB", names(p$strain_parameter$initial_state))),
-                    length(grep("PB", names(p$strain_parameter$initial_state)))))
+  PBs <- unlist(rep(
+    initial_total_PB / length(grep("PB", names(p$strain_parameter$initial_state))),
+    length(grep("PB", names(p$strain_parameter$initial_state)))
+  ))
   names(PBs) <- NULL
   p$strain_parameter$initial_state[grep("PB", names(p$strain_parameter$initial_state))] <- PBs
 
-  SBs <- unlist(rep(initial_total_SB/length(grep("SB", names(p$strain_parameter$initial_state))),
-                    length(grep("SB", names(p$strain_parameter$initial_state)))))
+  SBs <- unlist(rep(
+    initial_total_SB / length(grep("SB", names(p$strain_parameter$initial_state))),
+    length(grep("SB", names(p$strain_parameter$initial_state)))
+  ))
   names(SBs) <- NULL
   p$strain_parameter$initial_state[grep("SB", names(p$strain_parameter$initial_state))] <- SBs
 
   p
 }
-
-

### 2023-11-13 R/add_strain_var.R
merge symmetry project code (#39)
* Update bibliography and finalizing manuscript.

* Final draft

* Add event_definition = NULL

* Integrate project from @pascalbartschi

Functions were extended with the 'symmetric' parameter as well as a new ODI function and two mote plot functions.

* Delete manuscript_bibliography.bib

* v0.9.1.5 update

* v0.9.1.1: Changes from Pascal

* v0.9.1.1: Symmetric

* new_runsim_parameter list now holds sym_axis value

* added function

* added run_simulation_symmetric check and documentation

* added forcing of aO AND aS, as well as updated documentation

* even

@@ -26,6 +26,9 @@
 #'   mentioned variables. 
 #'
 #' @return the value of strain_parameter with added strain variability
+#' 
+#' @autoglobal
+#' 
 #' @export
 #'
 #' @md

### 2023-11-13 R/event_definition_symmetric.R

@@ -1,0 +1,83 @@
+#' Event definition for the symmetric simulation.
+#'
+#' This function contains events that can alter the state variables.
+#' In this event definition, the abundances of all functional groups have 1 added to their density at each event.
+#' @param times the time points in the simulation when events happen
+#' @param state current state variable values
+#' @param parms object of class `strain_parameter`
+#' @param log10aO_forcing_func function to change oxygen diffusivity `aO`
+#'   depending on `t`
+#' @param log10aS_forcing_func function. to change sulfide diffusivity `aS`
+#'    depending on `t`
+#' @param noise_sigma value of variation added to `SO`, `SR`, `O`, and `P` using
+#'   the formula `rnorm(1, 0, noise_sigma * X) * noise_sigma * X`
+#' @param minimum_abundances minimum abundances for CB, PB and SB. if the values
+#'   are lower, they will be set to these `minimum_abundances`
+#'
+#' @return the state vector, i.e. a vector containing the state variables.
+#'
+#' @md
+#'
+#' @autoglobal
+#'
+#' @export
+event_definition_symmetric <- function(
+    times,
+    state,
+    parms,
+    log10aO_forcing_func,
+    log10aS_forcing_func,
+    noise_sigma,
+    minimum_abundances) {
+  with(
+    as.list(state),
+    {
+      ## here adding a bit of noise to the abiotics
+      # if(noise_sigma!=0) {
+      SO <- SO + rnorm(1, 0, noise_sigma * SO) * noise_sigma * SO
+      SR <- SR + rnorm(1, 0, noise_sigma * SR) * noise_sigma * SR
+      O <- O + rnorm(1, 0, noise_sigma * O) * noise_sigma * O
+      P <- P + rnorm(1, 0, noise_sigma * P) * noise_sigma * P
+
+      ## and below setting the abundance to the minimum, in case it happens to be below it
+      CB <- state[grep("CB", names(state))]
+      names_CB <- names(CB)[order(names(CB))]
+      CB <- as.numeric(CB[order(names(CB))])
+      CB <- CB + minimum_abundances["CB"]
+
+      PB <- state[grep("PB", names(state))]
+      names_PB <- names(PB)[order(names(PB))]
+      PB <- as.numeric(PB[order(names(PB))])
+      PB <- PB + minimum_abundances["PB"]
+
+      SB <- state[grep("SB", names(state))]
+      names_SB <- names(SB)[order(names(SB))]
[…]

### 2023-11-13 R/get_final_states_a_N.R

@@ -5,12 +5,13 @@
 #' @param parameter  object of class \code{replication_ssfind_parameter} as generated by \code{new_replication_ssfind_parameter()}
 #'
 #' @return A vector of steady states
+#'
+#' @autoglobal
+#'
 #' @export
 get_final_states_a_N <- function(
-  x,
-  parameter
-){
-
+    x,
+    parameter) {
   ## set the oxygen diffusivity in the parameters to the required value
   parameter$a_O <- x["a_O"]
 
@@ -42,7 +43,7 @@ get_final_states_a_N <- function(
   names(SBs) <- NULL
   parameter$strain_parameter$initial_state[grep("SB", names(parameter$strain_parameter$initial_state))] <- SBs
 
- ## update forcing series with the required oxygen diffus. value
+  ## update forcing series with the required oxygen diffus. value
   parameter$log10a_series <- c(
     log10(parameter[["a_O"]]),
     log10(parameter[["a_O"]])
@@ -54,6 +55,5 @@ get_final_states_a_N <- function(
   )
 
   ## return only the final values of the state variables
-  simres$result[2,-1]
-
+  simres$result[2, -1]
 }

### 2023-11-13 R/get_stability_measures.R

@@ -1,5 +1,5 @@
 #' Get various measures of the stability
-#' 
+#'
 #' The measures include non-linearity and hysteresis measures, of an ecosystem
 #' response to environmental change. Takes steady state data as the input.
 #' @param ss_object object from which to calculate the stability measures. At
@@ -7,21 +7,21 @@
 #' @param threshold_diff_log10scale Amount of different between up and down
 #'   states to qualify as alternate stable states, on log10 scale
 #' @param ... additional arguments for methods
-#' 
+#'
 #' @return A data frame of stability measures of each state variable
 #'
 #' @md
 #'
 #' @rdname get_stability_measures
+#'
+#' @autoglobal
+#'
 #' @export
 #'
 #'
 get_stability_measures <- function(
-  ss_object,
-  threshold_diff_log10scale,
-  ...
-) {
-  
+    ss_object,
+    threshold_diff_log10scale,
+    ...) {
   UseMethod("get_stability_measures")
-  
 }

### 2023-11-13 R/get_stability_measures.replication_ssfind_result.R

@@ -1,87 +1,90 @@
-#' 
+#'
 #' The function `get_stability_measures.replication_ssfind_result` extracts the `result`
-#' `data.frame` or `tibble` from the object `x` and processes it. If one wants to extract. If 
+#' `data.frame` or `tibble` from the object `x` and processes it. If one wants to extract. If
 #' one extracts the `results` manually, the function `get_stability_measures_replication_ssfind_result`
 #' needs to be used.
 #' @rdname get_stability_measures
-#' 
-#' @global Quantity starts_with direction up down a_O a 
-#' 
+#'
+#' @autoglobal
+#'
 #' @md
-#' 
+#'
 #' @importFrom tidyr gather spread
-#' @importFrom dplyr summarise pull filter across
+#' @importFrom dplyr summarise pull filter across starts_with
 #' @importFrom stats na.omit
-#' 
+#'
+#' @autoglobal
+#'
 #' @export
 get_stability_measures.replication_ssfind_result <- function(
-  ss_object,
-  threshold_diff_log10scale = 3,
-  ...
-){
+    ss_object,
+    threshold_diff_log10scale = 3,
+    ...) {
   if (inherits(ss_object, "replication_ssfind_result")) {
     result <- ss_object$result
   } else {
     result <- ss_object
   }
-  
-  
+
+
   ## which initial condition varies?
   init_varying <- NA
-  if(length(unique(result$initial_N_CB)) > 1)
+  if (length(unique(result$initial_N_CB)) > 1) {
     init_varying <- "initial_N_CB"
-  if(length(unique(result$initial_N_SB)) > 1)
+  }
+  if (length(unique(result$initial_N_SB)) > 1) {
     init_varying <- "initial_N_SB"
-  if(length(unique(result$initial_N_PB)) > 1)
+  }
+  if (length(unique(result$initial_N_PB)) > 1) {
     init_varying <- "initial_N_PB"
-  
-  if(!is.na(init_varying)) {
-    
-    result$init_varying <- dplyr::pull(result[,init_varying],1)
-    
+  }
+
+  if (!is.na(init_varying)) {
+    result$init_varying <- dplyr::pull(result[, init_varying], 1)
+
     min_iniN <- min(result$init_varying)
     max_iniN <- max(result$init_varying)
-    
+
     ## The following is preparing the data
     these <- grep("B_", names(result))
     these <- c(these, which(names(result) %in% c("SO", "SR", "O", "P")))
[…]

### 2023-11-13 R/get_symmetry_measurements.R

@@ -1,0 +1,163 @@
+#' Computes measures that make comparisons between collapse and recovery
+#' trajectory and between antagonistic environmental variables possible.
+#'
+#' @param res Results from \code{run_temporal_ssfind()} or from \code{run_temporal_ssfind_symmetric()}
+#' @return returns a frame containing symmetry measures for each variable, such as hysteresis area,
+#'    shift magnitudes or distance between TP, as well as TP themselfes with their according state.
+#'
+#' @autoglobal
+#'
+#' @importFrom dplyr rename mutate select filter slice all_of starts_with
+#' @importFrom DescTools AUC
+#'
+#' @autoglobal
+#'
+#' @export
+get_symmetry_measurements <- function(res) {
+  ## check whether bush or symmetry simulation prepare dataframe
+  if ("a_O" %in% colnames(res)) {
+    res <- res %>%
+      rename(aO = a_O) %>%
+      mutate(recovery = ifelse(direction == "up", "oxic", "anoxic")) %>%
+      select(-direction)
+  } else {
+    res <- res %>%
+      select(-starts_with("PB"), -aS, -SO)
+  }
+
+  ## extract species names
+  all_species <- colnames(res)[!(colnames(res) %in% c("time", "aO", "recovery"))]
+
+  ## prepare storage df
+  measures <- data.frame(
+    anox_TP = rep(NA, length(all_species)),
+    anox_TPafter = rep(NA, length(all_species)),
+    state_anox_TP = rep(NA, length(all_species)),
+    state_anox_TPafter = rep(NA, length(all_species)),
+    ox_TP = rep(NA, length(all_species)),
+    ox_TPafter = rep(NA, length(all_species)),
+    state_ox_TP = rep(NA, length(all_species)),
+    state_ox_TPafter = rep(NA, length(all_species)),
+    hyst_area = rep(NA, length(all_species)),
+    hyst_range = rep(NA, length(all_species)),
+    abs_shift_recovery = rep(NA, length(all_species)),
+    abs_shift_catastrophy = rep(NA, length(all_species))
+  )
+
+  rownames(measures) <- all_species
+
+  ## for every species do
+  for (i in 1:length(all_species)) {
+    ## for every species in res calc sym measures
+    for (dir in c("oxic", "anoxic")) {
+      ## take local derivative at each point
+      # delta y
[…]

### 2023-11-13 R/globals.R

@@ -2,21 +2,28 @@
 
 utils::globalVariables(c(
   "Quantity", # <get_stability_measures.replication_ssfind_result>
-  "starts_with", # <get_stability_measures.replication_ssfind_result>
   "direction", # <get_stability_measures.replication_ssfind_result>
   "up", # <get_stability_measures.replication_ssfind_result>
   "down", # <get_stability_measures.replication_ssfind_result>
   "a_O", # <get_stability_measures.replication_ssfind_result>
   "a", # <get_stability_measures.replication_ssfind_result>
+  "recovery", # <get_stability_measures.temporal_ssfind_result>
+  "aO", # <get_stability_measures.temporal_ssfind_result>
+  "aS", # <get_stability_measures.temporal_ssfind_result>
   "Quantity", # <get_stability_measures.temporal_ssfind_result>
-  "starts_with", # <get_stability_measures.temporal_ssfind_result>
+  "time", # <get_stability_measures.temporal_ssfind_result>
   "direction", # <get_stability_measures.temporal_ssfind_result>
+  "Species", # <get_stability_measures.temporal_ssfind_result>
   "up", # <get_stability_measures.temporal_ssfind_result>
   "down", # <get_stability_measures.temporal_ssfind_result>
   "a_O", # <get_stability_measures.temporal_ssfind_result>
-  "a", # <get_stability_measures.temporal_ssfind_result>
-  "Species", # <get_stability_measures.temporal_ssfind_result>
-  "time", # <get_stability_measures.temporal_ssfind_result>
+  "a_O", # <get_symmetry_measurements>
+  "direction", # <get_symmetry_measurements>
+  "aS", # <get_symmetry_measurements>
+  "SO", # <get_symmetry_measurements>
+  "recovery", # <get_symmetry_measurements>
+  ".", # <get_symmetry_measurements>
+  "a", # <plot_dynamics>
   "species", # <plot_dynamics>
   "quantity", # <plot_dynamics>
   ".", # <plot_dynamics>
@@ -24,8 +31,59 @@ utils::globalVariables(c(
   "time", # <plot_dynamics>
   "log10_quantity", # <plot_dynamics>
   "var_type", # <plot_dynamics>
[…]

### 2023-11-13 R/new_PB_strain_parameter.R

@@ -2,13 +2,13 @@
 #'
 #' @param n number of strains
 #' @param values Allowed values are:
-#' - `"bush"`: default values from Bush et al (2017) \url{https://doi.org/10.1038/s41467-017-00912-x} will be used
+#' - `"bush"`: default values from Bush et al (2017) \doi{10.1038/s41467-017-00912-x} will be used
 #' - `"NA"`: all parameter will be set to `NA`. Usable for e.g. setting own parameter
 #'
 #' @return object of class \code{PB_strain_parameter}. The object contains a
 #'   `data.frame` with 8 columns and `n` rows. The columns are:
-#' 
-#' columns: 
+#'
+#' columns:
 #' - strain_name: the name of the strain
 #' - g_max_PB:
 #' - k_PB_SR:
@@ -18,23 +18,24 @@
 #' - y_P_PB:
 #' - m_PB:
 #' - i_CB:
-#' 
-#' @md 
+#'
+#' @md
+#'
+#' @autoglobal
+#'
 #' @export
 #'
 new_PB_strain_parameter <- function(
-  n = 1,
-  values = "bush"
-){
-  
+    n = 1,
+    values = "bush") {
   # Check for supported values of the argument "values"
   if (is.na(values)) {
     values <- "NA"
   }
-  if (!(values %in% c("NA", "bush"))) {
-    stop("Not supported value for `values`!\n", "Only NA, 'NA' and 'bush' supported!")
+  if (!(values %in% c("NA", "bush", "symmetric"))) {
+    stop("Not supported value for `values`!\n", "Only NA, 'NA', 'bush' and 'symmetric' supported!")
   }
-  
+
   ## Create object
   x <- rep(as.numeric(NA), n)
   nm <- paste0("PB_", 1:n)
@@ -49,18 +50,31 @@ new_PB_strain_parameter <- function(
     m_PB = x,
     i_PB = x
   )
-  
+
   ## Add values for the case of Bush et al 2017
   if (values == "bush") {
-    result$g_max_PB = rep(0.07, n)
-    result$k_PB_SR = rep(10, n)
-    result$k_PB_P = rep(0.5, n)
-    result$h_O_PB = rep(100, n)
-    result$y_SR_PB = rep(1.25e7, n)  ## Y_S_PB in the main text
-    result$y_P_PB = rep(1.67e8, n)
-    result$m_PB = rep(0.028, n)
[…]

### 2023-11-13 R/new_runsim_result.R

@@ -6,25 +6,27 @@
 #' @param result a `data.frame` containing the results of the simulation
 #'
 #' @return object of the class `runsim_result`. This object of class
-#'   `runsim_result` is identical to an object of class `runsim_parameter` 
-#'   plus an additional field `result` which contains the result of the simulation. 
+#'   `runsim_result` is identical to an object of class `runsim_parameter`
+#'   plus an additional field `result` which contains the result of the simulation.
 #' @md
+#'
+#' @autoglobal
+#'
 #' @export
 #'
 #' @examples
 new_runsim_results <- function(
-  parameter, 
-  result
-){
+    parameter,
+    result) {
   if (!inherits(parameter, "runsim_parameter")) {
     stop("`parameter` has to be of class runsim_parameter")
   }
-  
+
   p <- parameter
   p$result <- result
   if (!inherits(p, "runsim_result")) {
     class(p) <- append(class(p), "runsim_result")
   }
-  
+
   return(p)
 }

### 2023-11-13 R/plot_dynamics.R

@@ -1,84 +1,85 @@
 #' Plot the dynamics of a model run
-#' 
+#'
 #' This is a convenience function to plot the dynamics of a model run,
 #' with strains within functional groups displayed.
 #' @param simulation_result Object returned by the run_simulation function
 #' @param every_n Plot data of every other n sample.
-#' @return returns the ggplot object of the plot. If it is assigned to a 
+#' @return returns the ggplot object of the plot. If it is assigned to a
 #'    variable, the plot needs to be plotted, otherwise it is plotted.
-#' 
-#' @global species quantity . functional_group time log10_quantity var_type 
-#' 
+#'
+#' @autoglobal
+#'
 #' @importFrom dplyr filter mutate row_number case_when
 #' @importFrom ggplot2 ggplot aes facet_wrap xlab ylab geom_line scale_colour_manual guides guide_legend
 #' @importFrom tidyr gather
 #' @importFrom grDevices colorRampPalette
 #' @importFrom stringr str_detect
 #' @import patchwork
-#' 
+#'
 #' @export
 plot_dynamics <- function(
-  simulation_result, 
-  every_n = 1
-){
-  
+    simulation_result,
+    every_n = 1) {
   ## define colours
   colfunc_CB <- grDevices::colorRampPalette(c("#024F17", "#B5FFC9"))
   colfunc_SB <- grDevices::colorRampPalette(c("#7D1402", "#FCBEB3"))
   colfunc_PB <- grDevices::colorRampPalette(c("#6E0172", "#F9AEFC"))
-  
+
   ## data wrangling
   temp <- simulation_result$result %>%
-    dplyr::filter(dplyr::row_number() %% every_n == 0 ) %>%
+    dplyr::filter(dplyr::row_number() %% every_n == 0) %>%
     dplyr::mutate(a = 10^a) %>%
-    tidyr::gather(species, quantity, 2:ncol(.)) %>% 
-    dplyr::mutate(var_type=ifelse(grepl("B_", species), "Organism", "Substrate"),
-           functional_group = dplyr::case_when(str_detect(species, "CB_") ~ "CB",
-                                        str_detect(species, "SB_") ~ "SB",
-                                        str_detect(species, "PB_") ~ "PB"),
-           log10_quantity=log10(quantity))
+    tidyr::gather(species, quantity, 2:ncol(.)) %>%
[…]

### 2023-11-13 R/plot_symmetry_measurements.R

@@ -1,0 +1,76 @@
+#' Visualizes the symmetry measures obtained by the function \code{get_symmetry_measures()}
+#'
+#' @param res Results from \code{run_temporal_ssfind()} or from \code{run_temporal_ssfind_symmetric()}
+#' @param species Environmental varibale of which the symmetry measures shoudl be plotted
+#' @return ggplot object, with shifts visualized as arrows and hysteresis area ribbon
+#'
+#' @autoglobal
+#'
+#' @importFrom dplyr rename mutate select filter starts_with
+#' @importFrom ggplot2 geom_path geom_point geom_segment geom_ribbon scale_color_manual labs arrow scale_fill_manual unit
+#' @importFrom stats median
+#'
+#' @export
+plot_symmetry_measures <- function(res,
+                                   species) {
+  measures <- get_symmetry_measurements(res)
+  ## check whether bush or symmetry simulation prepare dataframe scale_fill_manual scale_color_manual labs
+  if ("a_O" %in% colnames(res)) {
+    res <- res %>%
+      rename(aO = a_O) %>%
+      mutate(recovery = ifelse(direction == "up", "oxic", "anoxic")) %>%
+      select(-direction)
+  } else {
+    res <- res %>%
+      select(-starts_with("PB"), -aS, -SO)
+  }
+  if (!(species %in% colnames(res))) {
+    stop(paste("Please select a species of", paste(colnames(res), collapse = ", "), "."))
+  }
+  # speciesmeasures <- gather(measures[species,], key = "measure", value = "value")
+  ## data wrangling to plotting-mature frame with xy coords in seperate columns
+
+  plot_df <- data.frame(
+    aO = as.numeric(select(measures[species, ], c(1, 2, 5, 6))),
+    state = as.numeric(select(measures[species, ], c(3, 4, 7, 8))),
+    measure = as.character(colnames(measures)[c(1, 2, 5, 6)])
+  )
+  ribbon_temp <- res %>%
+    filter((aO >= median(measures$ox_TP)), (aO <= median(measures$anox_TP))) %>%
+    select(c(species, recovery, aO))
+  ribbon_df <- data.frame(
+    oxic = rev(ribbon_temp %>% filter(recovery == "oxic") %>% pull(species)),
+    anoxic = ribbon_temp %>% filter(recovery == "anoxic") %>% pull(species),
[…]

### 2023-11-13 R/plot_trajectory_symmetry.R

@@ -1,0 +1,187 @@
+#' Visualizes the symmetry of antagonistic collapse or recovery trajectories, symmetry measures of shift
+#' are computed by the function \code{get_symmetry_measurements()}
+#'
+#' @param res Results from \code{run_temporal_ssfind()} or from \code{run_temporal_ssfind_symmetric()}
+#' @param trajectory Trajectory to be plotted. Either `collapse` or `recovery`.`.
+#' @param typ Type of environmental variable. Either `substrate` or `bacteria`.
+#' @param plot_log10 Logical. If \code{TRUE}, logarithmic plot of the data is done.
+#' @return ggplot object, with antagonistic shifts of related trajectory visualized as arrows.
+#'
+#' @autoglobal
+#'
+#' @importFrom dplyr rename mutate select filter recode starts_with
+#' @importFrom tidyr gather spread
+#' @importFrom ggplot2 geom_path geom_point geom_segment scale_color_manual labs arrow scale_x_continuous sec_axis unit
+#'
+#' @export
+
+plot_trajectory_symmetry <- function(res,
+                                     trajectory = "recovery",
+                                     typ = "substrate",
+                                     plot_log10 = FALSE) {
+  ## extract symmetry measure from result
+  measures <- get_symmetry_measurements(res)
+
+  ## check whether bush or symmetry simulation and prepare data frame accordingly
+  if ("a_O" %in% colnames(res)) {
+    res <- res %>%
+      rename(aO = a_O) %>%
+      mutate(recovery = ifelse(direction == "up", "oxic", "anoxic")) %>%
+      select(-direction)
+    sym.flag <- F
+  } else {
+    res <- res %>%
+      select(-starts_with("PB"), -SO)
+    sym.flag <- T
+  }
+
+  if (!(typ %in% c("bacteria", "substrate"))) {
+    stop("Type must be either 'bacteria' 'or substrate'")
+  }
+
+  # if (nrow(measures) != ncol(select(res, -c(time, aO, recovery)))){
+  #   stop("Ensure symmetry measure input is from according data ")
+  # }
+
+  # get measures in long format
+  measures <- measures %>%
+    mutate(species = rownames(.)) %>%
+    gather(key = "measure", value = "value", -species) %>%
[…]

### 2023-11-13 R/run_replication_ssfind.R

@@ -9,18 +9,18 @@
 #'
 #' @md
 #'
+#' @autoglobal
+#'
 #' @export
 
 run_replication_ssfind <- function(
-  parameter,
-  mc.cores = getOption("mc.cores", 0)
-){
-
+    parameter,
+    mc.cores = getOption("mc.cores", 0)) {
   if (!inherits(parameter, "replication_ssfind_parameter")) {
     stop("parameter has to be an object of type `replication_ssfind_parameter`!")
   }
 
-  if (Sys.info()['sysname'] == "Window") {
+  if (Sys.info()["sysname"] == "Window") {
     warning("On windows, mc.cores will always be set to one!")
     mc.cores <- 1
   }
@@ -35,8 +35,8 @@ run_replication_ssfind <- function(
         get_final_states_a_N(x, parameter)
       }
     )
-  ## multi core
-  ## mcapply over rows of ss_expt
+    ## multi core
+    ## mcapply over rows of ss_expt
   } else {
     temp_result <- parallel::mclapply(
       1:nrow(parameter$ss_expt),
@@ -55,10 +55,12 @@ run_replication_ssfind <- function(
   result <- temp_result %>%
     tibble::tibble() %>%
     tidyr::unnest(cols = 1) %>%
-    dplyr::mutate(initial_N_CB = parameter$ss_expt$N_CB,
-                  initial_N_PB = parameter$ss_expt$N_PB,
-                  initial_N_SB = parameter$ss_expt$N_SB,
-                  a_O = parameter$ss_expt$a_O)
+    dplyr::mutate(
+      initial_N_CB = parameter$ss_expt$N_CB,
+      initial_N_PB = parameter$ss_expt$N_PB,
+      initial_N_SB = parameter$ss_expt$N_SB,
+      a_O = parameter$ss_expt$a_O
+    )
 
-   return(new_replication_ssfind_results(parameter, result))
+  return(new_replication_ssfind_results(parameter, result))
 }

### 2023-11-13 R/set_diffusivities.R

@@ -1,0 +1,41 @@
+#' Used to update the asymmetry factor with logaO_series & log10aS_series accordingly
+#'
+#' The `set_diffusivities` enables manipulating `aS` forcing in asymmetric manner to decrease (<1) or increase (>1) stress on cyanobacteria. If the vector in
+#' `parameter$log10a_series` has multiple local minima, maxima respectively, it  is directly mirrored. ATTENTION: Mean value of the series has to be the axis at which the
+#' vector is mirrored!
+#'
+#' @param param         :parameter set of type `runsim_parameter`
+#' @param asym_factor   :asym_factor to update `runsim_parameter`
+#'
+#' @return updated parameterset
+#'
+#' @md
+#'
+#' @autoglobal
+#'
+#' @export
+
+
+set_diffusivities <- function(param,
+                              asym_factor = 1) {
+  if (!inherits(param, "runsim_parameter")) {
+    stop("Provide parameter object of class `runsim_parameter`")
+  }
+  # check whether vector is used for time dynamics
+  if (length(which(param$log10a_series == max(param$log10a_series))) > 1 | length(which(param$log10a_series == min(param$log10a_series))) > 1) {
+    param$log10aO_series <- param$log10a_series
+    param$log10aS_series <- 2 * mean(param$log10a_series) - param$log10a_series
+  } else {
+    param$asym_factor <- asym_factor
+    ## extract diffusivity series
+    sym.axis <- mean(param$log10a_series) # symmetry axis is between log10a_series limits
+    param$log10aO_series <- param$log10a_series
+    # mirror the log10_series to obtain symmetry between aO <--> aS(x) at sym.axis (y)
+    param$log10aS_series <- 2 * sym.axis - (seq(sym.axis - (abs(sym.axis - min(param$log10a_series)) * param$asym_factor),
+      sym.axis + (abs(sym.axis - min(param$log10a_series)) * param$asym_factor),
+      length = length(param$log10a_series)
+    ))
+  }
+
+  return(param)
+}

### 2023-11-13 R/symmetric_bushplus_dynamic_model.R

@@ -1,0 +1,105 @@
+#' The rate equations, modified from publication of Bush et al 2017
+#' \doi{10.1093/clinchem/39.5.766}. Equations of antagonistic
+#' environmental variables are identical, while the oxygen and sulfide
+#' diffusivities are forced in antisymmetric manner.
+#' Possibility to simulate multiple strain per functional group as published in
+#' \doi{10.1111/ele.14217} by Limberger et al 2023 remains.
+#'
+#' @param t The current time in the simulation
+#' @param state A vector containing the current (named) values of each state
+#'   variable
+#' @param parameters An object of class `runsim_parameter` as returned by
+#'   `new_runsim_parameter()``
+#' @param log10aO_forcing_func function to change oxygen diffusivity `aO` depending on `t`
+#' @param log10aS_forcing_func function to change sulfide diffusivity `aS` depending on `t`
+#' @param ... not used. Needed to catch additional parameters.
+#'
+#' @return a list containing two elements, namely the rate of change of the
+#'   strains, and also the current values of oxygen diffusivity `aO`, as well as sulfide diffusivity `aS`.
+#' @md
+#'
+#' @autoglobal
+#'
+#' @export
+symmetric_bushplus_dynamic_model <- function(
+    t,
+    state,
+    parameters,
+    log10aO_forcing_func,
+    log10aS_forcing_func,
+    ...) {
+  ## unpack state variables from the state object
+  CB <- state[grep("CB", names(state))]
+  names_CB <- names(CB)[order(names(CB))]
+  CB <- as.numeric(CB[order(names(CB))])
+  PB <- state[grep("PB", names(state))]
+  names_PB <- names(PB)[order(names(PB))]
+  PB <- as.numeric(PB[order(names(PB))])
+  SB <- state[grep("SB", names(state))]
+  names_SB <- names(SB)[order(names(SB))]
+  SB <- as.numeric(SB[order(names(SB))])
+  # print(c(CB, PB, SB))
+
+
+  # rates of change of CB and SB
+  CB_growth_rate <- growth1(state["P"], parameters$CB$g_max_CB, parameters$CB$k_CB_P) * inhibition(state["SR"], parameters$CB$h_SR_CB) * CB
[…]

### 2024-02-23 R/globals.R
update of trajectory symmetry visualizing funcions for symmetry paper revision 1

@@ -80,6 +80,23 @@ utils::globalVariables(c(
   "yend", # <plot_trajectory_symmetry>
   "plot_species", # <plot_trajectory_symmetry>
   "state", # <plot_trajectory_symmetry>
+  "a_O", # <plot_trajectory_symmetry_compact>
+  "direction", # <plot_trajectory_symmetry_compact>
+  "SO", # <plot_trajectory_symmetry_compact>
+  ".", # <plot_trajectory_symmetry_compact>
+  "species", # <plot_trajectory_symmetry_compact>
+  "type", # <plot_trajectory_symmetry_compact>
+  "measure", # <plot_trajectory_symmetry_compact>
+  "value", # <plot_trajectory_symmetry_compact>
+  "aO", # <plot_trajectory_symmetry_compact>
+  "aS", # <plot_trajectory_symmetry_compact>
+  "recovery", # <plot_trajectory_symmetry_compact>
+  "x", # <plot_trajectory_symmetry_compact>
+  "y", # <plot_trajectory_symmetry_compact>
+  "xend", # <plot_trajectory_symmetry_compact>
+  "yend", # <plot_trajectory_symmetry_compact>
+  "plot_species", # <plot_trajectory_symmetry_compact>
+  "state", # <plot_trajectory_symmetry_compact>
   "time", # <run_temporal_ssfind>
   "a", # <run_temporal_ssfind>
   "pars", # <run_temporal_ssfind_experiment>

### 2024-02-23 R/plot_symmetry_measures.R

@@ -69,7 +69,7 @@ plot_symmetry_measures <- function(res,
       mapping = aes(x = aO, ymin = oxic, ymax = anoxic, fill = "hysteresis area"), alpha = 0.3
     ) +
     scale_fill_manual(values = c("grey")) +
-    scale_color_manual(values = c("#e85050", "#5da1df")) +
+    scale_color_manual(values = c("#5da1df", "#e85050")) +
     labs(x = "aO", y = species, fill = "measures", color = "measures", linetype = "recovery")
 
   return(p)

### 2024-02-23 R/plot_trajectory_symmetry_compact.R

@@ -1,0 +1,116 @@
+#' Visualizes the symmetry of upards and downwards trajectories in the same figure, 
+#' symmetry measures of shift are computed by the function \code{get_symmetry_measurements()}
+#'
+#' @param res Results from \code{run_temporal_ssfind()} or from \code{run_temporal_ssfind_symmetric()}
+#' @param typ Type of environmental variable. Either `substrate` or `bacteria`.
+#' @param plot_log10 Logical. If \code{TRUE}, logarithmic plot of the data is done.
+#' @return ggplot object, with antagonistic shifts of related trajectory visualized as arrows.
+#'
+#' @autoglobal
+#'
+#' @importFrom dplyr rename mutate select filter recode starts_with
+#' @importFrom tidyr gather spread
+#' @importFrom ggplot2 geom_path geom_point geom_segment scale_color_manual labs arrow scale_x_continuous sec_axis unit
+#'
+#' @export
+
+plot_trajectory_symmetry_compact <- function (res, typ = "substrate", plot_log10 = FALSE) 
+{
+  measures <- get_symmetry_measurements(res)
+  if ("a_O" %in% colnames(res)) {
+    res <- res %>% rename(aO = a_O) %>% mutate(recovery = ifelse(direction == 
+                                                                   "up", "oxic", "anoxic")) %>% select(-direction)
+    sym.flag <- F
+  } else {
+    res <- res %>% select(-starts_with("PB"), -SO)
+    sym.flag <- T
+  }
+  if (!(typ %in% c("bacteria", "substrate"))) {
+    stop("Type must be either 'bacteria' 'or substrate'")
+  }
+  measures <- measures %>% mutate(species = rownames(.)) %>% 
+    gather(key = "measure", value = "value", 
+           -species) %>% mutate(type = ifelse(species %in% c("SO", 
+                                                             "SR", "P", "O"), "substrate", 
+                                              "bacteria")) %>% filter(type == typ, !(species %in% 
+                                                                                       c("P", "SO"))) %>% spread(key = measure, 
+                                                                                                                 value = value)
+  
+  
+  res_ox <- res %>% {
+    if (sym.flag) 
+      select(., starts_with("CB"), "O", 
+             "aO", "aS", "recovery")
+    else select(., starts_with("CB"), "O", 
+                "aO", "recovery")
+  } %>% {
+    if (sym.flag) 
+      gather(data = ., key = "species", value = "state", 
+             -c(aO, aS, recovery))
[…]
